### 2022-02-09 manuscript/draft-manuscript.Rmd
add manuscript first draft (very messy!)

---
title: "Effectiveness of COVID-19 Booster doses in England: a sequential trials study in OpenSAFELY"
output:
  bookdown::html_document2:
    number_sections: no
    toc: no
  html_document:
    keep_md: yes
    self_contained: yes
  word_document: default
  pdf_document:
    toc: no
bibliography: references.bib
csl: vancouver.csl
link-citations: yes
zotero: yes
header-includes:
- \usepackage{float} 
- \floatplacement{figure}{H}
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
---

```{r setup, include=FALSE}

library('tidyverse')
library('here')
library('glue')
library('lubridate')
library('gt')
library('patchwork')
library('scales')
source(here("lib", "functions", "utility.R"))

# where are the outputs (ie the inputs for this manuscript!) saved?
#output_dir_os <- here("output", "manuscript-objects")
output_dir_os <- here("released_output", "manuscript-objects")

# where should we put the objects created within this rmd script?
output_dir_rmd <- here("manuscript", "figures")
fs::dir_create(output_dir_rmd)

# only applicable if `self-contained: yes` option is enables in yaml header
knitr::opts_chunk$set(
  echo = TRUE,
  fig.path = paste0(output_dir_rmd, "/"),
  fig.pos = 'H' #to stop figures floating around when rendering to pdf
)

study_dates <-
  jsonlite::read_json(path=here("lib", "design", "study-dates.json")) %>%
  map(as.Date)

postbaselinecuts <- read_rds(here("lib", "design", "postbaselinecuts.rds"))

flowchart <- read_csv(fs::path(output_dir_os, "flowchart.csv"))
cohortsummary <- read_csv(fs::path(output_dir_os, "cohortsummary.csv"))
matchsummary <- read_csv(fs::path(output_dir_os, "matchsummary.csv"))
matchsummary_treated <- read_csv(fs::path(output_dir_os, "matchsummary_treated.csv")) %>% filter(treated == 1L)
matchsummary_control <- read_csv(fs::path(output_dir_os, "matchsummary_treated.csv")) %>% filter(treated == 0L)
matchcoverage <- read_csv(fs::path(output_dir_os, "matchcoverage.csv"))
incidences <- read_csv(fs::path(output_dir_os, "incidence.csv")) %>% mutate(events=events_1+events_0)

```

```{r, effects, echo=FALSE, message=FALSE, warning=FALSE}

  specify_decimal <- function(x, k, trim=FALSE) {

    fmtd <- format(round(x, k), nsmall = k)
    if (trim) {fmtd <- trimws(fmtd)}
    return(fmtd)
  }
  #
  # print_est1bracket <- function(x, b, round=1){
  #   paste0(specify_decimal(x, round), " (", specify_decimal(b, round), ")")
  # }
  #
  print_est2bracket <- function(x, b1, b2, round=1){
    paste0(specify_decimal(x, round), " (", specify_decimal(b1, round), ", ", specify_decimal(b2, round), ")")
  }

  outcomes <- 
    c(
      "postest",
      "covidemergency"#,
      #"covidadmitted",
      #"covidcc",
      #"coviddeath"
    ) %>% 
    set_names(., .)

# 
# cmlinc <- read_csv(fs::path(output_dir_os, "cmlinc.csv")) %>%
#   mutate(
#     treatment_descr = fct_inorder(treatment_descr),
#     outcome_descr = fct_inorder(outcome_descr),
#     model_name = fct_inorder(str_wrap(model_name, 15)),
#     inc_CI_to = paste0(label_number(0.1, scale=1000)(1-survival.ul), " to ", label_number(0.1, scale=1000)(1-survival.ll))
#   )

# riskdiff <- read_csv(fs::path(output_dir_os, "riskdiff.csv")) %>%
#   mutate(
#     treatment_descr = fct_inorder(treatment_descr),
#     outcome_descr = fct_inorder(outcome_descr),
#     model_name = fct_inorder(str_wrap(model_name, 15)),
#     diff_CI_to = paste0(label_number(0.01, scale=1000)(diff.ll), " to ", label_number(0.01, scale=1000)(diff.ul))
#   )

hazards <- read_csv(fs::path(output_dir_os, "effects.csv")) %>%
  mutate(
    treatment_descr = fct_inorder(treatment_descr),
    outcome_descr = fct_inorder(outcome_descr),
    #model_name = fct_inorder(str_wrap(model_name, 15)),
    hr_display = paste0(number_format(accuracy=0.01)(hr), " (", number_format(accuracy=0.01)(hr.ll), "-", number_format(accuracy=0.01)(hr.ul), ")")
  )

#cmlinc3 <- cmlinc %>% filter(model==3)
#riskdiff3 <- riskdiff %>% filter(model==3)
hazards3 <- hazards %>% filter(model==3)

```

William J Hulme^1^, Elizabeth J Williamson^2^, Amelia Green^1^, Elsie Horne^4,5^, Linda Nab^1^, Ruth Keogh^2^, Edward Parker^2^, Jonathan AC Sterne^4,5^, Miguel Hernán^6,7^, Ben Goldacre^1^ + OpenSAFELY collaborative.

<!-- Krishnan Bhaskaran^2^, Helen I McDonald^2^, Christopher T Rentsch^2^, Anna Schultze^2^, John Tazare^2^, Helen J Curtis^1^, Alex J Walker^1^, Laurie Tomlinson^2^, Tom Palmer^4,5^,  Brian MacKenna^1^, Caroline E Morton^1^, Amir Mehrkar^1^, Louis Fisher^1^, Seb Bacon^1^, Dave Evans^1^, Peter Inglesby^1^, George Hickman^1^, Simon Davy^1^, Tom Ward ^1^, Richard Croker^1^, Rosalind M Eggo^2^, Angel YS Wong^2^, Rohini Mathur^2^, Kevin Wing^2^, Harriet Forbes^2^, Daniel Grint^2^, Ian J Douglas^2^, Stephen JW Evans^2^, Liam Smeeth^2^, Chris Bates^3^, Jonathan Cockburn^3^, John Parry^3^, Frank Hester^3^, Sam Harper^3^,  -->

1\. The DataLab, Nuffield Department of Primary Care Health Sciences, University of Oxford, OX26GG, UK

2\. London School of Hygiene and Tropical Medicine, Keppel Street, London, WC1E 7HT, UK

3\. TPP, TPP House, 129 Low Lane, Horsforth, Leeds, LS18 5PX

4\. Population Health Sciences, University of Bristol, Oakfield House, Oakfield Grove, Bristol, BS8 2BN, UK

5\. NIHR Bristol, Biomedical Research Centre, Bristol, UK

6\. CAUSALab, Harvard T.H. Chan School of Public Health, Boston, MA 02115

7\. Departments of Epidemiology and Biostatistics, Harvard T.H. Chan School of Public Health, Boston, MA 02115

\newpage

## Abstract

<!-- ### Standard -->

*Background*

The UK COVID-19 vaccination programme delivered the first "booster" doses in mid September 2021, initially in groups at high risk of severe disease and then across the entire adult population. The BNT162b2 mRNA Pfizer-BioNTech and the Moderna mRNA-1273 vaccines were used.

*Methods*

With the approval of NHS England we used the OpenSAFELY-TPP database, covering 40% of GP practices in England and linked to national coronavirus surveillance, hospital episodes, and death registry data, to estimate the effectiveness of a booster dose in "fully vaccinated" people using a sequential trials approach. On each day of the recruitment period (`r format(as.Date(study_dates$studystart_date), "%d %B %Y")` to `r format(as.Date(study_dates$lastvax3_date), "%d %B %Y")`) people receiving a booster dose on that day were recruited to the treatment arm, with a matched sample of unboosted people recruited to the control arm. Controls were matched on geogrpahical region, JCVI priority group, and the week and type of the second vaccine dose. Trial arms across each day were pooled and analysed using a Cox model, controlling for additional confounders. Primary outcomes were recorded SARS-CoV-2 infection, COVID-19-related accident and emergency attendance, unplanned COVID-19-related hospital admission, and COVID-19-related death occurring within 12 weeks of recruitment.

*Results*

There were `r label_number(1, big.mark=",")(filter(matchsummary_treated, treatment=="pfizer") %>% pull(n))` people boosted with BNT162b2 and `r label_number(1, big.mark=",")(filter(matchsummary_treated, treatment=="moderna") %>% pull(n))` people boosted with mRNA-1273 who were successfully matched to an unboosted control in the study period.

*Conclusion*

In this cohort of fully vaccinated people in England, a booster dose was estimated to be highly effective at reducing.... , though there was evidence of waning effects. Heterologous boosting ...

<!-- **Keywords** COVID-19; Vaccine effectiveness; Target trial; EHR data -->

*Objectives* To estimate the effectiveness of the BNT162b2 and mRNA-1273 mRNA-1273 booster COVID-19 vaccines against infection and COVID-19 disease.

*Design* Matched sequential trial design, emulating a series of parallel two-arm RCTs.

*Setting* Linked primary care, hospital, and COVID-19 surveillance records available within the OpenSAFELY-TPP research platform.

*Participants* `r label_number(1, big.mark=",")(filter(flowchart, crit=="c4") %>% pull(n))` people with a complete primary COVID-19 vaccine course, registered with a GP practice using the TPP SystmOne clinical information system in England.

*Interventions* A booster vaccination with either BNT162b2 or mRNA-1273 administered as part of the national COVID-19 booster vaccine roll-out.

*Main outcome measures* Recorded SARS-CoV-2 infection, COVID-19-related accident and emergency attendance, unplanned COVID-19-related hospital admission, and COVID-19-related death.

*Results*

*Conclusions*

\newpage

# Introduction

The national COVID-19 vaccination programme in the United Kingdom began a new phase in September 2021 with the administration of the first booster COVID-19 vaccine doses in those who had already received their primary vaccination course (that is, "fully vaccinated" with a complete first and second dose). Initially offered to those at highest risk of severe disease, eligibility was extended to the entire adult population by the end of the year. In the vast majority of cases, BNT162b2 and mRNA-1273 vaccines were used and administered without consideration of the type of primary course.

We set out to estimate the effectiveness of booster doses amongst the fully vaccinated adult population in England using the OpenSAFELY-TPP linked primary care database covering around 40% of England's population.

# Methods

## Data source

The OpenSAFELY-TPP database (<https://opensafely.org>) covers 24 million people registered at GP practices that use TPP SystmOne electronic health record software. It includes pseudonymized data such as coded diagnoses, medications and physiological parameters. No free text data are included. This primary care data is linked (via NHS numbers) with A&E attendance and in-patient hospital spell records via NHS Digital's Hospital Episode Statistics (HES), national coronavirus testing records via the Second Generation Surveillance System (SGSS), and national death registry records from the Office for National Statistics (ONS). Vaccination status is available in the GP record directly via the National Immunisation Management System (NIMS). HCW status is recorded for all vaccine recipients at the time of vaccination, and this information is sent to OpenSAFELY-TPP from NHS Digital's COVID-19 data store.

## Study population and design

We studied adults (18 and over) in England who had received a complete primary COVID-19 vaccination course (two doses) of BNT162b2 or ChAdOx1-S and who were registered at a GP practice using TPP's SystmOne clinical information system. People were not eligible for inclusion if: they were a vaccinated health care worker, resident in a care or nursing home, medically housebound or receiving end-of-life care (due to distinct clustering of vaccine uptake and infection risk in these groups that could not be reasonably measured and controlled for); had received heterologous (i.e., mixed) first and second vaccine dose; information on sex, ethnicity, deprivation, and geographical region was not known.

After exclusions for missing values on demographic variables, there were no missing values in remaining variables as they were defined by the presence or absence of clinical codes or events in the patient record.

We used a sequential trials approach designed to emulate a sequence of two-arm parallel randomised controlled trials recruiting participants between `r format(as.Date(study_dates$studystart_date), "%d %B %Y")` to `r format(as.Date(study_dates$studyend_date), "%d %B %Y")` inclusive. For each trial, the treatment arm comprises people who received a booster dose on that day, and the control arm comprises the same number of unboosted people selected using exact 1:1 matching without replacement. Trial arms are matched on primary course vaccine type (BNT162b2, ChAdOx1-S, or mRNA-1273), the calendar week of the second vaccine dose, geographical region, prior infection status, and vaccine priority group (as determined by the Joint Committee for Vaccine and immunisation expert working group). People selected as controls are not eligible to be included again as a control in a subsequent trial, but are eligible for selection into the treatment arm of a subsequent trial. Unmatched booster recipients were not included.

Additional eligiblity criteria based on sufficiently many people being vaccinated on each day within region, priority group, and second dose week strata were also used to ensure unboosted people were both eligible and able to receive a booster dose at the time of potential recruitment. The trial sequence was terminated early if fewer than 10% of treated people were successfully matched. See supplementary materials for further details.

## Outcomes

The following five outcomes were considered: positive SARS-CoV-2 test; COVID-19 A&E attendance; unplanned COVID-19 hospital admission; COVID-19 critical care admission; COVID-19 death. Positive SARS-CoV-2 tests were identified using SGSS records and based on swab date. Both polymerase chain reaction (PCR) and lateral flow tests are included, without differentiation between symptomatic and asymptomatic infection. COVID-19 A&E attendances were identified using HES emergency care records with U07.1 ("COVID-19, virus identified") or U07.2 ("COVID-19, virus not identified") ICD-10 diagnosis codes @emergenc. Unplanned COVID-19 hospital admissions and critical care admissions were identified using HES in-patient hospital records with U07.1 or U07.2 reason for admission ICD-10 codes. COVID-19 deaths with ICD-10 codes (as above) mentioned anywhere on the death certificate (i.e., as an underlying or contributing cause of death) were included.

## Additional variables

Participant characteristics used describe the cohort and for confounder adjustment include: age, sex (male or female), English Index of Multiple Deprivation (IMD, grouped by quintiles), ethnicity (Black, Mixed, South Asian, White, Other, as per the UK census), NHS region (East of England, Midlands, London, North East and Yorkshire, North West, South East, South West), number of conditions in the clinically "at risk" classification, as per national prioritisation guidelines, the number of SARS-CoV-2 tests in the 6 months prior to the study start date (via SGSS), rurality (urban conurbation, urban city or town, rural town or village), evidence of prior SARS-CoV-2 infection (positive SARS-CoV-2 test, "probable" infection documented in primary care, or COVID-19 hospitalisation), learning disabilities, severe mental illness, undergoing an unplanned hospital admission, undergoing a planned hospital admission. All characteristics were ascertained as at the trial recruitment date (or study start date if assumed to be fixed from that date onwards).

## Statistical Analysis

Time zero was the trial recruitment day, and trial participants were followed up for no more than 12 weeks from this day. Follow-up was censored earlier than this at `r format(as.Date(study_dates$studyend_date), "%d %B %Y")`, death, or GP practice de-registration. Additionally for each matched pair, follow-up for both the boosted and unboosted person was censored if the control is boosted.

Data from each trial were pooled and analysed using Cox regression with period-specific hazard ratios for treatment, in two-week periods. Baseline hazards were allowed to differ for participants differing by recruitment day, region, priority group, and week and type of second vaccine dose, via Cox stratification. Other variables were included as covariates. A robust variance estimator was used to account for participant clustering.

[The primary estimand is the average treatment effect in the recruited population. To estimate this, expected cumulative incidence rates are computed for each participant assuming treatment was received by everyone in both trial arms at time zero. These curves are then averaged over all participants to give the marginal cumulative incidence. This is repeated assuming nobody is treated in either arm, and their difference taken. Calculating standard errors for these quantities is challenging, whether approached analytically (due to complex variance components) or computationally (via bootstrapping, due to the large datasets involved) and is there not reported. ]

## Subgroup analyses

We additionally estimated booster effectiveness in subgroups defined by primary vaccine course (both BNT162b2 or both ChAdOx1-S) using complete model stratification.

## Software, code, and reproducibility

Data management and analyses were conducted in Python 3.8 and R version 4.0.2. All code is available at <https://github.com/opensafely/booster-effectiveness>, and is shared openly for review and re-use under MIT open license. Codelists are available at <https://www.opencodelists.org/>. No person-level data is shared. Any reported figures based on counts below 6 are redacted or rounded for disclosure control.

This study followed the STROBE-RECORD reporting guidelines.

## Patient and public involvement

We have developed a publicly available website <https://opensafely.org/> through which we invite any patient or member of the public to contact us regarding this study or the broader OpenSAFELY project.

# Results

## Study population

A total of `r label_number(1, big.mark=",")(filter(flowchart, crit=="c0") %>% pull(n))` fully vaccinated (received a primary course of two COVID-19 vaccine doses) people aged 18 and over and registered at a TPP practice on `r format(as.Date(study_dates$studystart_date), "%d %B %Y")` were identified the database, with `r label_number(1, big.mark=",")(filter(flowchart, crit=="c4") %>% pull(n))` (`r label_percent(0.1)(filter(flowchart, crit=="c4") %>% pull(pct_all))`) eligible for inclusion (supplementary Tables S1 and S2). Of these, `r label_number(1, big.mark=",")(cohortsummary$n_boosted)` received a booster dose during the recruitment period (up to `r format(as.Date(study_dates$lastvax3_date), "%d %B %Y")`). `r label_number(1, big.mark=",")(cohortsummary$n_3pfizer)` ( `r label_percent(0.1)(cohortsummary$prop_3pfizer)`) for BNT162b2 and `r label_number(1, big.mark=",")(cohortsummary$n_3moderna)` ( `r label_percent(0.1)(cohortsummary$prop_3moderna)`) for mRNA-1273 (Figure 1).

## Matching

`r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n))` (`r label_percent(0.1)(filter(matchsummary, treatment=="pfizer") %>% pull(prop_recruited))`) BNT162b2 booster recipients were successfully matched to unboosted controls, contributing `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(fup_years))` person-years of follow-up across both trial arms. For mRNA-1273, this was `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="moderna") %>% pull(n))` (`r label_percent(0.1)(filter(matchsummary, treatment=="moderna") %>% pull(prop_recruited))`) and `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="moderna") %>% pull(fup_years))` person-years of follow-up (Table 2).

Characteristics at recruitment between treatment and controls arms are provided in supplementary materials (Table S2). Characteristics were (??) well-balanced between trial arms.

\newpage

#### Figure 1: Booster vaccination uptake, and matching success

Booster vaccination update over the duration of the study cohort, and the proportion of those that were matched for modelling.

```{r, vaxdate, echo=FALSE, message=FALSE, warning=FALSE, out.width='60%', out.height='50%', results='asis', out.extra=''}
#read_rds(fs::path(output_dir_os, "descriptive", "vaxdate", "plot_vaxdate_stack.rds"))

xmin <- min(matchcoverage$vax3_date )
xmax <- max(matchcoverage$vax3_date )+1
matchcoverage  %>%
  mutate(
    # vax12_type = fct_case_when(
    #   vax12_type == "pfizer-pfizer" ~ "BNT162b2",
    #   vax12_type == "az-az" ~ "ChAdOx1-S",
    #   vax12_type == "moderna-moderna" ~ "mRNA-1273",
    #   TRUE ~ NA_character_
    # ),
    course = fct_case_when(
      vax12_type=="pfizer-pfizer" & treatment=="pfizer" ~ "2x BNT162b2  / 1x BNT162b2",
      vax12_type=="az-az" & treatment=="pfizer" ~ "2 x ChAdOx1-S / 1 x BNT162b2",
      vax12_type=="pfizer-pfizer" & treatment=="moderna" ~ "2x BNT162b2 / 1x mRNA-1273",
      vax12_type=="az-az" & treatment=="moderna" ~ "2x ChAdOx1-S / 1x mRNA-1273",
      TRUE ~ NA_character_
    )
  ) %>%
  ggplot() +
  geom_col(
    aes(
      x=vax3_date+0.5,
      y=cumuln,
      group=fct_rev(matched),
      fill=fct_rev(matched),
      colour=NULL
    ),
    position=position_stack(),
    alpha=0.8,
    width=1
  )+
  geom_rect(xmin=xmin, xmax= xmax+1, ymin=0, ymax=6, fill="grey", colour="transparent")+
  facet_wrap(vars(course))+
  scale_x_date(
    breaks = unique(lubridate::ceiling_date(matchcoverage$vax3_date, "1 month")),
    limits = c(lubridate::floor_date((xmin), "1 month"), NA),
    labels = scales::label_date("%d/%m"),
    expand = expansion(add=1),
  )+
  scale_y_continuous(
    labels = scales::label_number(accuracy = 1, big.mark=","),
    expand = expansion(c(0, NA))
  )+
  scale_fill_brewer(type="qual", palette="Set2")+
  scale_colour_brewer(type="qual", palette="Set2")+
  labs(
    x="Date",
    y="Cumulative booster vaccines",
    colour=NULL,
    fill=NULL,
    alpha=NULL
  ) +
  theme_minimal()+
  theme(
    axis.line.x.bottom = element_line(),
    axis.text.x.top=element_text(hjust=0),
    strip.text.y.right = element_text(angle = 0),
    axis.ticks.x=element_line(),
    legend.position = "bottom"
  )+
  NULL
cat("  \n\n")
```

#### Table 1: Baseline characteristics

Participant characteristics as on the day of recruitment into the treatment or control arms, stratified by treatment and outcome type.

```{r table1, echo=FALSE, warning=FALSE, message=FALSE}
table1_body <- read_csv(fs::path(output_dir_os, "matchtable1.csv"))
table1_by <- read_csv(fs::path(output_dir_os, "matchtable1by.csv"))

table1 <- 
  table1_body %>%
  group_by(variable) %>%
  mutate(
    label = if_else(row_number()==1, label, str_c("&nbsp;&nbsp;&nbsp;", label)),
  ) %>%
  mutate(
    across(
      starts_with("stat_"),
      ~if_else(is.na(.x) & row_number()==1, "", .x)
    )
  ) %>%
  ungroup() %>%
  mutate(
    order = cumsum(lag(variable, 1, first(variable))!=variable)
  ) %>%
  select(
    variable, var_label, label, starts_with("stat_"), order
  ) %>%
  rename(!!!set_names(table1_by$by_col, table1_by$by_chr))

table1 %>%
  gt(groupname_col="treatment_descr") %>%
  fmt_markdown(columns = "label") %>%
  tab_style(
    style = list(
      cell_fill(color = "gray96")
    ),
    locations = cells_body(
      rows = order%%2 == 0
    )
  ) %>%
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_column_labels(everything())
  ) %>%
  cols_label(
     label=md("Characteristic")
  ) %>%
  cols_hide(c("variable", "var_label", "order"))
```

\newpage

#### Table 2: Recruitment, follow-up, and event incidence

```{r table2, echo=FALSE, warning=FALSE, message=FALSE}
incidences %>%
  filter(fup_period == "All") %>%
  left_join(
    matchsummary %>% select(treatment, eligible_1=total_treated, recruit_prop_1 = prop_recruited),
    by=c("treatment")
  ) %>%
  select(
    treatment_descr, 
    outcome_descr,
    eligible_1,
    n_1,
    recruit_prop_1,
    q_1,
    rate_1,
    q_0, 
    rate_0,
    rr, 
    rrCI
  ) %>%
  gt(
    groupname_col = c("treatment_descr"),
  ) %>%
  cols_label(
    #treatment_descr = "Treatment",
    outcome_descr = "Outcome",

    eligible_1 = "Eligible",
    n_1 = "Recruited",
    recruit_prop_1 = "(%)",

    q_0 = "Events / person-years",
    q_1   = "Events / person-years",

    rate_0 = "Incidence",
    rate_1 = "Incidence",

    rr = "Incidence rate ratio",
    rrCI = "95% CI"
  ) %>%
  tab_spanner(
    label = "Boosted",
    columns = ends_with("1")
  ) %>%
  tab_spanner(
    label = "Unboosted",
    columns = ends_with("0")
  ) %>%
  fmt_number(
    columns = c("rate_0", "rate_1", "rr"),
    decimals = 2
  ) %>%
  fmt_integer(
    columns = c("n_1", "eligible_1"),
    sep_mark = ","
  ) %>%
  fmt_number(
    columns = "recruit_prop_1",
    scale_by = 100,
    decimals = 1,

  ) %>%
  fmt_missing(
    everything(),
    missing_text="--"
  ) %>%
  cols_align(
    align = "right",
    columns = everything()
  )

```

## Estimated booster effectiveness

In the BNT162b2 and mRNA-1273 analyses respectively, matched participants contributed `r filter(incidences, treatment=="pfizer", outcome=="postest", fup_period=="all") %>% pull(events)` and `r filter(incidences, treatment=="moderna", outcome=="postest", fup_period=="all") %>% pull(events)` positive SARS-CoV-2 tests, `r filter(incidences, treatment=="pfizer", outcome=="covidattendance", fup_period=="all") %>% pull(events)` and `r filter(incidences, treatment=="moderna", outcome=="covidattendance", fup_period=="all") %>% pull(events)` COVID-19 A&E attendances, `r filter(incidences, treatment=="pfizer", outcome=="covidadmitted", fup_period=="all") %>% pull(events)` and `r filter(incidences, treatment=="moderna", outcome=="covidadmitted", fup_period=="all") %>% pull(events)` COVID-19 hospital admissions, `r filter(incidences, treatment=="pfizer", outcome=="covidcc", fup_period=="all") %>% pull(events)` and `r filter(incidences, treatment=="moderna", outcome=="covidcc", fup_period=="all") %>% pull(events)` COVID-19 ciritical care admissions, and `r filter(incidences, treatment=="pfizer", outcome=="coviddeath", fup_period=="all") %>% pull(events)` and `r filter(incidences, treatment=="moderna", outcome=="coviddeath", fup_period=="all") %>% pull(events)` COVID-19 deaths (Table 2).

During the fifth and sixth week after recruitment (days 28-41) the estimated hazard ratio comparing a BNT162b2 booster dose to no dose for a positive test was `r filter(hazards, outcome=="postest", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)`, for COVID-19 A&E attendances was `r filter(hazards, outcome=="covidattendance", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)`, for COVID-19 hospital admission was `r filter(hazards, outcome=="covidadmitted", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)`, for COVID-19 critical care admission was `r filter(hazards, outcome=="covidcc", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)`, and for COVID-19 death was `r filter(hazards, outcome=="coviddeath", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)` (Figure 2). For mRNA-1273, the corresponding estimated hazard ratios were `r filter(hazards, outcome=="postest", treatment=="moderna", model==3, term_left==28) %>% pull(hr_display)`, for COVID-19 A&E attendances was `r filter(hazards, outcome=="covidattendance", treatment=="moderna", model==3, term_left==28) %>% pull(hr_display)`, for COVID-19 hospital admission was `r filter(hazards, outcome=="covidadmitted", treatment=="moderna", term_left==28) %>% pull(hr_display)`, for COVID-19 critical care admission was `r filter(hazards, outcome=="covidcc", treatment=="moderna", model==3, term_left==28) %>% pull(hr_display)`, and for COVID-19 death was `r filter(hazards, outcome=="coviddeath", treatment=="moderna", model==3, term_left==28) %>% pull(hr_display)`.

Estimated effectiveness by primary vaccine course...

\newpage

#### Figure 2: Estimated booster effectiveness

For each outcome based on the fully adjusted model, the period-specific hazard ratio for boosting versus not boosting is shown. Models with less extensive confounder adjustment are provided in supplementary materials (supplementary Figure SX).

```{r plot_effects, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}

formatpercent100 <- function(x,accuracy){
  formatx <- scales::label_percent(accuracy)(x)

  if_else(
    formatx==scales::label_percent(accuracy)(1),
    paste0(">",scales::label_percent(1)((100-accuracy)/100)),
    formatx
  )
}

# plot_cmlinc <- cmlinc3 %>%
#     ggplot()+
#     geom_step(aes(x=tstop, y=(1-survival)*1000, group=vax1_az_descr, colour=vax1_az_descr))+
#     geom_rect(aes(xmin=tstop, xmax=lead_tstop, ymin=(1-survival.ll)*1000, ymax=(1-survival.ul)*1000, group=vax1_az_descr, fill=vax1_az_descr), alpha=0.1)+
#     geom_hline(aes(yintercept=0), colour='black')+
#     facet_wrap(facets=vars(outcome_descr), nrow=1, scales="free_y")+
#     scale_x_continuous(
#       breaks = seq(0,7*52,by=28),
#       expand = expansion(0)
#     )+
#     scale_y_continuous(
#       expand = expansion(0),
#       #limits = c(0L,NA)
#     )+
#     scale_colour_brewer(type="qual", palette="Set1", na.value="grey")+
#     scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey")+
#     labs(
#       x=NULL,
#       y="Marginalised\ncumul. incidence\nper 1000",
#       colour=NULL
#     )+
#     theme_minimal(base_size = 9)+
#     theme(
#       legend.position=c(0.05,0.1),
#       legend.justification = c(0,0),
#       legend.direction = "vertical",
#       axis.text.x.top=element_text(hjust=0),
# 
#       panel.border = element_blank(),
#       axis.line.y = element_line(colour = "black"),
#       axis.line.x.top = element_line(colour = "black"),
# 
#       panel.grid.minor.x = element_blank(),
#       panel.grid.minor.y = element_blank(),
#       strip.background = element_blank(),
# 
# 
#       panel.spacing = unit(0.8, "lines"),
# 
#     )+
#     NULL
#   
#   ## difference in cumulative incidence
#   
#   plot_riskdiff <-
#     riskdiff3 %>%
#     ggplot()+
#     geom_hline(aes(yintercept=0), colour='black')+
#     geom_step(aes(x=tstop, y=diff*1000), colour='orchid4')+
#     geom_rect(aes(xmin=tstop, xmax=lead_tstop, ymin=diff.ll*1000, ymax=diff.ul*1000), alpha=0.1, colour="transparent", fill='orchid4')+
#     facet_wrap(facets=vars(outcome_descr), nrow=1, scales="free_y")+
#     scale_x_continuous(
#       limits = c(0, NA),
#       breaks = seq(0,7*52,by=28),
#       expand = expansion(0)
#     )+
#     scale_y_continuous(
#       expand = expansion(0),
#       sec.axis = dup_axis(name="Higher incidence after \n <- ChAdOx1 / BNT162b2 ->", breaks = NULL)
#     )+
#     coord_cartesian(ylim= c(-3, 3))+
#     scale_colour_brewer(type="qual", palette="Set1", na.value="grey")+
#     scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey")+
#     labs(
#       x=NULL,
#       y="Difference\nin cumul. incidence\nper 1000",
#       colour=NULL
#     )+
#     theme_minimal(base_size = 9)+
#     theme(
#       axis.line.y.left = element_line(colour = "black"),
# 
#       panel.grid.minor.x = element_blank(),
#       panel.grid.minor.y = element_blank(),
#       strip.background = element_blank(),
#       strip.text = element_blank(),
# 
#       panel.spacing = unit(0.8, "lines"),
# 
#       plot.title = element_text(hjust = 0),
#       plot.title.position = "plot",
#       plot.caption.position = "plot",
#       plot.caption = element_text(hjust = 0, face= "italic")
# 
#      ) +
#    NULL
# 
#

  ## hazard ratio

  y_breaks <- c(0.01, 0.02, 0.05, 0.1, 0.2, 0.5, 1, 2)

  plot_hazards <-
    hazards3 %>%
    ggplot()+
    geom_point(aes(y=hr, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+
    geom_linerange(aes(ymin=hr.ll, ymax=hr.ul, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+
    geom_hline(aes(yintercept=1), colour='grey')+
    facet_grid(rows=vars(outcome_descr), cols=vars(treatment_descr), switch="y")+
    scale_y_log10(
      breaks=y_breaks,
      limits = c(0.01, 1.5),
      oob = scales::oob_keep,
      sec.axis = sec_axis(
        ~(1-.),
        name="Effectiveness",
        breaks = 1-(y_breaks),
        labels = function(x){formatpercent100(x, 1)}
      )
    )+
    scale_x_continuous(breaks=postbaselinecuts, limits=c(min(postbaselinecuts), max(postbaselinecuts)+1), expand = c(0, 0))+
    scale_colour_brewer(type="qual", palette="Set2", guide=guide_legend(ncol=1))+
    labs(
      y="Hazard ratio",
      x="Days since booster dose",
      colour=NULL
     ) +
    theme_bw()+
    theme(
      panel.border = element_blank(),
      axis.line.y = element_line(colour = "black"),
      axis.line.y.right = element_blank(),

      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      strip.background = element_blank(),
      strip.placement = "outside",
      strip.text.y.left = element_text(angle = 0),

      panel.spacing = unit(0.8, "lines"),

      plot.title = element_text(hjust = 0),
      plot.title.position = "plot",
      plot.caption.position = "plot",
      plot.caption = element_text(hjust = 0, face= "italic"),

      legend.position = "none"
     ) +
    NULL

  plot_mainresults <- 
    patchwork::wrap_plots(
      #plot_cmlinc, 
      #plot_riskdiff,
      plot_hazards, 
      ncol=1
    )
  ggsave(
    filename="mainresults.pdf",
    path=output_dir_rmd,
    plot=plot_mainresults,
  )

  plot_mainresults

```

\newpage

# Discussion

This observational study of `r label_number(1, big.mark=",")(flowchart[[nrow(flowchart), "n"]]*2)` fully vaccinated people living in England estimated high protective effects of booster vaccination against all studied outcomes. severe COVID-19 disease and death. Protection appeared to be highest in those receiving a X-X-X course, though . There were apparent waning effects after X weeks which is unlikely to explained by increased transmissibility of the Omicron variant given the low prevalence of Omicron at the end of the follow-up window.

### Strengths and weaknesses

We used routinely-collected health records with comprehensive coverage of primary care, hospital admissions, COVID-19 testing, COVID-19 vaccination, and death registrations to study fully vaccinated English residents. ...

The methodological approach used is robust but inefficient; many boosted individuals are not retained after matching, and those who are retained are censored when their matched control is boosted. Millions of eligible participants may be required to obtain precise estimates, particularly when event rates are low, and so extending the approach to smaller subgroups may not be feasible. FOr instance, due to small numbers, we did not study booster effectiveness in those who had received the mRNA-1273 vaccine as their primary course or those who had received a heterologous primary course.

This study considers effectiveness in a population where Alpha was the predominant circulating variant (ref gov.uk reports). Re-running the analysis against a backdrop of the now globally dominant Omicron variant will be undertaken where valid concurrent comparisons between boosted and unboosted people can be made.

Despite reasonable balance of baseline characteristics in the trial arms, and adjustment for a range of potential confounders, the possibility of unmeasured confounding remains. In particular, the apparent protective effect immediately after boosting is biologically implausible, and will be explained in part by the occurrence of COVID-19 symptoms at baseline not recorded in the health record.

### Findings in context

[pfizer phase 3 booster trial, in pfizer first course]
[ cov-boost trial]

A number of studies have estimated COVID-19 vaccine booster effectiveness in observational data.

### Conclusion

This study found high protection against severe COVID-19 disease

<!-- \newpage -->

<!-- # Summary -->

<!-- ### What is already known -->

<!-- ### What this study adds -->

\newpage

## Acknowledgements

We are very grateful for all the support received from the TPP Technical Operations team throughout this work, and for generous assistance from the information governance and database teams at NHS England / NHSX.

## Conflicts of Interest

All authors have completed the ICMJE uniform disclosure form and declare the following: BG has received research funding from Health Data Research UK (HDRUK), the Laura and John Arnold Foundation, the Wellcome Trust, the NIHR Oxford Biomedical Research Centre, the NHS National Institute for Health Research School of Primary Care Research, the Mohn-Westlake Foundation, the Good Thinking Foundation, the Health Foundation, and the World Health Organisation; he also receives personal income from speaking and writing for lay audiences on the misuse of science. IJD holds shares in GlaxoSmithKline (GSK).

## Funding

This work was supported by the Medical Research Council MR/V015737/1 and the Longitudinal Health and wellbeing strand of the National Core Studies programme. The OpenSAFELY platform is funded by the Wellcome Trust. TPP provided technical expertise and infrastructure within their data centre pro bono in the context of a national emergency. BGs work on clinical informatics is supported by the NIHR Oxford Biomedical Research Centre and the NIHR Applied Research Collaboration Oxford and Thames Valley. BG's work on better use of data in healthcare more broadly is currently funded in part by: NIHR Oxford Biomedical Research Centre, NIHR Applied Research Collaboration Oxford and Thames Valley, the Mohn-Westlake Foundation, NHS England, and the Health Foundation; all DataLab staff are supported by BG's grants on this work. LS reports grants from Wellcome, MRC, NIHR, UKRI, British Council, GSK, British Heart Foundation, and Diabetes UK outside this work. KB holds a Wellcome Senior Research Fellowship (220283/Z/20/Z). HIM is funded by the NIHR Health Protection Research Unit in Immunisation, a partnership between Public Health England and London School of Hygiene & Tropical Medicine. AYSW holds a fellowship from the British Heart Foundation. EJW holds grants from MRC. RM holds a Sir Henry Wellcome Fellowship funded by the Wellcome Trust (201375/Z/16/Z). HF holds a UKRI fellowship. IJD holds grants from NIHR and GSK.

Funders had no role in the study design, collection, analysis, and interpretation of data; in the writing of the report and in the decision to submit the article for publication.

The views expressed are those of the authors and not necessarily those of the NIHR, NHS England, Public Health England or the Department of Health and Social Care.

For the purpose of Open Access, the author has applied a CC BY public copyright licence to any Author Accepted Manuscript (AAM) version arising from this submission.

## Information governance and ethical approval

NHS England is the data controller; TPP is the data processor; and the key researchers on OpenSAFELY are acting with the approval of NHS England. This implementation of OpenSAFELY is hosted within the TPP environment which is accredited to the ISO 27001 information security standard and is NHS IG Toolkit compliant @betad @datasec ; Patient data has been pseudonymised for analysis and linkage using industry standard cryptographic hashing techniques; all pseudonymised datasets transmitted for linkage onto OpenSAFELY are encrypted; access to the platform is via a virtual private network (VPN) connection, restricted to a small group of researchers; the researchers hold contracts with NHS England and only access the platform to initiate database queries and statistical models; all database activity is logged; only aggregate statistical outputs leave the platform environment following best practice for anonymisation of results such as statistical disclosure control for low cell counts @isb1523 . The OpenSAFELY research platform adheres to the obligations of the UK General Data Protection Regulation (GDPR) and the Data Protection Act 2018. In March 2020, the Secretary of State for Health and Social Care used powers under the UK Health Service (Control of Patient Information) Regulations 2002 (COPI) to require organisations to process confidential patient information for the purposes of protecting public health, providing healthcare services to the public and monitoring and managing the COVID-19 outbreak and incidents of exposure; this sets aside the requirement for patient consent @coronavi2020 . Taken together, these provide the legal bases to link patient datasets on the OpenSAFELY platform. GP practices, from which the primary care data are obtained, are required to share relevant health information to support the public health response to the pandemic, and have been informed of the OpenSAFELY analytics platform.

This study was approved by the Health Research Authority (REC reference 20/LO/0651) and by the LSHTM Ethics Board (reference 21863).

## Guarantor

BG is guarantor.

\newpage

## References

### 2022-02-09 manuscript/draft-supplement.Rmd
add manuscript first draft (very messy!)

---
title: "Effectiveness of COVID-19 Booster doses in England: a sequential trials study in OpenSAFELY: Supplementary materials"
output:
  html_document:
    keep_md: yes
    self_contained: yes
  word_document: default
  bookdown::html_document2:
    number_sections: no
    toc: no
  pdf_document:
    toc: no
header-includes:
- \usepackage{float} 
- \floatplacement{figure}{H}
bibliography: references.bib
---

```{r setup, include=FALSE}

library('tidyverse')
library('here')
library('glue')
library('lubridate')
library('gt')
library('patchwork')
library('scales')
source(here("lib", "functions", "utility.R"))

# where are the outputs (ie the inputs for this manuscript!) saved?
output_dir_os <- here("output", "manuscript-objects")
#output_dir_os <- here("released_outputs", "output", "manuscript-objects")

# where should we put the objects created within this rmd script?
output_dir_rmd <- here("manuscript", "figures")
fs::dir_create(output_dir_rmd)

# only applicable if `self-contained: yes` option is enables in yaml header
knitr::opts_chunk$set(
  echo = TRUE,
  fig.path = paste0(output_dir_rmd, "/"),
  fig.pos = 'H' #to stop figures floating around when rendering to pdf
)

study_dates <-
  jsonlite::read_json(path=here("lib", "design", "study-dates.json")) %>%
  map(as.Date)

postbaselinecuts <- read_rds(here("lib", "design", "postbaselinecuts.rds"))

flowchart <- read_csv(fs::path(output_dir_os, "flowchart.csv"))
cohortsummary <- read_csv(fs::path(output_dir_os, "cohortsummary.csv"))
matchtable1 <- read_csv(fs::path(output_dir_os, "matchtable1.csv"))
matchtable1by <- read_csv(fs::path(output_dir_os, "matchtable1by.csv"))
matchsummary <- read_csv(fs::path(output_dir_os, "matchsummary.csv"))
matchsummary_treated <- read_csv(fs::path(output_dir_os, "matchsummary_treated.csv")) %>% filter(treated == 1L)
matchsummary_control <- read_csv(fs::path(output_dir_os, "matchsummary_treated.csv")) %>% filter(treated == 0L)
incidences <- read_csv(fs::path(output_dir_os, "ir.csv"))

```

\newpage

# Matching details

Booster eligibility was dependent on vaccine priority group, largely determined by clinical vulnerability and age. Booster doses were initially available 6 months after administration of the second dose, later reduced to 3 months following concerns over the surge in cases and the emergence of the seemingly more transmissible Omicron variant. Determining vaccine eligibility for a given person on a given day is therefore complex. To avoid these complexities, we determined trial eligibility for each person on each day of recruitment by asking if there were sufficiently many people with similar characteristics who were being vaccinated on that day. Specifically, if less than X people in the same region, priority group, and week of second vaccine dose had been vaccinated, then they were not eligible for recruitment.

### Table S1: Inclusion criteria

```{r flowchart, echo=FALSE, warning=FALSE, message=FALSE}
flowchart %>%
  transmute(
    Participants = str_wrap(criteria, 40),
    N = n,
    `N excluded` = n_exclude,
    #`% excluded` = pct_exclude,
    `% remaining` = pct_all
  ) %>%
  # mutate(
  #   Participants=if_else(row_number()==1,"HCWs aged 18-64 receiving BNT162b2 or ChAdOx1 between 4 Jan and 28 Feb",  Participants)
  # ) %>%
  gt() %>%
  fmt_percent(
    columns = starts_with(c("%")),
    decimals = 1
  ) %>%
  fmt_number(
    columns = starts_with(c("N")),
    decimals= 0
  ) %>%
  fmt_missing(
    everything(),
    missing_text="-"
  ) %>%
  tab_style(
    style=cell_text(whitespace ="pre-wrap"),
    #style=cell_text(style ="italic"),
    locations = cells_body(columns = "Participants")
  )
```

\newpage
### Table S2a: Characteristics at recruitment to Pfizer trial, by treatment status

```{r table1_pfizer, echo=FALSE, warning=FALSE, message=FALSE, out.extra=''}

  table1by <- 
    matchtable1by %>% 
    filter(treatment=="pfizer")

  table1 <- 
    matchtable1 %>%
    filter(treatment=="pfizer") %>%
    group_by(variable) %>%
    mutate(
      label = if_else(row_number()==1, label, str_c("&nbsp;&nbsp;&nbsp;", label)),
    ) %>%
    mutate(
      across(
        starts_with("stat_"),
        ~if_else(is.na(.x) & row_number()==1, "", .x)
      )
    ) %>%
    ungroup() %>%
    mutate(
      order = cumsum(lag(variable, 1, first(variable))!=variable)
    ) %>%
    select(
      variable, var_label, label, starts_with("stat_"), order
    ) %>%
    rename(!!!set_names(table1by$by_col, table1by$by_chr))

  table1_gt <-
    table1 %>%
    gt() %>%
    fmt_markdown(columns = "label") %>%
    tab_style(
      style = list(
        cell_fill(color = "gray96")
      ),
      locations = cells_body(
        rows = order%%2 == 0
      )
    ) %>%
    tab_style(
      style = list(
        cell_text(weight = "bold")
      ),
      locations = cells_column_labels(everything())
    ) %>%
    cols_label(
       label=md("Characteristic")
    ) %>%
    cols_hide(c("variable", "var_label", "order"))

  table1_gt

```

### Table S2b: Characteristics at recruitment to Moderna trial, by treatment status

```{r table1_moderna, echo=FALSE, warning=FALSE, message=FALSE, out.extra=''}

  table1by <- 
    matchtable1by %>% 
    filter(treatment=="moderna")

  table1 <- 
    matchtable1 %>%
    filter(treatment=="moderna") %>%
    group_by(variable) %>%
    mutate(
      label = if_else(row_number()==1, label, str_c("&nbsp;&nbsp;&nbsp;", label)),
    ) %>%
    mutate(
      across(
        starts_with("stat_"),
        ~if_else(is.na(.x) & row_number()==1, "", .x)
      )
    ) %>%
    ungroup() %>%
    mutate(
      order = cumsum(lag(variable, 1, first(variable))!=variable)
    ) %>%
    select(
      variable, var_label, label, starts_with("stat_"), order
    ) %>%
    rename(!!!set_names(table1by$by_col, table1by$by_chr))

  table1_gt <-
    table1 %>%
    gt() %>%
    fmt_markdown(columns = "label") %>%
    tab_style(
      style = list(
        cell_fill(color = "gray96")
      ),
      locations = cells_body(
        rows = order%%2 == 0
      )
    ) %>%
    tab_style(
      style = list(
        cell_text(weight = "bold")
      ),
      locations = cells_column_labels(everything())
    ) %>%
    cols_label(
       label=md("Characteristic")
    ) %>%
    cols_hide(c("variable", "var_label", "order"))

  table1_gt

```

\newpage

### 2022-02-09 analysis/data_process_long.R
create and use time-varying dataset before matching
This means we can now include variables that might change over the duration of the recruitment period as matching variables

@@ -13,6 +13,7 @@
 library('tidyverse')
 library('here')
 library('glue')
+library('survival')
 
 # Import custom user functions from lib
 source(here("lib", "functions", "utility.R"))
@@ -22,7 +23,16 @@ fs::dir_create(here("output", "data"))
 
 ## Import processed data ----
 
-data_processed <- read_rds(here("output", "data", "data_processed.rds"))
+data_processed <- read_rds(here("output", "data", "data_processed.rds")) %>%
+  mutate(
+    stop=10000L
+  )
+
+# import globally defined study dates and convert to "Date"
+study_dates <-
+  jsonlite::read_json(path=here("lib", "design", "study-dates.json")) %>%
+  map(as.Date)
+
 
 ## create one-row-per-event datasets ----
 # for vaccination, positive test, hospitalisation/discharge, covid in primary care, death
@@ -32,36 +42,26 @@ data_admitted_unplanned <- data_processed %>%
   select(patient_id, matches("^admitted\\_unplanned\\_\\d+\\_date"), matches("^discharged\\_unplanned\\_\\d+\\_date")) %>%
   pivot_longer(
     cols = -patient_id,
-    names_to = c(".value", "index"),
+    names_to = c("event", "index"),
     names_pattern = "^(.*)_(\\d+)_date",
+    values_to="date",
     values_drop_na = TRUE
   ) %>%
-  select(patient_id, index, admitted_date=admitted_unplanned, discharged_date = discharged_unplanned) %>%
-  arrange(patient_id, admitted_date)
+  #select(patient_id, index, admitted_date=admitted_unplanned, discharged_date = discharged_unplanned) %>%
+  arrange(patient_id, date)
 
 data_admitted_planned <- data_processed %>%
   select(patient_id, matches("^admitted\\_planned\\_\\d+\\_date"), matches("^discharged\\_planned\\_\\d+\\_date")) %>%
   pivot_longer(
     cols = -patient_id,
-    names_to = c(".value", "index"),
+    names_to = c("event", "index"),
     names_pattern = "^(.*)_(\\d+)_date",
+    values_to="date",
     values_drop_na = TRUE
   ) %>%
-  select(patient_id, index, admitted_date=admitted_planned, discharged_date = discharged_planned) %>%
-  arrange(patient_id, admitted_date)
[…]

### 2022-02-09 analysis/match_seqtrialcox.R

@@ -148,11 +148,14 @@ data_rollingstrata_vaxcount <-
   ) %>%
   filter(vax3_time>=0)
 
-data_matching_variables <-
+
+
+data_nontimevarying <-
   data_cohort %>%
   select(
     patient_id,
-    all_of(c(rolling_variables, matching_variables, "vax3_date"))
+    any_of(c(rolling_variables, matching_variables, "vax3_date")),
+    prior_covid_infection0
   ) %>%
   left_join(
     data_rollingstrata_vaxcount %>% select(all_of(c(rolling_variables, "vax3_date", "recruit"))),
@@ -164,6 +167,17 @@ data_matching_variables <-
   )
 
 
+data_timevarying <-
+  read_rds(here("output", "data", "data_timevarying.rds")) %>%
+  select(
+    patient_id,
+    tstart,
+    tstop,
+    any_of(matching_variables),
+    status_hospunplanned,
+    mostrecent_anycovid,
+  )
+
 ### event times ----
 
 ## tte = time-to-event, and always indicates time from study start date
@@ -171,7 +185,7 @@ data_tte <-
   data_cohort %>%
   transmute(
     patient_id,
-    day1_date = study_dates$studystart_date, # first trial date
+    day0_date = study_dates$studystart_date-1, # day before the first trial date
     treatment_date = if_else(vax3_type==treatment, vax3_date, as.Date(NA)),
     competingtreatment_date = if_else(vax3_type!=treatment, vax3_date, as.Date(NA)),
 
@@ -187,31 +201,22 @@ data_tte <-
 
     # assume vaccination occurs at the start of the day, and all other events occur at the end of the day.
 
-    tte_censor = tte(day1_date-1, censor_date, censor_date, na.censor=TRUE),
+    tte_censor = tte(day0_date, censor_date, censor_date, na.censor=TRUE),
 
-    tte_treatment = tte(day1_date, treatment_date, censor_date, na.censor=TRUE),
-
-
-    tte_anycovid0 = tte(day1_date-1, anycovid_0_date, censor_date, na.censor=TRUE),
-
-    tte_anycovid1 = tte(day1_date-1, anycovid_1_date, censor_date, na.censor=TRUE),
+    tte_treatment = tte(day0_date, treatment_date-1, censor_date, na.censor=TRUE),
 
     tte_stop = pmin(tte_censor, na.rm=TRUE),
 
   ) %>%
   filter(
[…]

### 2022-02-09 analysis/merge_seqtrialcox.R

@@ -91,6 +91,7 @@ matching_variables <- read_rds(here("lib", "design", "matching_variables.rds"))
 
 ## one pow per patient ----
 data_cohort <- read_rds(here("output", "data", "data_cohort.rds"))
+data_timevarying <- read_rds(here("output", "data", "data_timevarying.rds"))
 
 # compose modelling dataset ----
 
@@ -108,7 +109,6 @@ data_baseline <-
     region,
     jcvi_group,
     rural_urban_group,
-    prior_covid_infection,
     prior_tests_cat,
     multimorb,
     learndis,
@@ -176,83 +176,83 @@ logoutput_datasize(data_events)
 
 
 ## one row per time-varying covariate value change ----
-data_timevarying <- local({
-
-  data_join <-
-    data_tte %>%
-    select(patient_id, day1_date, censor_date, tte_stop)
-
-  data_postest <-
-    read_rds(here("output", "data", "data_long_postest_dates.rds")) %>%
-    inner_join(data_join, ., by =c("patient_id")) %>%
-    mutate(
-      tte = tte(day1_date-1, date, censor_date, na.censor=TRUE),
-    ) %>%
-    filter(!is.na(tte))
-
-  data_admitted_unplanned <-
-    read_rds(here("output", "data", "data_long_admitted_unplanned_dates.rds")) %>%
-    pivot_longer(
-      cols=c(admitted_date, discharged_date),
-      names_to="status",
-      values_to="date",
-      values_drop_na = TRUE
-    ) %>%
-    inner_join(data_join, ., by =c("patient_id")) %>%
-    mutate(
-      tte = tte(day1_date-1, date, as.Date(Inf), na.censor=TRUE) %>% as.integer(),
-      admittedunplanned_status = if_else(status=="admitted_date", 1L, 0L)
-    )
-
-  data_admitted_planned <-
-    read_rds(here("output", "data", "data_long_admitted_planned_dates.rds")) %>%
-    pivot_longer(
-      cols=c(admitted_date, discharged_date),
-      names_to="status",
-      values_to="date",
-      values_drop_na = TRUE
-    ) %>%
-    inner_join(data_join, ., by =c("patient_id")) %>%
-    mutate(
-      tte = tte(day1_date-1, date, as.Date(Inf), na.censor=TRUE) %>% as.integer(),
-      admittedplanned_status = if_else(status=="admitted_date", 1L, 0L)
-    )
-
-  data_timevarying <- data_tte %>%
[…]

### 2022-02-09 analysis/model_seqtrialcox.R

@@ -32,11 +32,6 @@ if(length(args)==0){
   subgroup <- args[[3]]
 }
 
-subgroup_variable <-  stringr::str_split_fixed(subgroup,"-",2)[,1]
-subgroup_level <- stringr::str_split_fixed(subgroup,"-",2)[,2]
-subgroup_dummy <- paste0(subgroup_variable,"_",subgroup_level)
-
-
 
 ## Import libraries ----
 library('tidyverse')
@@ -49,6 +44,12 @@ library('survival')
 source(here("lib", "functions", "utility.R"))
 source(here("lib", "functions", "survival.R"))
 source(here("lib", "functions", "redaction.R"))
+
+# derive subgroup info
+subgroup_variable <-  str_split_fixed(subgroup,"-",2)[,1]
+subgroup_level <- str_split_fixed(subgroup,"-",2)[,2]
+subgroup_dummy <- paste0(subgroup_variable,"_",subgroup_level)
+
 
 postbaselinecuts <- read_rds(here("lib", "design", "postbaselinecuts.rds"))
 matching_variables <- read_rds(here("lib", "design", "matching_variables.rds"))
@@ -122,13 +123,14 @@ data_timevaryingoutcomes <- local({
 
   data_join <-
     data_tte %>%
-    select(patient_id, day1_date, censor_date, tte_stop)
+    select(patient_id, day0_date, censor_date, tte_stop)
 
   data_outcome <-
-    read_rds(here("output", "data", glue("data_long_{outcome}_dates.rds"))) %>%
+    read_rds(here("output", "data", glue("data_allevents.rds"))) %>%
+    filter(event==outcome) %>%
     inner_join(data_join, ., by =c("patient_id")) %>%
     mutate(
-      tte = tte(day1_date-1, date, censor_date, na.censor=TRUE),
+      tte = tte(day0_date, date, censor_date, na.censor=TRUE),
     )
 
   data_timevaryingoutcomes <-
@@ -167,8 +169,8 @@ data_seqtrialcox <- local({
       by = c("patient_id")
     ) %>%
     filter(
-      tstart < tstop2,
-      tstart >= tstart2
+      tte_recruitment < tstop2,
+      tte_recruitment >= tstart2
     ) %>%
     select(-tstart2, -tstop2)
 
@@ -262,7 +264,6 @@ formula_demog <- . ~ . +
   ethnicity_combined
 
 formula_clinical <- . ~ . +
[…]

### 2022-02-09 analysis/combine_match.R
remove merge script (move code into match script)

@@ -103,7 +103,7 @@ write_csv(matchsummary_treated, fs::path(output_dir, "summary_treated.csv"))
 table1 <-
   match_metaparams %>%
   mutate(
-    table1 = map(treatment, ~read_csv(here("output", "models", "seqtrialcox", .x, glue("merge_table1.csv"))))
+    table1 = map(treatment, ~read_csv(here("output", "models", "seqtrialcox", .x, glue("match_table1.csv"))))
   ) %>%
   unnest(table1)
 
@@ -112,7 +112,7 @@ write_csv(table1, fs::path(output_dir, "table1.csv"))
 table1by <-
   match_metaparams %>%
   mutate(
-    table1by = map(treatment, ~read_csv(here("output", "models", "seqtrialcox", .x, glue("merge_table1by.csv"))))
+    table1by = map(treatment, ~read_csv(here("output", "models", "seqtrialcox", .x, glue("match_table1by.csv"))))
   ) %>%
   unnest(table1by)
 

### 2022-02-09 analysis/match_seqtrialcox.R

@@ -375,10 +375,14 @@ data_matched <- local({
     matched <- bind_rows(matched, matched_i)
 
   }
-  select(matched, -subclass)
+
+  matched %>%
+  select(-subclass) %>%
+  mutate(
+    treated_patient_id = paste0(treated, "_", patient_id),
+  )
 
 })
-
 
 
 write_rds(data_matched, fs::path(output_dir, "match_data_matched.rds"))
@@ -390,6 +394,78 @@ logoutput_table(controls_per_trial)
 # max trial date
 max_trial_day <- max(data_matched$trial_day, na.rm=TRUE)
 logoutput("max trial day is ", max_trial_day)
+
+
+
+# combine matching dataset with all other variables required for modelling ----
+
+## baseline variables ----
+
+data_baseline <-
+  data_cohort %>%
+  transmute(
+    patient_id,
+    age,
+    ageband,
+    sex,
+    ethnicity_combined,
+    imd_Q5,
+    region,
+    jcvi_group,
+    rural_urban_group,
+    prior_tests_cat,
+    multimorb,
+    learndis,
+    sev_mental,
+    cev,
+    sev_obesity,
+    chronic_heart_disease,
+    chronic_kidney_disease,
+    diabetes,
+    chronic_liver_disease,
+    chronic_resp_disease,
+    chronic_neuro_disease,
+    immunosuppressed,
+    asplenia,
+    immuno,
+
+    vax12_type,
+    vax2_week,
+    vax3_date,
+    vax3_type,
+
+  )
+logoutput_datasize(data_baseline)
+if(removeobjects) rm(data_cohort)
+
+
+data_tte <- read_rds(here("output", "models", "seqtrialcox", treatment, "match_data_tte.rds"))
+
+## create variables-at-time-zero dataset - one row per trial per arm per patient ----
+data_merged <-
+  # add time-varying info as at recruitment date (= tte_trial)
+  data_matched %>%
+  left_join(
+    data_timevarying %>% select(-any_of(matching_variables)),
+    by = c("patient_id")
+  ) %>%
+  filter(
+    tte_recruitment < tstop,
+    tte_recruitment >= tstart
+  ) %>%
+  # add time-non-varying info
+  left_join(
+    data_baseline %>% select(-any_of(matching_variables)),
+    by=c("patient_id")
+  ) %>%
[…]

### 2022-02-09 analysis/merge_seqtrialcox.R

@@ -1,370 +1,0 @@
-
-# # # # # # # # # # # # # # # # # # # # #
-# This script:
-# imports matching data
-# adds info as at recruitment date
-# reports baseline characteristics
-#
-# The script must be accompanied by one arguments:
-# `treatment` - the exposure in the regression model, pfizer or moderna
-
-# # # # # # # # # # # # # # # # # # # # #
-
-# Preliminaries ----
-
-
-# import command-line arguments ----
-
-args <- commandArgs(trailingOnly=TRUE)
-
-
-if(length(args)==0){
-  # use for interactive testing
-  removeobjects <- FALSE
-  treatment <- "pfizer"
-} else {
-  removeobjects <- TRUE
-  treatment <- args[[1]]
-}
-
-
-
-## Import libraries ----
-library('tidyverse')
-library('here')
-library('glue')
-library('survival')
-library('gt')
-library('gtsummary')
-
-## Import custom user functions from lib
-
-source(here("lib", "functions", "utility.R"))
-source(here("lib", "functions", "survival.R"))
-source(here("lib", "functions", "redaction.R"))
-
-## create output directories ----
-
-output_dir <- here("output", "models", "seqtrialcox", treatment)
-fs::dir_create(output_dir)
-
-## create special log file ----
-cat(glue("## script info for {treatment} ##"), "  \n", file = fs::path(output_dir, glue("merge_log.txt")), append = FALSE)
-
-## functions to pass additional log info to seperate file
-logoutput <- function(...){
-  cat(..., file = fs::path(output_dir, glue("merge_log.txt")), sep = "\n  ", append = TRUE)
-  cat("\n", file = fs::path(output_dir, glue("merge_log.txt")), sep = "\n  ", append = TRUE)
-}
-
-logoutput_datasize <- function(x){
-  nm <- deparse(substitute(x))
-  logoutput(
-    glue(nm, " data size = ", nrow(x)),
-    glue(nm, " memory usage = ", format(object.size(x), units="GB", standard="SI", digits=3L))
-  )
-}
-
-logoutput_table <- function(x){
-  capture.output(
-    x,
-    file=fs::path(output_dir, glue("merge_log.txt")),
-    append=TRUE
-  )
[…]

### 2022-02-09 analysis/model_seqtrialcox.R

@@ -108,7 +108,7 @@ var_labels <- read_rds(here("lib", "design", "variable-labels.rds"))
 
 
 data_merged <-
-  read_rds(here("output", "models", "seqtrialcox", treatment, "merge_data_merged.rds")) %>%
+  read_rds(here("output", "models", "seqtrialcox", treatment, "match_data_merged.rds")) %>%
   mutate(none="") %>%
   fastDummies::dummy_cols(select_columns = subgroup_variable, remove_selected_columns = FALSE) %>%
   filter(.[[subgroup_dummy]]==1L)
@@ -158,7 +158,9 @@ data_timevaryingoutcomes <- local({
 })
 
 
-## combine outcome variables with variables-at-baseline ----
+## combine all together ----
+# outcome variables with variables-at-baseline and add follow-up periods
+# create one row per trial per arm per patient per follow-up period
 
 data_seqtrialcox <- local({
 

### 2022-02-09 create-project.R

@@ -75,28 +75,7 @@ action_match <- function(treatment){
         #svg = glue("output/models/seqtrialcox/{treatment}/match_*.svg"),
         png = glue("output/models/seqtrialcox/{treatment}/match_*.png"),
         pdf = glue("output/models/seqtrialcox/{treatment}/match_*.pdf"),
-      )
-    ),
-
-    action(
-      name = glue("merge_seqtrialcox_{treatment}"),
-      run = glue("r:latest analysis/merge_seqtrialcox.R"),
-      arguments = c(treatment),
-      needs = list(
-        glue("match_seqtrialcox_{treatment}"),
-        "data_selection",
-        "data_process_long"
-      ),
-      highly_sensitive = lst(
-        rds = glue("output/models/seqtrialcox/{treatment}/merge_*.rds")
-      ),
-      moderately_sensitive = lst(
-        txt = glue("output/models/seqtrialcox/{treatment}/merge_*.txt"),
-        csv = glue("output/models/seqtrialcox/{treatment}/merge_*.csv"),
-        html = glue("output/models/seqtrialcox/{treatment}/merge_*.html")
-        #svg = glue("output/models/seqtrialcox/{treatment}/merge_*.svg"),
-        #png = glue("output/models/seqtrialcox/{treatment}/merge_*.png"),
-        #pdf = glue("output/models/seqtrialcox/{treatment}/merge_*.pdf"),
+        html = glue("output/models/seqtrialcox/{treatment}/match_*.html")
       )
     )
   )
@@ -117,7 +96,6 @@ action_model <- function(
       arguments = c(treatment, outcome, subgroup),
       needs = list(
         glue("match_seqtrialcox_{treatment}"),
-        glue("merge_seqtrialcox_{treatment}"),
         "data_selection",
         "data_process_long"
       ),
@@ -427,10 +405,9 @@ actions_list <- splice(
       as.list(
         glue_data(
           .x=expand_grid(
-            script=c("match", "merge"),
             treatment=c("pfizer", "moderna")
           ),
-          "{script}_seqtrialcox_{treatment}"
+          "match_seqtrialcox_{treatment}"
         )
       )
     ),
@@ -450,10 +427,9 @@ actions_list <- splice(
       as.list(
         glue_data(
           .x=expand_grid(
-            script=c("match", "merge"),
             treatment=c("pfizer", "moderna")
           ),
-          "{script}_seqtrialcox_{treatment}"
[…]

### 2022-02-09 analysis/data_process_long.R
fix previous commit

@@ -158,7 +158,7 @@ data_allevents <-
     time = as.integer(date - study_dates$studystart_date-1),
   )
 
-write_rds(data_allevents, here("output", "data", "data_allevents.rds"), compress="gz")
+write_rds(data_allevents, here("output", "data", "data_long_allevents.rds"), compress="gz")
 
 data_timevarying <-
   data_processed %>%
@@ -199,7 +199,7 @@ data_timevarying <-
   ) %>%
   mutate(id=NULL)
 
-write_rds(data_timevarying, here("output", "data", "data_timevarying.rds"), compress="gz")
+write_rds(data_timevarying, here("output", "data", "data_long_timevarying.rds"), compress="gz")
 
 # write_rds(data_pr_probable_covid, here("output", cohort, "data", "data_long_pr_probable_covid_dates.rds"), compress="gz")
 

### 2022-02-09 analysis/match_seqtrialcox.R

@@ -168,7 +168,7 @@ data_nontimevarying <-
 
 
 data_timevarying <-
-  read_rds(here("output", "data", "data_timevarying.rds")) %>%
+  read_rds(here("output", "data", "data_long_timevarying.rds")) %>%
   select(
     patient_id,
     tstart,
@@ -436,8 +436,8 @@ data_baseline <-
 
   )
 logoutput_datasize(data_baseline)
-if(removeobjects) rm(data_cohort)
 
+if(removeobjects) rm(data_cohort)
 
 data_tte <- read_rds(here("output", "models", "seqtrialcox", treatment, "match_data_tte.rds"))
 
@@ -473,7 +473,7 @@ write_rds(data_merged, fs::path(output_dir, "match_data_merged.rds"))
 data_coverage <-
   data_matched %>%
   filter(treated==1L) %>%
-  left_join(data_cohort %>% select(patient_id, vax3_date), by="patient_id") %>%
+  left_join(data_tte %>% transmute(patient_id, vax3_date=treatment_date), by="patient_id") %>%
   group_by(across(all_of(rolling_variables)), vax3_date) %>%
   summarise(
     n_matched=n(),

### 2022-02-09 analysis/model_seqtrialcox.R

@@ -126,7 +126,7 @@ data_timevaryingoutcomes <- local({
     select(patient_id, day0_date, censor_date, tte_stop)
 
   data_outcome <-
-    read_rds(here("output", "data", glue("data_allevents.rds"))) %>%
+    read_rds(here("output", "data", glue("data_long_allevents.rds"))) %>%
     filter(event==outcome) %>%
     inner_join(data_join, ., by =c("patient_id")) %>%
     mutate(

### 2022-02-09 create-project.R

@@ -65,7 +65,7 @@ action_match <- function(treatment){
       name = glue("match_seqtrialcox_{treatment}"),
       run = glue("r:latest analysis/match_seqtrialcox.R"),
       arguments = c(treatment),
-      needs = list("data_selection"),
+      needs = list("data_selection", "data_process_long"),
       highly_sensitive = lst(
         rds = glue("output/models/seqtrialcox/{treatment}/match_*.rds")
       ),

### 2022-02-09 analysis/data_process_long.R
fix outcome labelling

@@ -82,7 +82,8 @@ data_postest <- data_processed %>%
     values_to = "date",
     values_drop_na = TRUE
   ) %>%
-  arrange(patient_id, date)
+  arrange(patient_id, date) %>%
+  mutate(event="postest") # need to change name to match "outcome" argument
 
 data_covidemergency <- data_processed %>%
   select(patient_id, matches("^covidemergency\\_\\d+\\_date")) %>%
@@ -178,7 +179,7 @@ data_timevarying <-
     data2=data_allevents,
     id=patient_id,
     #postest = event(time, (event=="positive_test") *1L),
-    postest = event(if_else(event=="positive_test", time, NA_integer_)),
+    postest = event(if_else(event=="postest", time, NA_integer_)),
     covidemergency = event(time, (event=="covidemergency") *1L),
     covidadmitted = event(time, (event=="covidadmitted") *1L),
     covidcc = event(time, (event=="covidcc") *1L),

### 2022-02-09 analysis/combine_match.R
choose all relevent outcomes; tidy file paths

@@ -84,7 +84,7 @@ matchsummary <-
   ) %>%
   unnest(summary)
 
-write_csv(matchsummary, fs::path(output_dir, "matchsummary.csv"))
+write_csv(matchsummary, fs::path(output_dir, "summary.csv"))
 
 
 matchsummary_treated <-

### 2022-02-09 analysis/combine_model.R

@@ -13,8 +13,8 @@ args <- commandArgs(trailingOnly=TRUE)
 if(length(args)==0){
   # use for interactive testing
   removeobs <- FALSE
-  subgroup_variable <- "vax12_type"
-  #subgroup_variable <- "none"
+  #subgroup_variable <- "vax12_type"
+  subgroup_variable <- "none"
 } else {
   removeobs <- TRUE
   subgroup_variable <- args[[1]]
@@ -85,8 +85,8 @@ if(Sys.getenv("OPENSAFELY_BACKEND") %in% c("", "expectations")){
   model_metaparams <-
     expand_grid(
       treatment = factor(c("pfizer", "moderna")),
-      #outcome = factor(c("postest", "covidemergency", "covidadmission", "coviddeath")),
-      outcome = factor(c("postest", "covidadmission")),
+      outcome = factor(c("postest", "covidemergency", "covidadmission", "coviddeath")),
+      #outcome = factor(c("postest", "covidadmission")),
       subgroup = factor(recode_subgroup)
     ) %>%
     mutate(

### 2022-02-09 create-project.R

@@ -437,8 +437,8 @@ actions_list <- splice(
           .x=expand_grid(
             script=c("model", "report"),
             treatment=c("pfizer", "moderna"),
-            #outcome=c("postest", "covidemergency", "covidadmitted", "coviddeath"),
-            outcome=c("postest", "covidadmitted"),
+            outcome=c("postest", "covidemergency", "covidadmitted", "coviddeath"),
+            #outcome=c("postest", "covidadmitted"),
             subgroup_variable = c("none")
           ),
           "{script}_seqtrialcox_{treatment}_{outcome}_{subgroup_variable}"
@@ -473,8 +473,8 @@ actions_list <- splice(
           .x=expand_grid(
             script=c("model", "report"),
             treatment=c("pfizer", "moderna"),
-            #outcome=c("postest", "covidemergency", "covidadmitted", "covidcc", "coviddeath"),
-            outcome=c("postest", "covidadmitted"),
+            outcome=c("postest", "covidemergency", "covidadmitted",  "coviddeath"),
+            #outcome=c("postest", "covidadmitted"),
             subgroup = paste0("vax12_type-",c("pfizer-pfizer", "az-az"))
           ),
           "{script}_seqtrialcox_{treatment}_{outcome}_{subgroup}"

### 2022-02-09 create-project.R
include match outputs

@@ -501,6 +501,7 @@ actions_list <- splice(
       "descriptive_vaxdate",
       "match_seqtrialcox_pfizer",
       "match_seqtrialcox_moderna",
+      "combine_match",
       "combine_model",
       "combine_model_vax12_type"
     ),
@@ -509,7 +510,6 @@ actions_list <- splice(
       png = "output/manuscript-objects/*.png",
     )
   )
-
 
 )
 

### 2022-02-11 analysis/table1.R
use local var labels (not in design script)

@@ -39,8 +39,46 @@ data_cohort <- read_rds(here("output", "data", "data_cohort.rds")) %>%
     )
   )
 
-var_labels <-splice(N = N  ~ "Total N", var_labels)
-var_labels <-splice(allpop = allpop  ~ "All", var_labels)
+
+
+## variable labels
+var_labels <-
+  list(
+    N  ~ "Total N",
+    allpop  ~ "All",
+    vax12_type ~ "Primary vaccination course (doses 1 and 2)",
+    age ~ "Age",
+    ageband ~ "Age",
+    sex ~ "Sex",
+    ethnicity_combined ~ "Ethnicity",
+    imd_Q5 ~ "IMD",
+    region ~ "Region",
+    rural_urban_group ~ "Rural/urban category",
+    stp ~ "STP",
+    jcvi_group ~ "JCVI group",
+    cev ~ "Clinically extremely vulnerable",
+
+    sev_obesity ~ "Body Mass Index > 40 kg/m^2",
+
+    chronic_heart_disease ~ "Chronic heart disease",
+    chronic_kidney_disease ~ "Chronic kidney disease",
+    diabetes ~ "Diabetes",
+    chronic_liver_disease ~ "Chronic liver disease",
+    chronic_resp_disease ~ "Chronic respiratory disease",
+    chronic_neuro_disease ~ "Chronic neurological disease",
+    immunosuppressed ~ "Immunosuppressed",
+    asplenia ~ "Asplenia or poor spleen function",
+
+    learndis ~ "Learning disabilities",
+    sev_mental ~ "Serious mental illness",
+
+    multimorb ~ "Morbidity count",
+
+    prior_covid_infection0 ~ "Prior SARS-CoV-2 infection",
+    prior_tests_cat ~ "Number of SARS-CoV-2 tests in previous 3 months"
+
+  ) %>%
+  set_names(., map_chr(., all.vars))
 
 ## baseline variables
 tab_summary_baseline <- data_cohort %>%

### 2022-02-11 manuscript/draft-manuscript.Rmd
include ineligible treated people in match summary

@@
 
-## Study population
+## Study population and matching
 
-A total of `r label_number(1, big.mark=",")(filter(flowchart, crit=="c0") %>% pull(n))` fully vaccinated (received a primary course of two COVID-19 vaccine doses) people aged 18 and over and registered at a TPP practice on `r format(as.Date(study_dates$studystart_date), "%d %B %Y")` were identified the database, with `r label_number(1, big.mark=",")(filter(flowchart, crit=="c4") %>% pull(n))` (`r label_percent(0.1)(filter(flowchart, crit=="c4") %>% pull(pct_all))`) eligible for inclusion (supplementary Tables S1 and S2).
-Of these, `r label_number(1, big.mark=",")(cohortsummary$n_boosted)` received a booster dose during the recruitment period (up to `r format(as.Date(study_dates$lastvax3_date), "%d %B %Y")`).
-`r label_number(1, big.mark=",")(cohortsummary$n_3pfizer)` ( `r label_percent(0.1)(cohortsummary$prop_3pfizer)`) for BNT162b2 and `r label_number(1, big.mark=",")(cohortsummary$n_3moderna)` ( `r label_percent(0.1)(cohortsummary$prop_3moderna)`) for mRNA-1273 (Figure 1).
+A total of `r filter(matchsummary, treatment=="pfizer") %>% pull(n_treated))` BNT162b2 booster doses were administered between `r format(as.Date(study_dates$studystart_date), "%d %B %Y")` and `r format(as.Date(study_dates$studyend_date), "%d %B %Y")` in adults registered at a TPP practice, with `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_eligible))` (`r label_percent(0.1)(filter(matchsummary, treatment=="pfizer") %>% pull(prop_eligible_treated))`) eligible for inclusion into the treatment arm (supplementary Tables S1 and S2).
+Of these, `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` (`r label_percent(0.1)(filter(matchsummary, treatment=="pfizer") %>% pull(prop_matched_eligible))`) were successfully matched with unboosted controls (Figure 1), with `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(fup_years))` person-years of follow-up across both trial arms (Table 2).
 
-## Matching
-
-`r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n))` (`r label_percent(0.1)(filter(matchsummary, treatment=="pfizer") %>% pull(prop_recruited))`) BNT162b2 booster recipients were successfully matched to unboosted controls, contributing `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(fup_years))` person-years of follow-up across both trial arms.
-For mRNA-1273, this was `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="moderna") %>% pull(n))` (`r label_percent(0.1)(filter(matchsummary, treatment=="moderna") %>% pull(prop_recruited))`) and `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="moderna") %>% pull(fup_years))` person-years of follow-up (Table 2).
+<!-- For mRNA-1273, this was `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="moderna") %>% pull(n))` (`r label_percent(0.1)(filter(matchsummary, treatment=="moderna") %>% pull(prop_recruited))`) and `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="moderna") %>% pull(fup_years))` person-years of follow-up (Table 2).
+-->
 
@@
 
-Booster vaccination update over the duration of the study cohort, and the proportion of those that were matched for modelling.
+Booster vaccination uptake over the duration of the study period, and the proportion of those that were matched for modelling.
 
@@
       y=cumuln,
-      group=fct_rev(matched),
-      fill=fct_rev(matched),
+      group=status,
+      fill=status,
       colour=NULL

### 2022-02-11 analysis/match_seqtrialcox.R

@@ -34,6 +34,7 @@ library('tidyverse')
 library('here')
 library('glue')
 library('survival')
+library('MatchIt')
 
 ## Import custom user functions from lib
 
@@ -224,7 +225,7 @@ logoutput_datasize(data_tte)
 
 library("MatchIt")
 
-data_matched <- local({
+local({
 
   ## sequential trial matching routine is as follows:
   # each daily trial includes all n people who were vaccinated on that day (treated=1) and
@@ -244,11 +245,13 @@ data_matched <- local({
   candidate_ids0 <- data_tte$patient_id
 
   # initialise matching summary data
-  matched <- NULL
+  data_treated <- NULL
+  data_eligible <- NULL
+  data_matched <- NULL
 
   # matching formula
   matching_formula <- as.formula(str_c("treated ~ ", paste(matching_variables, collapse=" + ")))
-  trial=1
+  #trial=1
   for(trial in trials){
 
     #cat(trial, "\n")
@@ -260,6 +263,15 @@ data_matched <- local({
         data_timevarying,
         trial_time>=tstart,
         trial_time<tstop
+      )
+
+    data_treated_i <-
+      data_tte %>%
+      filter(tte_treatment %in% trial_time) %>%
+      transmute(
+        patient_id,
+        treated=1L,
+        trial_time=tte_treatment
       )
 
     # set of people boosted on trial day, + their candidate controls
@@ -282,20 +294,30 @@ data_matched <- local({
         status_hospunplanned==0L
       ) %>%
       mutate(
+        treated = tte_treatment %in% trial_time,
         # everyone treated on trial day and eligible
-        treated = ((tte_treatment %in% trial_time) & treated_within_recruitment_period)*1,
+        treated_candidate = ((tte_treatment %in% trial_time) & treated_within_recruitment_period)*1,
         # everyone not yet treated by trial day, not yet censored, and not already selected as a control
         control_candidate = (((tte_treatment > trial_time) | is.na(tte_treatment)) & (patient_id %in% candidate_ids0))*1
       ) %>%
-      filter((treated==1L) | (control_candidate==1L))
+      filter((treated_candidate==1L) | (control_candidate==1L))
+
+    data_eligible_i <-
+      matching_candidates_i %>%
[…]

### 2022-02-11 analysis/combine_model.R
fix subgroup labels

@@ -54,8 +54,8 @@ recode_outcome <- set_names(events_lookup$event, events_lookup$event_descr)
 
 if(subgroup_variable=="none"){
   recode_subgroup_variable <- c(`Main analysis` = "none")
-  recode_subgroup <- c(` `="")
-  subgroup <- c("")
+  recode_subgroup <- c(` `="none")
+  subgroup <- c("none")
 }
 
 if(subgroup_variable=="vax12_type"){
@@ -74,12 +74,14 @@ if(Sys.getenv("OPENSAFELY_BACKEND") %in% c("", "expectations")){
     expand_grid(
       treatment = factor(c("pfizer")),
       outcome = factor(c("postest")),
-      subgroup = factor(recode_subgroup)
+      subgroup = factor(recode_subgroup),
+      subgroup_variable = factor(subgroup_variable)
     ) %>%
     mutate(
       treatment_descr = fct_recode(treatment,  !!!recode_treatment),
       outcome_descr = fct_recode(outcome,  !!!recode_outcome),
-      subgroup_descr = fct_recode(recode_subgroup,  !!!recode_subgroup)
+      subgroup_descr = fct_recode(recode_subgroup,  !!!recode_subgroup),
+      subgroup_variable_descr = fct_recode(subgroup_variable,  !!!recode_subgroup_variable)
     )
 } else {
   model_metaparams <-
@@ -87,12 +89,14 @@ if(Sys.getenv("OPENSAFELY_BACKEND") %in% c("", "expectations")){
       treatment = factor(c("pfizer", "moderna")),
       outcome = factor(c("postest", "covidemergency", "covidadmitted", "coviddeath")),
       #outcome = factor(c("postest", "covidadmission")),
-      subgroup = factor(recode_subgroup)
+      subgroup = factor(recode_subgroup),
+      subgroup_variable = factor(subgroup_variable)
     ) %>%
     mutate(
       treatment_descr = fct_recode(treatment,  !!!recode_treatment),
       outcome_descr = fct_recode(outcome,  !!!recode_outcome),
-      subgroup_descr = fct_recode(subgroup,  !!!recode_subgroup)
+      subgroup_descr = fct_recode(subgroup,  !!!recode_subgroup),
+      subgroup_variable_descr = fct_recode(subgroup_variable,  !!!recode_subgroup_variable)
     )
 }
 

### 2022-02-11 analysis/combine_model.R
redo var labels

@@ -13,8 +13,8 @@ args <- commandArgs(trailingOnly=TRUE)
 if(length(args)==0){
   # use for interactive testing
   removeobs <- FALSE
-  #subgroup_variable <- "vax12_type"
-  subgroup_variable <- "none"
+  subgroup_variable <- "vax12_type"
+  #subgroup_variable <- "none"
 } else {
   removeobs <- TRUE
   subgroup_variable <- args[[1]]

### 2022-02-11 analysis/table1.R

@@ -19,8 +19,6 @@ source(here("lib", "functions", "redaction.R"))
 ## create output directories ----
 fs::dir_create(here("output", "descriptive", "table1"))
 
-## import metadata ----
-var_labels <- read_rds(here("lib", "design", "variable-labels.rds"))
 
 #list_formula <- read_rds(here("output", "data", "metadata_formulas.rds"))
 #list2env(list_formula, globalenv())

### 2022-02-11 analysis/match_seqtrialcox.R
Update match_seqtrialcox.R

@@ -468,8 +468,6 @@ data_baseline <-
   )
 logoutput_datasize(data_baseline)
 
-if(removeobjects) rm(data_cohort)
-
 
 ## create variables-at-time-zero dataset - one row per trial per arm per patient ----
 data_merged <-
@@ -507,7 +505,7 @@ data_coverage <-
   data_summary %>%
   filter(treated==1L) %>%
   #left_join(data_tte %>% transmute(patient_id, vax3_date=treatment_date), by="patient_id") %>%
-  left_join(data_cohort %>% select(patient_id, all_of(rolling_variables), vax3_date), by="patient_id") %>%
+  left_join(data_nontimevarying %>% select(patient_id, all_of(rolling_variables), vax3_date), by="patient_id") %>%
   group_by(across(all_of(rolling_variables)), vax3_date) %>%
   summarise(
     n_treated=sum(treated, na.rm=TRUE),

### 2022-02-11 analysis/match_seqtrialcox.R
Update match_seqtrialcox.R

@@ -468,6 +468,7 @@ data_baseline <-
   )
 logoutput_datasize(data_baseline)
 
+if(removeobjects) rm(data_cohort)
 
 ## create variables-at-time-zero dataset - one row per trial per arm per patient ----
 data_merged <-

### 2022-02-11 analysis/match_seqtrialcox.R
Update match_seqtrialcox.R

@@ -313,8 +313,8 @@ local({
 
 
     if(sum(matching_candidates_i$treated_candidate)<1) {
-      message("Terminating trial sequence at trial ", trial, " - No treated people eligible for inclusion.")
-      break
+      message("Skipping trial ", trial, " - No treated people eligible for inclusion.")
+      next
     }
 
     n_treated_all <- sum(matching_candidates_i$treated_candidate)

### 2022-02-13 manuscript/draft-manuscript.Rmd

@@
 # where are the outputs (ie the inputs for this manuscript!) saved?
-#output_dir_os <- here("output", "manuscript-objects")
-output_dir_os <- here("released_output", "manuscript-objects")
+output_dir_os <- here("output", "manuscript-objects")
+#output_dir_os <- here("released_output", "manuscript-objects")
 
@@
       "postest",
-      "covidemergency"#,
-      #"covidadmitted",
+      #"covidemergency",
+      "covidadmitted",
       #"covidcc",
-      #"coviddeath"
+      "coviddeath"
     ) %>% 
@@
 
-hazards <- read_csv(fs::path(output_dir_os, "effects.csv")) %>%
+hazards_none <- read_csv(fs::path(output_dir_os, "none", "effects.csv")) %>%
+  mutate(
+    treatment_descr = fct_inorder(treatment_descr),
+    outcome_descr = fct_inorder(outcome_descr),
+    #model_name = fct_inorder(str_wrap(model_name, 15)),
+    hr_display = paste0(number_format(accuracy=0.01)(hr), " (", number_format(accuracy=0.01)(hr.ll), "-", number_format(accuracy=0.01)(hr.ul), ")")
+  )
+
+hazards_vax12_type <- read_csv(fs::path(output_dir_os, "vax12_type", "effects.csv")) %>%
   mutate(
@@
 #riskdiff3 <- riskdiff %>% filter(model==3)
-hazards3 <- hazards %>% filter(model==3)
+hazards3_none <- hazards_none %>% filter(model==3)
+hazards3_vax12_type <- hazards_vax12_type %>% filter(model==3)
 
@@
 Initially offered to those at highest risk of severe disease, eligibility was extended to the entire adult population by the end of the year.
-In the vast majority of cases, BNT162b2 and mRNA-1273 vaccines were used and administered without consideration of the type of primary course.
 
 We set out to estimate the effectiveness of booster doses amongst the fully vaccinated adult population in England using the OpenSAFELY-TPP linked primary care database covering around 40% of England's population.
+
+In the vast majority of cases, BNT162b2 and mRNA-1273 vaccines were used, and administered without consideration of the type of primary course.
+BNT162b2 was used from 16 September 2022, and mRNA-1273 from X onwards.
+Consequently, available follow-up of BNT162b2 recipients is longer, and so in this early analysis we restrict to effectiveness of this vaccine.
+An analysis of mRNA-1273 booster effectiveness is planned when sufficient follow-up has accrued.
 
@@
 
-## Study population and design
+## Study design and population
 
-We studied adults (18 and over) in England who had received a complete primary COVID-19 vaccination course (two doses) of BNT162b2 or ChAdOx1-S and who were registered at a GP practice using TPP's SystmOne clinical information system.
-People were not eligible for inclusion if: they were a vaccinated health care worker, resident in a care or nursing home, medically housebound or receiving end-of-life care (due to distinct clustering of vaccine uptake and infection risk in these groups that could not be reasonably measured and controlled for); had received heterologous (i.e., mixed) first and second vaccine dose; information on sex, ethnicity, deprivation, and geographical region was not known.
+We used a sequential trials approach designed to emulate a sequence of daily two-arm parallel randomised controlled trials.
+For each trial, the treatment arm comprised people who received a BNT162b2 booster dose on that day, and the control arm comprised the same number of randomly selected unboosted people, matched on key characteristics described below.
 
-After exclusions for missing values on demographic variables, there were no missing values in remaining variables as they were defined by the presence or absence of clinical codes or events in the patient record.
-
-We used a sequential trials approach designed to emulate a sequence of two-arm parallel randomised controlled trials recruiting participants between `r format(as.Date(study_dates$studystart_date), "%d %B %Y")` to `r format(as.Date(study_dates$studyend_date), "%d %B %Y")` inclusive.
-For each trial, the treatment arm comprises people who received a booster dose on that day, and the control arm comprises the same number of unboosted people selected using exact 1:1 matching without replacement.
-Trial arms are matched on primary course vaccine type (BNT162b2, ChAdOx1-S, or mRNA-1273), the calendar week of the second vaccine dose, geographical region, prior infection status, and vaccine priority group (as determined by the Joint Committee for Vaccine and immunisation expert working group).
-People selected as controls are not eligible to be included again as a control in a subsequent trial, but are eligible for selection into the treatment arm of a subsequent trial.
-Unmatched booster recipients were not included.
+Trial recruitment was considered between `r format(as.Date(study_dates$studystart_date), "%d %B %Y")` and `r format(as.Date(study_dates$studyend_date), "%d %B %Y")` inclusive.
+Trial eligibility on each day of recruitment was as follows: adults (18 and over) who were registered at a GP practice using TPP's SystmOne clinical information system; received a primary COVID-19 vaccination course of either two BNT162b2 doses or two ChAdOx1-S doses (mRNA-1273 was not considered due to small numbers); not a vaccinated health care worker, not a resident in a care or nursing home, and not medically housebound or receiving end-of-life care (due to distinct clustering of vaccine uptake and infection risk in these groups that could not be reasonably measured and controlled for); no evidence of SARS-CoV-2 infection or COVID-19 disease within 90 days prior to recruitment; not undergoing an unplanned hospital admission; information on sex, ethnicity, deprivation, and geographical region was complete.
 
 Additional eligiblity criteria based on sufficiently many people being vaccinated on each day within region, priority group, and second dose week strata were also used to ensure unboosted people were both eligible and able to receive a booster dose at the time of potential recruitment.
-The trial sequence was terminated early if fewer than 10% of treated people were successfully matched.
+The trial sequence was terminated early if nobody was eligible for the treatment arm or fewer than 10% of eligible treated people were successfully matched.
 See supplementary materials for further details.
+
+The control arms were selected using exact 1:1 matching without replacement.
+Participants were matched on primary course vaccine type (BNT162b2 or ChAdOx1-S), the calendar week of the second vaccine dose, geographical region, prior infection, immunosuppression, and vaccine priority group (as determined by the Joint Committee for Vaccine and immunisation expert working group).
+People selected as controls were not eligible to be included again as a control in a subsequent trial, but were eligible for selection into the treatment arm of a subsequent trial.
+Any unmatched booster recipients were dropped.
+
+After exclusions for missing values on demographic variables, there were no missing values in remaining variables as they were each defined by the presence or absence of clinical codes or events in the patient record.
 
@@
 
-The following five outcomes were considered: positive SARS-CoV-2 test; COVID-19 A&E attendance; unplanned COVID-19 hospital admission; COVID-19 critical care admission; COVID-19 death.
+## Outcomes
+
+The following three outcomes were considered: positive SARS-CoV-2 test; unplanned COVID-19 hospital admission; COVID-19 death.
 Positive SARS-CoV-2 tests were identified using SGSS records and based on swab date.
 Both polymerase chain reaction (PCR) and lateral flow tests are included, without differentiation between symptomatic and asymptomatic infection.
-COVID-19 A&E attendances were identified using HES emergency care records with U07.1 ("COVID-19, virus identified") or U07.2 ("COVID-19, virus not identified") ICD-10 diagnosis codes @emergenc.
-Unplanned COVID-19 hospital admissions and critical care admissions were identified using HES in-patient hospital records with U07.1 or U07.2 reason for admission ICD-10 codes.
-COVID-19 deaths with ICD-10 codes (as above) mentioned anywhere on the death certificate (i.e., as an underlying or contributing cause of death) were included.
+Unplanned COVID-19 hospital admissions were identified using HES in-patient hospital records with U07.1 or U07.2 reason for admission ICD-10 codes.
+COVID-19 deaths with the same COVID-19 ICD-10 codes mentioned anywhere on the death certificate (i.e., as an underlying or contributing cause of death) were included.
+
+<!-- COVID-19 A&E attendances were identified using HES emergency care records with U07.1 ("COVID-19, virus identified") or U07.2 ("COVID-19, virus not identified") ICD-10 diagnosis codes @emergenc.
+-->
+
+COVID-19 related A&E attendance and critical/intensive care admission we also considered but there were doubts over data the reliability of event ascertainment in these cases.
 
@@
 
-Participant characteristics used describe the cohort and for confounder adjustment include: age, sex (male or female), English Index of Multiple Deprivation (IMD, grouped by quintiles), ethnicity (Black, Mixed, South Asian, White, Other, as per the UK census), NHS region (East of England, Midlands, London, North East and Yorkshire, North West, South East, South West), number of conditions in the clinically "at risk" classification, as per national prioritisation guidelines, the number of SARS-CoV-2 tests in the 6 months prior to the study start date (via SGSS), rurality (urban conurbation, urban city or town, rural town or village), evidence of prior SARS-CoV-2 infection (positive SARS-CoV-2 test, "probable" infection documented in primary care, or COVID-19 hospitalisation), learning disabilities, severe mental illness, undergoing an unplanned hospital admission, undergoing a planned hospital admission.
+Participant characteristics used describe trial participants and for confounder adjustment include: age, sex (male or female), English Index of Multiple Deprivation (IMD, grouped by quintiles), ethnicity (Black, Mixed, South Asian, White, Other, as per the UK census), NHS region (East of England, Midlands, London, North East and Yorkshire, North West, South East, South West), number of conditions in the clinically "at risk" classification, as per national prioritisation guidelines, the number of SARS-CoV-2 tests in the 6 months prior to the study start date (via SGSS), rurality (urban conurbation, urban city or town, rural town or village), evidence of prior SARS-CoV-2 infection (positive SARS-CoV-2 test, "probable" infection documented in primary care, or COVID-19 hospitalisation), learning disabilities, severe mental illness, undergoing an unplanned hospital admission, undergoing a planned hospital admission.
 All characteristics were ascertained as at the trial recruitment date (or study start date if assumed to be fixed from that date onwards).
@@
 
-Time zero was the trial recruitment day, and trial participants were followed up for no more than 12 weeks from this day.
+Time zero was the trial recruitment day, and trial participants were followed up for no more than 10 weeks from this day.
 Follow-up was censored earlier than this at `r format(as.Date(study_dates$studyend_date), "%d %B %Y")`, death, or GP practice de-registration.
@@
 
-Data from each trial were pooled and analysed using Cox regression with period-specific hazard ratios for treatment, in two-week periods.
-Baseline hazards were allowed to differ for participants differing by recruitment day, region, priority group, and week and type of second vaccine dose, via Cox stratification.
+Trials were pooled and analysed using Cox regression with period-specific hazard ratios for treatment, in two-week periods.
+Baseline hazards were allowed to differ for participants differing by recruitment day, region, and week and type of second vaccine dose, via Cox stratification.
 Other variables were included as covariates.
@@
 A total of `r filter(matchsummary, treatment=="pfizer") %>% pull(n_treated))` BNT162b2 booster doses were administered between `r format(as.Date(study_dates$studystart_date), "%d %B %Y")` and `r format(as.Date(study_dates$studyend_date), "%d %B %Y")` in adults registered at a TPP practice, with `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_eligible))` (`r label_percent(0.1)(filter(matchsummary, treatment=="pfizer") %>% pull(prop_eligible_treated))`) eligible for inclusion into the treatment arm (supplementary Tables S1 and S2).
-Of these, `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` (`r label_percent(0.1)(filter(matchsummary, treatment=="pfizer") %>% pull(prop_matched_eligible))`) were successfully matched with unboosted controls (Figure 1), with `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(fup_years))` person-years of follow-up across both trial arms (Table 2).
+Of these, `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` (`r label_percent(0.1)(filter(matchsummary, treatment=="pfizer") %>% pull(prop_matched_eligible))`) were successfully matched with unboosted controls (Figure 1).
 
@@
 
-Characteristics at recruitment between treatment and controls arms are provided in supplementary materials (Table S2). Characteristics were (??) well-balanced between trial arms.
+Participant characteristics at trial recruitment between trial arms were often well-balanced (Table 1), with some notable exceptions.
+Unboosted controls were typically...
+These unbalanced characteristics were adjusted for in the Cox regression models.
+
+There were `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(fup_years))` person-years of follow-up across both trial arms (Table 2).
 
@@
 
-#### Figure 1: Booster vaccination uptake, and matching success
+#### Figure 1: Booster vaccination uptake, treatment arm eligiblity and matching success
 
-Booster vaccination uptake over the duration of the study period, and the proportion of those that were matched for modelling.
+Booster vaccination uptake over the duration of the study period, and the number of boosted people who were eligible and matched.
 
@@
 
-#### Table 1: Baseline characteristics
+#### Table 1: Participant characteristics
 
-Participant characteristics as on the day of recruitment into the treatment or control arms, stratified by treatment and outcome type.
+Participant characteristics as on the day of recruitment into the treatment or control arm.
 
@@
   table1_body %>%
+  #filter(outcome %in% outcomes) %>%
   group_by(variable) %>%
@@
 incidences %>%
-  filter(fup_period == "All") %>%
+  filter(
+    fup_period == "All",
+    outcome %in% c("postest", "covidadmitted", "coviddeath")
+  ) %>%
   left_join(
@@
 
-In the BNT162b2 and mRNA-1273 analyses respectively, matched participants contributed `r filter(incidences, treatment=="pfizer", outcome=="postest", fup_period=="all") %>% pull(events)` and `r filter(incidences, treatment=="moderna", outcome=="postest", fup_period=="all") %>% pull(events)` positive SARS-CoV-2 tests, `r filter(incidences, treatment=="pfizer", outcome=="covidattendance", fup_period=="all") %>% pull(events)` and `r filter(incidences, treatment=="moderna", outcome=="covidattendance", fup_period=="all") %>% pull(events)` COVID-19 A&E attendances, `r filter(incidences, treatment=="pfizer", outcome=="covidadmitted", fup_period=="all") %>% pull(events)` and `r filter(incidences, treatment=="moderna", outcome=="covidadmitted", fup_period=="all") %>% pull(events)` COVID-19 hospital admissions, `r filter(incidences, treatment=="pfizer", outcome=="covidcc", fup_period=="all") %>% pull(events)` and `r filter(incidences, treatment=="moderna", outcome=="covidcc", fup_period=="all") %>% pull(events)` COVID-19 ciritical care admissions, and `r filter(incidences, treatment=="pfizer", outcome=="coviddeath", fup_period=="all") %>% pull(events)` and `r filter(incidences, treatment=="moderna", outcome=="coviddeath", fup_period=="all") %>% pull(events)` COVID-19 deaths (Table 2).
+Trial participants contributed `r filter(incidences, treatment=="pfizer", outcome=="postest", fup_period=="all") %>% pull(events)` positive SARS-CoV-2 tests, `r filter(incidences, treatment=="pfizer", outcome=="covidadmitted", fup_period=="all") %>% pull(events)` COVID-19 hospital admissions, and `r filter(incidences, treatment=="pfizer", outcome=="coviddeath", fup_period=="all") %>% pull(events)` COVID-19 deaths (Table 2).
 
-During the fifth and sixth week after recruitment (days 28-41) the estimated hazard ratio comparing a BNT162b2 booster dose to no dose for a positive test was `r filter(hazards, outcome=="postest", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)`, for COVID-19 A&E attendances was `r filter(hazards, outcome=="covidattendance", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)`, for COVID-19 hospital admission was `r filter(hazards, outcome=="covidadmitted", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)`, for COVID-19 critical care admission was `r filter(hazards, outcome=="covidcc", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)`, and for COVID-19 death was `r filter(hazards, outcome=="coviddeath", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)` (Figure 2).
-For mRNA-1273, the corresponding estimated hazard ratios were `r filter(hazards, outcome=="postest", treatment=="moderna", model==3, term_left==28) %>% pull(hr_display)`, for COVID-19 A&E attendances was `r filter(hazards, outcome=="covidattendance", treatment=="moderna", model==3, term_left==28) %>% pull(hr_display)`, for COVID-19 hospital admission was `r filter(hazards, outcome=="covidadmitted", treatment=="moderna", term_left==28) %>% pull(hr_display)`, for COVID-19 critical care admission was `r filter(hazards, outcome=="covidcc", treatment=="moderna", model==3, term_left==28) %>% pull(hr_display)`, and for COVID-19 death was `r filter(hazards, outcome=="coviddeath", treatment=="moderna", model==3, term_left==28) %>% pull(hr_display)`.
+<!-- `r filter(incidences, treatment=="pfizer", outcome=="covidattendance", fup_period=="all") %>% pull(events)` COVID-19 A&E attendances, `r filter(incidences, treatment=="pfizer", outcome=="covidcc", fup_period=="all") %>% pull(events)` COVID-19 ciritical care admissions, -->
 
-Estimated effectiveness by primary vaccine course...
+During the fifth and sixth week after recruitment (days 28-41) the estimated hazard ratio comparing a BNT162b2 booster dose to no dose was `r filter(hazards_none, outcome=="postest", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)` for a positive test, `r filter(hazards_none, outcome=="covidadmitted", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 hospital admission, and `r filter(hazards_none, outcome=="coviddeath", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 death (Figure 2).
+
+<!-- `r filter(hazards_none, outcome=="covidattendance", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 A&E attendances, `r filter(hazards_none, outcome=="covidcc", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 critical care admission,  -->
+
+<!-- For mRNA-1273, the corresponding estimated hazard ratios were `r filter(hazards, outcome=="postest", treatment=="moderna", model==3, term_left==28) %>% pull(hr_display)`, for COVID-19 A&E attendances was `r filter(hazards, outcome=="covidattendance", treatment=="moderna", model==3, term_left==28) %>% pull(hr_display)`, for COVID-19 hospital admission was `r filter(hazards, outcome=="covidadmitted", treatment=="moderna", term_left==28) %>% pull(hr_display)`, for COVID-19 critical care admission was `r filter(hazards, outcome=="covidcc", treatment=="moderna", model==3, term_left==28) %>% pull(hr_display)`, and for COVID-19 death was `r filter(hazards, outcome=="coviddeath", treatment=="moderna", model==3, term_left==28) %>% pull(hr_display)`.
+-->
+
+For the same period, the estimated effectiveness for BNT162b2 and ChAdOx1-S primary courses respectively were  `r filter(hazards_vax12_type, outcome=="postest", treatment=="pfizer", vax12_type=="pfizer-pfizer", model==3, term_left==28) %>% pull(hr_display)` and `r filter(hazards_vax12_type, outcome=="postest", treatment=="pfizer", vax12_type=="az-az", model==3, term_left==28) %>% pull(hr_display)` for a positive test, `r filter(hazards_vax12_type, outcome=="covidadmitted", treatment=="pfizer", vax12_type=="pfizer-pfizer", model==3, term_left==28) %>% pull(hr_display)` and `r filter(hazards_vax12_type, outcome=="covidadmitted", treatment=="pfizer", vax12_type=="az-az", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 hospital admission, and `r filter(hazards_vax12_type, outcome=="coviddeath", treatment=="pfizer", vax12_type=="pfizer-pfizer", model==3, term_left==28) %>% pull(hr_display)` and `r filter(hazards_vax12_type, outcome=="coviddeath", treatment=="pfizer", vax12_type=="az-az", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 death (Figure 2).
+
+<!--  `r filter(hazards_vax12_type, outcome=="covidemergency", treatment=="pfizer", vax12_type=="pfizer-pfizer", model==3, term_left==28) %>% pull(hr_display)` and `r filter(hazards_vax12_type, outcome=="emergency", treatment=="pfizer", vax12_type=="az-az", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 A&E attendances, -->
+<!-- `r filter(hazards_vax12_type, outcome=="covidcc", treatment=="pfizer", vax12_type=="pfizer-pfizer", model==3, term_left==28) %>% pull(hr_display)` and `r filter(hazards_vax12_type, outcome=="covidcc", treatment=="pfizer", vax12_type=="az-az", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 critical care admission, -->
 
@@
 
-For each outcome based on the fully adjusted model, the period-specific hazard ratio for boosting versus not boosting is shown.
+For each outcome based on the fully adjusted model, the period-specific hazard ratio for boosting with BNT162b2 versus not boosting is shown, stratified by primary course.
 Models with less extensive confounder adjustment are provided in supplementary materials (supplementary Figure SX).
@@
   plot_hazards <-
-    hazards3 %>%
+    hazards3_vax12_type %>%
+    filter(outcome %in% c("postest", "covidadmitted", "coviddeath")) %>%
     ggplot()+
@@
     geom_hline(aes(yintercept=1), colour='grey')+
-    facet_grid(rows=vars(outcome_descr), cols=vars(treatment_descr), switch="y")+
+    facet_grid(rows=vars(outcome_descr), cols=vars(subgroup_descr), switch="y")+
     scale_y_log10(

### 2022-02-14 analysis/manuscript_objects.R
update file names, manuscript

@@ -146,14 +146,14 @@ fs::file_copy(here("output", "descriptive", "vaxdate", "plot_vaxdate_stack_jcvi.
 
 ## matching ----
 
-fs::file_copy(here("output", "models", "seqtrialcox", "pfizer", "match_coverage_stack.png"), here("output", "manuscript-objects", "match_coverage_pfizer_stack.png"), overwrite = TRUE)
-fs::file_copy(here("output", "models", "seqtrialcox", "moderna", "match_coverage_stack.png"), here("output", "manuscript-objects", "match_coverage_moderna_stack.png"), overwrite = TRUE)
+#fs::file_copy(here("output", "models", "seqtrialcox", "pfizer", "match_coverage_stack.png"), here("output", "manuscript-objects", "match_coverage_pfizer_stack.png"), overwrite = TRUE)
+#fs::file_copy(here("output", "models", "seqtrialcox", "moderna", "match_coverage_stack.png"), here("output", "manuscript-objects", "match_coverage_moderna_stack.png"), overwrite = TRUE)
 
-fs::file_copy(here("output", "models", "seqtrialcox", "combined", "match", "table1.csv"), here("output", "manuscript-objects", "table1.csv"), overwrite = TRUE)
-fs::file_copy(here("output", "models", "seqtrialcox", "combined", "match", "table1by.csv"), here("output", "manuscript-objects", "table1by.csv"), overwrite = TRUE)
-fs::file_copy(here("output", "models", "seqtrialcox", "combined", "match", "coverage.csv"), here("output", "manuscript-objects", "coverage.csv"), overwrite = TRUE)
-fs::file_copy(here("output", "models", "seqtrialcox", "combined", "match", "summary.csv"), here("output", "manuscript-objects", "summary.csv"), overwrite = TRUE)
-fs::file_copy(here("output", "models", "seqtrialcox", "combined", "match", "summary_treated.csv"), here("output", "manuscript-objects", "summary_treated.csv"), overwrite = TRUE)
+fs::file_copy(here("output", "models", "seqtrialcox", "combined", "match", "table1.csv"), here("output", "manuscript-objects", "matchtable1.csv"), overwrite = TRUE)
+fs::file_copy(here("output", "models", "seqtrialcox", "combined", "match", "table1by.csv"), here("output", "manuscript-objects", "matchtable1by.csv"), overwrite = TRUE)
[…]

### 2022-02-14 manuscript/draft-manuscript.Rmd

@@
   mutate(
+    subgroup_descr = fct_inorder(subgroup_descr),
+    subgroup_level_descr = fct_inorder(subgroup_level_descr),
     treatment_descr = fct_inorder(treatment_descr),
@@
 
-## Outcomes
-
 The following three outcomes were considered: positive SARS-CoV-2 test; unplanned COVID-19 hospital admission; COVID-19 death.
@@
 
-A total of `r filter(matchsummary, treatment=="pfizer") %>% pull(n_treated))` BNT162b2 booster doses were administered between `r format(as.Date(study_dates$studystart_date), "%d %B %Y")` and `r format(as.Date(study_dates$studyend_date), "%d %B %Y")` in adults registered at a TPP practice, with `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_eligible))` (`r label_percent(0.1)(filter(matchsummary, treatment=="pfizer") %>% pull(prop_eligible_treated))`) eligible for inclusion into the treatment arm (supplementary Tables S1 and S2).
+A total of `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_treated))` BNT162b2 booster doses were administered between `r format(as.Date(study_dates$studystart_date), "%d %B %Y")` and `r format(as.Date(study_dates$studyend_date), "%d %B %Y")` in adults registered at a TPP practice, with `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_eligible))` (`r label_percent(0.1)(filter(matchsummary, treatment=="pfizer") %>% pull(prop_eligible_treated))`) eligible for inclusion into the treatment arm.
 Of these, `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` (`r label_percent(0.1)(filter(matchsummary, treatment=="pfizer") %>% pull(prop_matched_eligible))`) were successfully matched with unboosted controls (Figure 1).
 
-<!-- For mRNA-1273, this was `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="moderna") %>% pull(n))` (`r label_percent(0.1)(filter(matchsummary, treatment=="moderna") %>% pull(prop_recruited))`) and `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="moderna") %>% pull(fup_years))` person-years of follow-up (Table 2).
+<!-- For mRNA-1273, this was `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="moderna") %>% pull(n))` (`r label_percent(0.1)(filter(matchsummary, treatment=="moderna") %>% pull(prop_matched_eligible))`) and `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="moderna") %>% pull(fup_years))` person-years of follow-up (Table 2).
 -->
@@
 matchcoverage  %>%
+  filter(treatment=="pfizer") %>%
   mutate(
@@
       y=cumuln,
-      group=status,
-      fill=status,
+      group=fct_rev(status),
+      fill=fct_rev(status),
       colour=NULL
@@
   )+
-  scale_fill_brewer(type="qual", palette="Set2")+
-  scale_colour_brewer(type="qual", palette="Set2")+
+  scale_fill_brewer(type="qual", palette="Set2", direction = -1)+
+  scale_colour_brewer(type="qual", palette="Set2", direction = -1)+
   labs(
@@
 ```{r table1, echo=FALSE, warning=FALSE, message=FALSE}
-table1_body <- read_csv(fs::path(output_dir_os, "matchtable1.csv"))
-table1_by <- read_csv(fs::path(output_dir_os, "matchtable1by.csv"))
+table1_body <- read_csv(fs::path(output_dir_os, "matchtable1.csv")) %>% filter(treatment=="pfizer")
+table1_by <- read_csv(fs::path(output_dir_os, "matchtable1by.csv")) %>% filter(treatment=="pfizer")
 
@@
   left_join(
-    matchsummary %>% select(treatment, eligible_1=total_treated, recruit_prop_1 = prop_recruited),
+    matchsummary %>% select(treatment, eligible_1=n_eligible, recruit_prop_1 = prop_matched_eligible),
     by=c("treatment")
@@
 
-<!-- For mRNA-1273, the corresponding estimated hazard ratios were `r filter(hazards, outcome=="postest", treatment=="moderna", model==3, term_left==28) %>% pull(hr_display)`, for COVID-19 A&E attendances was `r filter(hazards, outcome=="covidattendance", treatment=="moderna", model==3, term_left==28) %>% pull(hr_display)`, for COVID-19 hospital admission was `r filter(hazards, outcome=="covidadmitted", treatment=="moderna", term_left==28) %>% pull(hr_display)`, for COVID-19 critical care admission was `r filter(hazards, outcome=="covidcc", treatment=="moderna", model==3, term_left==28) %>% pull(hr_display)`, and for COVID-19 death was `r filter(hazards, outcome=="coviddeath", treatment=="moderna", model==3, term_left==28) %>% pull(hr_display)`.
+<!-- For mRNA-1273, the corresponding estimated hazard ratios were `r filter(hazards_none, outcome=="postest", treatment=="moderna", model==3, term_left==28) %>% pull(hr_display)`, for COVID-19 A&E attendances was `r filter(hazards_none, outcome=="covidattendance", treatment=="moderna", model==3, term_left==28) %>% pull(hr_display)`, for COVID-19 hospital admission was `r filter(hazards_none, outcome=="covidadmitted", treatment=="moderna", term_left==28) %>% pull(hr_display)`, for COVID-19 critical care admission was `r filter(hazards_none, outcome=="covidcc", treatment=="moderna", model==3, term_left==28) %>% pull(hr_display)`, and for COVID-19 death was `r filter(hazards_none, outcome=="coviddeath", treatment=="moderna", model==3, term_left==28) %>% pull(hr_display)`.
 -->
 
-For the same period, the estimated effectiveness for BNT162b2 and ChAdOx1-S primary courses respectively were  `r filter(hazards_vax12_type, outcome=="postest", treatment=="pfizer", vax12_type=="pfizer-pfizer", model==3, term_left==28) %>% pull(hr_display)` and `r filter(hazards_vax12_type, outcome=="postest", treatment=="pfizer", vax12_type=="az-az", model==3, term_left==28) %>% pull(hr_display)` for a positive test, `r filter(hazards_vax12_type, outcome=="covidadmitted", treatment=="pfizer", vax12_type=="pfizer-pfizer", model==3, term_left==28) %>% pull(hr_display)` and `r filter(hazards_vax12_type, outcome=="covidadmitted", treatment=="pfizer", vax12_type=="az-az", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 hospital admission, and `r filter(hazards_vax12_type, outcome=="coviddeath", treatment=="pfizer", vax12_type=="pfizer-pfizer", model==3, term_left==28) %>% pull(hr_display)` and `r filter(hazards_vax12_type, outcome=="coviddeath", treatment=="pfizer", vax12_type=="az-az", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 death (Figure 2).
-
-<!--  `r filter(hazards_vax12_type, outcome=="covidemergency", treatment=="pfizer", vax12_type=="pfizer-pfizer", model==3, term_left==28) %>% pull(hr_display)` and `r filter(hazards_vax12_type, outcome=="emergency", treatment=="pfizer", vax12_type=="az-az", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 A&E attendances, -->
-<!-- `r filter(hazards_vax12_type, outcome=="covidcc", treatment=="pfizer", vax12_type=="pfizer-pfizer", model==3, term_left==28) %>% pull(hr_display)` and `r filter(hazards_vax12_type, outcome=="covidcc", treatment=="pfizer", vax12_type=="az-az", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 critical care admission, -->
+For the same period, the estimated effectiveness for BNT162b2 and ChAdOx1-S primary courses respectively were `r filter(hazards_vax12_type, outcome=="postest", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3, term_left==28) %>% pull(hr_display)` and `r filter(hazards_vax12_type, outcome=="postest", treatment=="pfizer", subgroup_level=="az-az", model==3, term_left==28) %>% pull(hr_display)` for a positive test, `r filter(hazards_vax12_type, outcome=="covidadmitted", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3, term_left==28) %>% pull(hr_display)` and `r filter(hazards_vax12_type, outcome=="covidadmitted", treatment=="pfizer", subgroup_level=="az-az", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 hospital admission, and `r filter(hazards_vax12_type, outcome=="coviddeath", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3, term_left==28) %>% pull(hr_display)` and `r filter(hazards_vax12_type, outcome=="coviddeath", treatment=="pfizer", subgroup_level=="az-az", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 death (Figure 2).
 
@@
     geom_hline(aes(yintercept=1), colour='grey')+
-    facet_grid(rows=vars(outcome_descr), cols=vars(subgroup_descr), switch="y")+
+    facet_grid(rows=vars(outcome_descr), cols=vars(subgroup_level_descr), switch="y")+
     scale_y_log10(
@@
 
-This observational study of `r label_number(1, big.mark=",")(flowchart[[nrow(flowchart), "n"]]*2)` fully vaccinated people living in England estimated high protective effects of booster vaccination against all studied outcomes.
-severe COVID-19 disease and death.
-Protection appeared to be highest in those receiving a X-X-X course, though .
-There were apparent waning effects after X weeks which is unlikely to explained by increased transmissibility of the Omicron variant given the low prevalence of Omicron at the end of the follow-up window.
+This observational study of fully vaccinated adults in England estimated high protection of BNT162b2 booster vaccination against all studied outcomes. 
+Protection appeared higher in those receiving a X-X primary course.
+There were apparent waning effects after X weeks which is unlikely to explained by increased transmissibility of the Omicron variant given Omicron was not yet dominant in the UK by the end of the follow-up window (ref).
 
-### Strengths and weaknesses
+### Limitations
 
-We used routinely-collected health records with comprehensive coverage of primary care, hospital admissions, COVID-19 testing, COVID-19 vaccination, and death registrations to study fully vaccinated English residents.
-...
-
-The methodological approach used is robust but inefficient; many boosted individuals are not retained after matching, and those who are retained are censored when their matched control is boosted.
-Millions of eligible participants may be required to obtain precise estimates, particularly when event rates are low, and so extending the approach to smaller subgroups may not be feasible.
-FOr instance, due to small numbers, we did not study booster effectiveness in those who had received the mRNA-1273 vaccine as their primary course or those who had received a heterologous primary course.
+We used routinely-collected health records with comprehensive coverage of primary care, hospital admissions, COVID-19 testing and vaccination, and death registrations to study over X million fully vaccinated English residents ...
 
@@
 
+Positive SARS-CoV-2 testing data likely underestimates the true incidence of infection.
+The mass availability of lateral flow tests in the UK during the study period, whose results are not routinely recorded in COVID-19 surveillance data, means many infections, including symptomatic infections, may be undocumented, despite encouragement to seek a confirmatory PCR test.
+Potential differences in testing behaviour between vaccinated and unvaccinated people also contributes to the unreliability of testing data as a means to assess effectiveness.
+
 Despite reasonable balance of baseline characteristics in the trial arms, and adjustment for a range of potential confounders, the possibility of unmeasured confounding remains.
@@
 
+The methodological approach used is robust but inefficient; many boosted individuals are not retained after matching, and those who are retained are censored when their matched control is boosted.
+Millions of eligible participants may be required to obtain precise estimates, particularly when event rates are low, and so extending the approach to smaller subgroups may not be feasible.
+For instance, due to small numbers, we did not study booster effectiveness in those who had received the mRNA-1273 vaccine as their primary course or those who had received a heterologous primary course, despite many thousands of such recipients in our records.
+
+Finally, mRNA-1273 boosting was rolled out later than for BNT162b2, and so we could not evaluate effectiveness of this vaccine due to insufficient follow-up.
+A analysis of mRNA-1273 boosting and comparative effectiveness between mRNA-1273 and BNT162b2 is planned.
+
 ### Findings in context
 
-[pfizer phase 3 booster trial, in pfizer first course]
-[ cov-boost trial]
+[pfizer phase 3 booster trial, in pfizer first course], [ cov-boost trial]
+
+A phase III trial assessing BNT162b2 booster efficacy in two-dose BNT162b2 recipients found.. (ref)
 

### 2022-02-14 create-project.R
remove dupe

@@ -498,7 +498,7 @@ actions_list <- splice(
     moderately_sensitive = lst(
       csv = "output/manuscript-objects/*.csv",
       png = "output/manuscript-objects/*.png",
-      csv = "output/manuscript-objects/*/*.csv",
+      csvsubgroup = "output/manuscript-objects/*/*.csv",
     )
   )
 

### 2022-02-14 analysis/report_seqtrialcox.R
redact events in KM table

@@ -257,14 +257,21 @@ dat_surv <-
   select(treated_descr, n_events, surv_obj_tidy) %>%
   unnest(surv_obj_tidy)
 
-data_surv_rounded <- dat_surv %>%
+data_surv_rounded <-
+  dat_surv %>%
   mutate(
     # Use ceiling not round. This is slightly biased upwards,
     # but means there's no disclosure risk at the boundaries (0 and 1) where masking would otherwise be threshold/2
     surv = ceiling_any(surv, 1/floor(max(n.risk, na.rm=TRUE)/(threshold+1))),
     surv.ll = ceiling_any(surv.ll, 1/floor(max(n.risk, na.rm=TRUE)/(threshold+1))),
     surv.ul = ceiling_any(surv.ul, 1/floor(max(n.risk, na.rm=TRUE)/(threshold+1))),
-  )
+    cml.event = ceiling_any(cumsum(replace_na(n.event, 0)), threshold+1),
+    cml.censor = ceiling_any(cumsum(replace_na(n.censor, 0)), threshold+1),
+    n.event = c(NA, diff(cml.event)),
+    n.censor = c(NA, diff(cml.censor)),
+    n.risk = ceiling_any(max(n.risk, na.rm=TRUE), threshold+1) - (cml.event + cml.censor)
+  ) %>%
+  select(treated_descr, time, leadtime, interval, surv, surv.ll, surv.ul, n.risk, n.event, n.censor)
 
 
 write_csv(data_surv_rounded, fs::path(output_dir, "report_km.csv"))

### 2022-02-15 manuscript/draft-manuscript.Rmd

@@
 output:
+  word_document: default
+  html_document:
+    keep_md: yes
+    self_contained: yes
   bookdown::html_document2:
@@
     toc: no
-  html_document:
-    keep_md: yes
-    self_contained: yes
-  word_document: default
   pdf_document:
@@
 # where are the outputs (ie the inputs for this manuscript!) saved?
-output_dir_os <- here("output", "manuscript-objects")
-#output_dir_os <- here("released_output", "manuscript-objects")
+#output_dir_os <- here("output", "manuscript-objects")
+output_dir_os <- here("released_output", "manuscript-objects")
 
@@
 matchcoverage <- read_csv(fs::path(output_dir_os, "matchcoverage.csv"))
-incidences <- read_csv(fs::path(output_dir_os, "incidence.csv")) %>% mutate(events=events_1+events_0)
+incidences <- read_csv(fs::path(output_dir_os, "none", "incidence.csv")) %>% mutate(events=events_1+events_0)
 
@@
     subgroup_descr = fct_inorder(subgroup_descr),
-    subgroup_level_descr = fct_inorder(subgroup_level_descr),
+    subgroup_level = fct_inorder(subgroup_level),
     treatment_descr = fct_inorder(treatment_descr),
@@
 
-William J Hulme^1^, Elizabeth J Williamson^2^, Amelia Green^1^, Elsie Horne^4,5^, Linda Nab^1^, Ruth Keogh^2^, Edward Parker^2^, Jonathan AC Sterne^4,5^, Miguel Hernán^6,7^, Ben Goldacre^1^ + OpenSAFELY collaborative.
+William J Hulme^1^, Elizabeth J Williamson^2^, Amelia Green^1^, Elsie Horne^3,4^, Helen Curtis^1^, Linda Nab^1^, Milan Wiedemann^1^, Ruth Keogh^2^, Edward Parker^2^, Venexia Walker^3,4^, Jonathan AC Sterne^3,4^, Miguel Hernán^6,7^, Ben Goldacre^1^ + OpenSAFELY collaborative.
 
-<!-- Krishnan Bhaskaran^2^, Helen I McDonald^2^, Christopher T Rentsch^2^, Anna Schultze^2^, John Tazare^2^, Helen J Curtis^1^, Alex J Walker^1^, Laurie Tomlinson^2^, Tom Palmer^4,5^,  Brian MacKenna^1^, Caroline E Morton^1^, Amir Mehrkar^1^, Louis Fisher^1^, Seb Bacon^1^, Dave Evans^1^, Peter Inglesby^1^, George Hickman^1^, Simon Davy^1^, Tom Ward ^1^, Richard Croker^1^, Rosalind M Eggo^2^, Angel YS Wong^2^, Rohini Mathur^2^, Kevin Wing^2^, Harriet Forbes^2^, Daniel Grint^2^, Ian J Douglas^2^, Stephen JW Evans^2^, Liam Smeeth^2^, Chris Bates^3^, Jonathan Cockburn^3^, John Parry^3^, Frank Hester^3^, Sam Harper^3^,  -->
+<!-- Krishnan Bhaskaran^2^, Helen I McDonald^2^, Christopher T Rentsch^2^, Anna Schultze^2^, John Tazare^2^, Helen J Curtis^1^, Alex J Walker^1^, Laurie Tomlinson^2^, Tom Palmer^3,4^,  Brian MacKenna^1^, Caroline E Morton^1^, Amir Mehrkar^1^, Louis Fisher^1^, Seb Bacon^1^, Dave Evans^1^, Peter Inglesby^1^, George Hickman^1^, Simon Davy^1^, Tom Ward ^1^, Richard Croker^1^, Rosalind M Eggo^2^, Angel YS Wong^2^, Rohini Mathur^2^, Kevin Wing^2^, Harriet Forbes^2^, Daniel Grint^2^, Ian J Douglas^2^, Stephen JW Evans^2^, Liam Smeeth^2^, Chris Bates^5^, Jonathan Cockburn^5^, John Parry^5^, Frank Hester^5^, Sam Harper^5^,  -->
 
@@
 
-3\. TPP, TPP House, 129 Low Lane, Horsforth, Leeds, LS18 5PX
+3\. Population Health Sciences, University of Bristol, Oakfield House, Oakfield Grove, Bristol, BS8 2BN, UK
 
-4\. Population Health Sciences, University of Bristol, Oakfield House, Oakfield Grove, Bristol, BS8 2BN, UK
+4\. NIHR Bristol, Biomedical Research Centre, Bristol, UK
 
-5\. NIHR Bristol, Biomedical Research Centre, Bristol, UK
+5\. TPP, TPP House, 129 Low Lane, Horsforth, Leeds, LS18 5PX
 
@@
 
-There were `r label_number(1, big.mark=",")(filter(matchsummary_treated, treatment=="pfizer") %>% pull(n))` people boosted with BNT162b2 and `r label_number(1, big.mark=",")(filter(matchsummary_treated, treatment=="moderna") %>% pull(n))` people boosted with mRNA-1273 who were successfully matched to an unboosted control in the study period.
+We matched a total of `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` BNT162b2 booster recipients to unboosted controls with only two doses.
+During the 28-41 days the estimated hazard ratio (95% CI) comparing a BNT162b2 booster dose to two doses only, was `r filter(hazards_none, outcome=="postest", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)` for a positive test, `r filter(hazards_none, outcome=="covidadmitted", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 hospital admission, and `r filter(hazards_none, outcome=="coviddeath", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 death.
 
@@
 
-In this cohort of fully vaccinated people in England, a booster dose was estimated to be highly effective at reducing.... , though there was evidence of waning effects. Heterologous boosting ...
+BNT162b2 boosting was estimated to be highly effective in protecting against COVID-19 related hospitalisation and death.
 
@@
 
-*Objectives* To estimate the effectiveness of the BNT162b2 and mRNA-1273 mRNA-1273 booster COVID-19 vaccines against infection and COVID-19 disease.
+<!-- *Objectives* To estimate the effectiveness of the BNT162b2 and mRNA-1273 mRNA-1273 booster COVID-19 vaccines against infection and COVID-19 disease. -->
 
-*Design* Matched sequential trial design, emulating a series of parallel two-arm RCTs.
+<!-- *Design* Matched sequential trial design, emulating a series of parallel two-arm RCTs. -->
 
-*Setting* Linked primary care, hospital, and COVID-19 surveillance records available within the OpenSAFELY-TPP research platform.
+<!-- *Setting* Linked primary care, hospital, and COVID-19 surveillance records available within the OpenSAFELY-TPP research platform. -->
 
-*Participants* `r label_number(1, big.mark=",")(filter(flowchart, crit=="c4") %>% pull(n))` people with a complete primary COVID-19 vaccine course, registered with a GP practice using the TPP SystmOne clinical information system in England.
+<!-- *Participants* `r label_number(1, big.mark=",")(filter(flowchart, crit=="c4") %>% pull(n))` people with a complete primary COVID-19 vaccine course, registered with a GP practice using the TPP SystmOne clinical information system in England.
+-->
 
-*Interventions* A booster vaccination with either BNT162b2 or mRNA-1273 administered as part of the national COVID-19 booster vaccine roll-out.
+<!-- *Interventions* A booster vaccination with either BNT162b2 or mRNA-1273 administered as part of the national COVID-19 booster vaccine roll-out. -->
 
-*Main outcome measures* Recorded SARS-CoV-2 infection, COVID-19-related accident and emergency attendance, unplanned COVID-19-related hospital admission, and COVID-19-related death.
+<!-- *Main outcome measures* Recorded SARS-CoV-2 infection, COVID-19-related accident and emergency attendance, unplanned COVID-19-related hospital admission, and COVID-19-related death. -->
 
-*Results*
+<!-- *Results* -->
 
-*Conclusions*
+<!-- *Conclusions* -->
 
@@
 In the vast majority of cases, BNT162b2 and mRNA-1273 vaccines were used, and administered without consideration of the type of primary course.
-BNT162b2 was used from 16 September 2022, and mRNA-1273 from X onwards.
-Consequently, available follow-up of BNT162b2 recipients is longer, and so in this early analysis we restrict to effectiveness of this vaccine.
+BNT162b2 was used from 16 September 2021, and mRNA-1273 from 29 October 2021 onwards.
+Consequently, available follow-up of BNT162b2 recipients is longer, and so we restrict the analysis to the effectiveness of BNT162b2boosting only.
 An analysis of mRNA-1273 booster effectiveness is planned when sufficient follow-up has accrued.
@@
 We used a sequential trials approach designed to emulate a sequence of daily two-arm parallel randomised controlled trials.
-For each trial, the treatment arm comprised people who received a BNT162b2 booster dose on that day, and the control arm comprised the same number of randomly selected unboosted people, matched on key characteristics described below.
+For each trial, the treatment arm comprised people who received a BNT162b2 booster dose on that day, and the control arm comprised the same number of randomly selected unboosted people matched on key characteristics, as described below.
+This closely follows the approach taken in previous vaccine effectiveness studies in Israeli registry data @dagan2021 @barda2021
 
@@
 
-## Outcomes
+## Exposures, outcomes, covariates
+
+<!--# COVID-19 A&E attendances were identified using HES emergency care records with U07.1 ("COVID-19, virus identified") or U07.2 ("COVID-19, virus not identified") ICD-10 diagnosis codes @emergenc.
+-->
 
@@
 
-<!-- COVID-19 A&E attendances were identified using HES emergency care records with U07.1 ("COVID-19, virus identified") or U07.2 ("COVID-19, virus not identified") ICD-10 diagnosis codes @emergenc.
--->
-
-COVID-19 related A&E attendance and critical/intensive care admission we also considered but there were doubts over data the reliability of event ascertainment in these cases.
-
-## Additional variables
+<!--# COVID-19 related A&E attendance and critical/intensive care admission we also considered but there were doubts over data the reliability of event ascertainment in these cases. -->
 
 Participant characteristics used describe trial participants and for confounder adjustment include: age, sex (male or female), English Index of Multiple Deprivation (IMD, grouped by quintiles), ethnicity (Black, Mixed, South Asian, White, Other, as per the UK census), NHS region (East of England, Midlands, London, North East and Yorkshire, North West, South East, South West), number of conditions in the clinically "at risk" classification, as per national prioritisation guidelines, the number of SARS-CoV-2 tests in the 6 months prior to the study start date (via SGSS), rurality (urban conurbation, urban city or town, rural town or village), evidence of prior SARS-CoV-2 infection (positive SARS-CoV-2 test, "probable" infection documented in primary care, or COVID-19 hospitalisation), learning disabilities, severe mental illness, undergoing an unplanned hospital admission, undergoing a planned hospital admission.
-All characteristics were ascertained as at the trial recruitment date (or study start date if assumed to be fixed from that date onwards).
+All characteristics were ascertained as at the trial recruitment date.
 
@@
 Follow-up was censored earlier than this at `r format(as.Date(study_dates$studyend_date), "%d %B %Y")`, death, or GP practice de-registration.
-Additionally for each matched pair, follow-up for both the boosted and unboosted person was censored if the control is boosted.
+Additionally for each matched pair, follow-up for both the boosted and unboosted person was censored if the control received a booster dose.
 
@@
 ]
-
-## Subgroup analyses
 
@@
 
-Participant characteristics at trial recruitment between trial arms were often well-balanced (Table 1), with some notable exceptions.
-Unboosted controls were typically...
+Participant characteristics at trial recruitment between trial arms (Table 1) were typically well-balanced (necessarily so on matched characteristics), though with some exceptions, for instance unboosted controls had higher levels of deprivation, higher rates of learning disabilities and severe mental illness, and lower SARS-CoV-2 testing frequency prior to the study.
 These unbalanced characteristics were adjusted for in the Cox regression models.
@@
     fup_period == "All",
-    outcome %in% c("postest", "covidadmitted", "coviddeath")
+    outcome %in% c("postest", "covidadmitted", "coviddeath"),
+    treatment == "pfizer"
   ) %>%
   left_join(
-    matchsummary %>% select(treatment, eligible_1=n_eligible, recruit_prop_1 = prop_matched_eligible),
+    matchsummary %>% select(treatment, n_eligible, prop_matched_eligible),
     by=c("treatment")
+  ) %>%
+  mutate(
+    treatment_decr = paste0(
+      treatment_descr, " - ",
+      label_number(1, big.mark=",")(n_eligible), "eligible, ",
+      label_number(1, big.mark=",")(n_1), "matched (",
+      label_percent(0.1)(prop_matched_eligible), ")."
+    )
   ) %>%
@@
     outcome_descr,
-    eligible_1,
-    n_1,
-    recruit_prop_1,
     q_1,
@@
     outcome_descr = "Outcome",
-
-    eligible_1 = "Eligible",
-    n_1 = "Recruited",
-    recruit_prop_1 = "(%)",
 
@@
     columns = c("rate_0", "rate_1", "rr"),
-    decimals = 2
-  ) %>%
-  fmt_integer(
-    columns = c("n_1", "eligible_1"),
-    sep_mark = ","
-  ) %>%
-  fmt_number(
-    columns = "recruit_prop_1",
-    scale_by = 100,
-    decimals = 1,
-
+    decimals = 3
   ) %>%
@@
 
-Trial participants contributed `r filter(incidences, treatment=="pfizer", outcome=="postest", fup_period=="all") %>% pull(events)` positive SARS-CoV-2 tests, `r filter(incidences, treatment=="pfizer", outcome=="covidadmitted", fup_period=="all") %>% pull(events)` COVID-19 hospital admissions, and `r filter(incidences, treatment=="pfizer", outcome=="coviddeath", fup_period=="all") %>% pull(events)` COVID-19 deaths (Table 2).
+Trial participants contributed `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="postest", fup_period=="All") %>% pull(events))` positive SARS-CoV-2 tests, `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="covidadmitted", fup_period=="All") %>% pull(events))` COVID-19 hospital admissions, and `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="coviddeath", fup_period=="All") %>% pull(events))` COVID-19 deaths (Table 2).
 
@@
 
-During the fifth and sixth week after recruitment (days 28-41) the estimated hazard ratio comparing a BNT162b2 booster dose to no dose was `r filter(hazards_none, outcome=="postest", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)` for a positive test, `r filter(hazards_none, outcome=="covidadmitted", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 hospital admission, and `r filter(hazards_none, outcome=="coviddeath", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 death (Figure 2).
+During the 28-41 days the estimated hazard ratio (95% CI) comparing a BNT162b2 booster dose to two doses only, was `r filter(hazards_none, outcome=="postest", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)` for a positive test, `r filter(hazards_none, outcome=="covidadmitted", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 hospital admission, and `r filter(hazards_none, outcome=="coviddeath", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 death (Figure 2).
 
@@
 
-For the same period, the estimated effectiveness for BNT162b2 and ChAdOx1-S primary courses respectively were `r filter(hazards_vax12_type, outcome=="postest", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3, term_left==28) %>% pull(hr_display)` and `r filter(hazards_vax12_type, outcome=="postest", treatment=="pfizer", subgroup_level=="az-az", model==3, term_left==28) %>% pull(hr_display)` for a positive test, `r filter(hazards_vax12_type, outcome=="covidadmitted", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3, term_left==28) %>% pull(hr_display)` and `r filter(hazards_vax12_type, outcome=="covidadmitted", treatment=="pfizer", subgroup_level=="az-az", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 hospital admission, and `r filter(hazards_vax12_type, outcome=="coviddeath", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3, term_left==28) %>% pull(hr_display)` and `r filter(hazards_vax12_type, outcome=="coviddeath", treatment=="pfizer", subgroup_level=="az-az", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 death (Figure 2).
+For the same period, the estimated hazard ratio (95% CI) for BNT162b2 and ChAdOx1-S primary courses respectively were `r filter(hazards_vax12_type, outcome=="postest", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3, term_left==28) %>% pull(hr_display)` and `r filter(hazards_vax12_type, outcome=="postest", treatment=="pfizer", subgroup_level=="az-az", model==3, term_left==28) %>% pull(hr_display)` for a positive test, `r filter(hazards_vax12_type, outcome=="covidadmitted", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3, term_left==28) %>% pull(hr_display)` and `r filter(hazards_vax12_type, outcome=="covidadmitted", treatment=="pfizer", subgroup_level=="az-az", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 hospital admission, and `r filter(hazards_vax12_type, outcome=="coviddeath", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3, term_left==28) %>% pull(hr_display)` and `r filter(hazards_vax12_type, outcome=="coviddeath", treatment=="pfizer", subgroup_level=="az-az", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 death (Figure 2).
 
@@
     hazards3_vax12_type %>%
-    filter(outcome %in% c("postest", "covidadmitted", "coviddeath")) %>%
+    filter(
+      outcome %in% outcomes,
+      treatment %in% c("pfizer")
+    ) %>%
     ggplot()+
@@
     geom_hline(aes(yintercept=1), colour='grey')+
-    facet_grid(rows=vars(outcome_descr), cols=vars(subgroup_level_descr), switch="y")+
+    facet_grid(rows=vars(outcome_descr), cols=vars(subgroup_level), switch="y")+
     scale_y_log10(
@@
 
-This observational study of fully vaccinated adults in England estimated high protection of BNT162b2 booster vaccination against all studied outcomes. 
-Protection appeared higher in those receiving a X-X primary course.
-There were apparent waning effects after X weeks which is unlikely to explained by increased transmissibility of the Omicron variant given Omicron was not yet dominant in the UK by the end of the follow-up window (ref).
+This observational study of over 8 million boosted adults in England and their matched controls estimated high protection of BNT162b2 booster vaccination against all studied outcomes using routinely-collected health records with comprehensive coverage of primary care, hospital admissions, COVID-19 testing and vaccination, and death registrations.
+Protection was similar between 2x BNT162b2 primary course recipients and 2x ChAdOx1-S recipients.
+There were apparent waning effects around 6 weeks after boosting, which is unlikely to explained by increased transmissibility of the Omicron variant given Omicron was not yet dominant in the UK by the end of the follow-up window @COVID19OmicronDaily, though changing testing behaviours over time between boosted and unboosted people may influence estimates.
 
 ### Limitations
-
-We used routinely-collected health records with comprehensive coverage of primary care, hospital admissions, COVID-19 testing and vaccination, and death registrations to study over X million fully vaccinated English residents ...
-
-This study considers effectiveness in a population where Alpha was the predominant circulating variant (ref gov.uk reports).
-Re-running the analysis against a backdrop of the now globally dominant Omicron variant will be undertaken where valid concurrent comparisons between boosted and unboosted people can be made.
 
@@
 In particular, the apparent protective effect immediately after boosting is biologically implausible, and will be explained in part by the occurrence of COVID-19 symptoms at baseline not recorded in the health record.
+
+This study considers effectiveness in a population where Delta was the predominant circulating variant (ref gov.uk reports).
+Re-running the analysis against a backdrop of the now globally dominant Omicron variant will be undertaken where valid concurrent comparisons between boosted and unboosted people can be made.
 
@@
 
-[pfizer phase 3 booster trial, in pfizer first course], [ cov-boost trial]
+<!-- [ cov-boost trial] -->
 
-A phase III trial assessing BNT162b2 booster efficacy in two-dose BNT162b2 recipients found.. (ref)
+A phase III trial assessing BNT162b2 booster efficacy in two-dose BNT162b2 recipients without prior infection reported a relative efficacy of 95.6% (95% CI 89.3-98.6) against Delta infection @PfizerBioNTechAnnounce2021.
 
-A number of studies have estimated COVID-19 vaccine booster effectiveness in observational data.
+Recent studies in the UK have provided observational evidence for increased protection from the booster vaccine against symptomatic COVID-19 infection in comparison to second dose recipients using a test negative design.
+Andrews et al.
+reported estimates of relative vaccine effectiveness in the 14 days after a BNT162b2 booster dose in those aged over 50 years of 87.4% (95%CI 84.9-89.4) where the primary course was ChAdOx1-S and 84.4% (95%CI 82.8-85.8) for BNT162b2 @andrews2021.
+Sheikh et al.
+reported comparable estimates of relative vaccine effectiveness for S gene positive symptomatic infection (83% (95%CI 81-84) 6-49 years, 88% (95%CI 86-89) ≥50 years) in the 14 days after a BNT162b2 or mRNA-1273 booster dose, but lower estimates for S gene negative symptomatic infection (56% (95%CI 51-60) 6-49 years, 57% (95%CI 52-62) ≥50 years).
+
+<!-- @https://www.research.ed.ac.uk/en/publications/severity-of-omicron-variant-of-concern-and-vaccine-effectiveness- -->
+
+<!-- Elsewhere, studies in Israel found .. -->
+
+<!-- @1050236/technical-briefing-34-14-january-2022.pdf -->
 
@@
 
-This study found high protection against severe COVID-19 disease
-
-<!-- \newpage -->
-
-<!-- # Summary -->
-
-<!-- ### What is already known -->
-
-<!-- ### What this study adds -->
+This study found high protection for BNT162b2 boosting against COVID-19 hospitalisation and death in the Delta era.
 
@@
 
-## Guarantor
-
 BG is guarantor.

### 2022-02-15 manuscript/draft-supplement.Rmd

@@
 cohortsummary <- read_csv(fs::path(output_dir_os, "cohortsummary.csv"))
-matchtable1 <- read_csv(fs::path(output_dir_os, "matchtable1.csv"))
-matchtable1by <- read_csv(fs::path(output_dir_os, "matchtable1by.csv"))
 matchsummary <- read_csv(fs::path(output_dir_os, "matchsummary.csv"))
@@
 matchsummary_control <- read_csv(fs::path(output_dir_os, "matchsummary_treated.csv")) %>% filter(treated == 0L)
-incidences <- read_csv(fs::path(output_dir_os, "ir.csv"))
+matchcoverage <- read_csv(fs::path(output_dir_os, "matchcoverage.csv"))
+incidences <- read_csv(fs::path(output_dir_os, "incidence.csv")) %>% mutate(events=events_1+events_0)
 
@@
 
+```{r, effects, echo=FALSE, message=FALSE, warning=FALSE}
+
+  specify_decimal <- function(x, k, trim=FALSE) {
+
+    fmtd <- format(round(x, k), nsmall = k)
+    if (trim) {fmtd <- trimws(fmtd)}
+    return(fmtd)
+  }
+  #
+  # print_est1bracket <- function(x, b, round=1){
+  #   paste0(specify_decimal(x, round), " (", specify_decimal(b, round), ")")
+  # }
+  #
+  print_est2bracket <- function(x, b1, b2, round=1){
+    paste0(specify_decimal(x, round), " (", specify_decimal(b1, round), ", ", specify_decimal(b2, round), ")")
+  }
+
+  outcomes <- 
+    c(
+      "postest",
+      #"covidemergency",
+      "covidadmitted",
+      #"covidcc",
+      "coviddeath"
+    ) %>% 
+    set_names(., .)
+
+# 
+# cmlinc <- read_csv(fs::path(output_dir_os, "cmlinc.csv")) %>%
+#   mutate(
+#     treatment_descr = fct_inorder(treatment_descr),
+#     outcome_descr = fct_inorder(outcome_descr),
+#     model_name = fct_inorder(str_wrap(model_name, 15)),
+#     inc_CI_to = paste0(label_number(0.1, scale=1000)(1-survival.ul), " to ", label_number(0.1, scale=1000)(1-survival.ll))
+#   )
+
+# riskdiff <- read_csv(fs::path(output_dir_os, "riskdiff.csv")) %>%
+#   mutate(
+#     treatment_descr = fct_inorder(treatment_descr),
+#     outcome_descr = fct_inorder(outcome_descr),
+#     model_name = fct_inorder(str_wrap(model_name, 15)),
+#     diff_CI_to = paste0(label_number(0.01, scale=1000)(diff.ll), " to ", label_number(0.01, scale=1000)(diff.ul))
+#   )
+
+hazards_none <- read_csv(fs::path(output_dir_os, "none", "effects.csv")) %>%
+  mutate(
+    treatment_descr = fct_inorder(treatment_descr),
+    outcome_descr = fct_inorder(outcome_descr),
+    #model_name = fct_inorder(str_wrap(model_name, 15)),
+    hr_display = paste0(number_format(accuracy=0.01)(hr), " (", number_format(accuracy=0.01)(hr.ll), "-", number_format(accuracy=0.01)(hr.ul), ")")
+  )
+
+hazards_vax12_type <- read_csv(fs::path(output_dir_os, "vax12_type", "effects.csv")) %>%
+  mutate(
+    subgroup_descr = fct_inorder(subgroup_descr),
+    subgroup_level_descr = fct_inorder(subgroup_level_descr),
+    treatment_descr = fct_inorder(treatment_descr),
+    outcome_descr = fct_inorder(outcome_descr),
+    #model_name = fct_inorder(str_wrap(model_name, 15)),
+    hr_display = paste0(number_format(accuracy=0.01)(hr), " (", number_format(accuracy=0.01)(hr.ll), "-", number_format(accuracy=0.01)(hr.ul), ")")
+  )
+
+```
 \newpage
@@
 \newpage
-### Table S2a: Characteristics at recruitment to Pfizer trial, by treatment status
 
-```{r table1_pfizer, echo=FALSE, warning=FALSE, message=FALSE, out.extra=''}
+```{r plot_effects, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}
 
-  table1by <- 
-    matchtable1by %>% 
-    filter(treatment=="pfizer")
+formatpercent100 <- function(x,accuracy){
+  formatx <- scales::label_percent(accuracy)(x)
 
-  table1 <- 
-    matchtable1 %>%
-    filter(treatment=="pfizer") %>%
-    group_by(variable) %>%
-    mutate(
-      label = if_else(row_number()==1, label, str_c("&nbsp;&nbsp;&nbsp;", label)),
-    ) %>%
-    mutate(
-      across(
-        starts_with("stat_"),
-        ~if_else(is.na(.x) & row_number()==1, "", .x)
+  if_else(
+    formatx==scales::label_percent(accuracy)(1),
+    paste0(">",scales::label_percent(1)((100-accuracy)/100)),
+    formatx
   )
-    ) %>%
-    ungroup() %>%
-    mutate(
-      order = cumsum(lag(variable, 1, first(variable))!=variable)
-    ) %>%
-    select(
-      variable, var_label, label, starts_with("stat_"), order
-    ) %>%
-    rename(!!!set_names(table1by$by_col, table1by$by_chr))
+}
 
-  table1_gt <-
-    table1 %>%
-    gt() %>%
-    fmt_markdown(columns = "label") %>%
-    tab_style(
-      style = list(
-        cell_fill(color = "gray96")
-      ),
-      locations = cells_body(
-        rows = order%%2 == 0
+# plot_cmlinc <- cmlinc3 %>%
+#     ggplot()+
+#     geom_step(aes(x=tstop, y=(1-survival)*1000, group=vax1_az_descr, colour=vax1_az_descr))+
+#     geom_rect(aes(xmin=tstop, xmax=lead_tstop, ymin=(1-survival.ll)*1000, ymax=(1-survival.ul)*1000, group=vax1_az_descr, fill=vax1_az_descr), alpha=0.1)+
+#     geom_hline(aes(yintercept=0), colour='black')+
+#     facet_wrap(facets=vars(outcome_descr), nrow=1, scales="free_y")+
+#     scale_x_continuous(
+#       breaks = seq(0,7*52,by=28),
+#       expand = expansion(0)
+#     )+
+#     scale_y_continuous(
+#       expand = expansion(0),
+#       #limits = c(0L,NA)
+#     )+
+#     scale_colour_brewer(type="qual", palette="Set1", na.value="grey")+
+#     scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey")+
+#     labs(
+#       x=NULL,
+#       y="Marginalised\ncumul. incidence\nper 1000",
+#       colour=NULL
+#     )+
+#     theme_minimal(base_size = 9)+
+#     theme(
+#       legend.position=c(0.05,0.1),
+#       legend.justification = c(0,0),
+#       legend.direction = "vertical",
+#       axis.text.x.top=element_text(hjust=0),
+# 
+#       panel.border = element_blank(),
+#       axis.line.y = element_line(colour = "black"),
+#       axis.line.x.top = element_line(colour = "black"),
+# 
+#       panel.grid.minor.x = element_blank(),
+#       panel.grid.minor.y = element_blank(),
+#       strip.background = element_blank(),
+# 
+# 
+#       panel.spacing = unit(0.8, "lines"),
+# 
+#     )+
+#     NULL
+#   
+#   ## difference in cumulative incidence
+#   
+#   plot_riskdiff <-
+#     riskdiff3 %>%
+#     ggplot()+
+#     geom_hline(aes(yintercept=0), colour='black')+
+#     geom_step(aes(x=tstop, y=diff*1000), colour='orchid4')+
+#     geom_rect(aes(xmin=tstop, xmax=lead_tstop, ymin=diff.ll*1000, ymax=diff.ul*1000), alpha=0.1, colour="transparent", fill='orchid4')+
+#     facet_wrap(facets=vars(outcome_descr), nrow=1, scales="free_y")+
+#     scale_x_continuous(
+#       limits = c(0, NA),
+#       breaks = seq(0,7*52,by=28),
+#       expand = expansion(0)
+#     )+
+#     scale_y_continuous(
+#       expand = expansion(0),
+#       sec.axis = dup_axis(name="Higher incidence after \n <- ChAdOx1 / BNT162b2 ->", breaks = NULL)
+#     )+
+#     coord_cartesian(ylim= c(-3, 3))+
+#     scale_colour_brewer(type="qual", palette="Set1", na.value="grey")+
+#     scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey")+
+#     labs(
+#       x=NULL,
+#       y="Difference\nin cumul. incidence\nper 1000",
+#       colour=NULL
+#     )+
+#     theme_minimal(base_size = 9)+
+#     theme(
+#       axis.line.y.left = element_line(colour = "black"),
+# 
+#       panel.grid.minor.x = element_blank(),
+#       panel.grid.minor.y = element_blank(),
+#       strip.background = element_blank(),
+#       strip.text = element_blank(),
+# 
+#       panel.spacing = unit(0.8, "lines"),
+# 
+#       plot.title = element_text(hjust = 0),
+#       plot.title.position = "plot",
+#       plot.caption.position = "plot",
+#       plot.caption = element_text(hjust = 0, face= "italic")
+# 
+#      ) +
+#    NULL
+# 
+#
+
+  ## hazard ratio
+
+  y_breaks <- c(0.01, 0.02, 0.05, 0.1, 0.2, 0.5, 1, 2)
+
+  plot_hazards <-
+    hazards_vax12_type %>%
+    filter(outcome %in% c("postest", "covidadmitted", "coviddeath")) %>%
+    ggplot()+
+    geom_point(aes(y=hr, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+
+    geom_linerange(aes(ymin=hr.ll, ymax=hr.ul, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+
+    geom_hline(aes(yintercept=1), colour='grey')+
+    facet_grid(rows=vars(outcome_descr), cols=vars(subgroup_level_descr), switch="y")+
+    scale_y_log10(
+      breaks=y_breaks,
+      limits = c(0.01, 1.5),
+      oob = scales::oob_keep,
+      sec.axis = sec_axis(
+        ~(1-.),
+        name="Effectiveness",
+        breaks = 1-(y_breaks),
+        labels = function(x){formatpercent100(x, 1)}
       )
-    ) %>%
-    tab_style(
-      style = list(
-        cell_text(weight = "bold")
-      ),
-      locations = cells_column_labels(everything())
-    ) %>%
-    cols_label(
-       label=md("Characteristic")
-    ) %>%
-    cols_hide(c("variable", "var_label", "order"))
+    )+
+    scale_x_continuous(breaks=postbaselinecuts, limits=c(min(postbaselinecuts), max(postbaselinecuts)+1), expand = c(0, 0))+
+    scale_colour_brewer(type="qual", palette="Set2", guide=guide_legend(ncol=1))+
+    labs(
+      y="Hazard ratio",
+      x="Days since booster dose",
+      colour=NULL
+     ) +
+    theme_bw()+
+    theme(
+      panel.border = element_blank(),
+      axis.line.y = element_line(colour = "black"),
+      axis.line.y.right = element_blank(),
+      panel.grid.minor.x = element_blank(),
+      panel.grid.minor.y = element_blank(),
+      strip.background = element_blank(),
+      strip.placement = "outside",
+      strip.text.y.left = element_text(angle = 0),
+      plot.title = element_text(hjust = 0),
+      plot.title.position = "plot",
+      plot.caption.position = "plot",
+      plot.caption = element_text(hjust = 0, face= "italic"),
+      legend.position = "bottom"
+     ) +
+    NULL
 
-  table1_gt
+  plot_results <- 
+    patchwork::wrap_plots(
+      #plot_cmlinc, 
+      #plot_riskdiff,
+      plot_hazards, 
+      ncol=1
+    )
 
-```
-
-### Table S2b: Characteristics at recruitment to Moderna trial, by treatment status
-
-```{r table1_moderna, echo=FALSE, warning=FALSE, message=FALSE, out.extra=''}
-
-  table1by <- 
-    matchtable1by %>% 
-    filter(treatment=="moderna")
-
-  table1 <- 
-    matchtable1 %>%
-    filter(treatment=="moderna") %>%
-    group_by(variable) %>%
-    mutate(
-      label = if_else(row_number()==1, label, str_c("&nbsp;&nbsp;&nbsp;", label)),
-    ) %>%
-    mutate(
-      across(
-        starts_with("stat_"),
-        ~if_else(is.na(.x) & row_number()==1, "", .x)
-      )
-    ) %>%
-    ungroup() %>%
-    mutate(
-      order = cumsum(lag(variable, 1, first(variable))!=variable)
-    ) %>%
-    select(
-      variable, var_label, label, starts_with("stat_"), order
-    ) %>%
-    rename(!!!set_names(table1by$by_col, table1by$by_chr))
-
-  table1_gt <-
-    table1 %>%
-    gt() %>%
-    fmt_markdown(columns = "label") %>%
-    tab_style(
-      style = list(
-        cell_fill(color = "gray96")
-      ),
-      locations = cells_body(
-        rows = order%%2 == 0
-      )
-    ) %>%
-    tab_style(
-      style = list(
-        cell_text(weight = "bold")
-      ),
-      locations = cells_column_labels(everything())
-    ) %>%
-    cols_label(
-       label=md("Characteristic")
-    ) %>%
-    cols_hide(c("variable", "var_label", "order"))
-
-  table1_gt
+  plot_results
 

### 2022-02-15 analysis/study_definition.py
only consider ordinary admissions for planned admissions (no day cases)

@@ -606,7 +606,7 @@ study = StudyDefinition(
     # see https://github.com/opensafely-core/cohort-extractor/pull/497 for codes
     # see https://docs.opensafely.org/study-def-variables/#sus for more info
     with_admission_method=["11", "12", "13", "81"],
-    with_patient_classification = ["1", "2"], # ordinary and day-case admissions only 
+    with_patient_classification = ["1"], # ordinary admissions only 
     date_format="YYYY-MM-DD",
     find_last_match_in_period=True,
   ),
@@ -616,7 +616,7 @@ study = StudyDefinition(
     # see https://github.com/opensafely-core/cohort-extractor/pull/497 for codes
     # see https://docs.opensafely.org/study-def-variables/#sus for more info
     with_admission_method=["11", "12", "13", "81"],
-    with_patient_classification = ["1", "2"], # ordinary and day-case admissions only
+    with_patient_classification = ["1"], # ordinary admissions only
     date_format="YYYY-MM-DD",
     find_last_match_in_period=True
   ), 
@@ -628,7 +628,7 @@ study = StudyDefinition(
     # see https://github.com/opensafely-core/cohort-extractor/pull/497 for codes
     # see https://docs.opensafely.org/study-def-variables/#sus for more info
     with_admission_method=["11", "12", "13", "81"],
-    with_patient_classification = ["1", "2"], # ordinary and day-case admissions only
+    with_patient_classification = ["1"], # ordinary and day-case admissions only
   ),
   
   

### 2022-02-15 analysis/design.R
change matching variables
* replace region with msoa
* add number of comorbidities
* replace exact week of second dose with 7-caliper day of second dose
* add 3-caliper age
* add cev

@@ -84,7 +84,28 @@ postbaselinecuts <- c(0,7,seq(14,7*10, 14))
 write_rds(postbaselinecuts, here("lib", "design", "postbaselinecuts.rds"))
 
 # what matching variables
-matching_variables <- c("jcvi_group", "vax12_type", "region", "vax2_week", "prior_covid_infection", "immunosuppressed", "status_hospplanned")
+exact_variables <- c(
+
+  "jcvi_group",
+  "vax12_type",
+  #"vax2_week",
+  "msoa",
+  "sex",
+  "cev",
+  "multimorb",
+  "prior_covid_infection",
+  "immunosuppressed",
+  "status_hospplanned"
+)
+write_rds(exact_variables, here("lib", "design", "exact_variables.rds"))
+
+caliper_variables <- c(
+  age = 3,
+  vax2_day = 7
+)
+write_rds(caliper_variables, here("lib", "design", "caliper_variables.rds"))
+
+matching_variables <- c(exact_variables, names(caliper_variables))
 write_rds(matching_variables, here("lib", "design", "matching_variables.rds"))
 
 # cut-off for rolling 7 day average, that determines recruitment period

### 2022-02-15 analysis/match_seqtrialcox.R

@@ -44,6 +44,8 @@ source(here("lib", "functions", "redaction.R"))
 
 postbaselinecuts <- read_rds(here("lib", "design", "postbaselinecuts.rds"))
 matching_variables <- read_rds(here("lib", "design", "matching_variables.rds"))
+exact_variables <- read_rds(here("lib", "design", "exact_variables.rds"))
+caliper_variables <- read_rds(here("lib", "design", "caliper_variables.rds"))
 
 if(Sys.getenv("OPENSAFELY_BACKEND") %in% c("", "expectations")){
   recruitment_period_cutoff <- 0.1
@@ -96,7 +98,8 @@ events <- read_rds(here("lib", "design", "event-variables.rds"))
 # Prepare data ----
 
 ## one pow per patient ----
-data_cohort <- read_rds(here("output", "data", "data_cohort.rds"))
+data_cohort <- read_rds(here("output", "data", "data_cohort.rds")) %>%
+  filter(!is.na(msoa))
 
 logoutput_datasize(data_cohort)
 
@@ -326,13 +329,29 @@ local({
       safely_matchit(
         formula = matching_formula,
         data = matching_candidates_i,
-        method = "exact",
+        method = "nearest",
         replace = FALSE,
         estimand = "ATE",
-        #m.order = "data", # data is sorted on (effectively random) patient ID
+        exact = exact_variables,
+        caliper = caliper_variables, std.caliper=FALSE,
+        m.order = "data", # data is sorted on (effectively random) patient ID
         #verbose = TRUE,
-        #ratio = 1L # irritatingly you can't set this for "exact" method, so have to filter later
+        ratio = 1L # irritatingly you can't set this for "exact" method, so have to filter later
       )[[1]]
+
+    matching_i <-
+      matchit(
+        formula = matching_formula,
+        data = matching_candidates_i,
+        method = "nearest",
+        replace = FALSE,
+        estimand = "ATT",
+        exact = exact_variables,
+        caliper = caliper_variables, std.caliper=FALSE,
+        m.order = "data", # data is sorted on (effectively random) patient ID
+        #verbose = TRUE,
+        ratio = 1L # irritatingly you can't set this for "exact" method, so have to filter later
+      )
 
     if(is.null(matching_i)) {
       message("Terminating trial sequence at trial ", trial, " - No exact matches found.")

### 2022-02-15 analysis/data_selection.R
swap region for msoa

@@ -42,6 +42,7 @@ data_criteria <- data_processed %>%
     has_imd = !is.na(imd),
     has_ethnicity = !is.na(ethnicity_combined),
     has_region = !is.na(region),
+    has_msoa = is.na(msoa),
     isnot_hscworker = !hscworker,
     isnot_carehomeresident = !care_home_combined,
     isnot_endoflife = !endoflife,
@@ -70,7 +71,7 @@ data_criteria <- data_processed %>%
       vax1_afterfirstvaxdate &
       vax2_beforelastvaxdate &
       vax3_afterstudystartdate &
-      has_age & has_sex & has_imd & has_ethnicity & has_region &
+      has_age & has_sex & has_imd & has_ethnicity & has_msoa &
       has_vaxgap12 & has_vaxgap23 & has_knownvax1 & has_knownvax2 & vax12_homologous &
       isnot_hscworker &
       isnot_carehomeresident & isnot_endoflife &
@@ -91,7 +92,7 @@ data_flowchart <- data_criteria %>%
   transmute(
     c0 = vax1_afterfirstvaxdate & vax2_beforelastvaxdate & vax3_afterstudystartdate,# & jcvi_group_6orhigher,
     #c1_1yearfup = c0_all & (has_follow_up_previous_year),
-    c1 = c0 & (has_age & has_sex & has_imd & has_ethnicity & has_region),
+    c1 = c0 & (has_age & has_sex & has_imd & has_ethnicity & has_msoa),
     c2 = c1 & (has_vaxgap12 & has_vaxgap23 & has_knownvax1 & has_knownvax2 & vax12_homologous),
     c3 = c2 & (isnot_hscworker ),
     c4 = c3 & (isnot_carehomeresident & isnot_endoflife & isnot_housebound),

### 2022-02-15 analysis/match_seqtrialcox.R
no need to filter to 1:1 ratio as already done

@@ -379,8 +379,7 @@ local({
       ) %>%
       group_by(subclass, treated) %>%
       filter(
-        row_number() <= pmin(n_treated, n_control), # 1:1 matching only
-        !is.na(subclass) # remove unmatchd people. equivalent to weight != 0
+        !is.na(subclass) # remove unmatched people. equivalent to weight != 0
       ) %>%
       group_by(treated) %>%
       mutate(

### 2022-02-15 analysis/match_seqtrialcox.R
tidy up matchit output

@@ -363,34 +363,19 @@ local({
     data_matched_i <-
       as.data.frame(matching_i$X) %>%
       add_column(
-        subclass = matching_i$subclass,
+        match_id = matching_i$subclass,
         treated = matching_i$treat,
         patient_id = matching_candidates_i$patient_id,
         weight = matching_i$weights,
         trial_day = trial,
       ) %>%
 
-      # filter treated and controls to ensure exact 1:1 matching
-      arrange(subclass, desc(treated), patient_id) %>%
-      group_by(subclass) %>%
-      mutate(
-        n_treated = sum(treated),
-        n_control = sum(1-treated)
-      ) %>%
-      group_by(subclass, treated) %>%
-      filter(
-        !is.na(subclass) # remove unmatched people. equivalent to weight != 0
-      ) %>%
-      group_by(treated) %>%
-      mutate(
-        match_id = row_number()
-      ) %>%
-      arrange(subclass, match_id, desc(treated)) %>%
+      filter(!is.na(match_id)) %>% # remove unmatched people. equivalent to weight != 0
+      arrange(match_id, desc(treated)) %>%
       left_join(
         matching_candidates_i %>% select(patient_id, starts_with("tte_")),
         by = "patient_id"
       ) %>%
-      select(-n_treated, -n_control) %>%
       group_by(match_id) %>%
       mutate(
         tte_recruitment = trial_time,
@@ -426,7 +411,6 @@ local({
 
   data_matched <<-
     data_matched %>%
-    select(-subclass) %>%
     mutate(
       treated_patient_id = paste0(treated, "_", patient_id),
       fup = pmin(tte_stop - tte_recruitment, last(postbaselinecuts))

### 2022-02-15 analysis/match_seqtrialcox.R
Update match_seqtrialcox.R

@@ -98,8 +98,7 @@ events <- read_rds(here("lib", "design", "event-variables.rds"))
 # Prepare data ----
 
 ## one pow per patient ----
-data_cohort <- read_rds(here("output", "data", "data_cohort.rds")) %>%
-  filter(!is.na(msoa))
+data_cohort <- read_rds(here("output", "data", "data_cohort.rds"))
 
 logoutput_datasize(data_cohort)
 

### 2022-02-15 analysis/data_process.R
bugs; swap to mahalanobid distance (to stop propensity calculation)

@@ -56,6 +56,7 @@ if(Sys.getenv("OPENSAFELY_BACKEND") %in% c("", "expectations")){
       admitted_covid_ccdays_2 = as.numeric(as.character(admitted_covid_ccdays_2)),
       admitted_covid_ccdays_3 = as.numeric(as.character(admitted_covid_ccdays_3)),
       admitted_covid_ccdays_4 = as.numeric(as.character(admitted_covid_ccdays_4)),
+      msoa = sample(factor(c("1", "2")), size=n(), replace=TRUE) # override msoa so matching success more likely
     )
 
 

### 2022-02-15 analysis/data_selection.R

@@ -42,7 +42,7 @@ data_criteria <- data_processed %>%
     has_imd = !is.na(imd),
     has_ethnicity = !is.na(ethnicity_combined),
     has_region = !is.na(region),
-    has_msoa = is.na(msoa),
+    has_msoa = !is.na(msoa),
     isnot_hscworker = !hscworker,
     isnot_carehomeresident = !care_home_combined,
     isnot_endoflife = !endoflife,

### 2022-02-15 analysis/match_seqtrialcox.R

@@ -328,21 +328,7 @@ local({
       safely_matchit(
         formula = matching_formula,
         data = matching_candidates_i,
-        method = "nearest",
-        replace = FALSE,
-        estimand = "ATE",
-        exact = exact_variables,
-        caliper = caliper_variables, std.caliper=FALSE,
-        m.order = "data", # data is sorted on (effectively random) patient ID
-        #verbose = TRUE,
-        ratio = 1L # irritatingly you can't set this for "exact" method, so have to filter later
-      )[[1]]
-
-    matching_i <-
-      matchit(
-        formula = matching_formula,
-        data = matching_candidates_i,
-        method = "nearest",
+        method = "nearest", distance = "mahalanobis", # these two options don't really do anything because we only want exact + caliper matching
         replace = FALSE,
         estimand = "ATT",
         exact = exact_variables,
@@ -350,7 +336,8 @@ local({
         m.order = "data", # data is sorted on (effectively random) patient ID
         #verbose = TRUE,
         ratio = 1L # irritatingly you can't set this for "exact" method, so have to filter later
-      )
+      )[[1]]
+
 
     if(is.null(matching_i)) {
       message("Terminating trial sequence at trial ", trial, " - No exact matches found.")
@@ -399,7 +386,7 @@ local({
     data_matched <- bind_rows(data_matched, data_matched_i)
   }
 
-  data_summary <<-
+   data_summary <<-
     data_treated %>%
     left_join(data_eligible, by=c("patient_id", "trial_time")) %>%
     left_join(data_matched %>% filter(treated==1L) %>% transmute(patient_id, matched=1L, trial_time=tte_treatment), by=c("patient_id", "trial_time")) %>%

### 2022-02-22 analysis/combine_model.R
add ve to effect table

@@ -53,7 +53,7 @@ recode_outcome <- set_names(events_lookup$event, events_lookup$event_descr)
 if(subgroup_variable=="none"){
   recode_subgroup <- c(` `="none")
   recode_subgroup_variable <- c(`Main analysis` = "none")
-  recode_subgroup_level <- c("none")
+  recode_subgroup_level <- c(`Any`="none")
   subgroup <- c("none")
 }
 
@@ -122,7 +122,10 @@ model_effects <-
     model_descr = fct_reorder(model_descr, model),
     hr = exp(estimate),
     hr.ll = exp(conf.low),
-    hr.ul = exp(conf.high)
+    hr.ul = exp(conf.high),
+    ve = 1-hr,
+    ve.ll = 1-hr.ul,
+    ve.ul = 1-hr.ll
   ) %>%
   filter(
     # this filtering is necessary for dummy data,

### 2022-02-22 analysis/data_process.R
update age cutoffs

@@ -114,8 +114,8 @@ data_processed <- data_extract %>%
 
     ageband = cut(
       age,
-      breaks=c(-Inf, 18, 30, 40, 50, 60, 65, Inf),
-      labels=c("under 18", "18-30", "30s", "40s", "50s", "60-64", "65+"),
+      breaks=c(-Inf, 18, 40, 50, 60, 70, 80, 90, Inf),
+      labels=c("under 18", "18-39", "40-49", "50-59", "60-69", "70-79", "80-89", "90+"),
       right=FALSE
     ),
 

### 2022-02-22 analysis/report_seqtrialcox.R
add estimate+CI

@@ -201,6 +201,7 @@ incidence_rate_redacted <- local({
       rr = rate_1 / rate_0,
       rrE = scales::label_number(accuracy=0.01, trim=FALSE)(rr),
       rrCI = rrCI_exact(events_1, yearsatrisk_1, events_0, yearsatrisk_0, 0.01),
+      rrECI = paste0(rrE, " ", rrCI)
     )
 
   redacted <-
@@ -215,6 +216,7 @@ incidence_rate_redacted <- local({
       rr = redactor2(pmin(events_1, events_0), 5, rr),
       rrE = redactor2(pmin(events_1, events_0), 5, rrE),
       rrCI = redactor2(pmin(events_1, events_0), 5, rrCI),
+      rrECI = redactor2(pmin(events_1, events_0), 5, rrECI),
 
       events_0 = redactor2(events_0, 5),
       events_1 = redactor2(events_1, 5),

### 2022-02-22 analysis/match_seqtrialcox.R
minor refactor to capture summary info for all boosted people

@@ -271,7 +271,7 @@ local({
       transmute(
         patient_id,
         treated=1L,
-        trial_time=tte_treatment
+        trial_time=trial_time,
       )
 
     # set of people boosted on trial day, + their candidate controls
@@ -308,7 +308,7 @@ local({
       transmute(
         patient_id,
         eligible=1L,
-        trial_time=tte_treatment
+        trial_time=trial_time,
       )
 
 
@@ -349,6 +349,7 @@ local({
         patient_id = matching_candidates_i$patient_id,
         weight = matching_i$weights,
         trial_day = trial,
+        trial_time = trial_time,
       ) %>%
 
       # filter treated and controls to ensure exact 1:1 matching
@@ -397,35 +398,47 @@ local({
     data_matched <- bind_rows(data_matched, data_matched_i)
   }
 
+  trial_time <- NULL
+
+  data_matched <-
+    data_matched %>%
+    transmute(patient_id, match_id, matched=1L, control=1L-treated, trial_time, tte_recruitment, tte_controlistreated, tte_stop)
+
   data_summary <<-
     data_treated %>%
     left_join(data_eligible, by=c("patient_id", "trial_time")) %>%
-    left_join(data_matched %>% filter(treated==1L) %>% transmute(patient_id, matched=1L, trial_time=tte_treatment), by=c("patient_id", "trial_time")) %>%
+    left_join(data_matched %>% filter(control==0L), by=c("patient_id", "trial_time")) %>%
     mutate(
-      eligible=replace_na(eligible, 0L),
-      matched=replace_na(matched, 0L)
-    )
-
-  data_matched <<-
-    data_matched %>%
-    select(-subclass) %>%
+      eligible = replace_na(eligible, 0L),
+      matched = replace_na(matched, 0L),
+      control = if_else(matched==1L, 0L, NA_integer_)
+    ) %>%
+    bind_rows(
+      data_matched %>% filter(control==1L)
+    ) %>%
     mutate(
-      treated_patient_id = paste0(treated, "_", patient_id),
-      fup = pmin(tte_stop - tte_recruitment, last(postbaselinecuts))
+      trial_day=trial_time+1,
+      status = fct_case_when(
+        treated & !eligible ~ "ineligible",
+        treated & eligible & !matched ~ "unmatched",
[…]

### 2022-02-22 manuscript/draft-manuscript.Rmd
edits from feedback

@@
 ---
-title: "Effectiveness of COVID-19 Booster doses in England: a sequential trials study in OpenSAFELY"
+title: "Effectiveness of BNT162b2 booster doses in England: a observational study in OpenSAFELY"
 output:
-  word_document: default
   html_document:
@@
     self_contained: yes
+  word_document: default
   bookdown::html_document2:
@@
 
+studydurationweeks <- as.numeric(as.duration(study_dates$studyend_date - study_dates$studystart_date +1) , "weeks")
 ```
@@
 
-hazards_none <- read_csv(fs::path(output_dir_os, "none", "effects.csv")) %>%
+hazards_none <- read_csv(fs::path(output_dir_os, "none", "effects.csv"))
+hazards_vax12_type <- read_csv(fs::path(output_dir_os, "vax12_type", "effects.csv"))
+
+hazards <- 
+  bind_rows(hazards_none, hazards_vax12_type) %>%
   mutate(
+
+    subgroup_descr = fct_inorder(subgroup_descr),
+    subgroup_level = fct_inorder(subgroup_level),
+    subgroup_level_descr = fct_inorder(fct_explicit_na(subgroup_level_descr, "All")),
     treatment_descr = fct_inorder(treatment_descr),
@@
     #model_name = fct_inorder(str_wrap(model_name, 15)),
-    hr_display = paste0(number_format(accuracy=0.01)(hr), " (", number_format(accuracy=0.01)(hr.ll), "-", number_format(accuracy=0.01)(hr.ul), ")")
-  )
-
-hazards_vax12_type <- read_csv(fs::path(output_dir_os, "vax12_type", "effects.csv")) %>%
-  mutate(
-    subgroup_descr = fct_inorder(subgroup_descr),
-    subgroup_level = fct_inorder(subgroup_level),
-    treatment_descr = fct_inorder(treatment_descr),
-    outcome_descr = fct_inorder(outcome_descr),
-    #model_name = fct_inorder(str_wrap(model_name, 15)),
-    hr_display = paste0(number_format(accuracy=0.01)(hr), " (", number_format(accuracy=0.01)(hr.ll), "-", number_format(accuracy=0.01)(hr.ul), ")")
+    hr_display = paste0(label_number(accuracy=0.01)(hr), " (", label_number(accuracy=0.01)(hr.ll), "-", label_number(accuracy=0.01)(hr.ul), ")"),
+    ve_display = paste0(label_percent(accuracy=0.1)(ve), " (", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
   )
@@
 #riskdiff3 <- riskdiff %>% filter(model==3)
-hazards3_none <- hazards_none %>% filter(model==3)
-hazards3_vax12_type <- hazards_vax12_type %>% filter(model==3)
+hazards3 <- hazards %>% filter(model==3)
 
@@
 
-William J Hulme^1^, Elizabeth J Williamson^2^, Amelia Green^1^, Elsie Horne^3,4^, Helen Curtis^1^, Linda Nab^1^, Milan Wiedemann^1^, Ruth Keogh^2^, Edward Parker^2^, Venexia Walker^3,4^, Jonathan AC Sterne^3,4^, Miguel Hernán^6,7^, Ben Goldacre^1^ + OpenSAFELY collaborative.
+William J Hulme^1^, Elizabeth J Williamson^2^, Amelia Green^1^, Elsie Horne^3,4^, Helen Curtis^1^, Linda Nab^1^, Milan Wiedemann^1^, Ruth Keogh^2^, Edward Parker^2^, Venexia Walker^3,4^, Jonathan AC Sterne^3,4^, Miguel A Hernán^6,7^, Ben Goldacre^1^ + OpenSAFELY collaborative.
 
@@
 
-With the approval of NHS England we used the OpenSAFELY-TPP database, covering 40% of GP practices in England and linked to national coronavirus surveillance, hospital episodes, and death registry data, to estimate the effectiveness of a booster dose in "fully vaccinated" people using a sequential trials approach.
-On each day of the recruitment period (`r format(as.Date(study_dates$studystart_date), "%d %B %Y")` to `r format(as.Date(study_dates$lastvax3_date), "%d %B %Y")`) people receiving a booster dose on that day were recruited to the treatment arm, with a matched sample of unboosted people recruited to the control arm.
-Controls were matched on geogrpahical region, JCVI priority group, and the week and type of the second vaccine dose.
-Trial arms across each day were pooled and analysed using a Cox model, controlling for additional confounders.
-Primary outcomes were recorded SARS-CoV-2 infection, COVID-19-related accident and emergency attendance, unplanned COVID-19-related hospital admission, and COVID-19-related death occurring within 12 weeks of recruitment.
+With the approval of NHS England we used the OpenSAFELY-TPP database, covering 40% of GP practices in England and linked to national coronavirus surveillance, hospital episodes, and death registry data, to estimate the effectiveness of a booster dose in "fully vaccinated" people.
+On each day of the recruitment period (`r format(as.Date(study_dates$studystart_date), "%d %B %Y")` to `r format(as.Date(study_dates$lastvax3_date), "%d %B %Y")`) each eligible person receiving a BNT162b2 booster dose on that day was recruited to the treatment group, and matched with an unboosted person recruited to the control group.
+Matching factors were region, JCVI priority group, and the day and type of the second vaccine dose.
+The treated and control groups across each day were pooled.
+Primary outcomes were documented SARS-CoV-2 infection, COVID-19-related accident and emergency attendance, unplanned COVID-19-related hospital admission, and COVID-19-related death occurring within 12 weeks of recruitment.
+Hazard ratios were estimated via a Cox model with additional covariates that were not used as matching factors.
 
@@
 
-We matched a total of `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` BNT162b2 booster recipients to unboosted controls with only two doses.
-During the 28-41 days the estimated hazard ratio (95% CI) comparing a BNT162b2 booster dose to two doses only, was `r filter(hazards_none, outcome=="postest", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)` for a positive test, `r filter(hazards_none, outcome=="covidadmitted", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 hospital admission, and `r filter(hazards_none, outcome=="coviddeath", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 death.
+We matched `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` BNT162b2 booster recipients to unboosted controls with only two doses.
+During days 14-20 after the booster, the estimated hazard ratio (95% CI) comparing a BNT162b2 booster dose to two doses only was `r filter(hazards_none, outcome=="postest", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for documented SARS-CoV-2 infection, `r filter(hazards_none, outcome=="covidadmitted", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(hazards_none, outcome=="coviddeath", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 death.
 
@@
 
+<!-- TODO -->
+
+<!-- * full description of booster eligiblity over time -->
+
+<!-- * figure 1 chart with SMD / %-point differences -->
+
+<!-- * flowchart incl matching -->
+
+<!-- * check for moderna references and remove if necessary -->
+
+<!-- * check final recruitment day and include in results -->
+
+<!-- * check if rolling average restrictions are ever used (need to output dataset) -->
+
+<!-- * JCVI group spell out -->
+
+<!-- * references -->
+
 # Introduction
 
-The national COVID-19 vaccination programme in the United Kingdom began a new phase in September 2021 with the administration of the first booster COVID-19 vaccine doses in those who had already received their primary vaccination course (that is, "fully vaccinated" with a complete first and second dose).
-Initially offered to those at highest risk of severe disease, eligibility was extended to the entire adult population by the end of the year.
+The national COVID-19 vaccination programme in England began a new phase on 16 September 2021 with the administration of the first booster COVID-19 vaccine doses in those who had already received their primary vaccination course (a completed first and second dose) (ref <https://www.england.nhs.uk/2021/09/nhs-begins-covid-19-booster-vaccination-campaign/>).
+Initially offered to those at highest risk of severe disease, eligibility was progressively extended and by 15 December 2021 every adult was eligible (<https://www.england.nhs.uk/2021/12/nhs-booster-bookings-open-to-every-eligible-adult/>).
+Booster doses were initially available no earlier than six months after dose two, but this was reduced to three months on 8 December 2021, following concerns over the surge in cases and the emergence of the Omicron variant (ref: <https://www.england.nhs.uk/2021/12/people-40-and-over-to-get-their-lifesaving-booster-jab-three-months-on-from-second-dose/>)
 
-We set out to estimate the effectiveness of booster doses amongst the fully vaccinated adult population in England using the OpenSAFELY-TPP linked primary care database covering around 40% of England's population.
+<!-- 13 december 30-39 become eligible (https://www.england.nhs.uk/2021/12/nhs-to-roll-out-life-saving-booster-jab-to-people-aged-30-plus/) -->
 
-In the vast majority of cases, BNT162b2 and mRNA-1273 vaccines were used, and administered without consideration of the type of primary course.
-BNT162b2 was used from 16 September 2021, and mRNA-1273 from 29 October 2021 onwards.
-Consequently, available follow-up of BNT162b2 recipients is longer, and so we restrict the analysis to the effectiveness of BNT162b2boosting only.
-An analysis of mRNA-1273 booster effectiveness is planned when sufficient follow-up has accrued.
+<!-- 15 december everyone eligible https://www.england.nhs.uk/2021/12/nhs-booster-bookings-open-to-every-eligible-adult/ -->
+
+[context]
+
+We set out to estimate the effectiveness of BNT162b2 mRNA Pfizer-BioNTech booster doses versus two doses only in adults in England using the OpenSAFELY-TPP linked primary care database covering around 40% of England's population.
+An analysis of effectiveness of Moderna mRNA-1273 boosters, which were administered from 29 October 2021, is planned when sufficient follow-up has accrued.
 
@@
 Vaccination status is available in the GP record directly via the National Immunisation Management System (NIMS).
-HCW status is recorded for all vaccine recipients at the time of vaccination, and this information is sent to OpenSAFELY-TPP from NHS Digital's COVID-19 data store.
+Healthcare worker status is recorded for all vaccine recipients at the time of vaccination, and this information is sent to OpenSAFELY-TPP from NHS Digital's COVID-19 data store.
 
-## Study design and population
+## Eligiblity criteria
 
-We used a sequential trials approach designed to emulate a sequence of daily two-arm parallel randomised controlled trials.
-For each trial, the treatment arm comprised people who received a BNT162b2 booster dose on that day, and the control arm comprised the same number of randomly selected unboosted people matched on key characteristics, as described below.
-This closely follows the approach taken in previous vaccine effectiveness studies in Israeli registry data @dagan2021 @barda2021
+We used observational data to emulate a sequence of target trials @dagan2021 @barda2021, starting on consecutive days between `r format(as.Date(study_dates$studystart_date), "%d %B %Y")` and `r format(as.Date(study_dates$lastvax3_date), "%d %B %Y")` inclusive.
 
-Trial recruitment was considered between `r format(as.Date(study_dates$studystart_date), "%d %B %Y")` and `r format(as.Date(study_dates$studyend_date), "%d %B %Y")` inclusive.
-Trial eligibility on each day of recruitment was as follows: adults (18 and over) who were registered at a GP practice using TPP's SystmOne clinical information system; received a primary COVID-19 vaccination course of either two BNT162b2 doses or two ChAdOx1-S doses (mRNA-1273 was not considered due to small numbers); not a vaccinated health care worker, not a resident in a care or nursing home, and not medically housebound or receiving end-of-life care (due to distinct clustering of vaccine uptake and infection risk in these groups that could not be reasonably measured and controlled for); no evidence of SARS-CoV-2 infection or COVID-19 disease within 90 days prior to recruitment; not undergoing an unplanned hospital admission; information on sex, ethnicity, deprivation, and geographical region was complete.
+Eligibility on each day was as follows: adults (18 years and over) who were registered at a GP practice using TPP's SystmOne clinical information system; received a primary COVID-19 vaccination course of either two BNT162b2 doses or two ChAdOx1-S doses (mRNA-1273 was not considered due to small numbers); not a vaccinated health care worker, not a resident in a care or nursing home, and not medically housebound or receiving end-of-life care (due to distinct clustering of vaccine uptake and infection risk in these groups that could not be reasonably measured and controlled for); no evidence of SARS-CoV-2 infection or COVID-19 disease within 90 days prior to recruitment; not undergoing an unplanned hospital admission; information on sex, ethnicity, deprivation, and NHS region was complete.
 
-Additional eligiblity criteria based on sufficiently many people being vaccinated on each day within region, priority group, and second dose week strata were also used to ensure unboosted people were both eligible and able to receive a booster dose at the time of potential recruitment.
-The trial sequence was terminated early if nobody was eligible for the treatment arm or fewer than 10% of eligible treated people were successfully matched.
-See supplementary materials for further details.
+Additional eligibility criteria, based on sufficiently many people being vaccinated on each day by region, priority group, and week of second dose.
+If the rolling weekly average count of BNT162b2 booster doses within these strata was below 50 then recruitment did not occur on that day.
+This helped to ensure that unboosted people were both eligible and able to receive a booster dose at the time of potential recruitment.
 
-The control arms were selected using exact 1:1 matching without replacement.
-Participants were matched on primary course vaccine type (BNT162b2 or ChAdOx1-S), the calendar week of the second vaccine dose, geographical region, prior infection, immunosuppression, and vaccine priority group (as determined by the Joint Committee for Vaccine and immunisation expert working group).
-People selected as controls were not eligible to be included again as a control in a subsequent trial, but were eligible for selection into the treatment arm of a subsequent trial.
-Any unmatched booster recipients were dropped.
+We excluded matched pairs from days in which fewer than 10% of eligible treated people were successfully matched. See supplementary materials for further details.
 
-After exclusions for missing values on demographic variables, there were no missing values in remaining variables as they were each defined by the presence or absence of clinical codes or events in the patient record.
+## Treatment groups
 
-## Exposures, outcomes, covariates
+For each trial, the treatment group comprised people who received a BNT162b2 booster dose on that day, and the control group comprised the same number of randomly selected unboosted people, matched on key characteristics as follows: primary course vaccine type (BNT162b2 or ChAdOx1-S); calendar week of the second vaccine dose; NHS region; prior infection; immunosuppression; vaccine priority group (as determined by the Joint Committee for Vaccine and Immunisation expert working group).
+People selected as controls were not eligible to be included again as a control in a subsequent trial, but were eligible for selection into the treatment group of a subsequent trial.
 
+## Outcomes
+
+The following three outcomes were considered: documented SARS-CoV-2 infection; COVID-19 hospital admission; and COVID-19 death.
+SARS-CoV-2 infections were identified using SGSS testing records and based on swab date.
+Both polymerase chain reaction (PCR) and lateral flow tests are included, without differentiation between symptomatic and asymptomatic infection.
+COVID-19 hospital admissions were identified using HES in-patient hospital records with U07.1 or U07.2 reason for admission ICD-10 codes.
+COVID-19 deaths with the same COVID-19 ICD-10 codes mentioned anywhere on the death certificate (i.e., as an underlying or contributing cause of death) were included.
+
+<!--# COVID-19 related A&E attendance and critical/intensive care admission we also considered but there were doubts over data the reliability of event ascertainment in these cases. -->
 <!--# COVID-19 A&E attendances were identified using HES emergency care records with U07.1 ("COVID-19, virus identified") or U07.2 ("COVID-19, virus not identified") ICD-10 diagnosis codes @emergenc.
@@
 
-The following three outcomes were considered: positive SARS-CoV-2 test; unplanned COVID-19 hospital admission; COVID-19 death.
-Positive SARS-CoV-2 tests were identified using SGSS records and based on swab date.
-Both polymerase chain reaction (PCR) and lateral flow tests are included, without differentiation between symptomatic and asymptomatic infection.
-Unplanned COVID-19 hospital admissions were identified using HES in-patient hospital records with U07.1 or U07.2 reason for admission ICD-10 codes.
-COVID-19 deaths with the same COVID-19 ICD-10 codes mentioned anywhere on the death certificate (i.e., as an underlying or contributing cause of death) were included.
+## Follow-up
 
-<!--# COVID-19 related A&E attendance and critical/intensive care admission we also considered but there were doubts over data the reliability of event ascertainment in these cases. -->
-
-Participant characteristics used describe trial participants and for confounder adjustment include: age, sex (male or female), English Index of Multiple Deprivation (IMD, grouped by quintiles), ethnicity (Black, Mixed, South Asian, White, Other, as per the UK census), NHS region (East of England, Midlands, London, North East and Yorkshire, North West, South East, South West), number of conditions in the clinically "at risk" classification, as per national prioritisation guidelines, the number of SARS-CoV-2 tests in the 6 months prior to the study start date (via SGSS), rurality (urban conurbation, urban city or town, rural town or village), evidence of prior SARS-CoV-2 infection (positive SARS-CoV-2 test, "probable" infection documented in primary care, or COVID-19 hospitalisation), learning disabilities, severe mental illness, undergoing an unplanned hospital admission, undergoing a planned hospital admission.
-All characteristics were ascertained as at the trial recruitment date.
+Each matched pair was followed from the day of recruitment (time zero or baseline) until the earliest of death, GP practice de-registration, booster receipt by the control, 10 weeks, or `r format(as.Date(study_dates$studyend_date), "%d %B %Y")`.
 
@@
 
-Time zero was the trial recruitment day, and trial participants were followed up for no more than 10 weeks from this day.
-Follow-up was censored earlier than this at `r format(as.Date(study_dates$studyend_date), "%d %B %Y")`, death, or GP practice de-registration.
-Additionally for each matched pair, follow-up for both the boosted and unboosted person was censored if the control received a booster dose.
+For each outcome, we estimated cumulative incidence curves using the Kaplan-Meier estimator separately in the treated and control groups.
 
-Trials were pooled and analysed using Cox regression with period-specific hazard ratios for treatment, in two-week periods.
-Baseline hazards were allowed to differ for participants differing by recruitment day, region, and week and type of second vaccine dose, via Cox stratification.
-Other variables were included as covariates.
+We estimated period-specific hazard ratios for treatment versus control in consecutive two-week periods using a Cox model.
+Baseline hazards were allowed to differ by recruitment day, region, and week and type of second vaccine dose.
+The following variables were included as covariates: age, sex (male or female), English Index of Multiple Deprivation (IMD, grouped by quintile), ethnicity (Black, Mixed, South Asian, White, Other, as per the UK census), NHS region (East of England, Midlands, London, North East and Yorkshire, North West, South East, South West), number of conditions in the clinically "at risk" classification, as per national prioritisation guidelines, the number of SARS-CoV-2 tests in the 6 months prior to the study start date (via SGSS), evidence of prior SARS-CoV-2 infection (positive SARS-CoV-2 test, "probable" infection documented in primary care, or COVID-19 hospitalisation), learning disabilities, severe mental illness, undergoing an unplanned hospital admission, undergoing a planned hospital admission.
 A robust variance estimator was used to account for participant clustering.
 
-[The primary estimand is the average treatment effect in the recruited population.
-To estimate this, expected cumulative incidence rates are computed for each participant assuming treatment was received by everyone in both trial arms at time zero.
+<!-- [The primary estimand is the average treatment effect in the recruited population.
+To estimate this, expected cumulative incidence rates are estimated for each participant assuming treatment was received by everyone in both treatment groups at time zero.
 These curves are then averaged over all participants to give the marginal cumulative incidence.
-This is repeated assuming nobody is treated in either arm, and their difference taken.
-Calculating standard errors for these quantities is challenging, whether approached analytically (due to complex variance components) or computationally (via bootstrapping, due to the large datasets involved) and is there not reported.
-]
+This is repeated assuming nobody is treated in either group, and their difference taken.
+Calculating standard errors for these quantities is challenging, whether approached analytically (due to complex variance components) or computationally (via bootstrapping, due to the large sample sizes involved) and are therefore not reported.
+] -->
 
-We additionally estimated booster effectiveness in subgroups defined by primary vaccine course (both BNT162b2 or both ChAdOx1-S) using complete model stratification.
+We additionally estimated booster effectiveness in subgroups defined by primary vaccine course (both BNT162b2 or both ChAdOx1-S) by repeating the analysis separately in each subgroup.
+
+Booster effectiveness is defined as one minus the hazard ratio, and expressed as a percentage.
+
+## Missing data
+
+After exclusions for missing values on demographic variables as described earlier, there were no missing values in remaining variables as they were each defined by the presence or absence of clinical codes or events in the patient record.
+This was therefore a complete-case analysis.
 
@@
 
-Data management and analyses were conducted in Python 3.8 and R version 4.0.2.
+Data management and analyses were conducted in Python version 3.8.10 and R version 4.0.2.
 All code is available at <https://github.com/opensafely/booster-effectiveness>, and is shared openly for review and re-use under MIT open license.
@@
 
-A total of `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_treated))` BNT162b2 booster doses were administered between `r format(as.Date(study_dates$studystart_date), "%d %B %Y")` and `r format(as.Date(study_dates$studyend_date), "%d %B %Y")` in adults registered at a TPP practice, with `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_eligible))` (`r label_percent(0.1)(filter(matchsummary, treatment=="pfizer") %>% pull(prop_eligible_treated))`) eligible for inclusion into the treatment arm.
+A total of `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_treated))` BNT162b2 booster doses were administered between `r format(as.Date(matchsummary_treated$firstrecruitdate), "%d %B %Y")` and `r format(as.Date(matchsummary_treated$lastrecruitdate), "%d %B %Y")` in adults registered at a TPP practice, with `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_eligible))` (`r label_percent(0.1)(filter(matchsummary, treatment=="pfizer") %>% pull(prop_eligible_treated))`) eligible for inclusion into the treatment group.
 Of these, `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` (`r label_percent(0.1)(filter(matchsummary, treatment=="pfizer") %>% pull(prop_matched_eligible))`) were successfully matched with unboosted controls (Figure 1).
@@
 
-Participant characteristics at trial recruitment between trial arms (Table 1) were typically well-balanced (necessarily so on matched characteristics), though with some exceptions, for instance unboosted controls had higher levels of deprivation, higher rates of learning disabilities and severe mental illness, and lower SARS-CoV-2 testing frequency prior to the study.
+Participant characteristics at trial recruitment between treatment groups (Table 1) were typically well-balanced (necessarily so on matched characteristics), though with some exceptions, for instance unboosted controls had higher levels of deprivation, higher rates of learning disabilities and severe mental illness, and lower SARS-CoV-2 testing frequency.
 These unbalanced characteristics were adjusted for in the Cox regression models.
 
-There were `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(fup_years))` person-years of follow-up across both trial arms (Table 2).
+There were `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(fup_years))` person-years of follow-up across both treatment groups (Table 2).
 
@@
 
-#### Figure 1: Booster vaccination uptake, treatment arm eligiblity and matching success
+#### Figure 1: Booster vaccination uptake, treatment group eligibility and matching success
 
@@
     course = fct_case_when(
-      vax12_type=="pfizer-pfizer" & treatment=="pfizer" ~ "2x BNT162b2  / 1x BNT162b2",
-      vax12_type=="az-az" & treatment=="pfizer" ~ "2 x ChAdOx1-S / 1 x BNT162b2",
-      vax12_type=="pfizer-pfizer" & treatment=="moderna" ~ "2x BNT162b2 / 1x mRNA-1273",
-      vax12_type=="az-az" & treatment=="moderna" ~ "2x ChAdOx1-S / 1x mRNA-1273",
+      vax12_type=="pfizer-pfizer" & treatment=="pfizer" ~ "2× BNT162b2  / 1× BNT162b2",
+      vax12_type=="az-az" & treatment=="pfizer" ~ "2× ChAdOx1-S / 1× BNT162b2",
+      vax12_type=="pfizer-pfizer" & treatment=="moderna" ~ "2× BNT162b2 / 1× mRNA-1273",
+      vax12_type=="az-az" & treatment=="moderna" ~ "2× ChAdOx1-S / 1× mRNA-1273",
       TRUE ~ NA_character_
@@
 
-Participant characteristics as on the day of recruitment into the treatment or control arm.
+Participant characteristics as on the day of recruitment into the treatment or control group.
 
@@
 
-#### Table 2: Recruitment, follow-up, and event incidence
+#### Table 2: Follow-up and outcomes
+
+Total follow-up and incidence rates for each outcome, by treatment group.
 
@@
     rate_0,
-    rr, 
-    rrCI
+    rrECI
   ) %>%
+  mutate(
+    rate_1=if_else(rate_1<0.001, "<0.001", label_number(0.001)(rate_1)),
+    rate_0=if_else(rate_0<0.001, "<0.001", label_number(0.001)(rate_0)),
+  )
   gt(
@@
 
-    rate_0 = "Incidence",
-    rate_1 = "Incidence",
-
-    rr = "Incidence rate ratio",
-    rrCI = "95% CI"
+    rate_0 = "Incidence rate",
+    rate_1 = "Incidence rate",
+    rrECI = "Incidence rate ratio (95% CI)"
   ) %>%
@@
     columns = ends_with("0")
-  ) %>%
-  fmt_number(
-    columns = c("rate_0", "rate_1", "rr"),
-    decimals = 3
   ) %>%
@@
 
-Trial participants contributed `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="postest", fup_period=="All") %>% pull(events))` positive SARS-CoV-2 tests, `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="covidadmitted", fup_period=="All") %>% pull(events))` COVID-19 hospital admissions, and `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="coviddeath", fup_period=="All") %>% pull(events))` COVID-19 deaths (Table 2).
+Participants in the emulated trial contributed `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="postest", fup_period=="All") %>% pull(events))` documented infection, `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="covidadmitted", fup_period=="All") %>% pull(events))` COVID-19 hospital admissions, and `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="coviddeath", fup_period=="All") %>% pull(events))` COVID-19 deaths (Table 2).
 
@@
 
-During the 28-41 days the estimated hazard ratio (95% CI) comparing a BNT162b2 booster dose to two doses only, was `r filter(hazards_none, outcome=="postest", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)` for a positive test, `r filter(hazards_none, outcome=="covidadmitted", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 hospital admission, and `r filter(hazards_none, outcome=="coviddeath", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 death (Figure 2).
+In the 14-20 days period of follow-up, the estimated effectiveness (95% CI) comparing a BNT162b2 booster dose to two doses only, was `r filter(hazards_none, outcome=="postest", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for documented infection, `r filter(hazards_none, outcome=="covidadmitted", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(hazards_none, outcome=="coviddeath", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 death (Figure 2).
 
-<!-- `r filter(hazards_none, outcome=="covidattendance", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 A&E attendances, `r filter(hazards_none, outcome=="covidcc", treatment=="pfizer", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 critical care admission,  -->
+<!-- `r filter(hazards_none, outcome=="covidattendance", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 A&E attendances, `r filter(hazards_none, outcome=="covidcc", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 critical care admission,  -->
 
-<!-- For mRNA-1273, the corresponding estimated hazard ratios were `r filter(hazards_none, outcome=="postest", treatment=="moderna", model==3, term_left==28) %>% pull(hr_display)`, for COVID-19 A&E attendances was `r filter(hazards_none, outcome=="covidattendance", treatment=="moderna", model==3, term_left==28) %>% pull(hr_display)`, for COVID-19 hospital admission was `r filter(hazards_none, outcome=="covidadmitted", treatment=="moderna", term_left==28) %>% pull(hr_display)`, for COVID-19 critical care admission was `r filter(hazards_none, outcome=="covidcc", treatment=="moderna", model==3, term_left==28) %>% pull(hr_display)`, and for COVID-19 death was `r filter(hazards_none, outcome=="coviddeath", treatment=="moderna", model==3, term_left==28) %>% pull(hr_display)`.
+<!-- For mRNA-1273, the corresponding estimated effectiveness were `r filter(hazards_none, outcome=="postest", treatment=="moderna", model==3, term_left==14) %>% pull(ve_display)`, for COVID-19 A&E attendances was `r filter(hazards_none, outcome=="covidattendance", treatment=="moderna", model==3, term_left==14) %>% pull(ve_display)`, for COVID-19 hospital admission was `r filter(hazards_none, outcome=="covidadmitted", treatment=="moderna", term_left==14) %>% pull(ve_display)`, for COVID-19 critical care admission was `r filter(hazards_none, outcome=="covidcc", treatment=="moderna", model==3, term_left==14) %>% pull(ve_display)`, and for COVID-19 death was `r filter(hazards_none, outcome=="coviddeath", treatment=="moderna", model==3, term_left==14) %>% pull(ve_display)`.
 -->
 
-For the same period, the estimated hazard ratio (95% CI) for BNT162b2 and ChAdOx1-S primary courses respectively were `r filter(hazards_vax12_type, outcome=="postest", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3, term_left==28) %>% pull(hr_display)` and `r filter(hazards_vax12_type, outcome=="postest", treatment=="pfizer", subgroup_level=="az-az", model==3, term_left==28) %>% pull(hr_display)` for a positive test, `r filter(hazards_vax12_type, outcome=="covidadmitted", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3, term_left==28) %>% pull(hr_display)` and `r filter(hazards_vax12_type, outcome=="covidadmitted", treatment=="pfizer", subgroup_level=="az-az", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 hospital admission, and `r filter(hazards_vax12_type, outcome=="coviddeath", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3, term_left==28) %>% pull(hr_display)` and `r filter(hazards_vax12_type, outcome=="coviddeath", treatment=="pfizer", subgroup_level=="az-az", model==3, term_left==28) %>% pull(hr_display)` for COVID-19 death (Figure 2).
+For the same period, the estimated effectiveness (95% CI) comparing a BNT162b2 booster dose to two doses only separately for 2× BNT162b2 and 2× ChAdOx1-S primary course recipients respectively were `r filter(hazards_vax12_type, outcome=="postest", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3, term_left==14) %>% pull(ve_display)` and `r filter(hazards_vax12_type, outcome=="postest", treatment=="pfizer", subgroup_level=="az-az", model==3, term_left==14) %>% pull(ve_display)` for documented infection, `r filter(hazards_vax12_type, outcome=="covidadmitted", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3, term_left==14) %>% pull(ve_display)` and `r filter(hazards_vax12_type, outcome=="covidadmitted", treatment=="pfizer", subgroup_level=="az-az", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(hazards_vax12_type, outcome=="coviddeath", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3, term_left==14) %>% pull(ve_display)` and `r filter(hazards_vax12_type, outcome=="coviddeath", treatment=="pfizer", subgroup_level=="az-az", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 death (Figure 2).
 
@@
 
-For each outcome based on the fully adjusted model, the period-specific hazard ratio for boosting with BNT162b2 versus not boosting is shown, stratified by primary course.
+For each outcome based on the fully adjusted model, the period-specific hazard ratios for boosting with BNT162b2 versus not boosting is shown, stratified by primary course.
 Models with less extensive confounder adjustment are provided in supplementary materials (supplementary Figure SX).
@@
 formatpercent100 <- function(x,accuracy){
-  formatx <- scales::label_percent(accuracy)(x)
+  formatx <- scales::label_percent(accuracy, suffix="")(x)
 
   if_else(
-    formatx==scales::label_percent(accuracy)(1),
-    paste0(">",scales::label_percent(1)((100-accuracy)/100)),
+    formatx==scales::label_percent(accuracy, suffix="")(1),
+    paste0(">",scales::label_percent(1, suffix="")((100-accuracy)/100)),
     formatx
@@
   plot_hazards <-
-    hazards3_vax12_type %>%
+    hazards3 %>%
     filter(
@@
     geom_hline(aes(yintercept=1), colour='grey')+
-    facet_grid(rows=vars(outcome_descr), cols=vars(subgroup_level), switch="y")+
+    facet_grid(rows=vars(outcome_descr), cols=vars(subgroup_level_descr), switch="y")+
     scale_y_log10(
@@
         ~(1-.),
-        name="Effectiveness",
+        name="Effectiveness (%)",
         breaks = 1-(y_breaks),
@@
 
-This observational study of over 8 million boosted adults in England and their matched controls estimated high protection of BNT162b2 booster vaccination against all studied outcomes using routinely-collected health records with comprehensive coverage of primary care, hospital admissions, COVID-19 testing and vaccination, and death registrations.
-Protection was similar between 2x BNT162b2 primary course recipients and 2x ChAdOx1-S recipients.
-There were apparent waning effects around 6 weeks after boosting, which is unlikely to explained by increased transmissibility of the Omicron variant given Omicron was not yet dominant in the UK by the end of the follow-up window @COVID19OmicronDaily, though changing testing behaviours over time between boosted and unboosted people may influence estimates.
+This observational study of over 8 million boosted adults in England between `r format(as.Date(matchsummary_treated$firstrecruitdate), "%d %B %Y")` and `r format(as.Date(matchsummary_treated$lastrecruitdate), "%d %B %Y")` and their matched controls estimated high protection of BNT162b2 booster vaccination against all studied outcomes using routinely-collected health records with comprehensive coverage of primary care, hospital admissions, COVID-19 testing and vaccination, and death registrations.
+Protection was similar between 2× BNT162b2 primary course recipients and 2× ChAdOx1-S recipients.
+There were apparent waning effects around 6 weeks after boosting, which is unlikely to be explained by increased transmissibility of the Omicron variant given Omicron was not yet dominant in the UK by the end of the follow-up window @COVID19OmicronDaily, though changing testing behaviours over time between boosted and unboosted people may influence estimates.
 
@@
 
-Positive SARS-CoV-2 testing data likely underestimates the true incidence of infection.
+Documented SARS-CoV-2 infection data likely underestimates the true incidence of infection.
 The mass availability of lateral flow tests in the UK during the study period, whose results are not routinely recorded in COVID-19 surveillance data, means many infections, including symptomatic infections, may be undocumented, despite encouragement to seek a confirmatory PCR test.
 Potential differences in testing behaviour between vaccinated and unvaccinated people also contributes to the unreliability of testing data as a means to assess effectiveness.
+SARS-CoV-2 testing was not widely available early in the pandemic so evidence of prior infection is likely to be under-ascertained.
+We may not have all records of negative tests so counts of number of tests are also incomplete.
 
-Despite reasonable balance of baseline characteristics in the trial arms, and adjustment for a range of potential confounders, the possibility of unmeasured confounding remains.
-In particular, the apparent protective effect immediately after boosting is biologically implausible, and will be explained in part by the occurrence of COVID-19 symptoms at baseline not recorded in the health record.
+We excluded a number of groups, such as health care workers and care home residents, where testing, vaccination, and infection risk characteristics were unusual or had substantial within-group heterogeneity that could not be adequately accounted for.
 
-This study considers effectiveness in a population where Delta was the predominant circulating variant (ref gov.uk reports).
-Re-running the analysis against a backdrop of the now globally dominant Omicron variant will be undertaken where valid concurrent comparisons between boosted and unboosted people can be made.
+Hospital admission records are only completed after discharge, so some longer hospital stays may not be counted.
+
+Despite reasonable balance of baseline characteristics between treatment groups, and adjustment for a range of potential confounders, the possibility of unmeasured confounding remains.
+In particular, the apparent protective effect immediately after boosting is biologically implausible, and will be explained in part by the occurrence of COVID-19 symptoms at baseline that were not recorded in the health record, and therefore could not be used to exclude from recruitment.
+
+This study considers effectiveness in a population where Delta was the predominant circulating variant @COVID19OmicronDaily.
+Re-running the analysis against a backdrop of the now globally-dominant Omicron variant will be undertaken where valid concurrent comparisons between boosted and unboosted people can be made.
 
@@
 Millions of eligible participants may be required to obtain precise estimates, particularly when event rates are low, and so extending the approach to smaller subgroups may not be feasible.
-For instance, due to small numbers, we did not study booster effectiveness in those who had received the mRNA-1273 vaccine as their primary course or those who had received a heterologous primary course, despite many thousands of such recipients in our records.
+For instance, due to small numbers, we did not study booster effectiveness in those who had received the mRNA-1273 vaccine as their primary course or those who had received a heterologous primary course, despite many thousands of such recipients in England.
 
 Finally, mRNA-1273 boosting was rolled out later than for BNT162b2, and so we could not evaluate effectiveness of this vaccine due to insufficient follow-up.
-A analysis of mRNA-1273 boosting and comparative effectiveness between mRNA-1273 and BNT162b2 is planned.
+An analysis of mRNA-1273 boosting and comparative effectiveness between mRNA-1273 and BNT162b2 is planned.
 
@@
 
-This study found high protection for BNT162b2 boosting against COVID-19 hospitalisation and death in the Delta era.
+This study of over 8 million patients in England found high protection for BNT162b2 boosting against COVID-19 hospitalisation and death during an `r studydurationweeks` week period in which the Delta variant was dominant.
 

### 2022-02-22 manuscript/draft-supplement.Rmd
edits from feedback

@@
 Determining vaccine eligibility for a given person on a given day is therefore complex.
-To avoid these complexities, we determined trial eligibility for each person on each day of recruitment by asking if there were sufficiently many people with similar characteristics who were being vaccinated on that day.
-Specifically, if less than X people in the same region, priority group, and week of second vaccine dose had been vaccinated, then they were not eligible for recruitment.
+To avoid these complexities, we determined trial eligibility for each person on each day of recruitment by ensuring that there were sufficiently many people with similar characteristics who were being vaccinated on that day.
+Specifically, a rolling weekly average was calculated within strata defined by region, priority group, and week of second vaccine dose.
+If fewer than 50 people had been vaccinated within each strata on each day, then that strata was not eligible for recruitment on that day.
+
+The Joint Committee on Vaccine and Immunisation (JCVI) priority groups were defined as follows:
+
+```{r, jcvi, echo=FALSE, message=FALSE, warning=FALSE}
+
+tibble(
+  `Priority group` = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10a", "10b"),
+  `Description` = c(
+    "Residents in a care home for older adults\nStaff working in care homes for older adults",
+    "All those 80 years of age and over\nFrontline health and social care workers",
+    "All those 75-79 years of age",
+    "All those 70-74 years of age\nIndividuals aged 16-69 in a high risk group",
+    "All those 65-69 years of age",
+    "Adults aged 16-64 years in an at-risk group",
+    "All those 60-64 years of age",
+    "All those 55-59 years of age",
+    "All those 50-54 years of age",
+    "All those 40-49 years of age",
+    "All those 18-39 years of age"
+  )
+) %>%
+mutate(Description = str_replace_all(Description, "\n", "<br>")) %>% 
+gt() %>% 
+fmt_markdown(columns = TRUE)
+
+```
+
+See original priority groups here: https://assets.publishing.service.gov.uk/government/uploads/system/uploads/attachment_data/file/1007737/Greenbook_chapter_14a_30July2021.pdf#page=15
+See revised priority groups here: https://www.england.nhs.uk/coronavirus/wp-content/uploads/sites/52/2021/07/C1327-covid-19-vaccination-autumn-winter-phase-3-planning.pdf
+
+Note that we excluded care home residents and health care workers in our analysis, so group 1 is not included and group 2 includes only those aged 80 and over.
+The original priority group list has 9 groups, with a 10th group implicitly defined as "everybody else".
+Here we explicitly define this group, and split into two (10a and 10b) because of the earlier booster eligiblity in the 40-49 group from 15 November 2021 onwards.
+(https://www.gov.uk/government/news/jcvi-issues-advice-on-covid-19-booster-vaccines-for-those-aged-40-to-49-and-second-doses-for-16-to-17-year-olds)
 

### 2022-02-22 analysis/design.R
reduce matching set

@@ -27,7 +27,7 @@ fs::dir_create(here("lib", "design"))
 study_dates <- lst(
   studystart_date = "2021-09-16", #start of follow-up thursday 16 sepetember first booster jabs administered in england
   studyend_date = "2021-12-15", # end of follow-up (no hosp dates after this)
-  lastvax2_date = "2021-08-01", # don't recruit anyone with second vaccination after this date
+  lastvax2_date = "2021-10-01", # don't recruit anyone with second vaccination after this date
   lastvax3_date = "2021-12-01", # end of recruitment
   firstpfizer_date = "2020-12-08", # first pfizer vaccination in national roll-out
   firstaz_date = "2021-01-04", # first az vaccination in national roll-out
@@ -88,20 +88,22 @@ exact_variables <- c(
 
   "jcvi_group",
   "vax12_type",
-  #"vax2_week",
+  "vax2_week",
   "region",
-  "sex",
+  #"sex",
   "cev",
   "multimorb",
   "prior_covid_infection",
   "immunosuppressed",
-  "status_hospplanned"
+  #"status_hospplanned"
+  NULL
 )
 write_rds(exact_variables, here("lib", "design", "exact_variables.rds"))
 
 caliper_variables <- c(
   age = 3,
-  vax2_day = 7
+  #vax2_day = 7,
+  NULL
 )
 write_rds(caliper_variables, here("lib", "design", "caliper_variables.rds"))
 

### 2022-02-22 analysis/combine_match.R
add matching eligiblity flowchart

@@ -118,3 +118,15 @@ table1by <-
 
 write_csv(table1by, fs::path(output_dir, "table1by.csv"))
 
+
+
+
+matchflowchart <-
+  match_metaparams %>%
+  mutate(
+    flowchart = map(treatment, ~read_csv(here("output", "models", "seqtrialcox", .x, glue("match_data_flowchart.csv"))))
+  ) %>%
+  unnest(flowchart)
+
+write_csv(matchflowchart, fs::path(output_dir, "flowchart.csv"))
+

### 2022-02-22 analysis/match_seqtrialcox.R

@@ -150,46 +150,6 @@ data_rollingstrata_vaxcount <-
   filter(vax3_time>=0)
 
 
-## baseline variables ----
-
-data_baseline <-
-  data_cohort %>%
-  transmute(
-    patient_id,
-    age,
-    ageband,
-    sex,
-    ethnicity_combined,
-    imd_Q5,
-    region,
-    jcvi_group,
-    jcvi_group_descr,
-    rural_urban_group,
-    prior_tests_cat,
-    multimorb,
-    learndis,
-    sev_mental,
-    cev,
-    sev_obesity,
-    chronic_heart_disease,
-    chronic_kidney_disease,
-    diabetes,
-    chronic_liver_disease,
-    chronic_resp_disease,
-    chronic_neuro_disease,
-    immunosuppressed,
-    asplenia,
-    immuno,
-
-    vax12_type,
-    vax12_type_descr,
-    vax2_week,
-    vax3_date,
-    vax3_type,
-
-  )
-logoutput_datasize(data_baseline)
-
 data_nontimevarying <-
   data_cohort %>%
   select(
@@ -207,8 +167,6 @@ data_nontimevarying <-
   )
 
 
-
-
 if(removeobjects) rm(data_rollingstrata_vaxcount)
 
 data_timevarying <-
@@ -219,6 +177,7 @@ data_timevarying <-
     tstop,
     any_of(matching_variables),
     status_hospunplanned,
+    status_hospplanned,
     mostrecent_anycovid,
   )
 
@@ -247,7 +206,7 @@ data_tte <-
 
     tte_censor = tte(day0_date, censor_date, censor_date, na.censor=TRUE),
 
-    tte_treatment = tte(day0_date, treatment_date-1, censor_date, na.censor=TRUE),
+    tte_treatment = tte(day0_date, treatment_date-1, censor_date-1, na.censor=TRUE), # -1 to move censoring forward by one day so that everybody has at least one day of follow up
 
     tte_stop = pmin(tte_censor, na.rm=TRUE),
 
@@ -282,8 +241,8 @@ local({
   # within the construct of the model, there are no time-dependent variables, only time-dependent treatment effects (modelled as piecewise constant hazards)
 
 
-  max_trial_day <- pmin(max(data_tte$tte_treatment, na.rm=TRUE), study_dates$lastvax3_date - study_dates$studystart_date, na.rm=TRUE)
-  trials <- seq_len(max_trial_day)
[…]

### 2022-02-22 manuscript/draft-manuscript.Rmd
addressing comments

@@
 matchsummary_control <- read_csv(fs::path(output_dir_os, "matchsummary_treated.csv")) %>% filter(treated == 0L)
+matchflowchart <- read_csv(fs::path(output_dir_os, "matchflowchart.csv"))
 matchcoverage <- read_csv(fs::path(output_dir_os, "matchcoverage.csv"))
@@
 We matched `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` BNT162b2 booster recipients to unboosted controls with only two doses.
-During days 14-20 after the booster, the estimated hazard ratio (95% CI) comparing a BNT162b2 booster dose to two doses only was `r filter(hazards_none, outcome=="postest", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for documented SARS-CoV-2 infection, `r filter(hazards_none, outcome=="covidadmitted", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(hazards_none, outcome=="coviddeath", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 death.
+During days 14-20 after the booster, the estimated hazard ratio (95% CI) comparing a BNT162b2 booster dose to two doses only was `r filter(hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for documented SARS-CoV-2 infection, `r filter(hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 death.
 
@@
 
-<!-- **Keywords** COVID-19; Vaccine effectiveness; Target trial; EHR data -->
+**Keywords** COVID-19; Booster vaccine effectiveness; Target trial; EHR data
 
@@
 
-<!-- * flowchart incl matching -->
-
 <!-- * check for moderna references and remove if necessary -->
@@
 <!-- * check if rolling average restrictions are ever used (need to output dataset) -->
-
-<!-- * JCVI group spell out -->
 
@@
 
-A total of `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_treated))` BNT162b2 booster doses were administered between `r format(as.Date(matchsummary_treated$firstrecruitdate), "%d %B %Y")` and `r format(as.Date(matchsummary_treated$lastrecruitdate), "%d %B %Y")` in adults registered at a TPP practice, with `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_eligible))` (`r label_percent(0.1)(filter(matchsummary, treatment=="pfizer") %>% pull(prop_eligible_treated))`) eligible for inclusion into the treatment group.
-Of these, `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` (`r label_percent(0.1)(filter(matchsummary, treatment=="pfizer") %>% pull(prop_matched_eligible))`) were successfully matched with unboosted controls (Figure 1).
+A total of `r label_number(1, big.mark=",")(filter(matchflowchart, crit=="c0", treatment=="pfizer") %>% pull(n))` BNT162b2 booster recipients were identified between `r format(as.Date(matchsummary_treated$firstrecruitdate), "%d %B %Y")` and `r format(as.Date(matchsummary_treated$lastrecruitdate), "%d %B %Y")` in adults registered at a TPP practice, with `r label_number(1, big.mark=",")(filter(matchflowchart, treatment=="pfizer", crit=="c7") %>% pull(n))` (`r label_percent(0.1)(filter(matchflowchart, treatment=="pfizer", crit=="c7") %>% pull(pct_all))`) eligible for inclusion into the treatment group.
+Of these, `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` (`r label_percent(0.1)(filter(matchsummary, treatment=="pfizer") %>% pull(prop_matched_eligible))`) were successfully matched with unboosted controls (Figure S1).
 
@@
 
-Participant characteristics at trial recruitment between treatment groups (Table 1) were typically well-balanced (necessarily so on matched characteristics), though with some exceptions, for instance unboosted controls had higher levels of deprivation, higher rates of learning disabilities and severe mental illness, and lower SARS-CoV-2 testing frequency.
-These unbalanced characteristics were adjusted for in the Cox regression models.
+Participant characteristics at trial recruitment between treatment groups (Table 1) were typically well-balanced though with some exceptions.
+For instance, unboosted controls had higher levels of deprivation, higher rates of learning disabilities and severe mental illness, and lower SARS-CoV-2 testing frequency.
 
@@
     rate_0=if_else(rate_0<0.001, "<0.001", label_number(0.001)(rate_0)),
-  )
+  ) %>%
   gt(
@@
 
-In the 14-20 days period of follow-up, the estimated effectiveness (95% CI) comparing a BNT162b2 booster dose to two doses only, was `r filter(hazards_none, outcome=="postest", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for documented infection, `r filter(hazards_none, outcome=="covidadmitted", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(hazards_none, outcome=="coviddeath", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 death (Figure 2).
+In the 14-20 days period of follow-up, the estimated effectiveness (95% CI) comparing a BNT162b2 booster dose to two doses only, was `r filter(hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for documented infection, `r filter(hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 death (Figure 2).
 
-<!-- `r filter(hazards_none, outcome=="covidattendance", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 A&E attendances, `r filter(hazards_none, outcome=="covidcc", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 critical care admission,  -->
+<!-- `r filter(hazards, subgroup=="none", outcome=="covidattendance", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 A&E attendances, `r filter(hazards, subgroup=="none", outcome=="covidcc", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 critical care admission,  -->
 
-<!-- For mRNA-1273, the corresponding estimated effectiveness were `r filter(hazards_none, outcome=="postest", treatment=="moderna", model==3, term_left==14) %>% pull(ve_display)`, for COVID-19 A&E attendances was `r filter(hazards_none, outcome=="covidattendance", treatment=="moderna", model==3, term_left==14) %>% pull(ve_display)`, for COVID-19 hospital admission was `r filter(hazards_none, outcome=="covidadmitted", treatment=="moderna", term_left==14) %>% pull(ve_display)`, for COVID-19 critical care admission was `r filter(hazards_none, outcome=="covidcc", treatment=="moderna", model==3, term_left==14) %>% pull(ve_display)`, and for COVID-19 death was `r filter(hazards_none, outcome=="coviddeath", treatment=="moderna", model==3, term_left==14) %>% pull(ve_display)`.
+<!-- For mRNA-1273, the corresponding estimated effectiveness were `r filter(hazards, subgroup=="none", outcome=="postest", treatment=="moderna", model==3, term_left==14) %>% pull(ve_display)`, for COVID-19 A&E attendances was `r filter(hazards, subgroup=="none", outcome=="covidattendance", treatment=="moderna", model==3, term_left==14) %>% pull(ve_display)`, for COVID-19 hospital admission was `r filter(hazards, subgroup=="none", outcome=="covidadmitted", treatment=="moderna", term_left==14) %>% pull(ve_display)`, for COVID-19 critical care admission was `r filter(hazards, subgroup=="none", outcome=="covidcc", treatment=="moderna", model==3, term_left==14) %>% pull(ve_display)`, and for COVID-19 death was `r filter(hazards, subgroup=="none", outcome=="coviddeath", treatment=="moderna", model==3, term_left==14) %>% pull(ve_display)`.
 -->
 
-For the same period, the estimated effectiveness (95% CI) comparing a BNT162b2 booster dose to two doses only separately for 2× BNT162b2 and 2× ChAdOx1-S primary course recipients respectively were `r filter(hazards_vax12_type, outcome=="postest", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3, term_left==14) %>% pull(ve_display)` and `r filter(hazards_vax12_type, outcome=="postest", treatment=="pfizer", subgroup_level=="az-az", model==3, term_left==14) %>% pull(ve_display)` for documented infection, `r filter(hazards_vax12_type, outcome=="covidadmitted", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3, term_left==14) %>% pull(ve_display)` and `r filter(hazards_vax12_type, outcome=="covidadmitted", treatment=="pfizer", subgroup_level=="az-az", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(hazards_vax12_type, outcome=="coviddeath", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3, term_left==14) %>% pull(ve_display)` and `r filter(hazards_vax12_type, outcome=="coviddeath", treatment=="pfizer", subgroup_level=="az-az", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 death (Figure 2).
+For the same period, the estimated effectiveness (95% CI) comparing a BNT162b2 booster dose to two doses only separately for 2× BNT162b2 and 2× ChAdOx1-S primary course recipients respectively were `r filter(hazards, outcome=="postest", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3, term_left==14) %>% pull(ve_display)` and `r filter(hazards, outcome=="postest", treatment=="pfizer", subgroup_level=="az-az", model==3, term_left==14) %>% pull(ve_display)` for documented infection, `r filter(hazards, outcome=="covidadmitted", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3, term_left==14) %>% pull(ve_display)` and `r filter(hazards, outcome=="covidadmitted", treatment=="pfizer", subgroup_level=="az-az", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(hazards, outcome=="coviddeath", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3, term_left==14) %>% pull(ve_display)` and `r filter(hazards, outcome=="coviddeath", treatment=="pfizer", subgroup_level=="az-az", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 death (Figure 2).
 

### 2022-02-22 analysis/combine_match.R

@@ -29,7 +29,7 @@ library('gt')
 source(here("lib", "functions", "utility.R"))
 source(here("lib", "functions", "survival.R"))
 
-output_dir <- here("output", "models", "seqtrialcox", "combined", "match")
+output_dir <- here("output", "match", "combined")
 fs::dir_create(output_dir)
 
 ## import globally defined study dates and convert to "Date"
@@ -47,6 +47,8 @@ match_metaparams <-
   tibble(
     treatment = factor(recode_treatment),
     treatment_descr = fct_recode(treatment,  !!!recode_treatment),
+  ) %>% filter(
+    treatment=="pfizer"
   )
 
 # combine matching ----
@@ -56,7 +58,7 @@ match_metaparams <-
 matchcoverage <-
   match_metaparams %>%
   mutate(
-    coverage = map(treatment, ~read_csv(here("output", "models", "seqtrialcox", .x, glue("match_data_coverage.csv"))))
+    coverage = map(treatment, ~read_csv(here("output", "match", .x, glue("merge_data_coverage.csv")), guess_max = 999999))
   ) %>%
   unnest(coverage) %>%
   arrange(treatment, vax12_type, status, vax3_date) %>%
@@ -80,7 +82,7 @@ write_csv(matchcoverage, fs::path(output_dir, "coverage.csv"))
 matchsummary <-
   match_metaparams %>%
   mutate(
-    summary = map(treatment, ~read_csv(here("output", "models", "seqtrialcox", .x, glue("match_summary.csv"))))
+    summary = map(treatment, ~read_csv(here("output",  "match", .x, glue("merge_summary.csv"))))
   ) %>%
   unnest(summary)
 
@@ -90,7 +92,7 @@ write_csv(matchsummary, fs::path(output_dir, "summary.csv"))
 matchsummary_treated <-
   match_metaparams %>%
   mutate(
-    summary = map(treatment, ~read_csv(here("output", "models", "seqtrialcox", .x, glue("match_summary_treated.csv"))))
+    summary = map(treatment, ~read_csv(here("output",  "match", .x, glue("merge_summary_treated.csv"))))
   ) %>%
   unnest(summary)
 
@@ -103,7 +105,7 @@ write_csv(matchsummary_treated, fs::path(output_dir, "summary_treated.csv"))
 table1 <-
   match_metaparams %>%
   mutate(
[…]

### 2022-02-22 analysis/data_process.R

@@ -220,7 +220,7 @@ data_processed <- data_extract %>%
       "55-59 years"="8",
       "50-54 years"="9",
       "40-49 years"="10a",
-      "16-39 years"="10b",
+      "16-39 years"="10b"
     ),
 
 

### 2022-02-22 analysis/match_seqtrialcox.R

@@ -3,7 +3,7 @@
 # This script:
 # imports processed data
 # chooses matching sets for each sequential trial
-# reports on matching quality
+# outputs matching summary
 #
 # The script must be accompanied by one arguments:
 # `treatment` - the exposure in the regression model, pfizer or moderna
@@ -56,7 +56,7 @@ if(Sys.getenv("OPENSAFELY_BACKEND") %in% c("", "expectations")){
 
 # create output directories ----
 
-output_dir <- here("output", "models", "seqtrialcox", treatment)
+output_dir <- here("output", "match", treatment)
 fs::dir_create(output_dir)
 
 ## create special log file ----
@@ -149,6 +149,7 @@ data_rollingstrata_vaxcount <-
   ) %>%
   filter(vax3_time>=0)
 
+write_rds(data_rollingstrata_vaxcount, fs::path(output_dir, "match_data_rollingstrata_vaxcount.rds"))
 
 data_nontimevarying <-
   data_cohort %>%
@@ -165,7 +166,6 @@ data_nontimevarying <-
     treated_within_recruitment_period = replace_na(recruit, FALSE),
     recruit=NULL,
   )
-
 
 if(removeobjects) rm(data_rollingstrata_vaxcount)
 
@@ -225,6 +225,48 @@ data_tte <-
 write_rds(data_tte, fs::path(output_dir, "match_data_tte.rds"))
 
 logoutput_datasize(data_tte)
+
+
+## baseline variables ----
+
+data_baseline <-
+  data_cohort %>%
+  transmute(
+    patient_id,
+    age,
+    ageband,
+    sex,
+    ethnicity_combined,
+    imd_Q5,
+    region,
+    jcvi_group,
+    jcvi_group_descr,
+    prior_tests_cat,
+    multimorb,
+    learndis,
+    sev_mental,
+    cev,
+    sev_obesity,
+    chronic_heart_disease,
+    chronic_kidney_disease,
+    diabetes,
+    chronic_liver_disease,
+    chronic_resp_disease,
+    asthma,
+    chronic_neuro_disease,
+    prior_covid_infection0,
+    immunosuppressed,
+    asplenia,
+    immuno,
+
+    vax12_type,
+    vax12_type_descr,
+    vax2_week,
+    vax3_date,
+    vax3_type,
+
+  )
+logoutput_datasize(data_baseline)
 
 if(removeobjects) rm(data_cohort)
 
@@ -400,7 +442,7 @@ local({
[…]

### 2022-02-22 analysis/model_seqtrialcox.R

@@ -22,7 +22,7 @@ if(length(args)==0){
   # use for interactive testing
   removeobjects <- FALSE
   treatment <- "pfizer"
-  outcome <- "covidemergency"
+  outcome <- "postest"
   subgroup <- "none"
   #subgroup <- "none"
 } else {
@@ -53,13 +53,6 @@ subgroup_dummy <- paste0(subgroup_variable,"_",subgroup_level)
 
 postbaselinecuts <- read_rds(here("lib", "design", "postbaselinecuts.rds"))
 matching_variables <- read_rds(here("lib", "design", "matching_variables.rds"))
-
-if(Sys.getenv("OPENSAFELY_BACKEND") %in% c("", "expectations")){
-  recruitment_period_cutoff <- 0.1
-} else{
-  recruitment_period_cutoff <- read_rds(here("lib", "design", "recruitment_period_cutoff.rds"))
-}
-
 
 # create output directories ----
 
@@ -105,13 +98,18 @@ outcome_var <- events$event_var[events$event==outcome]
 
 
 data_merged <-
-  read_rds(here("output", "models", "seqtrialcox", treatment, "match_data_merged.rds")) %>%
-  mutate(none="") %>%
+  read_rds(here("output", "match", treatment, "match_data_merged.rds")) %>%
+  filter(matched %in% 1L) %>%
+  mutate(
+    none="",
+    treated_patient_id = paste0(treated, "_", patient_id),
+    fup = pmin(tte_stop - tte_recruitment, last(postbaselinecuts)),
+  ) %>%
   fastDummies::dummy_cols(select_columns = subgroup_variable, remove_selected_columns = FALSE) %>%
   filter(.[[subgroup_dummy]]==1L)
 
 data_tte <-
-  read_rds(here("output", "models", "seqtrialcox", treatment, "match_data_tte.rds")) %>%
+  read_rds(here("output", "match", treatment, "match_data_tte.rds")) %>%
   filter(patient_id %in% data_merged$patient_id)
 
 ## create outcome variables ----
@@ -228,7 +226,6 @@ data_seqtrialcox <- local({
 
 logoutput_datasize(data_seqtrialcox)
 
-
 write_rds(data_seqtrialcox, fs::path(output_dir, "model_data_seqtrialcox.rds"))
 
 # outcome frequency
@@ -251,7 +248,7 @@ formula_vaxonly <- as.formula(
 
[…]

### 2022-02-22 analysis/report_seqtrialcox.R

@@ -22,7 +22,7 @@ if(length(args)==0){
   removeobs <- FALSE
   treatment <- "pfizer"
   outcome <- "postest"
-  subgroup <- "vax12_type-pfizer-pfizer"
+  subgroup <- "none"
 } else {
   removeobs <- TRUE
   treatment <- args[[1]]

### 2022-02-22 create-project.R

@@ -67,15 +67,29 @@ action_match <- function(treatment){
       arguments = c(treatment),
       needs = list("data_selection", "data_process_long"),
       highly_sensitive = lst(
-        rds = glue("output/models/seqtrialcox/{treatment}/match_*.rds")
+        rds = glue("output/match/{treatment}/match_*.rds")
       ),
       moderately_sensitive = lst(
-        txt = glue("output/models/seqtrialcox/{treatment}/match_*.txt"),
-        csv = glue("output/models/seqtrialcox/{treatment}/match_*.csv"),
-        #svg = glue("output/models/seqtrialcox/{treatment}/match_*.svg"),
-        png = glue("output/models/seqtrialcox/{treatment}/match_*.png"),
-        pdf = glue("output/models/seqtrialcox/{treatment}/match_*.pdf"),
-        html = glue("output/models/seqtrialcox/{treatment}/match_*.html")
+        txt = glue("output/match/{treatment}/match_*.txt"),
+        #csv = glue("output/match/{treatment}/match_*.csv"),
+      )
+    ),
+
+    action(
+      name = glue("merge_seqtrialcox_{treatment}"),
+      run = glue("r:latest analysis/merge_seqtrialcox.R"),
+      arguments = c(treatment),
+      needs = list("data_selection", "data_process_long", glue("match_seqtrialcox_{treatment}")),
+      highly_sensitive = lst(
+        rds = glue("output/match/{treatment}/merge_*.rds")
+      ),
+      moderately_sensitive = lst(
+        txt = glue("output/match/{treatment}/merge_*.txt"),
+        csv = glue("output/match/{treatment}/merge_*.csv"),
+        #svg = glue("output/match/{treatment}/merge_*.svg"),
+        png = glue("output/match/{treatment}/merge_*.png"),
+        pdf = glue("output/match/{treatment}/merge_*.pdf"),
+        html = glue("output/match/{treatment}/merge_*.html")
       )
     )
   )
@@ -395,17 +409,18 @@ actions_list <- splice(
       as.list(
         glue_data(
           .x=expand_grid(
+            script = c("match", "merge"),
             treatment=c("pfizer", "moderna")
           ),
-          "match_seqtrialcox_{treatment}"
+          "{script}_seqtrialcox_{treatment}",
         )
       )
     ),
     moderately_sensitive = lst(
-      csv = "output/models/seqtrialcox/combined/match/*.csv",
[…]

### 2022-02-22 analysis/model_seqtrialcox.R
add model 2

@@ -451,7 +451,7 @@ cox_model <- function(number, formula_cox){
 
 summary0 <- cox_model(0, formula0_pw)
 summary1 <- cox_model(1, formula1_pw)
-#summary2 <- cox_model(2, formula2_pw)
+summary2 <- cox_model(2, formula2_pw)
 summary3 <- cox_model(3, formula3_pw)
 
 # combine results
@@ -459,7 +459,7 @@ model_glance <-
   bind_rows(
     summary0$glance,
     summary1$glance,
-    #summary2$glance,
+    summary2$glance,
     summary3$glance,
   ) %>%
   mutate(
@@ -471,7 +471,7 @@ model_tidy <-
   bind_rows(
     summary0$tidy,
     summary1$tidy,
-    #summary2$tidy,
+    summary2$tidy,
     summary3$tidy,
   ) %>%
   mutate(

### 2022-02-22 manuscript/draft-supplement.Rmd

@@
 matchcoverage <- read_csv(fs::path(output_dir_os, "matchcoverage.csv"))
+matchflowchart <- read_csv(fs::path(output_dir_os, "matchflowchart.csv"))
 incidences <- read_csv(fs::path(output_dir_os, "incidence.csv")) %>% mutate(events=events_1+events_0)
@@
   mutate(
+    model_descr = fct_inorder(model_descr),
     subgroup_descr = fct_inorder(subgroup_descr),
@@
 ```{r flowchart, echo=FALSE, warning=FALSE, message=FALSE}
-flowchart %>%
+
+matchflowchart %>%
   transmute(
-    Participants = str_wrap(criteria, 40),
-    N = n,
-    `N excluded` = n_exclude,
-    #`% excluded` = pct_exclude,
-    `% remaining` = pct_all
+    treatment,
+    stage,
+    criteria = str_wrap(criteria, 40),
+    n,
+    n_exclude,
+    pct_all
   ) %>%
-  # mutate(
-  #   Participants=if_else(row_number()==1,"HCWs aged 18-64 receiving BNT162b2 or ChAdOx1 between 4 Jan and 28 Feb",  Participants)
+  pivot_wider(
+    id_cols = c(stage, criteria),
+    names_from = treatment,
+    values_from = c(n, n_exclude, pct_all)
+  ) %>%
+  gt() %>%
+  cols_label(
+    stage="",
+    criteria = "",
+
+    n_pfizer = "N",
+    #n_moderna = "N",
+
+    n_exclude_pfizer = "N excluded",
+    #n_exclude_moderna = "N excluded",
+
+    pct_all_pfizer = "% remaining",
+    #pct_all_moderna = "% remaining",
+  ) %>%
+  tab_spanner(
+    label = "BNT162b2",
+    columns = ends_with("pfizer")
+  ) %>%
+  # tab_spanner(
+  #   label = "mRNA-1273",
+  #   columns = ends_with("moderna")
   # ) %>%
-  gt() %>%
   fmt_percent(
-    columns = starts_with(c("%")),
+    columns = starts_with(c("pct_")),
     decimals = 1
@@
   fmt_number(
-    columns = starts_with(c("N")),
+    columns = starts_with(c("n_")),
     decimals= 0
@@
     #style=cell_text(style ="italic"),
-    locations = cells_body(columns = "Participants")
+    locations = cells_body(columns = "criteria")
   )

### 2022-02-23 analysis/combine_model.R
add meta-analysed HRs

@@ -33,7 +33,7 @@ library('gt')
 source(here("lib", "functions", "utility.R"))
 source(here("lib", "functions", "survival.R"))
 
-output_dir <- here("output", "models", "seqtrialcox", "combined", "model", subgroup_variable)
+output_dir <- here("output", "models", "seqtrialcox", "combined", subgroup_variable)
 fs::dir_create(output_dir)
 
 ## import globally defined study dates and convert to "Date"
@@ -136,9 +136,36 @@ model_effects <-
     hr.ul !=Inf, hr.ul !=0,
   )
 
-
-
 write_csv(model_effects, path = fs::path(output_dir, "effects.csv"))
+
+
+
+model_metaeffects <-
+  model_metaparams %>%
+  mutate(
+    metaeffects = pmap(list(treatment, outcome, subgroup), function(x, y, z) read_csv(here("output", "models", "seqtrialcox", x, y, z, glue("report_metaeffects.csv"))))
+  ) %>%
+  unnest(metaeffects) %>%
+  mutate(
+    model_descr = fct_reorder(model_descr, model),
+    hr = exp(estimate),
+    hr.ll = exp(conf.low),
+    hr.ul = exp(conf.high),
+    ve = 1-hr,
+    ve.ll = 1-hr.ul,
+    ve.ul = 1-hr.ll
+  ) %>%
+  filter(
+    # this filtering is necessary for dummy data,
+    # so that sec.axis doesn't think `1-x` is a non-monotonic function
+    # see - https://github.com/tidyverse/ggplot2/issues/3323#issuecomment-491421372
+    hr !=Inf, hr !=0,
+    hr.ll !=Inf, hr.ll !=0,
+    hr.ul !=Inf, hr.ul !=0,
+  )
+
+write_csv(model_metaeffects, path = fs::path(output_dir, "metaeffects.csv"))
+
 
 formatpercent100 <- function(x,accuracy){
   formatx <- scales::label_percent(accuracy)(x)

### 2022-02-23 analysis/manuscript_objects.R

@@ -162,9 +162,10 @@ fs::file_copy(here("output", "match", "combined", "flowchart.csv"), here("output
 
 for(subgroup_variable in c("none", "vax12_type")){
   fs::dir_create(here("output", "manuscript-objects", subgroup_variable))
-  fs::file_copy(here("output", "models", "seqtrialcox", "combined", "model", subgroup_variable, "incidence.csv"), here("output", "manuscript-objects", subgroup_variable, "incidence.csv"), overwrite = TRUE)
-  fs::file_copy(here("output", "models", "seqtrialcox", "combined", "model", subgroup_variable, "km.csv"), here("output", "manuscript-objects", subgroup_variable, "km.csv"), overwrite = TRUE)
-  fs::file_copy(here("output", "models", "seqtrialcox", "combined", "model", subgroup_variable, "effects.csv"), here("output", "manuscript-objects", subgroup_variable, "effects.csv"), overwrite = TRUE)
+  fs::file_copy(here("output", "models", "seqtrialcox", "combined", subgroup_variable, "incidence.csv"), here("output", "manuscript-objects", subgroup_variable, "incidence.csv"), overwrite = TRUE)
+  fs::file_copy(here("output", "models", "seqtrialcox", "combined", subgroup_variable, "km.csv"), here("output", "manuscript-objects", subgroup_variable, "km.csv"), overwrite = TRUE)
+  fs::file_copy(here("output", "models", "seqtrialcox", "combined", subgroup_variable, "effects.csv"), here("output", "manuscript-objects", subgroup_variable, "effects.csv"), overwrite = TRUE)
+  fs::file_copy(here("output", "models", "seqtrialcox", "combined", subgroup_variable, "metaeffects.csv"), here("output", "manuscript-objects", subgroup_variable, "metaeffects.csv"), overwrite = TRUE)
 }
 
 

### 2022-02-23 analysis/report_seqtrialcox.R

@@ -42,7 +42,7 @@ source(here("lib", "functions", "utility.R"))
 source(here("lib", "functions", "survival.R"))
 source(here("lib", "functions", "redaction.R"))
 
-output_dir_matched <- here("output", "models", "seqtrialcox", treatment, outcome)
+output_dir_matched <- here("output", "match", outcome)
 output_dir <- here("output", "models", "seqtrialcox", treatment, outcome, subgroup)
 
 data_seqtrialcox <- read_rds(fs::path(output_dir, "model_data_seqtrialcox.rds"))
@@ -146,10 +146,6 @@ format_ratio = function(numer,denom, width=7){
     str_pad(replace_na(scales::comma_format(accuracy=1)(denom),"--"), width=width, pad=" ")
   )
 }
-
-
-
-
 
 
 incidence_rate_redacted <- local({
@@ -407,5 +403,25 @@ if(any(data_tidy_surv$n.event==0)){
   ggsave(filename=fs::path(output_dir, "report_cumulriskplot.png"), plot_cumulrisk, width=20, height=15, units="cm")
 
 }
+
+
+## meta analysis of period-specific hazards ----
+
+
+model_metaeffects <- model_effects %>%
+  group_by(
+    model, model_descr
+  ) %>%
+  summarise(
+    estimate = weighted.mean(estimate, robust.se^-2),
+    robust.se = sqrt(1/sum(robust.se^-2)),
+    statistic = estimate/robust.se,
+    p.value = 2 * pmin(pnorm(statistic), pnorm(-statistic)),
+    conf.low = estimate + qnorm(0.025)*robust.se,
+    conf.high = estimate + qnorm(0.975)*robust.se,
+  )
+
+write_csv(model_metaeffects, path = fs::path(output_dir, "report_metaeffects.csv"))
+
 
 

### 2022-02-23 analysis/match_seqtrialcox.R
relabel trial status

@@ -470,7 +470,7 @@ logoutput("max trial day is ", max_trial_time)
 
 ## combine all data together as at recruitment date ----
 
-status_recode <- c(`Treated, ineligible` = "ineligible", `Treated, eligible, unmatched`= "unmatched", `Treated, eligible, matched` = "matched", `Control` = "control")
+status_recode <- c(`Boosted, ineligible` = "ineligible", `Boosted, eligible, unmatched`= "unmatched", `Boosted, eligible, matched` = "matched", `Control` = "control")
 
 ## create variables-at-time-zero dataset - one row per trial per arm per patient ----
 data_merged <-

### 2022-02-23 analysis/merge_seqtrialcox.R

@@ -104,7 +104,7 @@ logoutput_datasize(data_matched)
 # matching coverage per trial / day of follow up
 
 
-status_recode <- c(`Treated, ineligible` = "ineligible", `Treated, eligible, unmatched`= "unmatched", `Treated, eligible, matched` = "matched", `Control` = "control")
+status_recode <- c(`Boosted, ineligible` = "ineligible", `Boosted, eligible, unmatched`= "unmatched", `Boosted, eligible, matched` = "matched", `Control` = "control")
 
 
 # matching coverage for boosted people
@@ -435,7 +435,7 @@ data_smd <- tab_summary_baseline$meta_data %>%
   select(var_label, df_stats) %>%
   unnest(df_stats) %>%
   filter(
-    by %in% c("Treated, eligible, matched", "Control"),
+    by %in% c("Boosted, eligible, matched", "Control"),
     variable != "N"
   ) %>%
   group_by(var_label, variable_levels) %>%

### 2022-02-28 analysis/match_seqtrialcox.R
Update match_seqtrialcox.R

@@ -263,7 +263,7 @@ data_baseline <-
     vax12_type,
     vax12_type_descr,
     vax2_week,
-    va2_day,
+    vax2_day,
     vax3_date,
     vax3_type,
 

### 2022-02-28 manuscript/draft-supplement.Rmd
update supplement

@@
 incidences <- read_csv(fs::path(output_dir_os, "incidence.csv")) %>% mutate(events=events_1+events_0)
+smd <- read_csv(fs::path(output_dir_os, "matchsmd.csv"))
 
@@
 
-hazards_none <- read_csv(fs::path(output_dir_os, "none", "effects.csv")) %>%
+km_none <- read_csv(fs::path(output_dir_os, "none", "km.csv"))  
+km_vax12_type <- read_csv(fs::path(output_dir_os, "vax12_type", "km.csv")) 
+km <- 
+  bind_rows(km_none, km_vax12_type) %>%
   mutate(
+    subgroup_descr = fct_inorder(subgroup_descr),
+    subgroup_level = fct_inorder(subgroup_level),
+    subgroup_level_descr = fct_inorder(fct_explicit_na(subgroup_level_descr, "All")),
     treatment_descr = fct_inorder(treatment_descr),
     outcome_descr = fct_inorder(outcome_descr),
+    outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=20)),
     #model_name = fct_inorder(str_wrap(model_name, 15)),
-    hr_display = paste0(number_format(accuracy=0.01)(hr), " (", number_format(accuracy=0.01)(hr.ll), "-", number_format(accuracy=0.01)(hr.ul), ")")
   )
 
-hazards_vax12_type <- read_csv(fs::path(output_dir_os, "vax12_type", "effects.csv")) %>%
+hazards_none <- read_csv(fs::path(output_dir_os, "none", "effects.csv"))
+hazards_vax12_type <- read_csv(fs::path(output_dir_os, "vax12_type", "effects.csv"))
+hazards <- 
+  bind_rows(hazards_none, hazards_vax12_type) %>%
   mutate(
-    model_descr = fct_inorder(model_descr),
+
     subgroup_descr = fct_inorder(subgroup_descr),
-    subgroup_level_descr = fct_inorder(subgroup_level_descr),
+    subgroup_level = fct_inorder(subgroup_level),
+    subgroup_level_descr = fct_inorder(fct_explicit_na(subgroup_level_descr, "All")),
     treatment_descr = fct_inorder(treatment_descr),
     outcome_descr = fct_inorder(outcome_descr),
+    outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=20)),
     #model_name = fct_inorder(str_wrap(model_name, 15)),
-    hr_display = paste0(number_format(accuracy=0.01)(hr), " (", number_format(accuracy=0.01)(hr.ll), "-", number_format(accuracy=0.01)(hr.ul), ")")
+    hr_display = paste0(label_number(accuracy=0.01)(hr), " (", label_number(accuracy=0.01)(hr.ll), "-", label_number(accuracy=0.01)(hr.ul), ")"),
+    ve_display = paste0(label_percent(accuracy=0.1)(ve), " (", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
   )
@@
 
+### Figure S1: Standardised Mean Differences between Boosted and Unboosted groups
+
+```{r plot_smd, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}
+
+plot_smd <-
+  ggplot(data_smd)+
+  geom_point(aes(x=smd, y=level))+
+  geom_rect(aes(alpha = variable_card, ymin = rev(cardn)-0.5, ymax =rev(cardn+0.5)), xmin = -Inf, xmax = Inf, fill='grey', colour="transparent") +
+  scale_alpha_continuous(range=c(0,0.3), guide=FALSE)+
+  labs(
+    x="Standardised mean difference",
+    y=NULL,
+    alpha=NULL
+  )+
+  theme_minimal() +
+  theme(
+    strip.placement = "outside",
+    strip.background = element_rect(fill="transparent", colour="transparent"),
+    strip.text.y.left = element_text(angle = 0, hjust=1),
+    panel.grid.major.y = element_blank(),
+    panel.grid.minor.y = element_blank(),
+    panel.spacing = unit(0, "lines")
+  )
+
+plot_smd
+
+```
+
+### Figure S2: Kaplan-Meier cumulative incidence curves
+
+```{r plot_km, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}
+
+plot_km <- km %>%
+    ggplot()+
+    geom_step(aes(x=time, y=(1-surv)*1000, group=treated_descr, colour=treated_descr))+
+    geom_rect(aes(xmin=time, xmax=leadtime, ymin=(1-surv.ll)*1000, ymax=(1-surv.ul)*1000, group=treated_descr, fill=treated_descr), alpha=0.1)+
+    geom_hline(aes(yintercept=0), colour='black')+
+    facet_grid(rows=vars(outcome_descr_wrap), cols=vars(subgroup_level_descr), switch="y", scales="free_y")+
+    scale_x_continuous(
+      breaks = seq(0,7*max(postbaselinecuts),by=14),
+      expand = expansion(0)
+    )+
+    scale_y_continuous(
+      expand = expansion(0),
+      #limits = c(0L,NA)
+    )+
+    scale_colour_brewer(type="qual", palette="Set1", na.value="grey")+
+    scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey")+
+    labs(
+      x=NULL,
+      y="Cumulative incidence\nper 1000",
+      colour=NULL
+    )+
+    theme_minimal()+
+    theme(
+      legend.position=c(0.15,0.85),
+      legend.justification = c(1,0),
+      legend.direction = "vertical",
+      axis.text.x.top=element_text(hjust=0),
+
+      strip.placement = "outside",
+      strip.background = element_blank(),
+      strip.text.y.left = element_text(angle = 0),
+
+      panel.border = element_blank(),
+      axis.line.y = element_line(colour = "black"),
+      axis.line.x.top = element_line(colour = "black"),
+      panel.grid.minor.x = element_blank(),
+      panel.grid.minor.y = element_blank(),
+
+      panel.spacing = unit(0.8, "lines"),
+
+    )+
+    NULL
+
+plot_km
+
+```
+
+\newpage
+
+## Figure S3: Estimated booster effectiveness, all models
+
 ```{r plot_effects, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}
@@
 
-# plot_cmlinc <- cmlinc3 %>%
-#     ggplot()+
-#     geom_step(aes(x=tstop, y=(1-survival)*1000, group=vax1_az_descr, colour=vax1_az_descr))+
-#     geom_rect(aes(xmin=tstop, xmax=lead_tstop, ymin=(1-survival.ll)*1000, ymax=(1-survival.ul)*1000, group=vax1_az_descr, fill=vax1_az_descr), alpha=0.1)+
-#     geom_hline(aes(yintercept=0), colour='black')+
-#     facet_wrap(facets=vars(outcome_descr), nrow=1, scales="free_y")+
-#     scale_x_continuous(
-#       breaks = seq(0,7*52,by=28),
-#       expand = expansion(0)
-#     )+
-#     scale_y_continuous(
-#       expand = expansion(0),
-#       #limits = c(0L,NA)
-#     )+
-#     scale_colour_brewer(type="qual", palette="Set1", na.value="grey")+
-#     scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey")+
-#     labs(
-#       x=NULL,
-#       y="Marginalised\ncumul. incidence\nper 1000",
-#       colour=NULL
-#     )+
-#     theme_minimal(base_size = 9)+
-#     theme(
-#       legend.position=c(0.05,0.1),
-#       legend.justification = c(0,0),
-#       legend.direction = "vertical",
-#       axis.text.x.top=element_text(hjust=0),
-# 
-#       panel.border = element_blank(),
-#       axis.line.y = element_line(colour = "black"),
-#       axis.line.x.top = element_line(colour = "black"),
-# 
-#       panel.grid.minor.x = element_blank(),
-#       panel.grid.minor.y = element_blank(),
-#       strip.background = element_blank(),
-# 
-# 
-#       panel.spacing = unit(0.8, "lines"),
-# 
-#     )+
-#     NULL
-#   
-#   ## difference in cumulative incidence
-#   
-#   plot_riskdiff <-
-#     riskdiff3 %>%
-#     ggplot()+
-#     geom_hline(aes(yintercept=0), colour='black')+
-#     geom_step(aes(x=tstop, y=diff*1000), colour='orchid4')+
-#     geom_rect(aes(xmin=tstop, xmax=lead_tstop, ymin=diff.ll*1000, ymax=diff.ul*1000), alpha=0.1, colour="transparent", fill='orchid4')+
-#     facet_wrap(facets=vars(outcome_descr), nrow=1, scales="free_y")+
-#     scale_x_continuous(
-#       limits = c(0, NA),
-#       breaks = seq(0,7*52,by=28),
-#       expand = expansion(0)
-#     )+
-#     scale_y_continuous(
-#       expand = expansion(0),
-#       sec.axis = dup_axis(name="Higher incidence after \n <- ChAdOx1 / BNT162b2 ->", breaks = NULL)
-#     )+
-#     coord_cartesian(ylim= c(-3, 3))+
-#     scale_colour_brewer(type="qual", palette="Set1", na.value="grey")+
-#     scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey")+
-#     labs(
-#       x=NULL,
-#       y="Difference\nin cumul. incidence\nper 1000",
-#       colour=NULL
-#     )+
-#     theme_minimal(base_size = 9)+
-#     theme(
-#       axis.line.y.left = element_line(colour = "black"),
-# 
-#       panel.grid.minor.x = element_blank(),
-#       panel.grid.minor.y = element_blank(),
-#       strip.background = element_blank(),
-#       strip.text = element_blank(),
-# 
-#       panel.spacing = unit(0.8, "lines"),
-# 
-#       plot.title = element_text(hjust = 0),
-#       plot.title.position = "plot",
-#       plot.caption.position = "plot",
-#       plot.caption = element_text(hjust = 0, face= "italic")
-# 
-#      ) +
-#    NULL
-# 
-#
-
   ## hazard ratio
@@
   plot_hazards <-
-    hazards_vax12_type %>%
-    filter(outcome %in% c("postest", "covidadmitted", "coviddeath")) %>%
+    hazards %>%
+    filter(
+      outcome %in% outcomes,
+      treatment %in% c("pfizer")
+    ) %>%
     ggplot()+
@@
     geom_hline(aes(yintercept=1), colour='grey')+
-    facet_grid(rows=vars(outcome_descr), cols=vars(subgroup_level_descr), switch="y")+
+    facet_grid(rows=vars(outcome_descr_wrap), cols=vars(subgroup_level_descr), switch="y")+
     scale_y_log10(
@@
      ) +
-    theme_bw()+
+    theme_minimal()+
     theme(
@@
       axis.line.y.right = element_blank(),
+
       panel.grid.minor.x = element_blank(),
@@
       strip.text.y.left = element_text(angle = 0),
+
+      panel.spacing = unit(0.8, "lines"),
+
       plot.title = element_text(hjust = 0),
@@
       plot.caption = element_text(hjust = 0, face= "italic"),
-      legend.position = "bottom"
+
+      legend.position = "none"
      ) +
@@
 
-  plot_results <- 
-    patchwork::wrap_plots(
-      #plot_cmlinc, 
-      #plot_riskdiff,
-      plot_hazards, 
-      ncol=1
-    )
-
-  plot_results
+  plot_hazards
 

### 2022-02-28 analysis/manuscript_objects.R
remove png releases; add text file describing outputs

@@ -136,12 +136,12 @@ fs::file_copy(here("output", "descriptive", "table1", "table1by.csv"), here("out
 #   )
 # write_csv(cumulvax_jcvi_redacted, here("output", "descriptive", "vaxdate_jcvi_redacted.csv"))
 
-fs::file_copy(here("output", "descriptive", "vaxdate", "plot_vaxdate_count.png"), here("output", "manuscript-objects", "plot_vaxdate_count.png"), overwrite = TRUE)
-fs::file_copy(here("output", "descriptive", "vaxdate", "plot_vaxdate_step.png"), here("output", "manuscript-objects", "plot_vaxdate_step.png"), overwrite = TRUE)
-fs::file_copy(here("output", "descriptive", "vaxdate", "plot_vaxdate_stack.png"), here("output", "manuscript-objects", "plot_vaxdate_stack.png"), overwrite = TRUE)
-fs::file_copy(here("output", "descriptive", "vaxdate", "plot_vaxdate_count_jcvi.png"), here("output", "manuscript-objects", "plot_vaxdate_count_jcvi.png"), overwrite = TRUE)
-fs::file_copy(here("output", "descriptive", "vaxdate", "plot_vaxdate_step_jcvi.png"), here("output", "manuscript-objects", "plot_vaxdate_step_jcvi.png"), overwrite = TRUE)
-fs::file_copy(here("output", "descriptive", "vaxdate", "plot_vaxdate_stack_jcvi.png"), here("output", "manuscript-objects", "plot_vaxdate_stack_jcvi.png"), overwrite = TRUE)
+# fs::file_copy(here("output", "descriptive", "vaxdate", "plot_vaxdate_count.png"), here("output", "manuscript-objects", "plot_vaxdate_count.png"), overwrite = TRUE)
+# fs::file_copy(here("output", "descriptive", "vaxdate", "plot_vaxdate_step.png"), here("output", "manuscript-objects", "plot_vaxdate_step.png"), overwrite = TRUE)
+# fs::file_copy(here("output", "descriptive", "vaxdate", "plot_vaxdate_stack.png"), here("output", "manuscript-objects", "plot_vaxdate_stack.png"), overwrite = TRUE)
+# fs::file_copy(here("output", "descriptive", "vaxdate", "plot_vaxdate_count_jcvi.png"), here("output", "manuscript-objects", "plot_vaxdate_count_jcvi.png"), overwrite = TRUE)
[…]

### 2022-02-28 create-project.R

@@ -500,7 +500,8 @@ actions_list <- splice(
     ),
     moderately_sensitive = lst(
       csv = "output/manuscript-objects/*.csv",
-      png = "output/manuscript-objects/*.png",
+      #png = "output/manuscript-objects/*.png",
+      txt = "output/files-for-release.txt",
       csvsubgroup = "output/manuscript-objects/*/*.csv",
     )
   )

### 2022-02-28 manuscript/draft-manuscript.Rmd
Update draft-manuscript.Rmd

@@
 
+km_none <- read_csv(fs::path(output_dir_os, "none", "km.csv"))  
+km_vax12_type <- read_csv(fs::path(output_dir_os, "vax12_type", "km.csv"))
+
+km <- 
+  bind_rows(km_none, km_vax12_type) %>%
+  mutate(
+    subgroup_descr = fct_inorder(subgroup_descr),
+    subgroup_level = fct_inorder(subgroup_level),
+    subgroup_level_descr = fct_inorder(fct_explicit_na(subgroup_level_descr, "All")),
+    treatment_descr = fct_inorder(treatment_descr),
+    outcome_descr = fct_inorder(outcome_descr),
+    outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=20)),
+    #model_name = fct_inorder(str_wrap(model_name, 15)),
+  )
 # 
@@
     outcome_descr = fct_inorder(outcome_descr),
+    outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=20)),
+    #model_name = fct_inorder(str_wrap(model_name, 15)),
+    hr_display = paste0(label_number(accuracy=0.01)(hr), " (", label_number(accuracy=0.01)(hr.ll), "-", label_number(accuracy=0.01)(hr.ul), ")"),
+    ve_display = paste0(label_percent(accuracy=0.1)(ve), " (", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
+  )
+
+metahazards_none <- read_csv(fs::path(output_dir_os, "none", "metaeffects.csv"))
+metahazards_vax12_type <- read_csv(fs::path(output_dir_os, "vax12_type", "metaeffects.csv"))
+
+metahazards <- 
+  bind_rows(metahazards_none, metahazards_vax12_type) %>%
+  mutate(
+
+    subgroup_descr = fct_inorder(subgroup_descr),
+    subgroup_level = fct_inorder(subgroup_level),
+    subgroup_level_descr = fct_inorder(fct_explicit_na(subgroup_level_descr, "All")),
+    treatment_descr = fct_inorder(treatment_descr),
+    outcome_descr = fct_inorder(outcome_descr),
+    outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=20)),
     #model_name = fct_inorder(str_wrap(model_name, 15)),
@@
 
-The UK COVID-19 vaccination programme delivered the first "booster" doses in mid September 2021, initially in groups at high risk of severe disease and then across the entire adult population.
+The UK COVID-19 vaccination programme delivered the first "booster" doses in mid September 2021, initially in groups at high risk of severe disease and then across the adult population.
 The BNT162b2 mRNA Pfizer-BioNTech and the Moderna mRNA-1273 vaccines were used.
@@
 
-With the approval of NHS England we used the OpenSAFELY-TPP database, covering 40% of GP practices in England and linked to national coronavirus surveillance, hospital episodes, and death registry data, to estimate the effectiveness of a booster dose in "fully vaccinated" people.
-On each day of the recruitment period (`r format(as.Date(study_dates$studystart_date), "%d %B %Y")` to `r format(as.Date(study_dates$lastvax3_date), "%d %B %Y")`) each eligible person receiving a BNT162b2 booster dose on that day was recruited to the treatment group, and matched with an unboosted person recruited to the control group.
-Matching factors were region, JCVI priority group, and the day and type of the second vaccine dose.
-The treated and control groups across each day were pooled.
-Primary outcomes were documented SARS-CoV-2 infection, COVID-19-related accident and emergency attendance, unplanned COVID-19-related hospital admission, and COVID-19-related death occurring within 12 weeks of recruitment.
-Hazard ratios were estimated via a Cox model with additional covariates that were not used as matching factors.
+With the approval of NHS England we used the OpenSAFELY-TPP database, covering 40% of GP practices in England and linked to national coronavirus surveillance, hospital episodes, and death registry data, to estimate the effectiveness of a booster dose versus two doses only.
+On each day of the recruitment period (`r format(as.Date(study_dates$studystart_date), "%d %B %Y")` to `r format(as.Date(study_dates$lastvax3_date), "%d %B %Y")`) each eligible person receiving a BNT162b2 booster dose was matched with an unboosted control who had received two vaccine doses.
+Primary outcomes were documented SARS-CoV-2 infection, COVID-19-related accident and emergency attendance, COVID-19-related hospital admission, and COVID-19-related death occurring within 10 weeks of recruitment.
+Hazard ratios were estimated using Cox models adjusting for additional covariates that were not used as matching factors.
+Follow-up spanned a period in which the Delta variant was predominant.
 
@@
 
-We matched `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` BNT162b2 booster recipients to unboosted controls with only two doses.
+We matched `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` BNT162b2 booster recipients to unboosted controls.
 During days 14-20 after the booster, the estimated hazard ratio (95% CI) comparing a BNT162b2 booster dose to two doses only was `r filter(hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for documented SARS-CoV-2 infection, `r filter(hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 death.
@@
 
-BNT162b2 boosting was estimated to be highly effective in protecting against COVID-19 related hospitalisation and death.
+We estimated high protection of BNT162b2 boosting against infection, COVID-19 hospitalisation, and COVID-19 death, though with some evidence of waning over time.
 
@@
 
-The national COVID-19 vaccination programme in England began a new phase on 16 September 2021 with the administration of the first booster COVID-19 vaccine doses in those who had already received their primary vaccination course (a completed first and second dose) (ref <https://www.england.nhs.uk/2021/09/nhs-begins-covid-19-booster-vaccination-campaign/>).
-Initially offered to those at highest risk of severe disease, eligibility was progressively extended and by 15 December 2021 every adult was eligible (<https://www.england.nhs.uk/2021/12/nhs-booster-bookings-open-to-every-eligible-adult/>).
-Booster doses were initially available no earlier than six months after dose two, but this was reduced to three months on 8 December 2021, following concerns over the surge in cases and the emergence of the Omicron variant (ref: <https://www.england.nhs.uk/2021/12/people-40-and-over-to-get-their-lifesaving-booster-jab-three-months-on-from-second-dose/>)
+The national COVID-19 vaccination programme in England began a new phase in September 2021 with administration of the first booster COVID-19 vaccine doses in those who had already received their primary vaccination course (a completed first and second dose) (ref <https://www.england.nhs.uk/2021/09/nhs-begins-covid-19-booster-vaccination-campaign/>).
+Initially offered to those at highest risk of severe disease, eligibility was progressively extended and by 15 December 2021 every adult in England was eligible (<https://www.england.nhs.uk/2021/12/nhs-booster-bookings-open-to-every-eligible-adult/>).
+Booster doses were initially available no earlier than six months after dose two, but this was reduced to three months on 8 December 2021, following concerns over the surge in cases and the emergence of the Omicron variant (ref: <https://www.england.nhs.uk/2021/12/people-40-and-over-to-get-their-lifesaving-booster-jab-three-months-on-from-second-dose/>).
+Vaccine prioritisation schedules were largely determined by recommendations from the Joint Committee for Vaccine and Immunisation (JCVI) expert working group.
+
+In most cases, BNT162b2 and mRNA-1273 vaccines were used, and administered regardless of the vaccine brand used for the first and second doses.
+BNT162b2 was used from 16 September 2021, and mRNA-1273 from 29 October 2021 onwards.
 
@@
 
-We set out to estimate the effectiveness of BNT162b2 mRNA Pfizer-BioNTech booster doses versus two doses only in adults in England using the OpenSAFELY-TPP linked primary care database covering around 40% of England's population.
-An analysis of effectiveness of Moderna mRNA-1273 boosters, which were administered from 29 October 2021, is planned when sufficient follow-up has accrued.
+We estimated the effectiveness of booster vaccination with BNT162b2 versus two doses only in adults in England, using the OpenSAFELY-TPP linked primary care database covering around 40% of England's population.
+The study is conducted in an era where the Delta variant is dominant, with follow-up curtailed shortly after the emergence of Omicron.
 
@@
 
-The OpenSAFELY-TPP database (<https://opensafely.org>) covers 24 million people registered at GP practices that use TPP SystmOne electronic health record software.
+The OpenSAFELY-TPP database (<https://opensafely.org>) covers 24 million people registered at English General Practices that use TPP SystmOne electronic health record software.
 It includes pseudonymized data such as coded diagnoses, medications and physiological parameters.
@@
 Vaccination status is available in the GP record directly via the National Immunisation Management System (NIMS).
-Healthcare worker status is recorded for all vaccine recipients at the time of vaccination, and this information is sent to OpenSAFELY-TPP from NHS Digital's COVID-19 data store.
 
@@
 
-We used observational data to emulate a sequence of target trials @dagan2021 @barda2021, starting on consecutive days between `r format(as.Date(study_dates$studystart_date), "%d %B %Y")` and `r format(as.Date(study_dates$lastvax3_date), "%d %B %Y")` inclusive.
+We used observational data to emulate a target trial @dagan2021 @barda2021 of the effectiveness of booster vaccination in people who had received two COVID-19 vaccine doses and were eligible for booster vaccination between `r format(as.Date(study_dates$studystart_date), "%d %B)` and `r format(as.Date(study_dates$lastvax3_date), "%d %B %Y")` inclusive.
 
-Eligibility on each day was as follows: adults (18 years and over) who were registered at a GP practice using TPP's SystmOne clinical information system; received a primary COVID-19 vaccination course of either two BNT162b2 doses or two ChAdOx1-S doses (mRNA-1273 was not considered due to small numbers); not a vaccinated health care worker, not a resident in a care or nursing home, and not medically housebound or receiving end-of-life care (due to distinct clustering of vaccine uptake and infection risk in these groups that could not be reasonably measured and controlled for); no evidence of SARS-CoV-2 infection or COVID-19 disease within 90 days prior to recruitment; not undergoing an unplanned hospital admission; information on sex, ethnicity, deprivation, and NHS region was complete.
+Eligibility on each day was as follows: adults (18 years and over) who were registered at a GP practice using TPP's SystmOne clinical information system; received either two doses of BNT162b2 doses or two doses of ChAdOx1-S (mixed dosing and Moderna mRNA-1273 doses were not considered due to small numbers); not a health or social care worker, not a resident in a care or nursing home, and not medically housebound or receiving end-of-life care (due to distinct clustering of vaccine uptake and infection risk in these groups that could not be reasonably measured and controlled for); no evidence of SARS-CoV-2 infection or COVID-19 disease within the previous 90 days; not undergoing an unplanned hospital admission; information on sex, ethnicity, deprivation, and NHS region was complete.
 
-Additional eligibility criteria, based on sufficiently many people being vaccinated on each day by region, priority group, and week of second dose.
-If the rolling weekly average count of BNT162b2 booster doses within these strata was below 50 then recruitment did not occur on that day.
-This helped to ensure that unboosted people were both eligible and able to receive a booster dose at the time of potential recruitment.
+Additionally, if the rolling weekly average count of BNT162b2 booster doses within strata defined by region, JCVI group, and week of second dose was below 50 then recruitment did not occur for that strata on that day.
+This helped to ensure that recruitment was restricted to those who were boosted in line with national prioritisation schedules and that the matched controls were both eligible and able to receive a booster dose at the time of potential recruitment.
 
-We excluded matched pairs from days in which fewer than 10% of eligible treated people were successfully matched. See supplementary materials for further details.
+The supplementary materials provide further details of the matching process.
 
-## Treatment groups
+## Matching and treatment groups
 
-For each trial, the treatment group comprised people who received a BNT162b2 booster dose on that day, and the control group comprised the same number of randomly selected unboosted people, matched on key characteristics as follows: primary course vaccine type (BNT162b2 or ChAdOx1-S); calendar week of the second vaccine dose; NHS region; prior infection; immunosuppression; vaccine priority group (as determined by the Joint Committee for Vaccine and Immunisation expert working group).
+For each trial, the booster (treatment) group comprised people who received a BNT162b2 booster dose on that day, and the unboosted (control) group comprised the same number of randomly selected unboosted people, matched on the following characteristics: primary course vaccine type (BNT162b2 or ChAdOx1-S); date of the second vaccine dose, within 7 days; NHS region (East of England, Midlands, London, North East and Yorkshire, North West, South East, South West); any evidence of prior SARS-CoV-2 infection (positive SARS-CoV-2 test, "probable" infection documented in primary care, or COVID-19 hospitalisation); Clinical risk group, defined by JCVI (clinically extremely vulnerable, clinically at-risk, neither); age, within 3 years, and by age groups defined by JCVI.
+Note that the matching and exclusion criteria jointly ensure that treatment groups are also matched within JCVI priority group.
+
 People selected as controls were not eligible to be included again as a control in a subsequent trial, but were eligible for selection into the treatment group of a subsequent trial.
+
+<!-- We excluded matched pairs from days in which fewer than 10% of eligible treated people were successfully. -->
 
@@
 
-The following three outcomes were considered: documented SARS-CoV-2 infection; COVID-19 hospital admission; and COVID-19 death.
+The following three outcomes were considered: documented SARS-CoV-2 infection; COVID-19 hospital admission; COVID-19 death; and non-COVID-19 death as a negative control outcome.
 SARS-CoV-2 infections were identified using SGSS testing records and based on swab date.
@@
 
-Each matched pair was followed from the day of recruitment (time zero or baseline) until the earliest of death, GP practice de-registration, booster receipt by the control, 10 weeks, or `r format(as.Date(study_dates$studyend_date), "%d %B %Y")`.
+Each matched pair was followed from the day of recruitment (i.e., time zero) until the earliest of death, GP practice de-registration, booster receipt by the control, 10 weeks, or `r format(as.Date(study_dates$studyend_date), "%d %B %Y")`.
 
@@
 
-For each outcome, we estimated cumulative incidence curves using the Kaplan-Meier estimator separately in the treated and control groups.
+For each outcome, we estimated unadjusted cumulative incidence curves using the Kaplan-Meier estimator separately in the boosted and unboosted groups.
 
-We estimated period-specific hazard ratios for treatment versus control in consecutive two-week periods using a Cox model.
-Baseline hazards were allowed to differ by recruitment day, region, and week and type of second vaccine dose.
-The following variables were included as covariates: age, sex (male or female), English Index of Multiple Deprivation (IMD, grouped by quintile), ethnicity (Black, Mixed, South Asian, White, Other, as per the UK census), NHS region (East of England, Midlands, London, North East and Yorkshire, North West, South East, South West), number of conditions in the clinically "at risk" classification, as per national prioritisation guidelines, the number of SARS-CoV-2 tests in the 6 months prior to the study start date (via SGSS), evidence of prior SARS-CoV-2 infection (positive SARS-CoV-2 test, "probable" infection documented in primary care, or COVID-19 hospitalisation), learning disabilities, severe mental illness, undergoing an unplanned hospital admission, undergoing a planned hospital admission.
+We use Cox models, stratified by recruitment day, NHS region, and brand of second vaccine dose, to estimate hazard ratios comparing boosted with unboosted people.
+We estimated effectveness over all follow-up time (0-70 days), and separately in days 0-6, 7-13, 14-27, 28-41, 42-70.
+Models were adjusted for additional variables not used as matching factors, as described below.
 A robust variance estimator was used to account for participant clustering.
@@
 
-We additionally estimated booster effectiveness in subgroups defined by primary vaccine course (both BNT162b2 or both ChAdOx1-S) by repeating the analysis separately in each subgroup.
+We additionally estimated booster effectiveness in the following subgroups: primary vaccine course (both BNT162b2 or both ChAdOx1-S); age (18-64 versus 65 and over); clinically extremely vulnerable.
+Analyses were conducted independently in each subgroup.
 
@@
 
+## Covariates
+
+The following variables were included as covariates: sex (male or female); English Index of Multiple Deprivation (IMD, grouped by quintile); ethnicity (Black, Mixed, South Asian, White, Other, as per the UK census); morbidity count (diabetes, BMI > 40, chronic heart disease, chronic kidney disease, chronic liver disease, chronic respiratory disease or severe asthma, chronic neurological disease); learning disabilities; severe mental illness; the number of SARS-CoV-2 tests in the 6 months prior to the study start date; elective in-hospital episode at the time of recruitment.
+
+The supplementary materials provide further details of the codelists and data sources used for all variables in the study.
+
 ## Missing data
 
-After exclusions for missing values on demographic variables as described earlier, there were no missing values in remaining variables as they were each defined by the presence or absence of clinical codes or events in the patient record.
+After exclusions for missing values on demographic variables, there were no missing values in remaining variables as they were each defined by the presence or absence of clinical codes or events in the patient record.
 This was therefore a complete-case analysis.
@@
 
-A total of `r label_number(1, big.mark=",")(filter(matchflowchart, crit=="c0", treatment=="pfizer") %>% pull(n))` BNT162b2 booster recipients were identified between `r format(as.Date(matchsummary_treated$firstrecruitdate), "%d %B %Y")` and `r format(as.Date(matchsummary_treated$lastrecruitdate), "%d %B %Y")` in adults registered at a TPP practice, with `r label_number(1, big.mark=",")(filter(matchflowchart, treatment=="pfizer", crit=="c7") %>% pull(n))` (`r label_percent(0.1)(filter(matchflowchart, treatment=="pfizer", crit=="c7") %>% pull(pct_all))`) eligible for inclusion into the treatment group.
-Of these, `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` (`r label_percent(0.1)(filter(matchsummary, treatment=="pfizer") %>% pull(prop_matched_eligible))`) were successfully matched with unboosted controls (Figure S1).
+`r label_number(1, big.mark=",")(filter(matchflowchart, crit=="c0", treatment=="pfizer") %>% pull(n))` BNT162b2 booster recipients were identified between `r format(as.Date(matchsummary_treated$firstrecruitdate), "%d %B %Y")` and `r format(as.Date(matchsummary_treated$lastrecruitdate), "%d %B %Y")` in adults registered at a TPP practice, with `r label_number(1, big.mark=",")(filter(matchflowchart, treatment=="pfizer", crit=="c7") %>% pull(n))` (`r label_percent(0.1)(filter(matchflowchart, treatment=="pfizer", crit=="c7") %>% pull(pct_all))`) eligible for inclusion into the treatment group.
+Of these, `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` (`r label_percent(0.1)(filter(matchsummary, treatment=="pfizer") %>% pull(prop_matched_eligible))`) were successfully matched with unboosted controls (Table S1).
 
@@
 Participant characteristics at trial recruitment between treatment groups (Table 1) were typically well-balanced though with some exceptions.
-For instance, unboosted controls had higher levels of deprivation, higher rates of learning disabilities and severe mental illness, and lower SARS-CoV-2 testing frequency.
-
-There were `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(fup_years))` person-years of follow-up across both treatment groups (Table 2).
+For instance, unboosted controls had higher levels of deprivation, higher rates of learning disabilities and severe mental illness, and lower SARS-CoV-2 testing frequency, though the standardised mean difference was consistently below 0.1 (Supplementary Figure S1).
 
@@
 
-Booster vaccination uptake over the duration of the study period, and the number of boosted people who were eligible and matched.
+Cumulative booster vaccination over the duration of the study period, by matching eligibility and matching success.
 
@@
 xmax <- max(matchcoverage$vax3_date )+1
-matchcoverage  %>%
+data_coverage <- matchcoverage  %>%
+  mutate(
+    status = factor(status, levels=c("ineligible", "unmatched", "matched")),
+    status_descr = fct_relabel(status, str_to_title)
+  ) %>%
+  arrange(status) %>%
+  mutate(
+    status_descr = fct_inorder(status_descr)
+  ) %>%
   filter(treatment=="pfizer") %>%
@@
     )
+  )
+
+data_annotate <- 
+  data_coverage %>%
+  filter(vax12_type=="pfizer-pfizer") %>%
+  group_by(status_descr, course) %>%
+  summarise(
+    xpos = first(lubridate::floor_date((vax3_date), "1 month")),
+    ypos= max(cumuln)
   ) %>%
-  ggplot() +
+  ungroup() %>%
+  arrange(desc(status_descr)) %>%
+  mutate(
+    ypos=(cumsum(ypos) + lag(cumsum(ypos), 1, 0))/2
+  ) %>%
+  mutate(
+    course="2× ChAdOx1-S / 1× BNT162b2"
+  )
+
+ggplot(data_coverage) +
   geom_col(
@@
       y=cumuln,
-      group=fct_rev(status),
-      fill=fct_rev(status),
+      group=status_descr,
+      fill=status_descr,
       colour=NULL
@@
     position=position_stack(),
-    alpha=0.8,
+    alpha=0.5,
     width=1
@@
   geom_rect(xmin=xmin, xmax= xmax+1, ymin=0, ymax=6, fill="grey", colour="transparent")+
+  geom_text(data=data_annotate, mapping=aes(x = xpos, y = (ypos), label = status_descr, colour=status_descr), hjust=0, show.legend = FALSE)+
   facet_wrap(vars(course))+
@@
     axis.ticks.x=element_line(),
-    legend.position = "bottom"
+    legend.position = "none"
   )+
@@
 
-Total follow-up and incidence rates for each outcome, by treatment group.
+Total follow-up and incidence rates for each outcome, by treatment group. Cumulative incidences are provided in Supplementary Figure S2.
 
@@
   mutate(
-    treatment_decr = paste0(
+    treatment_descr = paste0(
       treatment_descr, " - ",
@@
 
-Participants in the emulated trial contributed `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="postest", fup_period=="All") %>% pull(events))` documented infection, `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="covidadmitted", fup_period=="All") %>% pull(events))` COVID-19 hospital admissions, and `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="coviddeath", fup_period=="All") %>% pull(events))` COVID-19 deaths (Table 2).
+Participants in the emulated trial experienced `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="postest", fup_period=="All") %>% pull(events))` documented infections, `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="covidadmitted", fup_period=="All") %>% pull(events))` COVID-19 hospital admissions, and `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="coviddeath", fup_period=="All") %>% pull(events))` COVID-19 deaths across `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(fup_years))` person-years of follow-up (Table 2).
 
@@
 
+For days 42-70, estimated effectiveness (95% CI) comparing a BNT162b2 booster dose to two doses only, was `r filter(hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_left==42) %>% pull(ve_display)` for documented infection, `r filter(hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_left==42) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_left==42) %>% pull(ve_display)` for COVID-19 death (Figure 2), suggesting a potential waning of booster effectiveness with time.
 <!-- `r filter(hazards, subgroup=="none", outcome=="covidattendance", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 A&E attendances, `r filter(hazards, subgroup=="none", outcome=="covidcc", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 critical care admission,  -->
@@
 For each outcome based on the fully adjusted model, the period-specific hazard ratios for boosting with BNT162b2 versus not boosting is shown, stratified by primary course.
-Models with less extensive confounder adjustment are provided in supplementary materials (supplementary Figure SX).
+Models with less extensive confounder adjustment are provided in supplementary materials (Figure S3).
 
@@
     geom_linerange(aes(ymin=hr.ll, ymax=hr.ul, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+
+    geom_point(data = filter(metahazards, model==3), mapping = aes(y=hr, x=max(postbaselinecuts+14)), colour="grey") +
+    geom_linerange(data = filter(metahazards, model==3), mapping = aes(ymin=hr.ll, ymax=hr.ul, x=max(postbaselinecuts+14)), colour="grey") +
     geom_hline(aes(yintercept=1), colour='grey')+
-    facet_grid(rows=vars(outcome_descr), cols=vars(subgroup_level_descr), switch="y")+
+    facet_grid(rows=vars(outcome_descr_wrap), cols=vars(subgroup_level_descr), switch="y")+
     scale_y_log10(
@@
     )+
-    scale_x_continuous(breaks=postbaselinecuts, limits=c(min(postbaselinecuts), max(postbaselinecuts)+1), expand = c(0, 0))+
+    scale_x_continuous(breaks=c(postbaselinecuts, max(postbaselinecuts+14)), labels=c(postbaselinecuts, "Overall"), limits=c(min(postbaselinecuts), max(postbaselinecuts+14)+1), expand = c(0, 0))+
     scale_colour_brewer(type="qual", palette="Set2", guide=guide_legend(ncol=1))+
@@
 
-  plot_mainresults <- 
-    patchwork::wrap_plots(
-      #plot_cmlinc, 
-      #plot_riskdiff,
-      plot_hazards, 
-      ncol=1
-    )
   ggsave(
-    filename="mainresults.pdf",
+    filename="plot_hazards.pdf",
     path=output_dir_rmd,
-    plot=plot_mainresults,
+    plot=plot_hazards,
   )
 
-  plot_mainresults
+  plot_hazards
 
@@
 
-This observational study of over 8 million boosted adults in England between `r format(as.Date(matchsummary_treated$firstrecruitdate), "%d %B %Y")` and `r format(as.Date(matchsummary_treated$lastrecruitdate), "%d %B %Y")` and their matched controls estimated high protection of BNT162b2 booster vaccination against all studied outcomes using routinely-collected health records with comprehensive coverage of primary care, hospital admissions, COVID-19 testing and vaccination, and death registrations.
+This observational study investigated the effectiveness of a BNT162b2 booster dose in `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` adults, and their matched controls, in England in the Delta era.
+We found high protection against all studied outcomes.
 Protection was similar between 2× BNT162b2 primary course recipients and 2× ChAdOx1-S recipients.
@@
 
-### Limitations
+### Strengths and Limitations
+
+We used routinely-collected health records with comprehensive coverage of primary care, hospital admissions, COVID-19 testing and vaccination, and death registrations, and with substantial power to estimate effectiveness of booster doses against severe but rare outcomes such as hospitalisation and death over time.
+
+We carefully matched booster recipients with unboosted controls using characteristics known at baseline to strengthen exchangeability between treatment groups and reduce dependence on modelling assumptions in the estimates of effectiveness.
+However, despite reasonable balance of important baseline characteristics and additional adjustment for a range of potential confounders, the possibility of unmeasured confounding remains.
+In particular, the apparent protective effect immediately after boosting is biologically implausible, and will be explained in part by the occurrence of COVID-19 symptoms at baseline that were not recorded in the health record, and so could not be identified and excluded.
+
+The matching approach robust but inefficient; a minority of boosted individuals are not retained after matching, and those who are retained are censored when their matched control is boosted, reducing the follow-up duration and therefore statistical power.
+Millions of eligible participants may be required to obtain precise estimates, particularly when event rates are low, and so extending the approach to smaller subgroups may not be feasible.
+For instance, due to small numbers, we did not study booster effectiveness in those who had received the mRNA-1273 vaccine as their primary course or those who had received a heterologous primary course, despite many thousands of such recipients in England.
 
@@
 The mass availability of lateral flow tests in the UK during the study period, whose results are not routinely recorded in COVID-19 surveillance data, means many infections, including symptomatic infections, may be undocumented, despite encouragement to seek a confirmatory PCR test.
-Potential differences in testing behaviour between vaccinated and unvaccinated people also contributes to the unreliability of testing data as a means to assess effectiveness.
+Potential differences in testing behaviour between boosted and unboosted people also contributes to the unreliability of testing data as a means to assess effectiveness.
 SARS-CoV-2 testing was not widely available early in the pandemic so evidence of prior infection is likely to be under-ascertained.
@@
 
-We excluded a number of groups, such as health care workers and care home residents, where testing, vaccination, and infection risk characteristics were unusual or had substantial within-group heterogeneity that could not be adequately accounted for.
-
 Hospital admission records are only completed after discharge, so some longer hospital stays may not be counted.
 
-Despite reasonable balance of baseline characteristics between treatment groups, and adjustment for a range of potential confounders, the possibility of unmeasured confounding remains.
-In particular, the apparent protective effect immediately after boosting is biologically implausible, and will be explained in part by the occurrence of COVID-19 symptoms at baseline that were not recorded in the health record, and therefore could not be used to exclude from recruitment.
+We excluded a number of groups, such as health care workers and care home residents, where testing, vaccination, and infection risk characteristics were unusual or had substantial within-group heterogeneity that could not be adequately accounted for.
 
@@
 
-The methodological approach used is robust but inefficient; many boosted individuals are not retained after matching, and those who are retained are censored when their matched control is boosted.
-Millions of eligible participants may be required to obtain precise estimates, particularly when event rates are low, and so extending the approach to smaller subgroups may not be feasible.
-For instance, due to small numbers, we did not study booster effectiveness in those who had received the mRNA-1273 vaccine as their primary course or those who had received a heterologous primary course, despite many thousands of such recipients in England.
-
-Finally, mRNA-1273 boosting was rolled out later than for BNT162b2, and so we could not evaluate effectiveness of this vaccine due to insufficient follow-up.
+Finally, mRNA-1273 boosting begun later than for BNT162b2, and so we could not evaluate effectiveness of this vaccine due to insufficient follow-up.
 An analysis of mRNA-1273 boosting and comparative effectiveness between mRNA-1273 and BNT162b2 is planned.
@@
 
-This study of over 8 million patients in England found high protection for BNT162b2 boosting against COVID-19 hospitalisation and death during an `r studydurationweeks` week period in which the Delta variant was dominant.
+This study of over 8 million people in England found high protection for BNT162b2 boosting against COVID-19 hospitalisation and death during an `r studydurationweeks` week period in which the Delta variant was dominant.
 

### 2022-02-28 analysis/combine_model.R
add non covid death

@@ -116,7 +116,7 @@ if(Sys.getenv("OPENSAFELY_BACKEND") %in% c("", "expectations")){
   model_metaparams <-
     expand_grid(
       treatment = factor(c("pfizer", "moderna")),
-      outcome = factor(c("postest", "covidemergency", "covidadmittedproxy1", "covidadmitted", "coviddeath")),
+      outcome = factor(c("postest", "covidemergency", "covidadmittedproxy1", "covidadmitted", "coviddeath", "noncovid")),
       #outcome = factor(c("postest", "covidadmission")),
       subgroup = factor(recode_subgroup),
       subgroup_variable = factor(subgroup_variable),

### 2022-02-28 create-project.R

@@ -369,6 +369,7 @@ actions_list <- splice(
   action_model("pfizer", "covidadmitted", "none"),
   action_model("pfizer", "covidcc", "none"),
   action_model("pfizer", "coviddeath", "none"),
+  action_model("pfizer", "noncoviddeath", "none"),
 
   comment("### Models by primary course ('vax12_type')"),
 
@@ -384,6 +385,8 @@ actions_list <- splice(
   action_model("pfizer", "covidcc", "vax12_type-pfizer-pfizer"),
   action_model("pfizer", "coviddeath", "vax12_type-az-az"),
   action_model("pfizer", "coviddeath", "vax12_type-pfizer-pfizer"),
+  action_model("pfizer", "noncoviddeath", "vax12_type-az-az"),
+  action_model("pfizer", "noncoviddeath", "vax12_type-pfizer-pfizer"),
 
 
   comment("### Models by clinical extremely vulnerable ('cev')"),
@@ -400,6 +403,8 @@ actions_list <- splice(
   action_model("pfizer", "covidcc", "cev-TRUE"),
   action_model("pfizer", "coviddeath", "cev-FALSE"),
   action_model("pfizer", "coviddeath", "cev-TRUE"),
+  action_model("pfizer", "noncoviddeath", "cev-FALSE"),
+  action_model("pfizer", "noncoviddeath", "cev-TRUE"),
 
   comment("### Models by age ('age65plus')"),
 
@@ -415,6 +420,8 @@ actions_list <- splice(
   action_model("pfizer", "covidcc", "age65plus-TRUE"),
   action_model("pfizer", "coviddeath", "age65plus-FALSE"),
   action_model("pfizer", "coviddeath", "age65plus-TRUE"),
+  action_model("pfizer", "noncoviddeath", "age65plus-FALSE"),
+  action_model("pfizer", "noncoviddeath", "age65plus-TRUE"),
 
   comment("# # # # # # # # # # # # # # # # # # #", "Moderna models", "# # # # # # # # # # # # # # # # # # #"),
 
@@ -426,6 +433,7 @@ actions_list <- splice(
   action_model("moderna", "covidadmitted", "none"),
   action_model("moderna", "covidcc", "none"),
   action_model("moderna", "coviddeath", "none"),
+  action_model("moderna", "noncoviddeath", "none"),
 
   comment("### Models by primary course ('vax12_type')"),
 
@@ -441,6 +449,8 @@ actions_list <- splice(
[…]

### 2022-02-28 analysis/manuscript_objects.R
remove uneeded plots

@@ -46,108 +46,7 @@ study_dates <-
   map(as.Date)
 
 
-## cohort ----
-
-fs::file_copy(here("output", "data", "flowchart.csv"), here("output", "manuscript-objects", "flowchart.csv"), overwrite = TRUE)
-
-data_cohort <- read_rds(here("output", "data", "data_cohort.rds"))
-
-cohort_summary <-
-  data_cohort %>%
-  mutate(
-    censor_date = pmin(
-      dereg_date,
-      vax4_date-1, # -1 because we assume vax occurs at the start of the day
-      if_else(!(vax3_type %in% c("pfizer", "moderna")), vax3_date-1, as.Date(NA)), # censor if third dose is not pfizer or moderna
-      death_date,
-      study_dates$studyend_date,
-      na.rm=TRUE
-    ),
-
-    # assume vaccination occurs at the start of the day, and all other events occur at the end of the day.
-
-    tte_censor = tte(study_dates$studystart_date-1, censor_date, censor_date, na.censor=TRUE),
-    #ind_censor = censor_indicator(censor_date, censor_date),
-  ) %>%
-  summarise(
-    n = n(),
-    n_boosted = sum(between(vax3_date, study_dates$studystart_date, study_dates$lastvax3_date) & vax3_type %in% c("pfizer", "moderna"), na.rm=TRUE),
-
-    n_12pfizer = sum(vax12_type=="pfizer-pfizer"),
-    prop_12pfizer = n_12pfizer/n,
-    n_12az = sum(vax12_type=="az-az"),
-    prop_12az = n_12az/n,
-    n_12moderna = sum(vax12_type=="moderna-moderna"),
-    prop_12moderna = n_12moderna/n,
-    n_3pfizer = sum(vax3_type=="pfizer" & vax3_date <= study_dates$lastvax3_date, na.rm=TRUE),
-    prop_3pfizer = n_3pfizer/n_boosted,
-    n_3moderna = sum(vax3_type=="moderna" & vax3_date <= study_dates$lastvax3_date, na.rm=TRUE),
-    prop_3moderna = n_3moderna/n_boosted,
-
-    age_median = median(age),
-    age_Q1 = quantile(age, 0.25),
-    age_Q3 = quantile(age, 0.75),
-    female = mean(sex=="Female"),
-    fu_years = sum(tte_censor)/365.25,
-    fu_median = median(tte_censor),
-    #priorinfection = mean(prior_covid_infection),
[…]

### 2022-02-28 analysis/merge_seqtrialcox.R
add some basic cohort descriptors

@@ -179,7 +179,8 @@ match_summary_trial <-
     n=n(),
     fup_sum = sum(fup),
     fup_years = sum(fup)/365.25,
-    fup_mean = mean(fup)
+    fup_mean = mean(fup),
+    fup_median = median(fup)
   ) %>%
   arrange(
     trial_time, treated
@@ -208,10 +209,24 @@ match_summary_treated <-
     lastrecruitdate = max(tte_recruitment) + day1_date,
     fup_sum = sum(fup),
     fup_years = sum(fup)/365.25,
-    fup_mean = mean(fup)
+    fup_mean = mean(fup),
+    fup_median = median(fup),
+
+    n_12pfizer = sum(vax12_type=="pfizer-pfizer"),
+    prop_12pfizer = n_12pfizer/n,
+    n_12az = sum(vax12_type=="az-az"),
+    prop_12az = n_12az/n,
+    # n_12moderna = sum(vax12_type=="moderna-moderna"),
+    # prop_12moderna = n_12moderna/n,
+
+    age_median = median(age),
+    age_Q1 = quantile(age, 0.25),
+    age_Q3 = quantile(age, 0.75),
+    female = mean(sex=="Female"),
   )
 
 write_csv(match_summary_treated, fs::path(output_dir, "merge_summary_treated.csv"))
+
 
 
 # summary of boosted people
@@ -234,7 +249,21 @@ match_summary <-
     lastrecruitdate = max(tte_recruitment) + day1_date,
     fup_sum = sum(fup),
     fup_years = sum(fup)/365.25,
-    fup_mean = mean(fup)
+    fup_mean = mean(fup),
+    fup_median = median(fup),
+
+    n_12pfizer = sum(vax12_type=="pfizer-pfizer"),
+    prop_12pfizer = n_12pfizer/n,
+    n_12az = sum(vax12_type=="az-az"),
+    prop_12az = n_12az/n,
+    # n_12moderna = sum(vax12_type=="moderna-moderna"),
+    # prop_12moderna = n_12moderna/n,
+
+    age_median = median(age),
+    age_Q1 = quantile(age, 0.25),
+    age_Q3 = quantile(age, 0.75),
+    female = mean(sex=="Female"),
+
   ) %>%
   bind_cols(candidate_summary) %>%
   mutate(

### 2022-02-28 analysis/combine_model.R
add split period meta analysis

@@ -216,6 +216,24 @@ model_metaeffects <-
 write_csv(model_metaeffects, path = fs::path(output_dir, "metaeffects.csv"))
 
 
+model_meta2effects <-
+  model_metaparams %>%
+  mutate(
+    meta2effects = pmap(list(treatment, outcome, subgroup), function(x, y, z) read_csv(here("output", "models", "seqtrialcox", x, y, z, glue("report_meta2effects.csv"))))
+  ) %>%
+  unnest(meta2effects) %>%
+  mutate(
+    model_descr = fct_reorder(model_descr, model),
+    hr = exp(estimate),
+    hr.ll = exp(conf.low),
+    hr.ul = exp(conf.high),
+    ve = 1-hr,
+    ve.ll = 1-hr.ul,
+    ve.ul = 1-hr.ll
+  )
+
+write_csv(model_meta2effects, path = fs::path(output_dir, "meta2effects.csv"))
+
 
 formatpercent100 <- function(x,accuracy){
   formatx <- scales::label_percent(accuracy)(x)

### 2022-02-28 analysis/manuscript_objects.R

@@ -65,6 +65,7 @@ for(subgroup_variable in c("none", "vax12_type", "cev", "age65plus")){
   fs::file_copy(here("output", "models", "seqtrialcox", "combined", subgroup_variable, "km.csv"), here("output", "manuscript-objects", subgroup_variable, "km.csv"), overwrite = TRUE)
   fs::file_copy(here("output", "models", "seqtrialcox", "combined", subgroup_variable, "effects.csv"), here("output", "manuscript-objects", subgroup_variable, "effects.csv"), overwrite = TRUE)
   fs::file_copy(here("output", "models", "seqtrialcox", "combined", subgroup_variable, "metaeffects.csv"), here("output", "manuscript-objects", subgroup_variable, "metaeffects.csv"), overwrite = TRUE)
+  fs::file_copy(here("output", "models", "seqtrialcox", "combined", subgroup_variable, "meta2effects.csv"), here("output", "manuscript-objects", subgroup_variable, "meta2effects.csv"), overwrite = TRUE)
   fs::file_copy(here("output", "models", "seqtrialcox", "combined", subgroup_variable, "overalleffects.csv"), here("output", "manuscript-objects", subgroup_variable, "overalleffects.csv"), overwrite = TRUE)
 }
 

### 2022-02-28 analysis/report_seqtrialcox.R

@@ -152,6 +152,34 @@ model_metaeffects <- model_effects %>%
 write_csv(model_metaeffects, path = fs::path(output_dir, "report_metaeffects.csv"))
 
 
+
+## meta analysis of period-specific hazards ----
+
+model_meta2effects <- model_effects %>%
+  mutate(
+    period = (term_right>28) + 1,
+  ) %>%
+  group_by(
+    period, model, model_descr,
+  ) %>%
+  summarise(
+    term_left = min(term_left),
+    term_right = max(term_right),
+    estimate = weighted.mean(estimate, robust.se^-2),
+    std.error = sqrt(1/sum(std.error^-2)),
+    robust.se = sqrt(1/sum(robust.se^-2)),
+    statistic = estimate/robust.se,
+    p.value = 2 * pmin(pnorm(statistic), pnorm(-statistic)),
+    conf.low = estimate + qnorm(0.025)*robust.se,
+    conf.high = estimate + qnorm(0.975)*robust.se,
+  ) %>%
+  mutate(
+    term_midpoint = (term_left+term_right)/2,
+  )
+
+write_csv(model_metaeffects, path = fs::path(output_dir, "report_meta2effects.csv"))
+
+
 ## incidence rates ----
 
 ### Event incidence following recruitment

### 2022-03-03 manuscript/draft-manuscript.Rmd
lots of updates to manuscript

@@
 ---
-title: "Effectiveness of BNT162b2 booster doses in England: a observational study in OpenSAFELY"
+title: "Effectiveness of BNT162b2 booster doses in England: a observational study in OpenSAFELY-TPP"
 output:
@@
 
-flowchart <- read_csv(fs::path(output_dir_os, "flowchart.csv"))
 cohortsummary <- read_csv(fs::path(output_dir_os, "cohortsummary.csv"))
@@
 matchcoverage <- read_csv(fs::path(output_dir_os, "matchcoverage.csv"))
-incidences <- read_csv(fs::path(output_dir_os, "none", "incidence.csv")) %>% mutate(events=events_1+events_0)
+incidences <- read_csv(fs::path(output_dir_os, "none", "incidence_all.csv")) %>% mutate(events=events_1+events_0)
 
@@
       #"covidcc",
-      "coviddeath"
+      "coviddeath",
+      "noncoviddeath"
     ) %>% 
@@
 
-km_none <- read_csv(fs::path(output_dir_os, "none", "km.csv"))  
-km_vax12_type <- read_csv(fs::path(output_dir_os, "vax12_type", "km.csv"))
-
 km <- 
-  bind_rows(km_none, km_vax12_type) %>%
+  bind_rows(
+    read_csv(fs::path(output_dir_os, "none", "km.csv")),
+    read_csv(fs::path(output_dir_os, "vax12_type", "km.csv")),
+    read_csv(fs::path(output_dir_os, "cev", "km.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
+    read_csv(fs::path(output_dir_os, "age65plus", "km.csv")) %>% mutate(subgroup_level=as.character(subgroup_level))
+  ) %>%
+  filter(
+    outcome %in% outcomes
+  ) %>%
   mutate(
@@
     outcome_descr = fct_inorder(outcome_descr),
-    outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=20)),
+    outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=15)),
     #model_name = fct_inorder(str_wrap(model_name, 15)),
@@
 
-hazards_none <- read_csv(fs::path(output_dir_os, "none", "effects.csv"))
-hazards_vax12_type <- read_csv(fs::path(output_dir_os, "vax12_type", "effects.csv"))
-
 hazards <- 
-  bind_rows(hazards_none, hazards_vax12_type) %>%
+  bind_rows(
+    read_csv(fs::path(output_dir_os, "none", "effects.csv")),
+    read_csv(fs::path(output_dir_os, "vax12_type", "effects.csv")),
+    read_csv(fs::path(output_dir_os, "cev", "effects.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
+    read_csv(fs::path(output_dir_os, "age65plus", "effects.csv"))  %>% mutate(subgroup_level=as.character(subgroup_level)),
+  ) %>%
+  filter(
+    outcome %in% outcomes
+  ) %>%
   mutate(
@@
 
-metahazards_none <- read_csv(fs::path(output_dir_os, "none", "metaeffects.csv"))
-metahazards_vax12_type <- read_csv(fs::path(output_dir_os, "vax12_type", "metaeffects.csv"))
-
-metahazards <- 
-  bind_rows(metahazards_none, metahazards_vax12_type) %>%
+overallhazards <- 
+  bind_rows(
+    read_csv(fs::path(output_dir_os, "none", "overalleffects.csv")),
+    read_csv(fs::path(output_dir_os, "vax12_type", "overalleffects.csv")),
+    read_csv(fs::path(output_dir_os, "cev", "overalleffects.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
+    read_csv(fs::path(output_dir_os, "age65plus", "overalleffects.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
+  ) %>%
+  filter(
+    outcome %in% outcomes
+  ) %>%
   mutate(
@@
   )
+
+meta2hazards <- 
+  # derive directly from hazards as meta2effects table needs to be re-run
+  bind_rows(
+    read_csv(fs::path(output_dir_os, "none", "effects.csv")),
+    read_csv(fs::path(output_dir_os, "vax12_type", "effects.csv")),
+    read_csv(fs::path(output_dir_os, "cev", "effects.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
+    read_csv(fs::path(output_dir_os, "age65plus", "effects.csv"))  %>% mutate(subgroup_level=as.character(subgroup_level)),
+  ) %>%
+  filter(
+    outcome %in% outcomes
+  ) %>%
+  mutate(
+    period = (term_right>28) + 1,
+    subgroup_descr = fct_inorder(subgroup_descr),
+    subgroup_level = fct_inorder(subgroup_level),
+    subgroup_level_descr = fct_inorder(fct_explicit_na(subgroup_level_descr, "All")),
+    treatment_descr = fct_inorder(treatment_descr),
+    outcome_descr = fct_inorder(outcome_descr),
+    outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=20)),
+  ) %>%
+  group_by(
+    treatment, treatment_descr, outcome, outcome_descr, outcome_descr_wrap, subgroup, subgroup_variable, subgroup_descr, subgroup_level, subgroup_level_descr, model, model_descr, period, 
+  ) %>%
+  summarise(
+    term_left = min(term_left),
+    term_right = max(term_right),
+    term_midpoint = (term_left+term_right)/2,
+    estimate = weighted.mean(estimate, robust.se^-2),
+    std.error = sqrt(1/sum(std.error^-2)),
+    robust.se = sqrt(1/sum(robust.se^-2)),
+    statistic = estimate/robust.se,
+    p.value = 2 * pmin(pnorm(statistic), pnorm(-statistic)),
+    conf.low = estimate + qnorm(0.025)*robust.se,
+    conf.high = estimate + qnorm(0.975)*robust.se,
+    hr = exp(estimate),
+    hr.ll = exp(conf.low),
+    hr.ul = exp(conf.high),
+    ve = 1-hr,
+    ve.ll = 1-hr.ul,
+    ve.ul = 1-hr.ll
+  ) %>%
+  ungroup() %>%
+  mutate(
+    hr_display = paste0(label_number(accuracy=0.01)(hr), " (", label_number(accuracy=0.01)(hr.ll), "-", label_number(accuracy=0.01)(hr.ul), ")"),
+    ve_display = paste0(label_percent(accuracy=0.1)(ve), " (", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
+  )
+
+# meta2hazards <- 
+#   bind_rows(
+#     read_csv(fs::path(output_dir_os, "none", "meta2effects.csv")),
+#     read_csv(fs::path(output_dir_os, "vax12_type", "meta2effects.csv")),
+#     read_csv(fs::path(output_dir_os, "cev", "meta2effects.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
+#     read_csv(fs::path(output_dir_os, "age65plus", "meta2effects.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
+#   ) %>%
+#   filter(
+#     outcome %in% outcomes
+#   ) %>%
+#   mutate(
+#     subgroup_descr = fct_inorder(subgroup_descr),
+#     subgroup_level = fct_inorder(subgroup_level),
+#     subgroup_level_descr = fct_inorder(fct_explicit_na(subgroup_level_descr, "All")),
+#     treatment_descr = fct_inorder(treatment_descr),
+#     outcome_descr = fct_inorder(outcome_descr),
+#     outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=20)),
+#     #model_name = fct_inorder(str_wrap(model_name, 15)),
+#     hr_display = paste0(label_number(accuracy=0.01)(hr), " (", label_number(accuracy=0.01)(hr.ll), "-", label_number(accuracy=0.01)(hr.ul), ")"),
+#     ve_display = paste0(label_percent(accuracy=0.1)(ve), " (", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
+#   )
 
@@
 
-With the approval of NHS England we used the OpenSAFELY-TPP database, covering 40% of GP practices in England and linked to national coronavirus surveillance, hospital episodes, and death registry data, to estimate the effectiveness of a booster dose versus two doses only.
-On each day of the recruitment period (`r format(as.Date(study_dates$studystart_date), "%d %B %Y")` to `r format(as.Date(study_dates$lastvax3_date), "%d %B %Y")`) each eligible person receiving a BNT162b2 booster dose was matched with an unboosted control who had received two vaccine doses.
-Primary outcomes were documented SARS-CoV-2 infection, COVID-19-related accident and emergency attendance, COVID-19-related hospital admission, and COVID-19-related death occurring within 10 weeks of recruitment.
+With the approval of NHS England we used the OpenSAFELY-TPP database, covering 40% of GP practices in England and linked to national coronavirus surveillance, hospital episodes, and death registry data, to estimate the effectiveness of BNT162b2 boosting in adults eligible for boosting between `r format(as.Date(study_dates$studystart_date), "%d %B %Y")` and `r format(as.Date(study_dates$lastvax3_date), "%d %B %Y")`.
+Each booster recipient was matched with an unboosted control on factors relating to booster priority status and prior immunisation.
+Primary outcomes were documented SARS-CoV-2 infection, COVID-19 hospital admission, and COVID-19 death up to 10 weeks after recruitment.
 Hazard ratios were estimated using Cox models adjusting for additional covariates that were not used as matching factors.
@@
 We matched `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` BNT162b2 booster recipients to unboosted controls.
-During days 14-20 after the booster, the estimated hazard ratio (95% CI) comparing a BNT162b2 booster dose to two doses only was `r filter(hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for documented SARS-CoV-2 infection, `r filter(hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 death.
+During days 15-28 after the booster, the estimated hazard ratio (95% CI) comparing a BNT162b2 booster dose to two doses only was `r filter(hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` for documented SARS-CoV-2 infection, `r filter(hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` for COVID-19 death.
 
@@
 
-<!-- *Participants* `r label_number(1, big.mark=",")(filter(flowchart, crit=="c4") %>% pull(n))` people with a complete primary COVID-19 vaccine course, registered with a GP practice using the TPP SystmOne clinical information system in England.
--->
+<!---# *Participants* `r label_number(1, big.mark=",")(filter(matchflowchart, crit=="c4") %>% pull(n))` people with a complete primary COVID-19 vaccine course, registered with a GP practice using the TPP SystmOne clinical information system in England.
+--->
 
@@
 <!-- TODO -->
-
-<!-- * full description of booster eligiblity over time -->
-
-<!-- * figure 1 chart with SMD / %-point differences -->
 
@@
 Booster doses were initially available no earlier than six months after dose two, but this was reduced to three months on 8 December 2021, following concerns over the surge in cases and the emergence of the Omicron variant (ref: <https://www.england.nhs.uk/2021/12/people-40-and-over-to-get-their-lifesaving-booster-jab-three-months-on-from-second-dose/>).
-Vaccine prioritisation schedules were largely determined by recommendations from the Joint Committee for Vaccine and Immunisation (JCVI) expert working group.
-
-In most cases, BNT162b2 and mRNA-1273 vaccines were used, and administered regardless of the vaccine brand used for the first and second doses.
-BNT162b2 was used from 16 September 2021, and mRNA-1273 from 29 October 2021 onwards.
+Vaccine prioritisation schedules were guided by recommendations from the Joint Committee for Vaccine and Immunisation (JCVI) expert working group.
 
@@
 
-[context]
+<!-- In most cases, BNT162b2 and half-dose mRNA-1273 vaccines were used for boosting, and administered regardless of the first and second dose brand.
+BNT162b2 was used from 16 September 2021, and mRNA-1273 from 29 October 2021 onwards.
+This provides an opportunity to investigate booster effectiveness of two different brands administered in overlapping periods and community contexts, as well as in subgroups defined by prior immunological and other clinical factors.-->
 
-We estimated the effectiveness of booster vaccination with BNT162b2 versus two doses only in adults in England, using the OpenSAFELY-TPP linked primary care database covering around 40% of England's population.
-The study is conducted in an era where the Delta variant is dominant, with follow-up curtailed shortly after the emergence of Omicron.
+We set out to emulate a target trial estimating the effectiveness of booster vaccination with BNT162b2 against COVID-19 outcomes in adults who had already received two COVID-19 vaccine doses.
+We used the OpenSAFELY-TPP linked primary care database covering around 40% of English residents.
+Recruitment covers an era where the Delta variant is dominant, with follow-up curtailed shortly after the emergence of Omicron.
 
 # Methods
+
+## Study design
+
+We used observational data to emulate a target trial of the effectiveness of booster vaccination @dagan2021 @barda2021, by comparing people who had received a BNT162b2 booster dose with matched people who had not.
 
@@
 
-We used observational data to emulate a target trial @dagan2021 @barda2021 of the effectiveness of booster vaccination in people who had received two COVID-19 vaccine doses and were eligible for booster vaccination between `r format(as.Date(study_dates$studystart_date), "%d %B)` and `r format(as.Date(study_dates$lastvax3_date), "%d %B %Y")` inclusive.
-
+We recruited adults who were eligible for booster vaccination between `r format(as.Date(study_dates$studystart_date), "%d %B")` and `r format(as.Date(study_dates$lastvax3_date), "%d %B %Y")` inclusive.
 Eligibility on each day was as follows: adults (18 years and over) who were registered at a GP practice using TPP's SystmOne clinical information system; received either two doses of BNT162b2 doses or two doses of ChAdOx1-S (mixed dosing and Moderna mRNA-1273 doses were not considered due to small numbers); not a health or social care worker, not a resident in a care or nursing home, and not medically housebound or receiving end-of-life care (due to distinct clustering of vaccine uptake and infection risk in these groups that could not be reasonably measured and controlled for); no evidence of SARS-CoV-2 infection or COVID-19 disease within the previous 90 days; not undergoing an unplanned hospital admission; information on sex, ethnicity, deprivation, and NHS region was complete.
 
-Additionally, if the rolling weekly average count of BNT162b2 booster doses within strata defined by region, JCVI group, and week of second dose was below 50 then recruitment did not occur for that strata on that day.
+Additionally, if the rolling weekly average count of BNT162b2 booster doses within strata defined by region, JCVI group, and the week of second dose was below 50 then recruitment did not occur for that strata on that day.
 This helped to ensure that recruitment was restricted to those who were boosted in line with national prioritisation schedules and that the matched controls were both eligible and able to receive a booster dose at the time of potential recruitment.
@@
 
-For each trial, the booster (treatment) group comprised people who received a BNT162b2 booster dose on that day, and the unboosted (control) group comprised the same number of randomly selected unboosted people, matched on the following characteristics: primary course vaccine type (BNT162b2 or ChAdOx1-S); date of the second vaccine dose, within 7 days; NHS region (East of England, Midlands, London, North East and Yorkshire, North West, South East, South West); any evidence of prior SARS-CoV-2 infection (positive SARS-CoV-2 test, "probable" infection documented in primary care, or COVID-19 hospitalisation); Clinical risk group, defined by JCVI (clinically extremely vulnerable, clinically at-risk, neither); age, within 3 years, and by age groups defined by JCVI.
-Note that the matching and exclusion criteria jointly ensure that treatment groups are also matched within JCVI priority group.
+On each day of recruitment, each eligible person receiving a booster dose on that day was recruited to the treatment group and matched with an unboosted control person who had received only two doses.
+Pairs were matched on the following characteristics: primary course vaccine brand (2× BNT162b2 or 2× ChAdOx1-S); date of the second vaccine dose, within 7 days; NHS region (East of England, Midlands, London, North East and Yorkshire, North West, South East, South West); any evidence of prior SARS-CoV-2 infection (positive SARS-CoV-2 test, "probable" infection documented in primary care, or COVID-19 hospitalisation); clinical risk groups used for prioritisation (clinically extremely vulnerable, clinically at-risk, neither); age, within 3 years, and by age groups used for prioritisation.
+The matching and exclusion criteria jointly ensure that treatment groups are also matched within JCVI priority group.
 
 People selected as controls were not eligible to be included again as a control in a subsequent trial, but were eligible for selection into the treatment group of a subsequent trial.
+Any unmatched boosted people were dropped.
 
@@
 
+## Additional variables
+
+The following additional variables were included as covariates in the Cox models (see below) or else to describe the treatment groups: sex (male or female); English Index of Multiple Deprivation (IMD, grouped by quintile); ethnicity (Black, Mixed, South Asian, White, Other, as per the UK census); morbidity count (diabetes, BMI > 40kg/m^2^, chronic heart disease, chronic kidney disease, chronic liver disease, chronic respiratory disease or severe asthma, chronic neurological disease); learning disabilities; severe mental illness; the number of SARS-CoV-2 tests in the 6 months prior to the study start date; elective in-hospital episode at the time of recruitment.
+
+The supplementary materials provide further details of the codelists and data sources used for all variables in the study.
+
 ## Statistical Analysis
@@
 We use Cox models, stratified by recruitment day, NHS region, and brand of second vaccine dose, to estimate hazard ratios comparing boosted with unboosted people.
-We estimated effectveness over all follow-up time (0-70 days), and separately in days 0-6, 7-13, 14-27, 28-41, 42-70.
+We estimated effectveness over all follow-up time (1-70 days), and separately in days 1-7, 8-14, 15-28, 29-42, 43-70.
 Models were adjusted for additional variables not used as matching factors, as described below.
@@
 
-We additionally estimated booster effectiveness in the following subgroups: primary vaccine course (both BNT162b2 or both ChAdOx1-S); age (18-64 versus 65 and over); clinically extremely vulnerable.
-Analyses were conducted independently in each subgroup.
+We also estimated booster effectiveness in the following subgroups: primary vaccine course (both BNT162b2 or both ChAdOx1-S); age (18-64 or 65 and over); clinically extremely vulnerable.
+Analyses were conducted separately in each subgroup.
 
-Booster effectiveness is defined as one minus the hazard ratio, and expressed as a percentage.
-
-## Covariates
-
-The following variables were included as covariates: sex (male or female); English Index of Multiple Deprivation (IMD, grouped by quintile); ethnicity (Black, Mixed, South Asian, White, Other, as per the UK census); morbidity count (diabetes, BMI > 40, chronic heart disease, chronic kidney disease, chronic liver disease, chronic respiratory disease or severe asthma, chronic neurological disease); learning disabilities; severe mental illness; the number of SARS-CoV-2 tests in the 6 months prior to the study start date; elective in-hospital episode at the time of recruitment.
-
-The supplementary materials provide further details of the codelists and data sources used for all variables in the study.
+Booster effectiveness is defined as one minus the hazard ratio, expressed as a percentage.
 
@@
 
-Total follow-up and incidence rates for each outcome, by treatment group. Cumulative incidences are provided in Supplementary Figure S2.
+Total follow-up and incidence rates for each outcome, by treatment group. The cumulative incidence of each outcome is provided in Supplementary Figure S2.
 
@@
     treatment_descr = paste0(
-      treatment_descr, " - ",
-      label_number(1, big.mark=",")(n_eligible), "eligible, ",
-      label_number(1, big.mark=",")(n_1), "matched (",
+      #treatment_descr, " - ",
+      label_number(1, big.mark=",")(n_eligible), " eligible, ",
+      label_number(1, big.mark=",")(n_1), " matched (",
       label_percent(0.1)(prop_matched_eligible), ")."
@@
 
+### Main analysis
+
 Participants in the emulated trial experienced `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="postest", fup_period=="All") %>% pull(events))` documented infections, `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="covidadmitted", fup_period=="All") %>% pull(events))` COVID-19 hospital admissions, and `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="coviddeath", fup_period=="All") %>% pull(events))` COVID-19 deaths across `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(fup_years))` person-years of follow-up (Table 2).
 
-<!-- `r filter(incidences, treatment=="pfizer", outcome=="covidattendance", fup_period=="all") %>% pull(events)` COVID-19 A&E attendances, `r filter(incidences, treatment=="pfizer", outcome=="covidcc", fup_period=="all") %>% pull(events)` COVID-19 ciritical care admissions, -->
+The estimated effectiveness (95% CI) comparing a BNT162b2 booster dose to two doses only over all follow-up time was `r filter(overallhazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3) %>% pull(ve_display)` for documented infection, `r filter(overallhazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(overallhazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 death.
+Effectiveness before and after 28 days since boosting is also estimated (Figure 2, Table 3), as well as for narrower periods (Supplementary Table S3).
 
-In the 14-20 days period of follow-up, the estimated effectiveness (95% CI) comparing a BNT162b2 booster dose to two doses only, was `r filter(hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for documented infection, `r filter(hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 death (Figure 2).
-
-For days 42-70, estimated effectiveness (95% CI) comparing a BNT162b2 booster dose to two doses only, was `r filter(hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_left==42) %>% pull(ve_display)` for documented infection, `r filter(hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_left==42) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_left==42) %>% pull(ve_display)` for COVID-19 death (Figure 2), suggesting a potential waning of booster effectiveness with time.
-<!-- `r filter(hazards, subgroup=="none", outcome=="covidattendance", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 A&E attendances, `r filter(hazards, subgroup=="none", outcome=="covidcc", treatment=="pfizer", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 critical care admission,  -->
-
-<!-- For mRNA-1273, the corresponding estimated effectiveness were `r filter(hazards, subgroup=="none", outcome=="postest", treatment=="moderna", model==3, term_left==14) %>% pull(ve_display)`, for COVID-19 A&E attendances was `r filter(hazards, subgroup=="none", outcome=="covidattendance", treatment=="moderna", model==3, term_left==14) %>% pull(ve_display)`, for COVID-19 hospital admission was `r filter(hazards, subgroup=="none", outcome=="covidadmitted", treatment=="moderna", term_left==14) %>% pull(ve_display)`, for COVID-19 critical care admission was `r filter(hazards, subgroup=="none", outcome=="covidcc", treatment=="moderna", model==3, term_left==14) %>% pull(ve_display)`, and for COVID-19 death was `r filter(hazards, subgroup=="none", outcome=="coviddeath", treatment=="moderna", model==3, term_left==14) %>% pull(ve_display)`.
+<!-- In the first 28 days of follow-up the estimated effectiveness (95% CI) comparing a BNT162b2 booster dose to two doses only, was `r filter(meta2hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` for documented infection, `r filter(meta2hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(meta2hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` for COVID-19 death (Figure 2).
 -->
 
-For the same period, the estimated effectiveness (95% CI) comparing a BNT162b2 booster dose to two doses only separately for 2× BNT162b2 and 2× ChAdOx1-S primary course recipients respectively were `r filter(hazards, outcome=="postest", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3, term_left==14) %>% pull(ve_display)` and `r filter(hazards, outcome=="postest", treatment=="pfizer", subgroup_level=="az-az", model==3, term_left==14) %>% pull(ve_display)` for documented infection, `r filter(hazards, outcome=="covidadmitted", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3, term_left==14) %>% pull(ve_display)` and `r filter(hazards, outcome=="covidadmitted", treatment=="pfizer", subgroup_level=="az-az", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(hazards, outcome=="coviddeath", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3, term_left==14) %>% pull(ve_display)` and `r filter(hazards, outcome=="coviddeath", treatment=="pfizer", subgroup_level=="az-az", model==3, term_left==14) %>% pull(ve_display)` for COVID-19 death (Figure 2).
+<!-- For days 28-70, estimated effectiveness (95% CI) comparing a BNT162b2 booster dose to two doses only, was `r filter(meta2hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` for documented infection, `r filter(meta2hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(meta2hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` for COVID-19 death (Figure 2), suggesting a potential waning of booster effectiveness with time.
+-->
+
+### Subgroup analyses
+
+The estimated effectiveness (95% CI) for 2× BNT162b2 and 2× ChAdOx1-S primary course recipients respectively were `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3) %>% pull(ve_display)` and `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup_level=="az-az", model==3) %>% pull(ve_display)` for documented infection, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3) %>% pull(ve_display)` and `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup_level=="az-az", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3) %>% pull(ve_display)` and `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup_level=="az-az", model==3) %>% pull(ve_display)` for COVID-19 death (Figure 2).
+
+Period-specific estimates are also provided.
 
@@
 
-For each outcome based on the fully adjusted model, the period-specific hazard ratios for boosting with BNT162b2 versus not boosting is shown, stratified by primary course.
+For each outcome based on the fully adjusted model, the hazard ratios for boosting with BNT162b2 versus not boosting is shown, stratified by primary course and time since boosting.
 Models with less extensive confounder adjustment are provided in supplementary materials (Figure S3).
@@
 
+  overallhazards_plotdata <- 
+    overallhazards %>%
+    filter(
+      model==3, 
+      outcome %in% outcomes,
+      treatment %in% c("pfizer"),
+      subgroup_variable %in% c("none", "vax12_type")
+    )
+
+  postbaselinecuts28 <- c(0,28,70)
+
   plot_hazards <-
-    hazards3 %>%
+    meta2hazards %>%
     filter(
+      model==3,
       outcome %in% outcomes,
-      treatment %in% c("pfizer")
+      treatment %in% c("pfizer"),
+      subgroup_variable %in% c("none", "vax12_type")
     ) %>%
@@
     geom_linerange(aes(ymin=hr.ll, ymax=hr.ul, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+
-    geom_point(data = filter(metahazards, model==3), mapping = aes(y=hr, x=max(postbaselinecuts+14)), colour="grey") +
-    geom_linerange(data = filter(metahazards, model==3), mapping = aes(ymin=hr.ll, ymax=hr.ul, x=max(postbaselinecuts+14)), colour="grey") +
+    geom_point(data = overallhazards_plotdata, mapping = aes(y=hr, x=max(postbaselinecuts28+14)), colour="darkgrey") +
+    geom_linerange(data = overallhazards_plotdata, mapping = aes(ymin=hr.ll, ymax=hr.ul, x=max(postbaselinecuts28+14)), colour="darkgrey") +
+    geom_rect(data = overallhazards_plotdata, ymin=-Inf, ymax=Inf, xmin=max(postbaselinecuts28), xmax=max(postbaselinecuts28+28), fill="grey95", alpha=0.5, colour="transparent") +
     geom_hline(aes(yintercept=1), colour='grey')+
     facet_grid(rows=vars(outcome_descr_wrap), cols=vars(subgroup_level_descr), switch="y")+
-    scale_y_log10(
-      breaks=y_breaks,
-      limits = c(0.01, 1.5),
+    # scale_y_log10(
+    #   breaks=y_breaks,
+    #   limits = c(0.03, 1.5),
+    #   oob = scales::oob_keep,
+    #   sec.axis = sec_axis(
+    #     ~(1-.),
+    #     name="Effectiveness (%)",
+    #     breaks = 1-(y_breaks),
+    #     labels = function(x){formatpercent100(x, 1)}
+    #   )
+    # )+
+    scale_y_reverse(
+      limits = c(1.1, 0),
+      breaks = c(1, 0.75, 0.50, 0.25, 0),
+      #expand = expansion(c(NA, 0)),
       oob = scales::oob_keep,
@@
         name="Effectiveness (%)",
-        breaks = 1-(y_breaks),
-        labels = function(x){formatpercent100(x, 1)}
+        labels= scales::label_percent(1, suffix="")
       )
     )+
-    scale_x_continuous(breaks=c(postbaselinecuts, max(postbaselinecuts+14)), labels=c(postbaselinecuts, "Overall"), limits=c(min(postbaselinecuts), max(postbaselinecuts+14)+1), expand = c(0, 0))+
+    scale_x_continuous(breaks=c(postbaselinecuts28, max(postbaselinecuts28+14)), labels=c(postbaselinecuts28, "All"), limits=c(min(postbaselinecuts28), max(postbaselinecuts28+14)+14), expand = c(0, 0))+
     scale_colour_brewer(type="qual", palette="Set2", guide=guide_legend(ncol=1))+
@@
 
+## Table 3: Estimated booster effectiveness
+
+Main and subgroup analyses
+```{r table_effects, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}
+
+bind_rows(
+  meta2hazards,
+  overallhazards %>% mutate(term_left = 0, term_right=70)
+) %>%
+  filter(model==3, treatment=="pfizer") %>%
+  mutate(
+    period = fct_inorder(paste(term_left+1, "-", term_right)),
+    ve_display = str_remove(ve_display, fixed("%"))
+  ) %>%
+  arrange(treatment_descr, outcome_descr, subgroup_level_descr, period) %>%
+  group_by(outcome, treatment, subgroup_level_descr, model) %>%
+  mutate(
+    subgroup_level_descr=if_else(row_number()==1, subgroup_level_descr, factor(NA, levels=levels(subgroup_level_descr)))
+  ) %>%
+  ungroup() %>%
+  select(outcome_descr, subgroup_level_descr, period, hr_display, ve_display) %>%
+  gt(
+    groupname_col = c("outcome_descr"),
+  ) %>%
+  cols_label(
+    #treatment_descr = "Treatment",
+    subgroup_level_descr="",
+    outcome_descr = "Outcome",
+
+    period = "Days since booster",
+    hr_display   = "HR (95% CI)",
+    ve_display   = "VE (95% CI)",
+  ) %>%
+  cols_align(
+    align = "right",
+    columns = everything()
+  ) %>%
+  fmt_missing(
+    c("subgroup_level_descr"),
+    missing_text=""
+  )
+
+```
+
+\newpage
+
 # Discussion
@@
 
+We have considered effectiveness in a population where Delta was the predominant circulating variant @COVID19OmicronDaily.
+Re-running the analysis against a backdrop of the now globally-dominant Omicron variant will be undertaken where valid concurrent comparisons between boosted and unboosted people can be made.
+Further, mRNA-1273 boosting begun later than for BNT162b2 so we did not evaluate effectiveness of this vaccine due to insufficient follow-up.
+An analysis of mRNA-1273 boosting and comparative effectiveness between mRNA-1273 and BNT162b2 is planned.
+
 ### Strengths and Limitations
 
-We used routinely-collected health records with comprehensive coverage of primary care, hospital admissions, COVID-19 testing and vaccination, and death registrations, and with substantial power to estimate effectiveness of booster doses against severe but rare outcomes such as hospitalisation and death over time.
+We used routinely-collected health records with comprehensive coverage of primary care, hospital admissions, COVID-19 testing and vaccination, and death registrations.
+This provides substantial power to estimate effectiveness of booster doses against severe but rare outcomes such as hospitalisation and death, including period-specific effects to investigate waning protection.
 
@@
 We may not have all records of negative tests so counts of number of tests are also incomplete.
-
-Hospital admission records are only completed after discharge, so some longer hospital stays may not be counted.
+Hospital admission records are only completed after discharge, so some very long hospital stays commencing within the follow-up period may not have been included.
 
 We excluded a number of groups, such as health care workers and care home residents, where testing, vaccination, and infection risk characteristics were unusual or had substantial within-group heterogeneity that could not be adequately accounted for.
-
-This study considers effectiveness in a population where Delta was the predominant circulating variant @COVID19OmicronDaily.
-Re-running the analysis against a backdrop of the now globally-dominant Omicron variant will be undertaken where valid concurrent comparisons between boosted and unboosted people can be made.
-
-Finally, mRNA-1273 boosting begun later than for BNT162b2, and so we could not evaluate effectiveness of this vaccine due to insufficient follow-up.
-An analysis of mRNA-1273 boosting and comparative effectiveness between mRNA-1273 and BNT162b2 is planned.
+The generalisability of our results to these excluded groups is unclear.
 
@@
 Andrews et al.
-reported estimates of relative vaccine effectiveness in the 14 days after a BNT162b2 booster dose in those aged over 50 years of 87.4% (95%CI 84.9-89.4) where the primary course was ChAdOx1-S and 84.4% (95%CI 82.8-85.8) for BNT162b2 @andrews2021.
+reported estimates of relative vaccine effectiveness against syptomatic disease in the 14 days after a BNT162b2 booster dose in those aged over 50 years of 87.4% (95%CI 84.9-89.4) where the primary course was ChAdOx1-S and 84.4% (95%CI 82.8-85.8) for BNT162b2 @andrews2021.
 Sheikh et al.
@@
 
-<!-- Elsewhere, studies in Israel found .. -->
+Elsewhere, a study in Israeli health registry data @barda2021 found strong protection against admission (1 – risk ratio (95%CI) = 93% (88-97)), severe disease (92% (82-97)), and death (81% (59-97)), with similar results in specific demographic and clinical subgroups.
+but did not consider differences @barda2021
 
@@
 Taken together, these provide the legal bases to link patient datasets on the OpenSAFELY platform.
-GP practices, from which the primary care data are obtained, are required to share relevant health information to support the public health response to the pandemic, and have been informed of the OpenSAFELY analytics platform.
+General Practices, from which the primary care data are obtained, are required to share relevant health information to support the public health response to the pandemic, and have been informed of the OpenSAFELY analytics platform.
 

### 2022-03-03 manuscript/draft-supplement.Rmd
lots of updates to manuscript

@@
 # where are the outputs (ie the inputs for this manuscript!) saved?
-output_dir_os <- here("output", "manuscript-objects")
-#output_dir_os <- here("released_outputs", "output", "manuscript-objects")
+#output_dir_os <- here("output", "manuscript-objects")
+output_dir_os <- here("released_output",  "manuscript-objects")
 
@@
 
-flowchart <- read_csv(fs::path(output_dir_os, "flowchart.csv"))
-cohortsummary <- read_csv(fs::path(output_dir_os, "cohortsummary.csv"))
 matchsummary <- read_csv(fs::path(output_dir_os, "matchsummary.csv"))
@@
 matchflowchart <- read_csv(fs::path(output_dir_os, "matchflowchart.csv"))
-incidences <- read_csv(fs::path(output_dir_os, "incidence.csv")) %>% mutate(events=events_1+events_0)
+incidences <- read_csv(fs::path(output_dir_os, "none", "incidence_all.csv")) %>% mutate(events=events_1+events_0)
 smd <- read_csv(fs::path(output_dir_os, "matchsmd.csv"))
@@
 
-km_none <- read_csv(fs::path(output_dir_os, "none", "km.csv"))  
-km_vax12_type <- read_csv(fs::path(output_dir_os, "vax12_type", "km.csv")) 
 km <- 
-  bind_rows(km_none, km_vax12_type) %>%
+  bind_rows(
+    read_csv(fs::path(output_dir_os, "none", "km.csv")),
+    read_csv(fs::path(output_dir_os, "vax12_type", "km.csv")) ,
+    #read_csv(fs::path(output_dir_os, "cev", "km.csv")) ,
+    #read_csv(fs::path(output_dir_os, "age65plus", "km.csv")) ,
+  ) %>%
   mutate(
@@
 
-hazards_none <- read_csv(fs::path(output_dir_os, "none", "effects.csv"))
-hazards_vax12_type <- read_csv(fs::path(output_dir_os, "vax12_type", "effects.csv"))
 hazards <- 
-  bind_rows(hazards_none, hazards_vax12_type) %>%
+  bind_rows(
+    read_csv(fs::path(output_dir_os, "none", "effects.csv")),
+    read_csv(fs::path(output_dir_os, "vax12_type", "effects.csv")),
+    #read_csv(fs::path(output_dir_os, "cev", "effects.csv")),
+    #read_csv(fs::path(output_dir_os, "age65plus", "effects.csv")),
+  ) %>%
   mutate(
@@
 
-Note that we excluded care home residents and health care workers in our analysis, so group 1 is not included and group 2 includes only those aged 80 and over.
+Note that we excluded care home residents and health care workers in our analysis, so members of JCVI group 1 are not included and JCVI group 2 includes only those aged 80 and over.
 The original priority group list has 9 groups, with a 10th group implicitly defined as "everybody else".
@@
 (https://www.gov.uk/government/news/jcvi-issues-advice-on-covid-19-booster-vaccines-for-those-aged-40-to-49-and-second-doses-for-16-to-17-year-olds)
+
+# Variables and codelists
+
+```{r, codelists, echo=FALSE, message=FALSE, warning=FALSE}
+
+variable_table <- readxl::read_xlsx(here("lib", "design", "variable-descriptions.xlsx"))
+
+variable_table %>%
+  gt(groupname_col = "category") %>%
+  cols_label(
+    name = "Variable",
+    source = "Data source",
+    description = "Notes",
+    support = "Values ",
+    codelistID = "Codelist identifier"
+  ) %>%
+   fmt_missing(
+    everything(),
+    missing_text=""
+  )
+
+```
+
+Codelists can be found at `https://codelists.opensafely.org/codelist/<ID>`, substituting `<ID>` for the codelist identifier in the table above.
 
@@
 plot_smd <-
-  ggplot(data_smd)+
+  smd %>%
+  mutate(
+    level=fct_rev(fct_inorder(level))
+  ) %>%
+  ggplot()+
   geom_point(aes(x=smd, y=level))+
@@
 plot_km <- km %>%
+  filter(
+    outcome %in% outcomes,
+    treatment=="pfizer"
+  ) %>%
     ggplot()+
@@
     geom_point(aes(y=hr, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+
-    geom_linerange(aes(ymin=hr.ll, ymax=hr.ul, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+
+    geom_linerange(aes(ymin=hr.ll, ymax=hr.ul, x=term_midpoint, colour=model_descr), position = position_dodge(width = 4))+
     geom_hline(aes(yintercept=1), colour='grey')+

### 2022-03-03 manuscript/reports.Rmd
lots of updates to manuscript

---
title: "Effectiveness of BNT162b2 booster doses in England: a observational study in OpenSAFELY-TPP: Materials"
output:
  html_document:
    keep_md: yes
    self_contained: yes
  word_document: default
  bookdown::html_document2:
    number_sections: no
    toc: no
  pdf_document:
    toc: no
bibliography: references.bib
csl: vancouver.csl
link-citations: yes
zotero: yes
header-includes:
- \usepackage{float} 
- \floatplacement{figure}{H}
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
---

```{r setup, include=FALSE}

library('tidyverse')
library('here')
library('glue')
library('lubridate')
library('gt')
library('patchwork')
library('scales')
source(here("lib", "functions", "utility.R"))

# where are the outputs (ie the inputs for this manuscript!) saved?
#output_dir_os <- here("output", "manuscript-objects")
output_dir_os <- here("released_output", "manuscript-objects")

# where should we put the objects created within this rmd script?
output_dir_rmd <- here("manuscript", "figures")
fs::dir_create(output_dir_rmd)

# only applicable if `self-contained: yes` option is enables in yaml header
knitr::opts_chunk$set(
  echo = TRUE,
  fig.path = paste0(output_dir_rmd, "/"),
  fig.pos = 'H' #to stop figures floating around when rendering to pdf
)

study_dates <-
  jsonlite::read_json(path=here("lib", "design", "study-dates.json")) %>%
  map(as.Date)

postbaselinecuts <- read_rds(here("lib", "design", "postbaselinecuts.rds"))

cohortsummary <- read_csv(fs::path(output_dir_os, "cohortsummary.csv"))
matchsummary <- read_csv(fs::path(output_dir_os, "matchsummary.csv"))
matchsummary_treated <- read_csv(fs::path(output_dir_os, "matchsummary_treated.csv")) %>% filter(treated == 1L)
matchsummary_control <- read_csv(fs::path(output_dir_os, "matchsummary_treated.csv")) %>% filter(treated == 0L)
matchflowchart <- read_csv(fs::path(output_dir_os, "matchflowchart.csv"))
matchcoverage <- read_csv(fs::path(output_dir_os, "matchcoverage.csv"))
incidences <- read_csv(fs::path(output_dir_os, "none", "incidence_all.csv")) %>% mutate(events=events_1+events_0)

studydurationweeks <- as.numeric(as.duration(study_dates$studyend_date - study_dates$studystart_date +1) , "weeks")
```

```{r, effects, echo=FALSE, message=FALSE, warning=FALSE}

  specify_decimal <- function(x, k, trim=FALSE) {

    fmtd <- format(round(x, k), nsmall = k)
    if (trim) {fmtd <- trimws(fmtd)}
    return(fmtd)
  }
  #
  # print_est1bracket <- function(x, b, round=1){
  #   paste0(specify_decimal(x, round), " (", specify_decimal(b, round), ")")
  # }
  #
  print_est2bracket <- function(x, b1, b2, round=1){
    paste0(specify_decimal(x, round), " (", specify_decimal(b1, round), ", ", specify_decimal(b2, round), ")")
  }

  outcomes <- 
    c(
      "postest",
      #"covidemergency",
      "covidadmitted",
      #"covidcc",
      "coviddeath",
      "noncoviddeath"
    ) %>% 
    set_names(., .)

km <- 
  bind_rows(
    read_csv(fs::path(output_dir_os, "none", "km.csv")),
    read_csv(fs::path(output_dir_os, "vax12_type", "km.csv")),
    read_csv(fs::path(output_dir_os, "cev", "km.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
    read_csv(fs::path(output_dir_os, "age65plus", "km.csv")) %>% mutate(subgroup_level=as.character(subgroup_level))
  ) %>%
  filter(
    outcome %in% outcomes
  ) %>%
  mutate(
    subgroup_descr = fct_inorder(subgroup_descr),
    subgroup_level = fct_inorder(subgroup_level),
    subgroup_level_descr = fct_inorder(fct_explicit_na(subgroup_level_descr, "All")),
    treatment_descr = fct_inorder(treatment_descr),
    outcome_descr = fct_inorder(outcome_descr),
    outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=12)),
    #model_name = fct_inorder(str_wrap(model_name, 15)),
  )
# 
# cmlinc <- read_csv(fs::path(output_dir_os, "cmlinc.csv")) %>%
#   mutate(
#     treatment_descr = fct_inorder(treatment_descr),
#     outcome_descr = fct_inorder(outcome_descr),
#     model_name = fct_inorder(str_wrap(model_name, 15)),
#     inc_CI_to = paste0(label_number(0.1, scale=1000)(1-survival.ul), " to ", label_number(0.1, scale=1000)(1-survival.ll))
#   )

# riskdiff <- read_csv(fs::path(output_dir_os, "riskdiff.csv")) %>%
#   mutate(
#     treatment_descr = fct_inorder(treatment_descr),
#     outcome_descr = fct_inorder(outcome_descr),
#     model_name = fct_inorder(str_wrap(model_name, 15)),
#     diff_CI_to = paste0(label_number(0.01, scale=1000)(diff.ll), " to ", label_number(0.01, scale=1000)(diff.ul))
#   )

hazards <- 
  bind_rows(
    read_csv(fs::path(output_dir_os, "none", "effects.csv")),
    read_csv(fs::path(output_dir_os, "vax12_type", "effects.csv")),
    read_csv(fs::path(output_dir_os, "cev", "effects.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
    read_csv(fs::path(output_dir_os, "age65plus", "effects.csv"))  %>% mutate(subgroup_level=as.character(subgroup_level)),
  ) %>%
  filter(
    outcome %in% outcomes
  ) %>%
  mutate(

    subgroup_descr = fct_inorder(subgroup_descr),
    subgroup_level = fct_inorder(subgroup_level),
    subgroup_level_descr = fct_inorder(fct_explicit_na(subgroup_level_descr, "All")),
    treatment_descr = fct_inorder(treatment_descr),
    outcome_descr = fct_inorder(outcome_descr),
    outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=15)),
    #model_name = fct_inorder(str_wrap(model_name, 15)),
    hr_display = paste0(label_number(accuracy=0.01)(hr), " (", label_number(accuracy=0.01)(hr.ll), "-", label_number(accuracy=0.01)(hr.ul), ")"),
    ve_display = paste0(label_percent(accuracy=0.1)(ve), " (", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),

  )

overallhazards <- 
  bind_rows(
    read_csv(fs::path(output_dir_os, "none", "overalleffects.csv")),
    read_csv(fs::path(output_dir_os, "vax12_type", "overalleffects.csv")),
    read_csv(fs::path(output_dir_os, "cev", "overalleffects.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
    read_csv(fs::path(output_dir_os, "age65plus", "overalleffects.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
  ) %>%
  filter(
    outcome %in% outcomes
  ) %>%
  mutate(

    subgroup_descr = fct_inorder(subgroup_descr),
    subgroup_level = fct_inorder(subgroup_level),
    subgroup_level_descr = fct_inorder(fct_explicit_na(subgroup_level_descr, "All")),
    treatment_descr = fct_inorder(treatment_descr),
    outcome_descr = fct_inorder(outcome_descr),
    outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=15)),
    #model_name = fct_inorder(str_wrap(model_name, 15)),
    hr_display = paste0(label_number(accuracy=0.01)(hr), " (", label_number(accuracy=0.01)(hr.ll), "-", label_number(accuracy=0.01)(hr.ul), ")"),
    ve_display = paste0(label_percent(accuracy=0.1)(ve), " (", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
  )

meta28hazards <- 
  # derive directly from hazards as meta28effects table needs to be re-run
  bind_rows(
    read_csv(fs::path(output_dir_os, "none", "effects.csv")),
    read_csv(fs::path(output_dir_os, "vax12_type", "effects.csv")),
    read_csv(fs::path(output_dir_os, "cev", "effects.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
    read_csv(fs::path(output_dir_os, "age65plus", "effects.csv"))  %>% mutate(subgroup_level=as.character(subgroup_level)),
  ) %>%
  filter(
    outcome %in% outcomes
  ) %>%
  mutate(
    period = (term_right>28) + 1,
    subgroup_descr = fct_inorder(subgroup_descr),
    subgroup_level = fct_inorder(subgroup_level),
    subgroup_level_descr = fct_inorder(fct_explicit_na(subgroup_level_descr, "All")),
    treatment_descr = fct_inorder(treatment_descr),
    outcome_descr = fct_inorder(outcome_descr),
    outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=15)),
  ) %>%
  group_by(
    treatment, treatment_descr, outcome, outcome_descr, outcome_descr_wrap, subgroup, subgroup_variable, subgroup_descr, subgroup_level, subgroup_level_descr, model, model_descr, period, 
  ) %>%
  summarise(
    term_left = min(term_left),
    term_right = max(term_right),
    term_midpoint = (term_left+term_right)/2,
    estimate = weighted.mean(estimate, robust.se^-2),
    std.error = sqrt(1/sum(std.error^-2)),
    robust.se = sqrt(1/sum(robust.se^-2)),
    statistic = estimate/robust.se,
    p.value = 2 * pmin(pnorm(statistic), pnorm(-statistic)),
    conf.low = estimate + qnorm(0.025)*robust.se,
    conf.high = estimate + qnorm(0.975)*robust.se,
    hr = exp(estimate),
    hr.ll = exp(conf.low),
    hr.ul = exp(conf.high),
    ve = 1-hr,
    ve.ll = 1-hr.ul,
    ve.ul = 1-hr.ll
  ) %>%
  ungroup() %>%
  mutate(
    hr_display = paste0(label_number(accuracy=0.01)(hr), " (", label_number(accuracy=0.01)(hr.ll), "-", label_number(accuracy=0.01)(hr.ul), ")"),
    ve_display = paste0(label_percent(accuracy=0.1)(ve), " (", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
  )

meta42hazards <- 
  # derive directly from hazards as meta28effects table needs to be re-run
  bind_rows(
    read_csv(fs::path(output_dir_os, "none", "effects.csv")),
    read_csv(fs::path(output_dir_os, "vax12_type", "effects.csv")),
    read_csv(fs::path(output_dir_os, "cev", "effects.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
    read_csv(fs::path(output_dir_os, "age65plus", "effects.csv"))  %>% mutate(subgroup_level=as.character(subgroup_level)),
  ) %>%
  filter(
    outcome %in% outcomes
  ) %>%
  mutate(
    period = (term_right>42) + 1,
    subgroup_descr = fct_inorder(subgroup_descr),
    subgroup_level = fct_inorder(subgroup_level),
    subgroup_level_descr = fct_inorder(fct_explicit_na(subgroup_level_descr, "All")),
    treatment_descr = fct_inorder(treatment_descr),
    outcome_descr = fct_inorder(outcome_descr),
    outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=15)),
  ) %>%
  group_by(
    treatment, treatment_descr, outcome, outcome_descr, outcome_descr_wrap, subgroup, subgroup_variable, subgroup_descr, subgroup_level, subgroup_level_descr, model, model_descr, period, 
  ) %>%
  summarise(
    term_left = min(term_left),
    term_right = max(term_right),
    term_midpoint = (term_left+term_right)/2,
    estimate = weighted.mean(estimate, robust.se^-2),
    std.error = sqrt(1/sum(std.error^-2)),
    robust.se = sqrt(1/sum(robust.se^-2)),
    statistic = estimate/robust.se,
    p.value = 2 * pmin(pnorm(statistic), pnorm(-statistic)),
    conf.low = estimate + qnorm(0.025)*robust.se,
    conf.high = estimate + qnorm(0.975)*robust.se,
    hr = exp(estimate),
    hr.ll = exp(conf.low),
    hr.ul = exp(conf.high),
    ve = 1-hr,
    ve.ll = 1-hr.ul,
    ve.ul = 1-hr.ll
  ) %>%
  ungroup() %>%
  mutate(
    hr_display = paste0(label_number(accuracy=0.01)(hr), " (", label_number(accuracy=0.01)(hr.ll), "-", label_number(accuracy=0.01)(hr.ul), ")"),
    ve_display = paste0(label_percent(accuracy=0.1)(ve), " (", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
  )

hazards3 <- hazards %>% filter(model==3)

```

\newpage

<!-- TODO -->

<!-- * full description of booster eligiblity over time -->

<!-- * figure 1 chart with SMD / %-point differences -->

<!-- * check for moderna references and remove if necessary -->

<!-- * check final recruitment day and include in results -->

<!-- * check if rolling average restrictions are ever used (need to output dataset) -->

<!-- * references -->

#### Table 0: Inclusion criteria

```{r flowchart, echo=FALSE, warning=FALSE, message=FALSE}

matchflowchart %>%
  transmute(
    treatment,
    stage,
    criteria = str_wrap(criteria, 40),
    n,
    n_exclude,
    pct_all
  ) %>%
  pivot_wider(
    id_cols = c(stage, criteria),
    names_from = treatment,
    values_from = c(n, n_exclude, pct_all)
  ) %>%
  gt() %>%
  cols_label(
    stage="",
    criteria = "",

    n_pfizer = "N",
    #n_moderna = "N",

    n_exclude_pfizer = "N excluded",
    #n_exclude_moderna = "N excluded",

    pct_all_pfizer = "% remaining",
    #pct_all_moderna = "% remaining",
  ) %>%
  tab_spanner(
    label = "BNT162b2",
    columns = ends_with("pfizer")
  ) %>%
  # tab_spanner(
  #   label = "mRNA-1273",
  #   columns = ends_with("moderna")
  # ) %>%
  fmt_percent(
    columns = starts_with(c("pct_")),
    decimals = 1
  ) %>%
  fmt_number(
    columns = starts_with(c("n_")),
    decimals= 0
  ) %>%
  fmt_missing(
    everything(),
    missing_text="-"
  ) %>%
  tab_style(
    style=cell_text(whitespace ="pre-wrap"),
    #style=cell_text(style ="italic"),
    locations = cells_body(columns = "criteria")
  )
```

#### Table 1: Participant characteristics

Participant characteristics as on the day of recruitment into the treatment or control group.

```{r table1, echo=FALSE, warning=FALSE, message=FALSE}
table1_body <- read_csv(fs::path(output_dir_os, "matchtable1.csv")) %>% filter(treatment=="pfizer")
table1_by <- read_csv(fs::path(output_dir_os, "matchtable1by.csv")) %>% filter(treatment=="pfizer")

table1 <- 
  table1_body %>%
  #filter(outcome %in% outcomes) %>%
  group_by(variable) %>%
  mutate(
    label = if_else(row_number()==1, label, str_c("&nbsp;&nbsp;&nbsp;", label)),
  ) %>%
  mutate(
    across(
      starts_with("stat_"),
      ~if_else(is.na(.x) & row_number()==1, "", .x)
    )
  ) %>%
  ungroup() %>%
  mutate(
    order = cumsum(lag(variable, 1, first(variable))!=variable)
  ) %>%
  select(
    variable, var_label, label, starts_with("stat_"), order
  ) %>%
  rename(!!!set_names(table1_by$by_col, table1_by$by_chr))

table1 %>%
  gt(groupname_col="treatment_descr") %>%
  fmt_markdown(columns = "label") %>%
  tab_style(
    style = list(
      cell_fill(color = "gray96")
    ),
    locations = cells_body(
      rows = order%%2 == 0
    )
  ) %>%
  tab_style(
    style = list(
      cell_text(weight = "bold")
    ),
    locations = cells_column_labels(everything())
  ) %>%
  cols_label(
     label=md("Characteristic")
  ) %>%
  cols_hide(c("variable", "var_label", "order"))
```

\newpage

#### Table 2: Follow-up and outcomes

Total follow-up and incidence rates for each outcome, by treatment group. The cumulative incidence of each outcome is provided in Supplementary Figure S2.

```{r table2, echo=FALSE, warning=FALSE, message=FALSE}
incidences %>%
  filter(
    fup_period == "All",
    outcome %in% c("postest", "covidadmitted", "coviddeath"),
    treatment == "pfizer"
  ) %>%
  left_join(
    matchsummary %>% select(treatment, n_eligible, prop_matched_eligible),
    by=c("treatment")
  ) %>%
  mutate(
    treatment_descr = paste0(
      #treatment_descr, " - ",
      label_number(1, big.mark=",")(n_eligible), " eligible, ",
      label_number(1, big.mark=",")(n_1), " matched (",
      label_percent(0.1)(prop_matched_eligible), ")."
    ),
  ) %>%
  select(
    treatment_descr, 
    outcome_descr,
    q_1,
    rate_1,
    q_0, 
    rate_0,
    rrECI
  ) %>%
  mutate(
    rate_1=if_else(rate_1<0.001, "<0.001", label_number(0.001)(rate_1)),
    rate_0=if_else(rate_0<0.001, "<0.001", label_number(0.001)(rate_0)),
  ) %>%
  gt(
   groupname_col = c("treatment_descr"),
  ) %>%
  cols_label(
    #treatment_descr = "Treatment",
    outcome_descr = "Outcome",

    q_0 = "Events / person-years",
    q_1   = "Events / person-years",

    rate_0 = "Incidence rate",
    rate_1 = "Incidence rate",
    rrECI = "Incidence rate ratio (95% CI)"
  ) %>%
  tab_spanner(
    label = "Boosted",
    columns = ends_with("1")
  ) %>%
  tab_spanner(
    label = "Unboosted",
    columns = ends_with("0")
  ) %>%
  fmt_missing(
    everything(),
    missing_text="--"
  ) %>%
  cols_align(
    align = "right",
    columns = everything()
  )

```
\newpage

## Table 3a: Estimated booster effectiveness (5 period)

Main and subgroup analyses
```{r table_effects, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}

hazards %>%
  filter(model==3, treatment=="pfizer") %>%
  mutate(
    period = paste(term_left+1, "-", term_right),
    ve_display = str_remove(ve_display, fixed("%"))
  ) %>%
  arrange(treatment_descr, outcome_descr, subgroup_level_descr, term_left) %>%
  group_by(outcome, treatment, subgroup_level_descr, model) %>%
  mutate(
    subgroup_level_descr=if_else(row_number()==1, subgroup_level_descr, factor(NA, levels=levels(subgroup_level_descr)))
  ) %>%
  ungroup() %>%
  select(outcome_descr, subgroup_level_descr, period, hr_display, ve_display) %>%
  gt(
    groupname_col = c("outcome_descr"),
  ) %>%
  cols_label(
    #treatment_descr = "Treatment",
    subgroup_level_descr="",
    outcome_descr = "Outcome",

    period = "Days since booster",
    hr_display   = "HR (95% CI)",
    ve_display   = "VE (95% CI)",
  ) %>%
  cols_align(
    align = "right",
    columns = everything()
  ) %>%
  fmt_missing(
    c("subgroup_level_descr"),
    missing_text=""
  )

```

\newpage

## Table 3b: Estimated booster effectiveness (2 period split at 28 days)

```{r table_effects28, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}

meta28hazards %>%
  filter(model==3, treatment=="pfizer") %>%
  mutate(
    period = paste(term_left+1, "-", term_right),
    ve_display = str_remove(ve_display, fixed("%"))
  ) %>%
  arrange(treatment_descr, outcome_descr, subgroup_level_descr, term_left) %>%
  group_by(outcome, treatment, subgroup_level_descr, model) %>%
  mutate(
    subgroup_level_descr=if_else(row_number()==1, subgroup_level_descr, factor(NA, levels=levels(subgroup_level_descr)))
  ) %>%
  ungroup() %>%
  select(outcome_descr, subgroup_level_descr, period, hr_display, ve_display) %>%
  gt(
    groupname_col = c("outcome_descr"),
  ) %>%
  cols_label(
    #treatment_descr = "Treatment",
    subgroup_level_descr="",
    outcome_descr = "Outcome",

    period = "Days since booster",
    hr_display   = "HR (95% CI)",
    ve_display   = "VE (95% CI)",
  ) %>%
  cols_align(
    align = "right",
    columns = everything()
  ) %>%
  fmt_missing(
    c("subgroup_level_descr"),
    missing_text=""
  )

```

\newpage

## Table 3c: Estimated booster effectiveness (2 period split at 42 days)

```{r table_effects42, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}

meta28hazards %>%
  filter(model==3, treatment=="pfizer") %>%
  mutate(
    period = paste(term_left+1, "-", term_right),
    ve_display = str_remove(ve_display, fixed("%"))
  ) %>%
  arrange(treatment_descr, outcome_descr, subgroup_level_descr, term_left) %>%
  group_by(outcome, treatment, subgroup_level_descr, model) %>%
  mutate(
    subgroup_level_descr=if_else(row_number()==1, subgroup_level_descr, factor(NA, levels=levels(subgroup_level_descr)))
  ) %>%
  ungroup() %>%
  select(outcome_descr, subgroup_level_descr, period, hr_display, ve_display) %>%
  gt(
    groupname_col = c("outcome_descr"),
  ) %>%
  cols_label(
    #treatment_descr = "Treatment",
    subgroup_level_descr="",
    outcome_descr = "Outcome",

    period = "Days since booster",
    hr_display   = "HR (95% CI)",
    ve_display   = "VE (95% CI)",
  ) %>%
  cols_align(
    align = "right",
    columns = everything()
  ) %>%
  fmt_missing(
    c("subgroup_level_descr"),
    missing_text=""
  )

```

\newpage

#### Figure 1: Booster vaccination uptake, treatment group eligibility and matching success

Cumulative booster vaccination over the duration of the study period, by matching eligibility and matching success. Separately by primary vaccination course.

```{r, vaxdate, echo=FALSE, message=FALSE, warning=FALSE, out.width='60%', out.height='50%', results='asis', out.extra=''}
#read_rds(fs::path(output_dir_os, "descriptive", "vaxdate", "plot_vaxdate_stack.rds"))

xmin <- min(matchcoverage$vax3_date )
xmax <- max(matchcoverage$vax3_date )+1
data_coverage <- matchcoverage  %>%
  mutate(
    status = factor(status, levels=c("ineligible", "unmatched", "matched")),
    status_descr = fct_relabel(status, str_to_title)
  ) %>%
  arrange(status) %>%
  mutate(
    status_descr = fct_inorder(status_descr)
  ) %>%
  filter(treatment=="pfizer") %>%
  mutate(
    # vax12_type = fct_case_when(
    #   vax12_type == "pfizer-pfizer" ~ "BNT162b2",
    #   vax12_type == "az-az" ~ "ChAdOx1-S",
    #   vax12_type == "moderna-moderna" ~ "mRNA-1273",
    #   TRUE ~ NA_character_
    # ),
    course = fct_case_when(
      vax12_type=="pfizer-pfizer" & treatment=="pfizer" ~ "2× BNT162b2  / 1× BNT162b2",
      vax12_type=="az-az" & treatment=="pfizer" ~ "2× ChAdOx1-S / 1× BNT162b2",
      vax12_type=="pfizer-pfizer" & treatment=="moderna" ~ "2× BNT162b2 / 1× mRNA-1273",
      vax12_type=="az-az" & treatment=="moderna" ~ "2× ChAdOx1-S / 1× mRNA-1273",
      TRUE ~ NA_character_
    )
  )

data_annotate <- 
  data_coverage %>%
  filter(vax12_type=="pfizer-pfizer") %>%
  group_by(status_descr, course) %>%
  summarise(
    xpos = first(lubridate::floor_date((vax3_date), "1 month")),
    ypos= max(cumuln)
  ) %>%
  ungroup() %>%
  arrange(desc(status_descr)) %>%
  mutate(
    ypos=(cumsum(ypos) + lag(cumsum(ypos), 1, 0))/2
  ) %>%
  mutate(
    course="2× ChAdOx1-S / 1× BNT162b2"
  )

ggplot(data_coverage) +
  geom_col(
    aes(
      x=vax3_date+0.5,
      y=cumuln,
      group=status_descr,
      fill=status_descr,
      colour=NULL
    ),
    position=position_stack(),
    alpha=0.5,
    width=1
  )+
  geom_rect(xmin=xmin, xmax= xmax+1, ymin=0, ymax=6, fill="grey", colour="transparent")+
  geom_text(data=data_annotate, mapping=aes(x = xpos, y = (ypos), label = status_descr, colour=status_descr), hjust=0, show.legend = FALSE)+
  facet_wrap(vars(course))+
  scale_x_date(
    breaks = unique(lubridate::ceiling_date(matchcoverage$vax3_date, "1 month")),
    limits = c(lubridate::floor_date((xmin), "1 month"), NA),
    labels = scales::label_date("%d/%m"),
    expand = expansion(add=1),
  )+
  scale_y_continuous(
    labels = scales::label_number(accuracy = 1, big.mark=","),
    expand = expansion(c(0, NA))
  )+
  scale_fill_brewer(type="qual", palette="Set2", direction = -1)+
  scale_colour_brewer(type="qual", palette="Set2", direction = -1)+
  labs(
    x="Date",
    y="Cumulative booster vaccines",
    colour=NULL,
    fill=NULL,
    alpha=NULL
  ) +
  theme_minimal()+
  theme(
    axis.line.x.bottom = element_line(),
    axis.text.x.top=element_text(hjust=0),
    strip.text.y.right = element_text(angle = 0),
    axis.ticks.x=element_line(),
    legend.position = "none"
  )+
  NULL
cat("  \n\n")
```

\newpage

#### Figure 2a: Estimated booster effectiveness (5 period)

For each outcome based on the fully adjusted model, the period-specific hazard ratios for boosting with BNT162b2 versus not boosting is shown, stratified by primary course.

```{r plot_effects, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}

formatpercent100 <- function(x,accuracy){
  formatx <- scales::label_percent(accuracy, suffix="")(x)

  if_else(
    formatx==scales::label_percent(accuracy, suffix="")(1),
    paste0(">",scales::label_percent(1, suffix="")((100-accuracy)/100)),
    formatx
  )
}

  ## hazard ratio

  y_breaks <- c(0.01, 0.02, 0.05, 0.1, 0.2, 0.5, 1, 2)

  overallhazards_plotdata <- 
    overallhazards %>%
    filter(
      model==3, 
      outcome %in% outcomes,
      treatment %in% c("pfizer"),
      subgroup_variable %in% c("none", "vax12_type")
    )

  plot_hazards <-
    hazards %>%
    filter(
      model==3,
      outcome %in% outcomes,
      treatment %in% c("pfizer"),
      subgroup_variable %in% c("none", "vax12_type")
    ) %>%
    ggplot()+
    geom_point(aes(y=hr, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+
    geom_linerange(aes(ymin=hr.ll, ymax=hr.ul, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+
    geom_point(data = overallhazards_plotdata, mapping = aes(y=hr, x=max(postbaselinecuts+14)), colour="darkgrey") +
    geom_linerange(data = overallhazards_plotdata, mapping = aes(ymin=hr.ll, ymax=hr.ul, x=max(postbaselinecuts+14)), colour="darkgrey") +
    geom_rect(data = overallhazards_plotdata, ymin=-Inf, ymax=Inf, xmin=max(postbaselinecuts+14)-7, xmax=max(postbaselinecuts+14)+7, fill="grey95", alpha=0.5, colour="transparent") +
    geom_hline(aes(yintercept=1), colour='grey')+
    facet_grid(rows=vars(outcome_descr_wrap), cols=vars(subgroup_level_descr), switch="y")+
    scale_y_log10(
      breaks=y_breaks,
      limits = c(0.01, 1.5),
      oob = scales::oob_keep,
      sec.axis = sec_axis(
        ~(1-.),
        name="Effectiveness (%)",
        breaks = 1-(y_breaks),
        labels = function(x){formatpercent100(x, 1)}
      )
    )+
    scale_x_continuous(breaks=c(postbaselinecuts, max(postbaselinecuts+14)), labels=c(postbaselinecuts, "Overall"), limits=c(min(postbaselinecuts), max(postbaselinecuts+14)+14), expand = c(0, 0))+
    scale_colour_brewer(type="qual", palette="Set2", guide=guide_legend(ncol=1))+
    labs(
      y="Hazard ratio",
      x="Days since booster dose",
      colour=NULL
     ) +
    theme_minimal()+
    theme(
      panel.border = element_blank(),
      axis.line.y = element_line(colour = "black"),
      axis.line.y.right = element_blank(),

      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      strip.background = element_blank(),
      strip.placement = "outside",
      strip.text.y.left = element_text(angle = 0),

      panel.spacing = unit(0.8, "lines"),

      plot.title = element_text(hjust = 0),
      plot.title.position = "plot",
      plot.caption.position = "plot",
      plot.caption = element_text(hjust = 0, face= "italic"),

      legend.position = "none"
     ) +
    NULL

  plot_hazards

```

\newpage

#### Figure 2b: Estimated booster effectiveness (2 period split at 28 days)

For each outcome based on the fully adjusted model, the period-specific hazard ratios for boosting with BNT162b2 versus not boosting is shown, stratified by primary course.

```{r plot_effects28, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}

formatpercent100 <- function(x,accuracy){
  formatx <- scales::label_percent(accuracy, suffix="")(x)

  if_else(
    formatx==scales::label_percent(accuracy, suffix="")(1),
    paste0(">",scales::label_percent(1, suffix="")((100-accuracy)/100)),
    formatx
  )
}
 ## hazard ratio

  y_breaks <- c(0.01, 0.02, 0.05, 0.1, 0.2, 0.5, 1, 2)

  overallhazards_plotdata <- 
    overallhazards %>%
    filter(
      model==3, 
      outcome %in% outcomes,
      treatment %in% c("pfizer"),
      subgroup_variable %in% c("none", "vax12_type")
    )

  postbaselinecuts28 <- c(0,28,70)

  plot_hazards <-
    meta28hazards %>%
    filter(
      model==3,
      outcome %in% outcomes,
      treatment %in% c("pfizer"),
      subgroup_variable %in% c("none", "vax12_type")
    ) %>%
    ggplot()+
    geom_point(aes(y=hr, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+
    geom_linerange(aes(ymin=hr.ll, ymax=hr.ul, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+
    geom_point(data = overallhazards_plotdata, mapping = aes(y=hr, x=max(postbaselinecuts28+14)), colour="darkgrey") +
    geom_linerange(data = overallhazards_plotdata, mapping = aes(ymin=hr.ll, ymax=hr.ul, x=max(postbaselinecuts28+14)), colour="darkgrey") +
    geom_rect(data = overallhazards_plotdata, ymin=-Inf, ymax=Inf, xmin=max(postbaselinecuts28+14)-7, xmax=max(postbaselinecuts28+14)+7, fill="grey95", alpha=0.5, colour="transparent") +
    geom_hline(aes(yintercept=1), colour='grey')+
    facet_grid(rows=vars(outcome_descr_wrap), cols=vars(subgroup_level_descr), switch="y")+
    scale_y_log10(
      breaks=y_breaks,
      limits = c(0.01, 1.5),
      oob = scales::oob_keep,
      sec.axis = sec_axis(
        ~(1-.),
        name="Effectiveness (%)",
        breaks = 1-(y_breaks),
        labels = function(x){formatpercent100(x, 1)}
      )
    )+
    scale_x_continuous(breaks=c(postbaselinecuts28, max(postbaselinecuts28+14)), labels=c(postbaselinecuts28, "Overall"), limits=c(min(postbaselinecuts28), max(postbaselinecuts28+14)+14), expand = c(0, 0))+
    scale_colour_brewer(type="qual", palette="Set2", guide=guide_legend(ncol=1))+
    labs(
      y="Hazard ratio",
      x="Days since booster dose",
      colour=NULL
     ) +
    theme_minimal()+
    theme(
      panel.border = element_blank(),
      axis.line.y = element_line(colour = "black"),
      axis.line.y.right = element_blank(),

      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      strip.background = element_blank(),
      strip.placement = "outside",
      strip.text.y.left = element_text(angle = 0),

      panel.spacing = unit(0.8, "lines"),

      plot.title = element_text(hjust = 0),
      plot.title.position = "plot",
      plot.caption.position = "plot",
      plot.caption = element_text(hjust = 0, face= "italic"),

      legend.position = "none"
     ) +
    NULL
  plot_hazards

```

\newpage

#### Figure 2c: Estimated booster effectiveness (2 period split at 42 days)

For each outcome based on the fully adjusted model, the period-specific hazard ratios for boosting with BNT162b2 versus not boosting is shown, stratified by primary course.

```{r plot_effects42, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}

formatpercent100 <- function(x,accuracy){
  formatx <- scales::label_percent(accuracy, suffix="")(x)

  if_else(
    formatx==scales::label_percent(accuracy, suffix="")(1),
    paste0(">",scales::label_percent(1, suffix="")((100-accuracy)/100)),
    formatx
  )
}
 ## hazard ratio

  y_breaks <- c(0.01, 0.02, 0.05, 0.1, 0.2, 0.5, 1, 2)

  overallhazards_plotdata <- 
    overallhazards %>%
    filter(
      model==3, 
      outcome %in% outcomes,
      treatment %in% c("pfizer"),
      subgroup_variable %in% c("none", "vax12_type")
    )

  postbaselinecuts42 <- c(0,42,70)

  plot_hazards <-
    meta42hazards %>%
    filter(
      model==3,
      outcome %in% outcomes,
      treatment %in% c("pfizer"),
      subgroup_variable %in% c("none", "vax12_type")
    ) %>%
    ggplot()+
    geom_point(aes(y=hr, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+
    geom_linerange(aes(ymin=hr.ll, ymax=hr.ul, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+
    geom_point(data = overallhazards_plotdata, mapping = aes(y=hr, x=max(postbaselinecuts42+14)), colour="darkgrey") +
    geom_linerange(data = overallhazards_plotdata, mapping = aes(ymin=hr.ll, ymax=hr.ul, x=max(postbaselinecuts42+14)), colour="darkgrey") +
    geom_rect(data = overallhazards_plotdata, ymin=-Inf, ymax=Inf, xmin=max(postbaselinecuts42+14)-7, xmax=max(postbaselinecuts42+14)+7, fill="grey95", alpha=0.5, colour="transparent") +
    geom_hline(aes(yintercept=1), colour='grey')+
    facet_grid(rows=vars(outcome_descr_wrap), cols=vars(subgroup_level_descr), switch="y")+
    scale_y_log10(
      breaks=y_breaks,
      limits = c(0.01, 1.5),
      oob = scales::oob_keep,
      sec.axis = sec_axis(
        ~(1-.),
        name="Effectiveness (%)",
        breaks = 1-(y_breaks),
        labels = function(x){formatpercent100(x, 1)}
      )
    )+
    scale_x_continuous(breaks=c(postbaselinecuts42, max(postbaselinecuts42+14)), labels=c(postbaselinecuts42, "Overall"), limits=c(min(postbaselinecuts42), max(postbaselinecuts42+14)+14), expand = c(0, 0))+
    scale_colour_brewer(type="qual", palette="Set2", guide=guide_legend(ncol=1))+
    labs(
      y="Hazard ratio",
      x="Days since booster dose",
      colour=NULL
     ) +
    theme_minimal()+
    theme(
      panel.border = element_blank(),
      axis.line.y = element_line(colour = "black"),
      axis.line.y.right = element_blank(),

      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      strip.background = element_blank(),
      strip.placement = "outside",
      strip.text.y.left = element_text(angle = 0),

      panel.spacing = unit(0.8, "lines"),

      plot.title = element_text(hjust = 0),
      plot.title.position = "plot",
      plot.caption.position = "plot",
      plot.caption = element_text(hjust = 0, face= "italic"),

      legend.position = "none"
     ) +
    NULL

  plot_hazards

```

\newpage

## Figure 3: Kaplan-Meier incidence curves

```{r plot_km, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}

plot_km <- km %>%
  filter(
    outcome %in% outcomes,
    treatment=="pfizer",
    subgroup_variable %in% c("none", "vax12_type")
  ) %>%
    ggplot()+
    geom_step(aes(x=time, y=(1-surv)*1000, group=treated_descr, colour=treated_descr))+
    geom_rect(aes(xmin=time, xmax=leadtime, ymin=(1-surv.ll)*1000, ymax=(1-surv.ul)*1000, group=treated_descr, fill=treated_descr), alpha=0.1)+
    geom_hline(aes(yintercept=0), colour='black')+
    facet_grid(rows=vars(outcome_descr_wrap), cols=vars(subgroup_level_descr), switch="y", scales="free_y")+
    scale_x_continuous(
      breaks = seq(0,7*max(postbaselinecuts),by=14),
      expand = expansion(0)
    )+
    scale_y_continuous(
      expand = expansion(0),
      #limits = c(0L,NA)
    )+
    scale_colour_brewer(type="qual", palette="Set1", na.value="grey")+
    scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey")+
    labs(
      x=NULL,
      y="Cumulative incidence\nper 1000",
      colour=NULL
    )+
    theme_minimal()+
    theme(
      legend.position=c(0.05,0.85),
      legend.justification = c(0,0),
      legend.direction = "vertical",
      axis.text.x.top=element_text(hjust=0),

      strip.placement = "outside",
      strip.background = element_blank(),
      strip.text.y.left = element_text(angle = 0),

      panel.border = element_blank(),
      axis.line.y = element_line(colour = "black"),
      axis.line.x.top = element_line(colour = "black"),
      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),

      panel.spacing = unit(0.8, "lines"),

    )+
    NULL

plot_km

```

### 2022-03-07 manuscript/draft-manuscript.Rmd
refs

@@
 
-The national COVID-19 vaccination programme in England began a new phase in September 2021 with administration of the first booster COVID-19 vaccine doses in those who had already received their primary vaccination course (a completed first and second dose) (ref <https://www.england.nhs.uk/2021/09/nhs-begins-covid-19-booster-vaccination-campaign/>).
-Initially offered to those at highest risk of severe disease, eligibility was progressively extended and by 15 December 2021 every adult in England was eligible (<https://www.england.nhs.uk/2021/12/nhs-booster-bookings-open-to-every-eligible-adult/>).
-Booster doses were initially available no earlier than six months after dose two, but this was reduced to three months on 8 December 2021, following concerns over the surge in cases and the emergence of the Omicron variant (ref: <https://www.england.nhs.uk/2021/12/people-40-and-over-to-get-their-lifesaving-booster-jab-three-months-on-from-second-dose/>).
+The national COVID-19 vaccination programme in England began a new phase in September 2021 with administration of the first booster COVID-19 vaccine doses in those who had already received their primary vaccination course of a first and second dose @NHSEnglandNHS.
+Initially offered to those at highest risk of severe disease, eligibility was progressively extended and by 15 December 2021 every adult in England was eligible @NHSEnglandNHSb.
+Booster doses were initially available no earlier than six months after dose two, but this was reduced to three months on 8 December 2021, following concerns over the surge in cases and the emergence of the Omicron variant @NHSEnglandNHSa.
 Vaccine prioritisation schedules were guided by recommendations from the Joint Committee for Vaccine and Immunisation (JCVI) expert working group.
+
+<!-- https://www.england.nhs.uk/2021/09/nhs-begins-covid-19-booster-vaccination-campaign/ -->
 
@@
 
-We set out to emulate a target trial estimating the effectiveness of booster vaccination with BNT162b2 against COVID-19 outcomes in adults who had already received two COVID-19 vaccine doses.
+We set out to emulate a target trial assessing the effectiveness of booster vaccination with BNT162b2 against COVID-19 outcomes in adults who had already received two COVID-19 vaccine doses.
 We used the OpenSAFELY-TPP linked primary care database covering around 40% of English residents.
-Recruitment covers an era where the Delta variant is dominant, with follow-up curtailed shortly after the emergence of Omicron.
+Recruitment and follow-up covers an era where the Delta variant is dominant, prior to the dominance of the Omicron variant.
 
@@
 Protection was similar between 2× BNT162b2 primary course recipients and 2× ChAdOx1-S recipients.
-There were apparent waning effects around 6 weeks after boosting, which is unlikely to be explained by increased transmissibility of the Omicron variant given Omicron was not yet dominant in the UK by the end of the follow-up window @COVID19OmicronDaily, though changing testing behaviours over time between boosted and unboosted people may influence estimates.
+There were apparent waning effects around 6 weeks after boosting, which is unlikely to be explained by increased transmissibility of the Omicron variant given Omicron was not yet dominant in the UK by the end of the follow-up window @COVID19OmicronDaily.
+Changing testing behaviours over time between boosted and unboosted people may influence estimates.
 
@@
 A phase III trial assessing BNT162b2 booster efficacy in two-dose BNT162b2 recipients without prior infection reported a relative efficacy of 95.6% (95% CI 89.3-98.6) against Delta infection @PfizerBioNTechAnnounce2021.
-
-Recent studies in the UK have provided observational evidence for increased protection from the booster vaccine against symptomatic COVID-19 infection in comparison to second dose recipients using a test negative design.
+Observational studies in the UK have provided evidence for increased protection from the booster vaccine against symptomatic COVID-19 infection in comparison to second dose recipients using a test negative design.
 Andrews et al.
@@
 
-Elsewhere, a study in Israeli health registry data @barda2021 found strong protection against admission (1 – risk ratio (95%CI) = 93% (88-97)), severe disease (92% (82-97)), and death (81% (59-97)), with similar results in specific demographic and clinical subgroups.
-but did not consider differences @barda2021
+Elsewhere, a study in Israeli health registry data @barda2021 found strong protection against admission (1 – risk ratio (95%CI) = 93% (88-97)), severe disease (92% (82-97)), and death (81% (59-97)), with similar results in specific demographic and clinical subgroups, but did not investigate ossible waning @barda2021.
 

### 2022-03-07 manuscript/draft-manuscript.Rmd
update boilerplate

@@
 
-William J Hulme^1^, Elizabeth J Williamson^2^, Amelia Green^1^, Elsie Horne^3,4^, Helen Curtis^1^, Linda Nab^1^, Milan Wiedemann^1^, Ruth Keogh^2^, Edward Parker^2^, Venexia Walker^3,4^, Jonathan AC Sterne^3,4^, Miguel A Hernán^6,7^, Ben Goldacre^1^ + OpenSAFELY collaborative.
-
-<!-- Krishnan Bhaskaran^2^, Helen I McDonald^2^, Christopher T Rentsch^2^, Anna Schultze^2^, John Tazare^2^, Helen J Curtis^1^, Alex J Walker^1^, Laurie Tomlinson^2^, Tom Palmer^3,4^,  Brian MacKenna^1^, Caroline E Morton^1^, Amir Mehrkar^1^, Louis Fisher^1^, Seb Bacon^1^, Dave Evans^1^, Peter Inglesby^1^, George Hickman^1^, Simon Davy^1^, Tom Ward ^1^, Richard Croker^1^, Rosalind M Eggo^2^, Angel YS Wong^2^, Rohini Mathur^2^, Kevin Wing^2^, Harriet Forbes^2^, Daniel Grint^2^, Ian J Douglas^2^, Stephen JW Evans^2^, Liam Smeeth^2^, Chris Bates^5^, Jonathan Cockburn^5^, John Parry^5^, Frank Hester^5^, Sam Harper^5^,  -->
+William J Hulme^1^, Elizabeth J Williamson^2^, Amelia Green^1^, Linda Nab^1^, Elsie Horne^3,4^, Ruth Keogh^2^, Edward Parker^2^, Venexia Walker^3,4^, Helen Curtis^1^, Milan Wiedemann^1^, Christine Cunningham^1^, Alex J Walker^1^, Krishnan Bhaskaran^2^, Helen I McDonald^2^, Christopher T Rentsch^2^, Anna Schultze^2^, John Tazare^2^, Laurie Tomlinson^2^, Tom Palmer^3,4^, Brian MacKenna^1^, Caroline E Morton^1^, Amir Mehrkar^1^, Louis Fisher^1^, Seb Bacon^1^, Dave Evans^1^, Peter Inglesby^1^, George Hickman^1^, Simon Davy^1^, Tom Ward^1^, Richard Croker^1^, Rosalind M Eggo^2^, Angel YS Wong^2^, Rohini Mathur^2^, Kevin Wing^2^, Harriet Forbes^2^, Daniel Grint^2^, Viyasaan  Mahalingasivam^2^, Bang Zheng^2^, Ian J Douglas^2^, Stephen JW Evans^2^, Liam Smeeth^2^, Chris Bates^5^, Jonathan Cockburn^5^, John Parry^5^, Frank Hester^5^, Sam Harper^5^, Jonathan AC Sterne^3,4^, Miguel A Hernán^6,7^, Ben Goldacre^1^.
 
@@
 We matched `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` BNT162b2 booster recipients to unboosted controls.
-During days 15-28 after the booster, the estimated hazard ratio (95% CI) comparing a BNT162b2 booster dose to two doses only was `r filter(hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` for documented SARS-CoV-2 infection, `r filter(hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` for COVID-19 death.
+During the first 28 days after the booster, the estimated hazard ratio (95% CI) comparing a BNT162b2 booster dose to two doses only was `r filter(hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` for documented SARS-CoV-2 infection, `r filter(hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` for COVID-19 death.
 
@@
 
-The OpenSAFELY-TPP database (<https://opensafely.org>) covers 24 million people registered at English General Practices that use TPP SystmOne electronic health record software.
-It includes pseudonymized data such as coded diagnoses, medications and physiological parameters.
+All data were linked, stored and analysed securely within the OpenSAFELY platform: <https://opensafely.org/>. 
+Primary care records managed by the GP software provider TPP were linked, using NHS numbers, to A&E attendance and in-patient hospital spell records via NHS Digital's Hospital Episode Statistics (HES), national coronavirus testing records via the Second Generation Surveillance System (SGSS), and national death registry records from the Office for National Statistics (ONS).
+Vaccination status is available in the GP record directly via the National Immunisation Management System (NIMS).
+Data include pseudonymized data such as coded diagnoses, medications and physiological parameters.
 No free text data are included.
-This primary care data is linked (via NHS numbers) with A&E attendance and in-patient hospital spell records via NHS Digital's Hospital Episode Statistics (HES), national coronavirus testing records via the Second Generation Surveillance System (SGSS), and national death registry records from the Office for National Statistics (ONS).
-Vaccination status is available in the GP record directly via the National Immunisation Management System (NIMS).
 
@@
 For each outcome, we estimated unadjusted cumulative incidence curves using the Kaplan-Meier estimator separately in the boosted and unboosted groups.
+Rates of competing events (due to all-cause or cause-specific death, depending on the outcome) were anticipated to be low and so were not accounted for.
 
 We use Cox models, stratified by recruitment day, NHS region, and brand of second vaccine dose, to estimate hazard ratios comparing boosted with unboosted people.
-We estimated effectveness over all follow-up time (1-70 days), and separately in days 1-7, 8-14, 15-28, 29-42, 43-70.
-Models were adjusted for additional variables not used as matching factors, as described below.
+We estimated effectveness over all follow-up time (up to 70 days), and separately in days 1-28 and 29-70.
+Models were adjusted for additional variables not used as matching factors, described above.
 A robust variance estimator was used to account for participant clustering.
@@
 
-We also estimated booster effectiveness in the following subgroups: primary vaccine course (both BNT162b2 or both ChAdOx1-S); age (18-64 or 65 and over); clinically extremely vulnerable.
+We also estimated booster effectiveness in the following subgroups: primary vaccine course (both BNT162b2 or both ChAdOx1-S); age (18-64 or 65 and over); clinically extremely vulnerable or not.
 Analyses were conducted separately in each subgroup.
@@
 Data management and analyses were conducted in Python version 3.8.10 and R version 4.0.2.
-All code is available at <https://github.com/opensafely/booster-effectiveness>, and is shared openly for review and re-use under MIT open license.
+All code is shared openly for review and re-sue under MIT open license at <https://github.com/opensafely/booster-effectiveness>.
 Codelists are available at <https://www.opencodelists.org/>.
-No person-level data is shared.
+Detailed pseudonymised patient data is potentially re-identifiable and therefore not shared.
 Any reported figures based on counts below 6 are redacted or rounded for disclosure control.
@@
 
-NHS England is the data controller; TPP is the data processor; and the key researchers on OpenSAFELY are acting with the approval of NHS England.
+NHS England is the data controller for OpenSAFELY-TPP; TPP is the data processor; and study authors using OpenSAFELY have the approval of NHS England.
 This implementation of OpenSAFELY is hosted within the TPP environment which is accredited to the ISO 27001 information security standard and is NHS IG Toolkit compliant @betad @datasec ; Patient data has been pseudonymised for analysis and linkage using industry standard cryptographic hashing techniques; all pseudonymised datasets transmitted for linkage onto OpenSAFELY are encrypted; access to the platform is via a virtual private network (VPN) connection, restricted to a small group of researchers; the researchers hold contracts with NHS England and only access the platform to initiate database queries and statistical models; all database activity is logged; only aggregate statistical outputs leave the platform environment following best practice for anonymisation of results such as statistical disclosure control for low cell counts @isb1523 .
@@
 Taken together, these provide the legal bases to link patient datasets on the OpenSAFELY platform.
-General Practices, from which the primary care data are obtained, are required to share relevant health information to support the public health response to the pandemic, and have been informed of the OpenSAFELY analytics platform.
+General practices, from which the primary care data are obtained, are required to share relevant health information to support the public health response to the pandemic, and have been informed of the OpenSAFELY analytics platform.
 

### 2022-03-07 manuscript/draft-manuscript.Rmd
update

@@
 output:
+  word_document: default
+  pdf_document:
+    toc: no
+  bookdown::html_document2:
+    number_sections: no
+    toc: no
   html_document:
@@
     self_contained: yes
-  word_document: default
-  bookdown::html_document2:
-    number_sections: no
-    toc: no
-  pdf_document:
-    toc: no
 bibliography: references.bib
@@
     outcome_descr = fct_inorder(outcome_descr),
-    outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=20)),
+    outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=15)),
   ) %>%
@@
 
-William J Hulme^1^, Elizabeth J Williamson^2^, Amelia Green^1^, Linda Nab^1^, Elsie Horne^3,4^, Ruth Keogh^2^, Edward Parker^2^, Venexia Walker^3,4^, Helen Curtis^1^, Milan Wiedemann^1^, Christine Cunningham^1^, Alex J Walker^1^, Krishnan Bhaskaran^2^, Helen I McDonald^2^, Christopher T Rentsch^2^, Anna Schultze^2^, John Tazare^2^, Laurie Tomlinson^2^, Tom Palmer^3,4^, Brian MacKenna^1^, Caroline E Morton^1^, Amir Mehrkar^1^, Louis Fisher^1^, Seb Bacon^1^, Dave Evans^1^, Peter Inglesby^1^, George Hickman^1^, Simon Davy^1^, Tom Ward^1^, Richard Croker^1^, Rosalind M Eggo^2^, Angel YS Wong^2^, Rohini Mathur^2^, Kevin Wing^2^, Harriet Forbes^2^, Daniel Grint^2^, Viyasaan  Mahalingasivam^2^, Bang Zheng^2^, Ian J Douglas^2^, Stephen JW Evans^2^, Liam Smeeth^2^, Chris Bates^5^, Jonathan Cockburn^5^, John Parry^5^, Frank Hester^5^, Sam Harper^5^, Jonathan AC Sterne^3,4^, Miguel A Hernán^6,7^, Ben Goldacre^1^.
+William J Hulme^1^, Elizabeth J Williamson^2^, Elsie Horne^3,4^, Amelia Green^1^, Linda Nab^1^, Ruth Keogh^2^, Edward Parker^2^, Venexia Walker^3,4^, Helen Curtis^1^, Milan Wiedemann^1^, Christine Cunningham^1^, Alex J Walker^1^, Louis Fisher^1^, Christopher T Rentsch^2^, Anna Schultze^2^, Krishnan Bhaskaran^2^, Helen I McDonald^2^, John Tazare^2^, Laurie Tomlinson^2^, Tom Palmer^3,4^, Brian MacKenna^1^, Richard Croker^1^, Caroline E Morton^1^, Colm Andrews^1^, Robin Parks^1^, Lisa Hopcroft^1^, Jon Massey^1^, Jessica Morley^1^, Amir Mehrkar^1^, Seb Bacon^1^, Dave Evans^1^, Peter Inglesby^1^, George Hickman^1^, Simon Davy^1^, Tom Ward^1^, Rosalind M Eggo^2^, Angel YS Wong^2^, Rohini Mathur^2^, Kevin Wing^2^, Harriet Forbes^2^, Daniel Grint^2^, Viyasaan Mahalingasivam^2^, Bang Zheng^2^, Ian J Douglas^2^, Stephen JW Evans^2^, Liam Smeeth^2^, Chris Bates^5^, Jonathan Cockburn^5^, John Parry^5^, Frank Hester^5^, Sam Harper^5^, Jonathan AC Sterne^3,4^, Miguel A Hernán^6,7^, Ben Goldacre^1^.
 
@@
 Primary care records managed by the GP software provider TPP were linked, using NHS numbers, to A&E attendance and in-patient hospital spell records via NHS Digital's Hospital Episode Statistics (HES), national coronavirus testing records via the Second Generation Surveillance System (SGSS), and national death registry records from the Office for National Statistics (ONS).
-Vaccination status is available in the GP record directly via the National Immunisation Management System (NIMS).
+COVID-19 vaccination history is available in the GP record directly via the National Immunisation Management System (NIMS).
 Data include pseudonymized data such as coded diagnoses, medications and physiological parameters.
@@
 
-## Eligiblity criteria
+## Eligibility criteria
 
@@
 
-The following three outcomes were considered: documented SARS-CoV-2 infection; COVID-19 hospital admission; COVID-19 death; and non-COVID-19 death as a negative control outcome.
+The following three outcomes were considered: documented SARS-CoV-2 infection; COVID-19 hospital admission; COVID-19 death; and non-COVID-19 death as a negative control outcome @Lipsitch2010.
 SARS-CoV-2 infections were identified using SGSS testing records and based on swab date.
@@
 <!--# COVID-19 related A&E attendance and critical/intensive care admission we also considered but there were doubts over data the reliability of event ascertainment in these cases. -->
+
 <!--# COVID-19 A&E attendances were identified using HES emergency care records with U07.1 ("COVID-19, virus identified") or U07.2 ("COVID-19, virus not identified") ICD-10 diagnosis codes @emergenc.
@@
 
-The following additional variables were included as covariates in the Cox models (see below) or else to describe the treatment groups: sex (male or female); English Index of Multiple Deprivation (IMD, grouped by quintile); ethnicity (Black, Mixed, South Asian, White, Other, as per the UK census); morbidity count (diabetes, BMI > 40kg/m^2^, chronic heart disease, chronic kidney disease, chronic liver disease, chronic respiratory disease or severe asthma, chronic neurological disease); learning disabilities; severe mental illness; the number of SARS-CoV-2 tests in the 6 months prior to the study start date; elective in-hospital episode at the time of recruitment.
+The following additional variables were included as covariates in the Cox models (see below) or else to describe the treatment groups: sex (male or female); English Index of Multiple Deprivation (IMD, grouped by quintile); ethnicity (Black, Mixed, South Asian, White, Other, as per the UK census); morbidity count (diabetes, BMI \> 40kg/m^2^, chronic heart disease, chronic kidney disease, chronic liver disease, chronic respiratory disease or severe asthma, chronic neurological disease); learning disabilities; severe mental illness; the number of SARS-CoV-2 tests in the 6 months prior to the study start date; elective in-hospital episode at the time of recruitment.
 
@@
 
-#### Figure 1: Booster vaccination uptake, treatment group eligibility and matching success
+#### Figure 1: Booster uptake and matching success
 
@@
 
-```{r, vaxdate, echo=FALSE, message=FALSE, warning=FALSE, out.width='60%', out.height='50%', results='asis', out.extra=''}
+```{r, vaxdate, echo=FALSE, message=FALSE, warning=FALSE, out.width='90%', out.height='50%', results='asis', out.extra=''}
 #read_rds(fs::path(output_dir_os, "descriptive", "vaxdate", "plot_vaxdate_stack.rds"))
@@
 
-Total follow-up and incidence rates for each outcome, by treatment group. The cumulative incidence of each outcome is provided in Supplementary Figure S2.
+Total follow-up and incidence rates for each outcome, by treatment group. The cumulative incidence of each outcome is provided in Figure 2.
 
@@
 
+### Figure 2: Cumulative outcome incidence
+
+Kaplan-Meier estimates of cumulative incidence in matched boosted and unboosted treatment groups, stratified by primary course and without further adjustment for potential confounders.
+
+```{r cmlinc, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', out.width='90%', out.height='50%', results='asis', out.extra=''}
+
+plot_km <- km %>%
+  filter(
+    outcome %in% outcomes,
+    treatment=="pfizer",
+    subgroup_variable %in% c("none", "vax12_type")
+  ) %>%
+    ggplot()+
+    geom_step(aes(x=time, y=(1-surv)*1000, group=treated_descr, colour=treated_descr))+
+    geom_rect(aes(xmin=time, xmax=leadtime, ymin=(1-surv.ll)*1000, ymax=(1-surv.ul)*1000, group=treated_descr, fill=treated_descr), alpha=0.1)+
+    geom_hline(aes(yintercept=0), colour='black')+
+    facet_grid(rows=vars(outcome_descr_wrap), cols=vars(subgroup_level_descr), switch="y", scales="free_y")+
+    scale_x_continuous(
+      breaks = seq(0,7*max(postbaselinecuts),by=14),
+      expand = expansion(0)
+    )+
+    scale_y_continuous(
+      expand = expansion(0),
+      #limits = c(0L,NA)
+    )+
+    scale_colour_brewer(type="qual", palette="Set1", na.value="grey")+
+    scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey")+
+    labs(
+      x=NULL,
+      y="Cumulative incidence\nper 1000",
+      colour=NULL
+    )+
+    theme_minimal()+
+    theme(
+      legend.position = "bottom",
+      legend.direction = "horizontal",
+      #legend.position=c(0.01,0.85),
+      #legend.justification = c(0,0),
+      #legend.direction = "vertical",
+      axis.text.x.top=element_text(hjust=0),
+
+      strip.placement = "outside",
+      strip.background = element_blank(),
+      strip.text.y.left = element_text(angle = 0),
+
+      panel.border = element_blank(),
+      axis.line.y = element_line(colour = "black"),
+      axis.line.x.top = element_line(colour = "black"),
+      panel.grid.minor.x = element_blank(),
+      panel.grid.minor.y = element_blank(),
+
+      panel.spacing = unit(0.8, "lines"),
+
+    )+
+    NULL
+
+plot_km
+```
+
 ## Estimated booster effectiveness
@@
 
-Participants in the emulated trial experienced `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="postest", fup_period=="All") %>% pull(events))` documented infections, `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="covidadmitted", fup_period=="All") %>% pull(events))` COVID-19 hospital admissions, and `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="coviddeath", fup_period=="All") %>% pull(events))` COVID-19 deaths across `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(fup_years))` person-years of follow-up (Table 2).
+Participants in the emulated trial experienced `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="postest", fup_period=="All") %>% pull(events))` documented infections, `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="covidadmitted", fup_period=="All") %>% pull(events))` COVID-19 hospital admissions, and `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="coviddeath", fup_period=="All") %>% pull(events))` COVID-19 deaths across `r label_number(1, big.mark=",")((filter(matchsummary, treatment=="pfizer") %>% pull(fup_years)) *365.25/7)` person-weeks of follow-up (Table 2, Figure 2).
 
 The estimated effectiveness (95% CI) comparing a BNT162b2 booster dose to two doses only over all follow-up time was `r filter(overallhazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3) %>% pull(ve_display)` for documented infection, `r filter(overallhazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(overallhazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 death.
-Effectiveness before and after 28 days since boosting is also estimated (Figure 2, Table 3), as well as for narrower periods (Supplementary Table S3).
-
-<!-- In the first 28 days of follow-up the estimated effectiveness (95% CI) comparing a BNT162b2 booster dose to two doses only, was `r filter(meta2hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` for documented infection, `r filter(meta2hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(meta2hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` for COVID-19 death (Figure 2).
+Effectiveness before and after 28 days since boosting is also estimated (Figure 3, Table 3), as well as for narrower periods (Supplementary Table S3).
+
+<!-- In the first 28 days of follow-up the estimated effectiveness (95% CI) comparing a BNT162b2 booster dose to two doses only, was `r filter(meta2hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` for documented infection, `r filter(meta2hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(meta2hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` for COVID-19 death (Figure 3).
 -->
 
-<!-- For days 28-70, estimated effectiveness (95% CI) comparing a BNT162b2 booster dose to two doses only, was `r filter(meta2hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` for documented infection, `r filter(meta2hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(meta2hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` for COVID-19 death (Figure 2), suggesting a potential waning of booster effectiveness with time.
+<!-- For days 28-70, estimated effectiveness (95% CI) comparing a BNT162b2 booster dose to two doses only, was `r filter(meta2hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` for documented infection, `r filter(meta2hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(meta2hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` for COVID-19 death (Figure 3), suggesting a potential waning of booster effectiveness with time.
 -->
@@
 
-Period-specific estimates are also provided.
-
 \newpage
 
-#### Figure 2: Estimated booster effectiveness
+#### Figure 3: Estimated booster effectiveness
 
@@
 Main and subgroup analyses
+
 ```{r table_effects, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}
@@
 We found high protection against all studied outcomes.
-Protection was similar between 2× BNT162b2 primary course recipients and 2× ChAdOx1-S recipients.
-There were apparent waning effects around 6 weeks after boosting, which is unlikely to be explained by increased transmissibility of the Omicron variant given Omicron was not yet dominant in the UK by the end of the follow-up window @COVID19OmicronDaily.
-Changing testing behaviours over time between boosted and unboosted people may influence estimates.
-
-We have considered effectiveness in a population where Delta was the predominant circulating variant @COVID19OmicronDaily.
+There were apparent waning protection against documented infection around 6 weeks after boosting, which is unlikely to be explained by increased transmissibility of the Omicron variant given Omicron was not yet dominant in the UK by the end of the follow-up window @COVID19OmicronDaily.
+Changing testing behaviours over time between boosted and unboosted people may influence these estimates, if for example persistently unboosted participants, who will contribute more person-time in later follow-up periods, were less likely to get tested.
+
+Protection was similar between 2× BNT162b2 primary course recipients and 2× ChAdOx1-S recipients, except for COVID-19 death in the first 28 days where protected appeared higher for those whose primary course was 2× BNT162b2 than for 2× ChAdOx1-S.
+For severe outcomes, protection was lower in the clinical extremely vulnerable sub-population.
+
+We considered effectiveness in a population where Delta was the predominant circulating variant @COVID19OmicronDaily.
 Re-running the analysis against a backdrop of the now globally-dominant Omicron variant will be undertaken where valid concurrent comparisons between boosted and unboosted people can be made.
@@
 reported estimates of relative vaccine effectiveness against syptomatic disease in the 14 days after a BNT162b2 booster dose in those aged over 50 years of 87.4% (95%CI 84.9-89.4) where the primary course was ChAdOx1-S and 84.4% (95%CI 82.8-85.8) for BNT162b2 @andrews2021.
-Sheikh et al.
-reported comparable estimates of relative vaccine effectiveness for S gene positive symptomatic infection (83% (95%CI 81-84) 6-49 years, 88% (95%CI 86-89) ≥50 years) in the 14 days after a BNT162b2 or mRNA-1273 booster dose, but lower estimates for S gene negative symptomatic infection (56% (95%CI 51-60) 6-49 years, 57% (95%CI 52-62) ≥50 years).
-
-<!-- @https://www.research.ed.ac.uk/en/publications/severity-of-omicron-variant-of-concern-and-vaccine-effectiveness- -->
-
-Elsewhere, a study in Israeli health registry data @barda2021 found strong protection against admission (1 – risk ratio (95%CI) = 93% (88-97)), severe disease (92% (82-97)), and death (81% (59-97)), with similar results in specific demographic and clinical subgroups, but did not investigate ossible waning @barda2021.
+A study in Scotland reported comparable estimates of relative vaccine effectiveness for S gene positive symptomatic infection (83% (95%CI 81-84) 6-49 years, 88% (95%CI 86-89)  over 50 years) in the 14 days after a BNT162b2 or mRNA-1273 booster dose, but lower estimates for S gene negative symptomatic infection (56% (95%CI 51-60) 6-49 years, 57% (95%CI 52-62) over 50 years) @sheikhSeverityOmicronVariant2021.
+
+Elsewhere, a study in Israeli health registry data @barda2021 found strong protection against admission (1 -- risk ratio (95%CI) = 93% (88-97)), severe disease (92% (82-97)), and COVID-19 death (81% (59-97)), with similar results in specific demographic and clinical subgroups @barda2021.
+Our results broadly agree with these findings and provide important additional insights into effectiveness in specific clinical subgroups and over longer follow-up.
 
@@
 
-This study of over 8 million people in England found high protection for BNT162b2 boosting against COVID-19 hospitalisation and death during an `r studydurationweeks` week period in which the Delta variant was dominant.
+This study of over 8 million people in England found high protection for BNT162b2 boosting against documented infection, COVID-19 hospitalisation and death during a period in which the Delta variant was dominant.
 

### 2022-03-07 manuscript/draft-supplement.Rmd
update

@@
 \newpage
+
+# COVID-19 vaccination history
+
+The dates and brands of any COVID-19 vaccines administered as part of the national roll-out (e.g., excluding in vaccine trials and vaccines administered abroad) are available in the primary care records via transfer from the National Immunisation Management System (NIMS).
+Vaccine cardinality (dose 1, dose 2, ...) is determined by vaccination dates alone.
+We cannot distinguish between "third" doses in immunosuppressed or otherwise vulnerable people which were administered ahead of the national booster roll-out, and "booster" doses, i.e., indication for 3rd vaccine dose.
+However, we match on characteristics that approximate determinants of third dose vaccination, and the recruitment period excludes a majority of early third dose recipients.
 

### 2022-03-07 analysis/combine_model.R
add competing-risk cumulative incidences

@@ -454,3 +454,14 @@ ggsave(filename=fs::path(output_dir, "kmplot.png"), plot_km, width=20, height=15
 ggsave(filename=fs::path(output_dir, "kmplot.pdf"), plot_km, width=20, height=15, units="cm")
 
 
+
+## CIF rates ----
+
+cif <- model_metaparams %>%
+  mutate(
+    cif = pmap(list(treatment, outcome, subgroup), function(x, y, z) read_csv(here("output", "models", "seqtrialcox", x, y, z, glue("report_cif.csv"))))
+  ) %>%
+  unnest(cif)
+
+write_csv(cif, fs::path(output_dir, "cif.csv"))
+

### 2022-03-07 analysis/match_seqtrialcox.R

@@ -202,7 +202,20 @@ data_tte <-
       na.rm=TRUE
     ),
 
+    noncompetingcensor_date = pmin(
+      dereg_date,
+      competingtreatment_date-1, # -1 because we assume vax occurs at the start of the day
+      vax4_date-1, # -1 because we assume vax occurs at the start of the day
+      study_dates$studyend_date,
+      na.rm=TRUE
+    ),
+
     # assume vaccination occurs at the start of the day, and all other events occur at the end of the day.
+
+    # possible competing events
+    tte_coviddeath = tte(day0_date, coviddeath_date, noncompetingcensor_date, na.censor=TRUE),
+    tte_noncoviddeath = tte(day0_date, noncoviddeath_date, noncompetingcensor_date, na.censor=TRUE),
+    tte_death = tte(day0_date, death_date, noncompetingcensor_date, na.censor=TRUE),
 
     tte_censor = tte(day0_date, censor_date, censor_date, na.censor=TRUE),
 

### 2022-03-07 analysis/report_seqtrialcox.R

@@ -361,6 +361,153 @@ plot_km
 ggsave(filename=fs::path(output_dir, "report_kmplot.svg"), plot_km, width=20, height=15, units="cm")
 ggsave(filename=fs::path(output_dir, "report_kmplot.png"), plot_km, width=20, height=15, units="cm")
 
+## Cumulative incidence function (accountign for competing risks ----
+
+library('cmprsk')
+
+
+# TODO once matching script re-run, remove this and import data_tte instead
+data_tte <- read_rds(here("output", "data", "data_cohort.rds")) %>%
+  filter(patient_id %in% data_seqtrialcox$patient_id) %>%
+  transmute(
+    patient_id,
+    day0_date = study_dates$studystart_date-1, # day before the first trial date
+
+    treatment_date = if_else(vax3_type==treatment, vax3_date, as.Date(NA)),
+    competingtreatment_date = if_else(vax3_type!=treatment, vax3_date, as.Date(NA)),
+
+    # person-time is up to and including censor date
+    censor_date = pmin(
+      dereg_date,
+      competingtreatment_date-1, # -1 because we assume vax occurs at the start of the day
+      vax4_date-1, # -1 because we assume vax occurs at the start of the day
+      study_dates$studyend_date,
+      na.rm=TRUE
+    ),
+
+    # possible competing events
+    tte_coviddeath = tte(day0_date, coviddeath_date, censor_date, na.censor=TRUE),
+    tte_noncoviddeath = tte(day0_date, noncoviddeath_date, censor_date, na.censor=TRUE),
+    tte_death = tte(day0_date, death_date, censor_date, na.censor=TRUE),
+
+    tte_censor = tte(day0_date, censor_date, censor_date, na.censor=TRUE),
+  ) %>%
+  mutate(
+    # convert tte variables (days since day0), to integer to save space
+    across(starts_with("tte_"), as.integer),
+  )
+
+
+competing_event<-"death"
+if(outcome == "death"){
+  competing_event<-NULL
+}
+if(outcome=="coviddeath"){
+  competing_event<-"noncoviddeath"
+}
+if(outcome=="noncoviddeath"){
+  competing_event<-"coviddeath"
+}
+
+data_cif0 <-
+data_seqtrialcox %>%
[…]

### 2022-03-07 analysis/report_seqtrialcox.R
move library statement

@@ -36,6 +36,7 @@ library('tidyverse')
 library('here')
 library('glue')
 library('survival')
+library('cmprsk')
 
 ## Import custom user functions
 source(here("lib", "functions", "utility.R"))
@@ -362,9 +363,6 @@ ggsave(filename=fs::path(output_dir, "report_kmplot.svg"), plot_km, width=20, he
 ggsave(filename=fs::path(output_dir, "report_kmplot.png"), plot_km, width=20, height=15, units="cm")
 
 ## Cumulative incidence function (accountign for competing risks ----
-
-library('cmprsk')
-
 
 # TODO once matching script re-run, remove this and import data_tte instead
 data_tte <- read_rds(here("output", "data", "data_cohort.rds")) %>%

### 2022-03-07 analysis/data_process_long.R
add non-covid admission outcome

@@ -131,6 +131,18 @@ data_covidadmitted <- data_processed %>%
   arrange(patient_id, date) %>%
   mutate(event="covidadmitted") # need to change name to match "outcome" argument
 
+data_noncovidadmitted <-
+  anti_join(
+    data_admitted_unplanned %>% filter(event=="admitted_unplanned"), data_covidadmitted, by=c("patient_id", "date")
+  ) %>%
+  arrange(patient_id, date) %>%
+  group_by(patient_id) %>%
+  mutate(
+    event="noncovidadmitted",
+    index=row_number()
+  ) %>%
+  ungroup()
+
 data_covidcc <- data_processed %>%
   select(patient_id, matches("^covidcc\\_\\d+\\_date")) %>%
   pivot_longer(
@@ -178,6 +190,7 @@ data_allevents <-
     data_covidemergency,
     data_covidemergencyhosp,
     data_covidadmitted,
+    data_noncovidadmitted,
     data_covidcc,
     data_coviddeath,
     data_noncoviddeath,

### 2022-03-07 analysis/design.R

@@ -50,6 +50,7 @@ events_lookup <- tribble(
   "postest", "positive_test_date", "Positive SARS-CoV-2 test",
   "covidemergency", "covidemergency_date", "COVID-19 A&E attendance",
   "covidadmitted", "covidadmitted_date", "COVID-19 hospitalisation",
+  "noncovidadmitted", "noncovidadmitted_date", "Non-COVID-19 hospitalisation",
   "covidadmittedproxy1", "covidadmittedproxy1_date", "COVID-19 hospitalisation (A&E proxy)",
   "covidadmittedproxy2", "covidadmittedproxy2_date", "COVID-19 hospitalisation (A&E proxy v2)",
   "covidcc", "covidcc_date", "COVID-19 critical care",

### 2022-03-07 create-project.R

@@ -161,7 +161,7 @@ action_combine_model <- function(
           .x=expand_grid(
             subgroups = paste0(subgroup,dash,subgroup_levels),
             treatment=c("pfizer", "moderna"),
-            outcome=c("postest", "covidemergency", "covidadmittedproxy1", "covidadmitted",  "coviddeath", "noncoviddeath"),
+            outcome=c("postest", "covidemergency", "covidadmittedproxy1", "covidadmitted", "noncovidadmitted", "coviddeath", "noncoviddeath"),
             script=c("model", "report"),
 
           ),
@@ -366,6 +366,7 @@ actions_list <- splice(
   action_model("pfizer", "postest", "none"),
   action_model("pfizer", "covidemergency", "none"),
   action_model("pfizer", "covidadmittedproxy1", "none"),
+  action_model("pfizer", "nonovidadmitted", "none"),
   action_model("pfizer", "covidadmitted", "none"),
   action_model("pfizer", "covidcc", "none"),
   action_model("pfizer", "coviddeath", "none"),
@@ -381,6 +382,8 @@ actions_list <- splice(
   action_model("pfizer", "covidadmittedproxy1", "vax12_type-pfizer-pfizer"),
   action_model("pfizer", "covidadmitted", "vax12_type-az-az"),
   action_model("pfizer", "covidadmitted", "vax12_type-pfizer-pfizer"),
+  action_model("pfizer", "noncovidadmitted", "vax12_type-az-az"),
+  action_model("pfizer", "noncovidadmitted", "vax12_type-pfizer-pfizer"),
   action_model("pfizer", "covidcc", "vax12_type-az-az"),
   action_model("pfizer", "covidcc", "vax12_type-pfizer-pfizer"),
   action_model("pfizer", "coviddeath", "vax12_type-az-az"),
@@ -399,6 +402,8 @@ actions_list <- splice(
   action_model("pfizer", "covidadmittedproxy1", "cev-TRUE"),
   action_model("pfizer", "covidadmitted", "cev-FALSE"),
   action_model("pfizer", "covidadmitted", "cev-TRUE"),
+  action_model("pfizer", "noncovidadmitted", "cev-FALSE"),
+  action_model("pfizer", "noncovidadmitted", "cev-TRUE"),
   action_model("pfizer", "covidcc", "cev-FALSE"),
   action_model("pfizer", "covidcc", "cev-TRUE"),
[…]

### 2022-03-14 manuscript/draft-manuscript.Rmd
update manuscript + supplement

@@
 output:
-  word_document: default
+  html_document:
+    keep_md: yes
+    self_contained: yes
   pdf_document:
@@
     toc: no
-  html_document:
-    keep_md: yes
-    self_contained: yes
+  rtf_document:
+    toc: no
+  word_document: default
 bibliography: references.bib
@@
 
-studydurationweeks <- as.numeric(as.duration(study_dates$studyend_date - study_dates$studystart_date +1) , "weeks")
+pfizerdurationweeks <- as.numeric(as.duration(study_dates$pfizerend_date - study_dates$pfizerstart_date +1) , "weeks")
 ```
@@
 #cmlinc3 <- cmlinc %>% filter(model==3)
-#riskdiff3 <- riskdiff %>% filter(model==3)
 hazards3 <- hazards %>% filter(model==3)
 
+riskdiff <- km %>% 
+  group_by(treatment, treatment_descr, outcome, outcome_descr, outcome_descr_wrap, subgroup, subgroup_descr, subgroup_variable, subgroup_variable_descr, subgroup_level, subgroup_level_descr, time) %>%
+  summarise(
+    survdiff = diff(surv),
+  ) %>% 
+  mutate(
+    survdiff_display = label_number(accuracy=0.1, scale=1000, big.mark =",")(survdiff),
+    survdiff_inv_display = label_number(accuracy=0.1, scale=1000, big.mark =",")(-survdiff),
+  ) %>%
+  ungroup()
 ```
 
-William J Hulme^1^, Elizabeth J Williamson^2^, Elsie Horne^3,4^, Amelia Green^1^, Linda Nab^1^, Ruth Keogh^2^, Edward Parker^2^, Venexia Walker^3,4^, Helen Curtis^1^, Milan Wiedemann^1^, Christine Cunningham^1^, Alex J Walker^1^, Louis Fisher^1^, Christopher T Rentsch^2^, Anna Schultze^2^, Krishnan Bhaskaran^2^, Helen I McDonald^2^, John Tazare^2^, Laurie Tomlinson^2^, Tom Palmer^3,4^, Brian MacKenna^1^, Richard Croker^1^, Caroline E Morton^1^, Colm Andrews^1^, Robin Parks^1^, Lisa Hopcroft^1^, Jon Massey^1^, Jessica Morley^1^, Amir Mehrkar^1^, Seb Bacon^1^, Dave Evans^1^, Peter Inglesby^1^, George Hickman^1^, Simon Davy^1^, Tom Ward^1^, Rosalind M Eggo^2^, Angel YS Wong^2^, Rohini Mathur^2^, Kevin Wing^2^, Harriet Forbes^2^, Daniel Grint^2^, Viyasaan Mahalingasivam^2^, Bang Zheng^2^, Ian J Douglas^2^, Stephen JW Evans^2^, Liam Smeeth^2^, Chris Bates^5^, Jonathan Cockburn^5^, John Parry^5^, Frank Hester^5^, Sam Harper^5^, Jonathan AC Sterne^3,4^, Miguel A Hernán^6,7^, Ben Goldacre^1^.
+William J Hulme^1^, Elizabeth J Williamson^2^, Elsie Horne^3,4^, Amelia Green^1^, Linda Nab^1^, Ruth Keogh^2^, Edward PK Parker^2^, Venexia Walker^3,4^, Helen Curtis^1^, Milan Wiedemann^1^, Christine Cunningham^1^, Alex J Walker^1^, Louis Fisher^1^, Christopher T Rentsch^2^, Anna Schultze^2^, Krishnan Bhaskaran^2^, Helen I McDonald^2^, John Tazare^2^, Laurie Tomlinson^2^, Tom Palmer^3,4^, Brian MacKenna^1^, Richard Croker^1^, Caroline E Morton^1^, Colm Andrews^1^, Robin Parks^1^, Lisa Hopcroft^1^, Jon Massey^1^, Jessica Morley^1^, Amir Mehrkar^1^, Seb Bacon^1^, Dave Evans^1^, Peter Inglesby^1^, George Hickman^1^, Simon Davy^1^, Tom Ward^1^, Rosalind M Eggo^2^, Angel YS Wong^2^, Rohini Mathur^2^, Kevin Wing^2^, Harriet Forbes^2^, Daniel Grint^2^, Viyasaan Mahalingasivam^2^, Bang Zheng^2^, Ian J Douglas^2^, Stephen JW Evans^2^, Liam Smeeth^2^, Chris Bates^5^, Jonathan Cockburn^5^, John Parry^5^, Frank Hester^5^, Sam Harper^5^, Jonathan AC Sterne^3,4^, Miguel A Hernán^6,7^, Ben Goldacre^1^.
 
@@
 
-4\. NIHR Bristol, Biomedical Research Centre, Bristol, UK
+4\. NIHR Bristol Biomedical Research Centre, Bristol, UK
 
@@
 
-The UK COVID-19 vaccination programme delivered the first "booster" doses in mid September 2021, initially in groups at high risk of severe disease and then across the adult population.
-The BNT162b2 mRNA Pfizer-BioNTech and the Moderna mRNA-1273 vaccines were used.
+The UK COVID-19 vaccination programme delivered its first "booster" doses in mid September 2021, initially in groups at high risk of severe disease and then across the adult population.
+The BNT162b2 mRNA Pfizer-BioNTech vaccine was used initially, with Moderna mRNA-1273 subsequently also used.
 
@@
 
-With the approval of NHS England we used the OpenSAFELY-TPP database, covering 40% of GP practices in England and linked to national coronavirus surveillance, hospital episodes, and death registry data, to estimate the effectiveness of BNT162b2 boosting in adults eligible for boosting between `r format(as.Date(study_dates$studystart_date), "%d %B %Y")` and `r format(as.Date(study_dates$lastvax3_date), "%d %B %Y")`.
+With the approval of NHS England we used the OpenSAFELY-TPP database, covering 40% of GP practices in England and linked to national coronavirus surveillance, hospital episodes, and death registry data, to estimate the effectiveness of BNT162b2 in eligible adults between `r format(as.Date(study_dates$pfizerstart_date), "%e %B")` and `r format(as.Date(study_dates$pfizerend_date), "%e %B %Y")`.
+Follow up was for up to 10 weeks, during a period when the Delta variant of SARS-CoV-2 was dominant.
 Each booster recipient was matched with an unboosted control on factors relating to booster priority status and prior immunisation.
-Primary outcomes were documented SARS-CoV-2 infection, COVID-19 hospital admission, and COVID-19 death up to 10 weeks after recruitment.
-Hazard ratios were estimated using Cox models adjusting for additional covariates that were not used as matching factors.
-Follow-up spanned a period in which the Delta variant was predominant.
+Outcomes were positive SARS-CoV-2 test, COVID-19 hospital admission, COVID-19 death and non-COVID-9 death.
+Vaccine effectiveness, defined as 1 minus the adjusted hazard ratio, was estimated using Cox models.
 
@@
 We matched `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` BNT162b2 booster recipients to unboosted controls.
-During the first 28 days after the booster, the estimated hazard ratio (95% CI) comparing a BNT162b2 booster dose to two doses only was `r filter(hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` for documented SARS-CoV-2 infection, `r filter(hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` for COVID-19 death.
+During the first 28 days after the booster, estimated booster effectiveness (95% CI) comparing a BNT162b2 booster dose to two doses only was `r filter(overallhazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for non-COVID-19 death.
 
@@
 
-We estimated high protection of BNT162b2 boosting against infection, COVID-19 hospitalisation, and COVID-19 death, though with some evidence of waning over time.
+We estimated high protection of BNT162b2 boosting against positive SARS-CoV-2 test, COVID-19 hospital admission, and COVID-19 death.
 
@@
 
-The national COVID-19 vaccination programme in England began a new phase in September 2021 with administration of the first booster COVID-19 vaccine doses in those who had already received their primary vaccination course of a first and second dose @NHSEnglandNHS.
-Initially offered to those at highest risk of severe disease, eligibility was progressively extended and by 15 December 2021 every adult in England was eligible @NHSEnglandNHSb.
-Booster doses were initially available no earlier than six months after dose two, but this was reduced to three months on 8 December 2021, following concerns over the surge in cases and the emergence of the Omicron variant @NHSEnglandNHSa.
+The national COVID-19 vaccination programme in England administered its first booster doses in September 2021 in adults who had already received their two-dose primary vaccination course @NHSEnglandNHS.
+Eligibility was initially restricted to those at highest risk of severe disease, then progressively extended.
+By 15 December 2021 every adult was eligible @NHSEnglandNHSb.
+Booster doses were initially available no earlier than six months after dose two, but this was reduced to three months on 8 December 2021, following concerns over the emergence of the Omicron variant @NHSEnglandNHSa1 @NHSEnglandNHSa2.
 Vaccine prioritisation schedules were guided by recommendations from the Joint Committee for Vaccine and Immunisation (JCVI) expert working group.
@@
 
-We set out to emulate a target trial assessing the effectiveness of booster vaccination with BNT162b2 against COVID-19 outcomes in adults who had already received two COVID-19 vaccine doses.
+We aimed to emulate a target trial assessing the effectiveness of booster vaccination with BNT162b2 against COVID-19 outcomes @dagan2021 @barda2021, by comparing fully vaccinated adults who did and did not receive a booster vaccine dose.
 We used the OpenSAFELY-TPP linked primary care database covering around 40% of English residents.
-Recruitment and follow-up covers an era where the Delta variant is dominant, prior to the dominance of the Omicron variant.
+Recruitment and follow-up covers an era where the Delta variant is dominant.
 
 # Methods
-
-## Study design
-
-We used observational data to emulate a target trial of the effectiveness of booster vaccination @dagan2021 @barda2021, by comparing people who had received a BNT162b2 booster dose with matched people who had not.
 
@@
 
-We recruited adults who were eligible for booster vaccination between `r format(as.Date(study_dates$studystart_date), "%d %B")` and `r format(as.Date(study_dates$lastvax3_date), "%d %B %Y")` inclusive.
-Eligibility on each day was as follows: adults (18 years and over) who were registered at a GP practice using TPP's SystmOne clinical information system; received either two doses of BNT162b2 doses or two doses of ChAdOx1-S (mixed dosing and Moderna mRNA-1273 doses were not considered due to small numbers); not a health or social care worker, not a resident in a care or nursing home, and not medically housebound or receiving end-of-life care (due to distinct clustering of vaccine uptake and infection risk in these groups that could not be reasonably measured and controlled for); no evidence of SARS-CoV-2 infection or COVID-19 disease within the previous 90 days; not undergoing an unplanned hospital admission; information on sex, ethnicity, deprivation, and NHS region was complete.
+We included adults aged 18 years and over who were eligible for booster vaccination between `r format(as.Date(study_dates$pfizerstart_date), "%e %B")` and `r format(as.Date(study_dates$pfizerend_date), "%e %B %Y")` inclusive (the "recruitment period").
+On each day within this period, eligibility for the study was determined as follows: registered at a GP practice using TPP's SystmOne clinical information system; received a two-dose primary vaccination course of either BNT162b2 or ChAdOx1-S (mixed dosing and Moderna mRNA-1273 were not considered due to small numbers); not a health or social care worker, not resident in a care or nursing home; not medically housebound or receiving end-of-life care; no evidence of SARS-CoV-2 infection or COVID-19 disease within the previous 90 days; not undergoing an unplanned hospital admission; complete information on sex, ethnicity, deprivation, and NHS region.
 
@@
 
-On each day of recruitment, each eligible person receiving a booster dose on that day was recruited to the treatment group and matched with an unboosted control person who had received only two doses.
-Pairs were matched on the following characteristics: primary course vaccine brand (2× BNT162b2 or 2× ChAdOx1-S); date of the second vaccine dose, within 7 days; NHS region (East of England, Midlands, London, North East and Yorkshire, North West, South East, South West); any evidence of prior SARS-CoV-2 infection (positive SARS-CoV-2 test, "probable" infection documented in primary care, or COVID-19 hospitalisation); clinical risk groups used for prioritisation (clinically extremely vulnerable, clinically at-risk, neither); age, within 3 years, and by age groups used for prioritisation.
-The matching and exclusion criteria jointly ensure that treatment groups are also matched within JCVI priority group.
-
-People selected as controls were not eligible to be included again as a control in a subsequent trial, but were eligible for selection into the treatment group of a subsequent trial.
-Any unmatched boosted people were dropped.
+On each day of the recruitment period, each eligible person receiving a booster dose was recruited to the treatment group and matched, if possible, with an unboosted control person who had received only two doses.
+Pairs were matched on the following characteristics: primary course vaccine brand; date of second vaccine dose, within 7 days; NHS region (East of England, Midlands, London, North East and Yorkshire, North West, South East, South West); evidence of prior SARS-CoV-2 infection (positive SARS-CoV-2 test, "probable" infection documented in primary care, or COVID-19 hospital admission); clinical risk groups used for prioritisation (clinically extremely vulnerable, clinically at-risk, neither); age, within 3 years; and age groups used by JCVI for prioritisation.
+<!-- The matching and exclusion criteria jointly ensure that treatment groups are also matched within JCVI priority group. -->
+
+People selected as controls were not eligible to be included again as a control, but were eligible for subsequent selection into the treatment group. Any unmatched boosted people were excluded.
 
@@
 
-The following three outcomes were considered: documented SARS-CoV-2 infection; COVID-19 hospital admission; COVID-19 death; and non-COVID-19 death as a negative control outcome @Lipsitch2010.
+Four outcomes were considered: positive SARS-CoV-2 test; COVID-19 hospital admission; COVID-19 death; and non-COVID-19 death.
 SARS-CoV-2 infections were identified using SGSS testing records and based on swab date.
@@
 
-Each matched pair was followed from the day of recruitment (i.e., time zero) until the earliest of death, GP practice de-registration, booster receipt by the control, 10 weeks, or `r format(as.Date(study_dates$studyend_date), "%d %B %Y")`.
+Each matched pair was followed from the day of recruitment (i.e., time zero) until the earliest of death, GP practice de-registration, booster receipt by the control, 10 weeks, or `r format(as.Date(study_dates$studyend_date), "%e %B %Y")`.
 
@@
 
-The following additional variables were included as covariates in the Cox models (see below) or else to describe the treatment groups: sex (male or female); English Index of Multiple Deprivation (IMD, grouped by quintile); ethnicity (Black, Mixed, South Asian, White, Other, as per the UK census); morbidity count (diabetes, BMI \> 40kg/m^2^, chronic heart disease, chronic kidney disease, chronic liver disease, chronic respiratory disease or severe asthma, chronic neurological disease); learning disabilities; severe mental illness; the number of SARS-CoV-2 tests in the 6 months prior to the study start date; elective in-hospital episode at the time of recruitment.
+The following variables were included as covariates in Cox models (see below): sex (male or female); English Index of Multiple Deprivation (IMD, grouped by quintile); ethnicity (Black, Mixed, South Asian, White, Other, as per the UK census); morbidity count (diabetes, BMI \> 40kg/m^2^, chronic heart disease, chronic kidney disease, chronic liver disease, chronic respiratory disease or severe asthma, chronic neurological disease); learning disabilities; severe mental illness; the number of SARS-CoV-2 tests in the 6 months prior to the study start date; elective in-hospital episode at the time of recruitment.
 
@@
 
-For each outcome, we estimated unadjusted cumulative incidence curves using the Kaplan-Meier estimator separately in the boosted and unboosted groups.
-Rates of competing events (due to all-cause or cause-specific death, depending on the outcome) were anticipated to be low and so were not accounted for.
-
-We use Cox models, stratified by recruitment day, NHS region, and brand of second vaccine dose, to estimate hazard ratios comparing boosted with unboosted people.
-We estimated effectveness over all follow-up time (up to 70 days), and separately in days 1-28 and 29-70.
-Models were adjusted for additional variables not used as matching factors, described above.
+We estimated Kaplan-Meier cumulative incidence curves separately in the boosted and unboosted groups.
+Rates of competing events (due to all-cause or cause-specific death) were low and so were not explicitly accounted for.
+
+We used Cox models, stratified by recruitment day, NHS region, and brand of second vaccine dose, to estimate adjusted hazard ratios comparing boosted with unboosted people, overall and separately for days 1-28 and 29-70.
 A robust variance estimator was used to account for participant clustering.
@@
 
-We also estimated booster effectiveness in the following subgroups: primary vaccine course (both BNT162b2 or both ChAdOx1-S); age (18-64 or 65 and over); clinically extremely vulnerable or not.
-Analyses were conducted separately in each subgroup.
-
-Booster effectiveness is defined as one minus the hazard ratio, expressed as a percentage.
+We also estimated booster effectiveness separately in the following subgroups: primary vaccine course; age (18-64 or 65 and over); clinically extremely vulnerable or not.
+
+Booster vaccine effectiveness was calculated as one minus the hazard ratio, expressed as a percentage.
 
@@
 
-After exclusions for missing values on demographic variables, there were no missing values in remaining variables as they were each defined by the presence or absence of clinical codes or events in the patient record.
-This was therefore a complete-case analysis.
+After exclusions for missing values on demographic variables, there were no missing values in the remaining variables as they were each defined by the presence or absence of clinical codes or events in the patient record.
 
@@
 
-`r label_number(1, big.mark=",")(filter(matchflowchart, crit=="c0", treatment=="pfizer") %>% pull(n))` BNT162b2 booster recipients were identified between `r format(as.Date(matchsummary_treated$firstrecruitdate), "%d %B %Y")` and `r format(as.Date(matchsummary_treated$lastrecruitdate), "%d %B %Y")` in adults registered at a TPP practice, with `r label_number(1, big.mark=",")(filter(matchflowchart, treatment=="pfizer", crit=="c7") %>% pull(n))` (`r label_percent(0.1)(filter(matchflowchart, treatment=="pfizer", crit=="c7") %>% pull(pct_all))`) eligible for inclusion into the treatment group.
-Of these, `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` (`r label_percent(0.1)(filter(matchsummary, treatment=="pfizer") %>% pull(prop_matched_eligible))`) were successfully matched with unboosted controls (Table S1).
+`r label_number(1, big.mark=",")(filter(matchflowchart, crit=="c0", treatment=="pfizer") %>% pull(n))` adults registered at a TPP pracitce received a BNT162b2 booster vaccination during the recruitment period, with `r label_number(1, big.mark=",")(filter(matchflowchart, treatment=="pfizer", crit=="c7") %>% pull(n))` (`r label_percent(0.1)(filter(matchflowchart, treatment=="pfizer", crit=="c7") %>% pull(pct_all))`) eligible for inclusion into the treatment group, of whom `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` (`r label_percent(0.1)(filter(matchsummary, treatment=="pfizer") %>% pull(prop_matched_eligible))`) were matched with unboosted controls (Figure 1).
 
@@
 
-Participant characteristics at trial recruitment between treatment groups (Table 1) were typically well-balanced though with some exceptions.
-For instance, unboosted controls had higher levels of deprivation, higher rates of learning disabilities and severe mental illness, and lower SARS-CoV-2 testing frequency, though the standardised mean difference was consistently below 0.1 (Supplementary Figure S1).
+As expected, the matching factors has similar distributions in the boosted and unboosted groups at the start of follow up (Table 1), and the distribution of comorbidities was generally similar between the groups.
+However, unboosted controls had higher levels of deprivation, higher rates of learning disabilities and severe mental illness, and lower SARS-CoV-2 testing frequency, though the standardised mean difference was consistently below 0.1 (Supplementary Figure S2).
+
+## Estimated booster effectiveness
+
+### Main analysis
+
+There were `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="postest", fup_period=="All") %>% pull(events))` positive SARS-CoV-2 tests, `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="covidadmitted", fup_period=="All") %>% pull(events))` COVID-19 hospital admissions, and `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="coviddeath", fup_period=="All") %>% pull(events))` COVID-19 deaths and `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="noncoviddeath", fup_period=="All") %>% pull(events))` non-COVID-19 deaths across `r label_number(1, big.mark=",")((filter(matchsummary, treatment=="pfizer") %>% pull(fup_years)) *365.25/7)` person-weeks of follow-up (Table 2).
+
+The cumulative incidence of positive SARS-CoV-2 tests in boosted people appears to taper substantially around 7 days after boosting (Figure 2), something not observed in unboosted controls (Figure 2).
+The marginal risk difference at 70 days for boosted versus unboosted was `r riskdiff %>% filter(treatment=="pfizer", outcome=="postest", time==70, subgroup=="none") %>% pull(survdiff_inv_display)` per 1000 people.
+For more severe outcomes the cumulative incidence between treatment groups also diverged substantially.
+At 70 days, there were  `r (riskdiff %>% filter(treatment=="pfizer", outcome=="covidadmitted", time==70, subgroup=="none") %>% pull(survdiff_display))` fewer COVID-19 admissions, `r (riskdiff %>% filter(treatment=="pfizer", outcome=="coviddeath", time==70, subgroup=="none") %>% pull(survdiff_display))` fewer COVID-19 deaths, and  `r (riskdiff %>% filter(treatment=="pfizer", outcome=="covidadmitted", time==70, subgroup=="none") %>% pull(survdiff_display))` fewer non-COVID-19 deaths per 1000 people in the boosted group compared with the unboosted group.
+
+The overall estimated effectiveness (95% CI) comparing a BNT162b2 booster dose to two doses only was `r filter(overallhazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for non-COVID-19 death.
+Estimated effectiveness of BNT162b2 boosting in days 1-28 and days 29-70 was `r filter(meta2hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` and `r filter(meta2hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(meta2hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` and `r filter(meta2hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(meta2hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` and `r filter(meta2hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` for COVID-19 death, and `r filter(meta2hazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` and `r filter(meta2hazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` for non-COVID-19 death.
+
+Estimated effectiveness for narrower time periods are provided in supplementary materials (Table S3).
+
+### Subgroup analyses
+
+Estimated booster effectiveness was similar between subgroups defined by primary vaccine course.
+For those receiving 2× BNT162b2, estimated effectiveness was `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3) %>% pull(ve_display)` for non-COVID-19 death.
+For 2× ChAdOx1-S recipients this was `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup_level=="az-az", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup_level=="az-az", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup_level=="az-az", model==3) %>% pull(ve_display)`, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup_level=="az-az", model==3) %>% pull(ve_display)` respectively.
+
+For people aged 65 years and over, estimated effectiveness was `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for non-COVID-19 death.
+Estimated effectiveness was slightly lower in people aged under 65 years, at `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` respectively.
+
+For clinically extremely vulnerable people estimated effectiveness was `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for non-COVID-19 death.
+For people who were not clinically extremely vulnerable effectiveness was slightly higher, at `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="cev-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="cev-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="cev-FALSE", model==3) %>% pull(ve_display)`, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="cev-FALSE", model==3) %>% pull(ve_display)` respectively.
+
+# Discussion
+
+This observational study investigated the effectiveness of a BNT162b2 booster dose in `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` adults, and their matched controls, in England in the Delta era.
+We found high protection against all studied outcomes.
+There were apparent waning protection against positive SARS-CoV-2 tests around 6 weeks after boosting, which is unlikely to be explained by increased transmissibility of the Omicron variant given Omicron was not yet dominant in the UK by the end of the follow-up window @COVID19OmicronDaily.
+Changing testing behaviours over time between boosted and unboosted people may influence these estimates, if for example persistently unboosted participants, who will contribute more person-time in later follow-up periods, were less likely to get tested.
+
+Protection was similar between 2× BNT162b2 primary course recipients and 2× ChAdOx1-S recipients, except for COVID-19 death in the first 28 days where protected appeared higher for those whose primary course was 2× BNT162b2 than for 2× ChAdOx1-S.
+For severe outcomes, protection was lower in the clinical extremely vulnerable sub-population than in those who were not vulnerable.
+
+We considered effectiveness in a population where Delta was the predominant circulating variant @COVID19OmicronDaily.
+Re-running the analysis against a backdrop of the now globally-dominant Omicron variant will be undertaken where valid concurrent comparisons between boosted and unboosted people can be made.
+Further, mRNA-1273 boosting began later than for BNT162b2 so we did not evaluate effectiveness of this vaccine due to insufficient follow-up.
+An analysis of mRNA-1273 boosting and comparative effectiveness between mRNA-1273 and BNT162b2 is planned.
+
+### Strengths and Limitations
+
+We used routinely-collected health records with comprehensive coverage of primary care, hospital admissions, COVID-19 testing and vaccination, and death registrations.
+This provides substantial power to estimate effectiveness of booster doses against severe but rare outcomes such as hospital admission and death, including period-specific effects to investigate waning protection.
+
+We carefully matched booster recipients with unboosted controls using characteristics known at baseline to strengthen exchangeability between treatment groups and reduce dependence on modelling assumptions in the estimates of effectiveness.
+However, despite reasonable balance of important baseline characteristics and additional adjustment for a range of potential confounders, the possibility of unmeasured confounding remains.
+In particular, the apparent protective effect immediately after boosting is biologically implausible, and will be explained in part by the occurrence at baseline of COVID-19 symptoms in the unboosted controls that were not recorded in the health record, and so could not be identified and excluded.
+Such symptoms are much less likely in booster recipients as symptomatic people, or those with a recent positive test, were ineligible for boosting.
+
+The matching approach is robust but inefficient; a minority of boosted individuals are not retained after matching, and those who are retained are censored when their matched control is boosted, reducing the follow-up duration and therefore statistical power.
+Millions of eligible participants may be required to obtain precise estimates, particularly when event rates are low, and so extending the approach to smaller subgroups may not be feasible.
+For instance, due to small numbers, we did not study booster effectiveness in those who had received the mRNA-1273 vaccine as their primary course or those who had received a heterologous primary course, despite many thousands of such recipients in England.
+
+Positive SARS-CoV-2 test data likely underestimates the true incidence of infection.
+The mass availability of lateral flow tests in the UK during the study period, whose results are not routinely recorded in COVID-19 surveillance data, means many infections, including symptomatic infections, may be undocumented, despite encouragement to seek a confirmatory PCR test.
+Potential differences in testing behaviour between boosted and unboosted people also contributes to the unreliability of testing data as a means to assess effectiveness @glasziou2022.
+SARS-CoV-2 testing was not widely available early in the pandemic so evidence of prior infection is likely to be under-ascertained.
+We may not have all records of negative tests so counts of number of tests are also incomplete.
+Hospital admission records are only completed after discharge, so some very long hospital stays commencing within the follow-up period may not have been included.
+
+We excluded a number of groups, such as health care workers and care home residents, where testing use, vaccination uptake, and infection risk were unusual or had substantial within-group heterogeneity that could not adequately measured and controlled for.
+The generalisability of our results to these excluded groups is unclear.
+
+### Findings in context
+
+<!-- [ cov-boost trial] -->
+
+A phase III trial assessing BNT162b2 booster efficacy in two-dose BNT162b2 recipients without prior infection reported a relative efficacy of 95.6% (95% CI 89.3-98.6) against Delta infection @PfizerBioNTechAnnounce2021.
+Observational studies in the UK have provided evidence for increased protection from the booster vaccine against symptomatic COVID-19 infection in comparison to second dose recipients using a test negative design.
+Andrews et al.
+reported estimates of relative vaccine effectiveness against symptomatic disease in the 14 days after a BNT162b2 booster dose in those aged over 50 years of 87.4% (95%CI 84.9-89.4) where the primary course was ChAdOx1-S and 84.4% (95%CI 82.8-85.8) for BNT162b2 @andrews2021.
+A study in Scotland reported comparable estimates of relative vaccine effectiveness for S gene positive symptomatic infection (83% (95%CI 81-84) 6-49 years, 88% (95%CI 86-89) over 50 years) in the 14 days after a BNT162b2 or mRNA-1273 booster dose, but lower estimates for S gene negative symptomatic infection (56% (95%CI 51-60) 6-49 years, 57% (95%CI 52-62) over 50 years) @sheikhSeverityOmicronVariant2021.
+Elsewhere, a study in Israeli health registry data found strong protection against admission (1 -- risk ratio (95%CI) = 93% (88-97)), severe disease (92% (82-97)), and COVID-19 death (81% (59-97)), with similar results in specific demographic and clinical subgroups @barda2021.
+
+Our results broadly agree with these findings and provide important additional insights into effectiveness in specific clinical subgroups, younger age groups, and more severe outcomes over a longer period of follow up.
+Encouragingly, we found effectiveness was broadly similar irrespective of whether BNT162b2 or ChAdOx1-S was given for the primary vaccination course, and the were only slight reductions in effectiveness in those who were clinically extremely vulnerable.
+
+<!-- @1050236/technical-briefing-34-14-january-2022.pdf -->
+
+### Conclusion
+
+This study of over 8 million people in England found high protection for BNT162b2 boosting against positive SARS-CoV-2 tests, COVID-19 hospital admission, and death during a period in which the Delta variant was dominant.
 
@@
 
-#### Figure 1: Booster uptake and matching success
-
-Cumulative booster vaccination over the duration of the study period, by matching eligibility and matching success.
-
-```{r, vaxdate, echo=FALSE, message=FALSE, warning=FALSE, out.width='90%', out.height='50%', results='asis', out.extra=''}
-#read_rds(fs::path(output_dir_os, "descriptive", "vaxdate", "plot_vaxdate_stack.rds"))
-
-xmin <- min(matchcoverage$vax3_date )
-xmax <- max(matchcoverage$vax3_date )+1
-data_coverage <- matchcoverage  %>%
-  mutate(
-    status = factor(status, levels=c("ineligible", "unmatched", "matched")),
-    status_descr = fct_relabel(status, str_to_title)
-  ) %>%
-  arrange(status) %>%
-  mutate(
-    status_descr = fct_inorder(status_descr)
-  ) %>%
-  filter(treatment=="pfizer") %>%
-  mutate(
-    # vax12_type = fct_case_when(
-    #   vax12_type == "pfizer-pfizer" ~ "BNT162b2",
-    #   vax12_type == "az-az" ~ "ChAdOx1-S",
-    #   vax12_type == "moderna-moderna" ~ "mRNA-1273",
-    #   TRUE ~ NA_character_
-    # ),
-    course = fct_case_when(
-      vax12_type=="pfizer-pfizer" & treatment=="pfizer" ~ "2× BNT162b2  / 1× BNT162b2",
-      vax12_type=="az-az" & treatment=="pfizer" ~ "2× ChAdOx1-S / 1× BNT162b2",
-      vax12_type=="pfizer-pfizer" & treatment=="moderna" ~ "2× BNT162b2 / 1× mRNA-1273",
-      vax12_type=="az-az" & treatment=="moderna" ~ "2× ChAdOx1-S / 1× mRNA-1273",
-      TRUE ~ NA_character_
-    )
-  )
-
-data_annotate <- 
-  data_coverage %>%
-  filter(vax12_type=="pfizer-pfizer") %>%
-  group_by(status_descr, course) %>%
-  summarise(
-    xpos = first(lubridate::floor_date((vax3_date), "1 month")),
-    ypos= max(cumuln)
-  ) %>%
-  ungroup() %>%
-  arrange(desc(status_descr)) %>%
-  mutate(
-    ypos=(cumsum(ypos) + lag(cumsum(ypos), 1, 0))/2
-  ) %>%
-  mutate(
-    course="2× ChAdOx1-S / 1× BNT162b2"
-  )
-
-ggplot(data_coverage) +
-  geom_col(
-    aes(
-      x=vax3_date+0.5,
-      y=cumuln,
-      group=status_descr,
-      fill=status_descr,
-      colour=NULL
-    ),
-    position=position_stack(),
-    alpha=0.5,
-    width=1
-  )+
-  geom_rect(xmin=xmin, xmax= xmax+1, ymin=0, ymax=6, fill="grey", colour="transparent")+
-  geom_text(data=data_annotate, mapping=aes(x = xpos, y = (ypos), label = status_descr, colour=status_descr), hjust=0, show.legend = FALSE)+
-  facet_wrap(vars(course))+
-  scale_x_date(
-    breaks = unique(lubridate::ceiling_date(matchcoverage$vax3_date, "1 month")),
-    limits = c(lubridate::floor_date((xmin), "1 month"), NA),
-    labels = scales::label_date("%d/%m"),
-    expand = expansion(add=1),
-  )+
-  scale_y_continuous(
-    labels = scales::label_number(accuracy = 1, big.mark=","),
-    expand = expansion(c(0, NA))
-  )+
-  scale_fill_brewer(type="qual", palette="Set2", direction = -1)+
-  scale_colour_brewer(type="qual", palette="Set2", direction = -1)+
-  labs(
-    x="Date",
-    y="Cumulative booster vaccines",
-    colour=NULL,
-    fill=NULL,
-    alpha=NULL
-  ) +
-  theme_minimal()+
-  theme(
-    axis.line.x.bottom = element_line(),
-    axis.text.x.top=element_text(hjust=0),
-    strip.text.y.right = element_text(angle = 0),
-    axis.ticks.x=element_line(),
-    legend.position = "none"
-  )+
-  NULL
-cat("  \n\n")
-```
+## Acknowledgements
+
+We are very grateful for all the support received from the TPP Technical Operations team throughout this work, and for generous assistance from the information governance and database teams at NHS England / NHSX.
+
+Our thanks to Susana Monge-Corella for helpful discussions on optimising the matching algorithm.
+
+## Conflicts of Interest
+
+All authors have completed the ICMJE uniform disclosure form and declare the following: BG has received research funding from Health Data Research UK (HDRUK), the Laura and John Arnold Foundation, the Wellcome Trust, the NIHR Oxford Biomedical Research Centre, the NHS National Institute for Health Research School of Primary Care Research, the Mohn-Westlake Foundation, the Good Thinking Foundation, the Health Foundation, and the World Health Organisation; he also receives personal income from speaking and writing for lay audiences on the misuse of science.
+IJD holds shares in GlaxoSmithKline (GSK).
+
+## Funding
+
+This work was supported by the UKRI Medical Research Council MR/V015737/1 and the Longitudinal Health and Wellbeing COVID-19 National Core Study (UKRI MRC MC_PC_20030 and MC_PC_20059).
+Data assets were supported by the Data and Connectivity National Core Study, led by Health Data Research UK in partnership with the Office for National Statistics and funded by UKRI (MC_PC_20029 and MC_PC_20058).
+The OpenSAFELY platform is funded by the Wellcome Trust.
+TPP provided technical expertise and infrastructure within their data centre pro bono in the context of a national emergency.
+BGs work on clinical informatics is supported by the NIHR Oxford Biomedical Research Centre and the NIHR Applied Research Collaboration Oxford and Thames Valley.
+BG's work on better use of data in healthcare more broadly is currently funded in part by: NIHR Oxford Biomedical Research Centre, NIHR Applied Research Collaboration Oxford and Thames Valley, the Mohn-Westlake Foundation, NHS England, and the Health Foundation; all DataLab staff are supported by BG's grants on this work.
+LS reports grants from Wellcome, MRC, NIHR, UKRI, British Council, GSK, British Heart Foundation, and Diabetes UK outside this work.
+KB holds a Wellcome Senior Research Fellowship (220283/Z/20/Z).
+HIM is funded by the NIHR Health Protection Research Unit in Immunisation, a partnership between Public Health England and London School of Hygiene & Tropical Medicine.
+AYSW holds a fellowship from the British Heart Foundation.
+EJW holds grants from MRC.
+RM holds a Sir Henry Wellcome Fellowship funded by the Wellcome Trust (201375/Z/16/Z).
+HF holds a UKRI fellowship.
+IJD holds grants from NIHR and GSK.
+JACS is funded by the NIHR Bristol Biomedical Research Centre and by Health Data Research UK.
+
+Funders had no role in the study design, collection, analysis, and interpretation of data; in the writing of the report and in the decision to submit the article for publication.
+
+The views expressed are those of the authors and not necessarily those of the NIHR, NHS England, Public Health England or the Department of Health and Social Care.
+
+For the purpose of Open Access, the author has applied a CC BY public copyright licence to any Author Accepted Manuscript (AAM) version arising from this submission.
+
+## Information governance and ethical approval
+
+NHS England is the data controller for OpenSAFELY-TPP; TPP is the data processor; and study authors using OpenSAFELY have the approval of NHS England.
+This implementation of OpenSAFELY is hosted within the TPP environment which is accredited to the ISO 27001 information security standard and is NHS IG Toolkit compliant @betad @datasec ; Patient data has been pseudonymised for analysis and linkage using industry standard cryptographic hashing techniques; all pseudonymised datasets transmitted for linkage onto OpenSAFELY are encrypted; access to the platform is via a virtual private network (VPN) connection, restricted to a small group of researchers; the researchers hold contracts with NHS England and only access the platform to initiate database queries and statistical models; all database activity is logged; only aggregate statistical outputs leave the platform environment following best practice for anonymisation of results such as statistical disclosure control for low cell counts @isb1523 .
+The OpenSAFELY research platform adheres to the obligations of the UK General Data Protection Regulation (GDPR) and the Data Protection Act 2018.
+In March 2020, the Secretary of State for Health and Social Care used powers under the UK Health Service (Control of Patient Information) Regulations 2002 (COPI) to require organisations to process confidential patient information for the purposes of protecting public health, providing healthcare services to the public and monitoring and managing the COVID-19 outbreak and incidents of exposure; this sets aside the requirement for patient consent @coronavi2020 .
+Taken together, these provide the legal bases to link patient datasets on the OpenSAFELY platform.
+General practices, from which the primary care data are obtained, are required to share relevant health information to support the public health response to the pandemic, and have been informed of the OpenSAFELY analytics platform.
+
+This study was approved by the Health Research Authority (REC reference 20/LO/0651) and by the LSHTM Ethics Board (reference 21863).
+
+BG is guarantor.
+
+\newpage
+
+## Tables and Figures
 
@@
 table1 %>%
+  select(-`Boosted, ineligible`, -`Boosted, eligible, unmatched`) %>%
   gt(groupname_col="treatment_descr") %>%
@@
     fup_period == "All",
-    outcome %in% c("postest", "covidadmitted", "coviddeath"),
+    outcome %in% c("postest", "covidadmitted", "coviddeath", "noncoviddeath"),
     treatment == "pfizer"
@@
   select(
-    treatment_descr, 
+    #treatment_descr, 
     outcome_descr,
+    #fup_period,
     q_1,
@@
     rate_0,
-    rrECI
+    #rrECI
   ) %>%
@@
   gt(
-   groupname_col = c("treatment_descr"),
+   #groupname_col = c("treatment_descr"),
   ) %>%
@@
     outcome_descr = "Outcome",
-
+    #fup_period = "Days since booster",
     q_0 = "Events / person-years",
@@
     rate_1 = "Incidence rate",
-    rrECI = "Incidence rate ratio (95% CI)"
+    #rrECI = "Incidence rate ratio (95% CI)"
   ) %>%
@@
 
-### Figure 2: Cumulative outcome incidence
+\newpage
+
+#### Table 3: Estimated booster effectiveness
+
+Vaccine effectiveness, $(1-\text{hazard ratio})\times100\%$, for the main and subgroup analyses across all outcomes.
+
+```{r table_effects, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}
+
+bind_rows(
+  meta2hazards,
+  overallhazards %>% mutate(term_left = 0, term_right=70)
+) %>%
+  filter(model==3, treatment=="pfizer") %>%
+  mutate(
+    period = fct_inorder(paste(term_left+1, "-", term_right)),
+    ve_display = str_remove(ve_display, fixed("%"))
+  ) %>%
+  arrange(treatment_descr, outcome_descr, subgroup_level_descr, period) %>%
+  group_by(outcome, treatment, subgroup_level_descr, model) %>%
+  mutate(
+    #subgroup_level_descr=if_else(row_number()==1, subgroup_level_descr, factor(NA, levels=levels(subgroup_level_descr))),
+    period = paste0(period, " days")
+  ) %>%
+  ungroup() %>%
+  select(outcome_descr, subgroup_level_descr, period, ve_display) %>%
+  pivot_wider(
+    id_cols = c("outcome_descr", "subgroup_level_descr"),
+    names_from = "period",
+    #names_prefix ="period",
+    values_from = "ve_display"
+  ) %>%
+  gt(
+    groupname_col = c("outcome_descr"),
+  ) %>%
+  cols_label(
+    #treatment_descr = "Treatment",
+    subgroup_level_descr="",
+    outcome_descr = "Outcome",
+
+    #hr_display   = "HR (95% CI)",
+    #ve_display   = "VE (95% CI)",
+  ) %>%
+  cols_align(
+    align = "right",
+    columns = everything()
+  ) %>%
+  tab_spanner(
+    label = "Booster vaccine effectiveness (95% CI)",
+    columns = ends_with("days")
+  ) %>%
+  fmt_missing(
+    c("subgroup_level_descr"),
+    missing_text=""
+  )
+
+```
+
+\newpage
+
+#### Figure 1: Inclusion criteria
+
+(need to convert to figure)
+
+```{r flowchart, echo=FALSE, warning=FALSE, message=FALSE}
+
+matchflowchart %>%
+  transmute(
+    treatment,
+    stage,
+    criteria = str_wrap(criteria, 40),
+    n,
+    n_exclude,
+    pct_all
+  ) %>%
+  pivot_wider(
+    id_cols = c(stage, criteria),
+    names_from = treatment,
+    values_from = c(n, n_exclude, pct_all)
+  ) %>%
+  gt() %>%
+  cols_label(
+    stage="",
+    criteria = "",
+
+    n_pfizer = "N",
+    #n_moderna = "N",
+
+    n_exclude_pfizer = "N excluded",
+    #n_exclude_moderna = "N excluded",
+
+    pct_all_pfizer = "% remaining",
+    #pct_all_moderna = "% remaining",
+  ) %>%
+  tab_spanner(
+    label = "BNT162b2",
+    columns = ends_with("pfizer")
+  ) %>%
+  # tab_spanner(
+  #   label = "mRNA-1273",
+  #   columns = ends_with("moderna")
+  # ) %>%
+  fmt_percent(
+    columns = starts_with(c("pct_")),
+    decimals = 1
+  ) %>%
+  fmt_number(
+    columns = starts_with(c("n_")),
+    decimals= 0
+  ) %>%
+  fmt_missing(
+    everything(),
+    missing_text="-"
+  ) %>%
+  tab_style(
+    style=cell_text(whitespace ="pre-wrap"),
+    #style=cell_text(style ="italic"),
+    locations = cells_body(columns = "criteria")
+  )
+```
+
+\newpage
+
+#### Figure 2: Cumulative outcome incidence
 
@@
 
-```{r cmlinc, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', out.width='90%', out.height='50%', results='asis', out.extra=''}
-
-plot_km <- km %>%
+```{r cmlinc, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', out.width='90%', fig.height=7.5, results='asis'}
+
+plot_km <- 
+  km %>%
   filter(
@@
 
+  ggsave(
+    filename="plot_km.svg",
+    path=output_dir_rmd,
+    plot=plot_km,
+  )
+
 plot_km
+
 ```
-
-## Estimated booster effectiveness
-
-### Main analysis
-
-Participants in the emulated trial experienced `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="postest", fup_period=="All") %>% pull(events))` documented infections, `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="covidadmitted", fup_period=="All") %>% pull(events))` COVID-19 hospital admissions, and `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="coviddeath", fup_period=="All") %>% pull(events))` COVID-19 deaths across `r label_number(1, big.mark=",")((filter(matchsummary, treatment=="pfizer") %>% pull(fup_years)) *365.25/7)` person-weeks of follow-up (Table 2, Figure 2).
-
-The estimated effectiveness (95% CI) comparing a BNT162b2 booster dose to two doses only over all follow-up time was `r filter(overallhazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3) %>% pull(ve_display)` for documented infection, `r filter(overallhazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(overallhazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 death.
-Effectiveness before and after 28 days since boosting is also estimated (Figure 3, Table 3), as well as for narrower periods (Supplementary Table S3).
-
-<!-- In the first 28 days of follow-up the estimated effectiveness (95% CI) comparing a BNT162b2 booster dose to two doses only, was `r filter(meta2hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` for documented infection, `r filter(meta2hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(meta2hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` for COVID-19 death (Figure 3).
--->
-
-<!-- For days 28-70, estimated effectiveness (95% CI) comparing a BNT162b2 booster dose to two doses only, was `r filter(meta2hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` for documented infection, `r filter(meta2hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(meta2hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` for COVID-19 death (Figure 3), suggesting a potential waning of booster effectiveness with time.
--->
-
-### Subgroup analyses
-
-The estimated effectiveness (95% CI) for 2× BNT162b2 and 2× ChAdOx1-S primary course recipients respectively were `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3) %>% pull(ve_display)` and `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup_level=="az-az", model==3) %>% pull(ve_display)` for documented infection, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3) %>% pull(ve_display)` and `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup_level=="az-az", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3) %>% pull(ve_display)` and `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup_level=="az-az", model==3) %>% pull(ve_display)` for COVID-19 death (Figure 2).
 
@@
   ggsave(
-    filename="plot_hazards.pdf",
+    filename="plot_hazards.svg",
     path=output_dir_rmd,
@@
 
-## Table 3: Estimated booster effectiveness
-
-Main and subgroup analyses
-
-```{r table_effects, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}
-
-bind_rows(
-  meta2hazards,
-  overallhazards %>% mutate(term_left = 0, term_right=70)
-) %>%
-  filter(model==3, treatment=="pfizer") %>%
-  mutate(
-    period = fct_inorder(paste(term_left+1, "-", term_right)),
-    ve_display = str_remove(ve_display, fixed("%"))
-  ) %>%
-  arrange(treatment_descr, outcome_descr, subgroup_level_descr, period) %>%
-  group_by(outcome, treatment, subgroup_level_descr, model) %>%
-  mutate(
-    subgroup_level_descr=if_else(row_number()==1, subgroup_level_descr, factor(NA, levels=levels(subgroup_level_descr)))
-  ) %>%
-  ungroup() %>%
-  select(outcome_descr, subgroup_level_descr, period, hr_display, ve_display) %>%
-  gt(
-    groupname_col = c("outcome_descr"),
-  ) %>%
-  cols_label(
-    #treatment_descr = "Treatment",
-    subgroup_level_descr="",
-    outcome_descr = "Outcome",
-
-    period = "Days since booster",
-    hr_display   = "HR (95% CI)",
-    ve_display   = "VE (95% CI)",
-  ) %>%
-  cols_align(
-    align = "right",
-    columns = everything()
-  ) %>%
-  fmt_missing(
-    c("subgroup_level_descr"),
-    missing_text=""
-  )
-
-```
-
-\newpage
-
-# Discussion
-
-This observational study investigated the effectiveness of a BNT162b2 booster dose in `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` adults, and their matched controls, in England in the Delta era.
-We found high protection against all studied outcomes.
-There were apparent waning protection against documented infection around 6 weeks after boosting, which is unlikely to be explained by increased transmissibility of the Omicron variant given Omicron was not yet dominant in the UK by the end of the follow-up window @COVID19OmicronDaily.
-Changing testing behaviours over time between boosted and unboosted people may influence these estimates, if for example persistently unboosted participants, who will contribute more person-time in later follow-up periods, were less likely to get tested.
-
-Protection was similar between 2× BNT162b2 primary course recipients and 2× ChAdOx1-S recipients, except for COVID-19 death in the first 28 days where protected appeared higher for those whose primary course was 2× BNT162b2 than for 2× ChAdOx1-S.
-For severe outcomes, protection was lower in the clinical extremely vulnerable sub-population.
-
-We considered effectiveness in a population where Delta was the predominant circulating variant @COVID19OmicronDaily.
-Re-running the analysis against a backdrop of the now globally-dominant Omicron variant will be undertaken where valid concurrent comparisons between boosted and unboosted people can be made.
-Further, mRNA-1273 boosting begun later than for BNT162b2 so we did not evaluate effectiveness of this vaccine due to insufficient follow-up.
-An analysis of mRNA-1273 boosting and comparative effectiveness between mRNA-1273 and BNT162b2 is planned.
-
-### Strengths and Limitations
-
-We used routinely-collected health records with comprehensive coverage of primary care, hospital admissions, COVID-19 testing and vaccination, and death registrations.
-This provides substantial power to estimate effectiveness of booster doses against severe but rare outcomes such as hospitalisation and death, including period-specific effects to investigate waning protection.
-
-We carefully matched booster recipients with unboosted controls using characteristics known at baseline to strengthen exchangeability between treatment groups and reduce dependence on modelling assumptions in the estimates of effectiveness.
-However, despite reasonable balance of important baseline characteristics and additional adjustment for a range of potential confounders, the possibility of unmeasured confounding remains.
-In particular, the apparent protective effect immediately after boosting is biologically implausible, and will be explained in part by the occurrence of COVID-19 symptoms at baseline that were not recorded in the health record, and so could not be identified and excluded.
-
-The matching approach robust but inefficient; a minority of boosted individuals are not retained after matching, and those who are retained are censored when their matched control is boosted, reducing the follow-up duration and therefore statistical power.
-Millions of eligible participants may be required to obtain precise estimates, particularly when event rates are low, and so extending the approach to smaller subgroups may not be feasible.
-For instance, due to small numbers, we did not study booster effectiveness in those who had received the mRNA-1273 vaccine as their primary course or those who had received a heterologous primary course, despite many thousands of such recipients in England.
-
-Documented SARS-CoV-2 infection data likely underestimates the true incidence of infection.
-The mass availability of lateral flow tests in the UK during the study period, whose results are not routinely recorded in COVID-19 surveillance data, means many infections, including symptomatic infections, may be undocumented, despite encouragement to seek a confirmatory PCR test.
-Potential differences in testing behaviour between boosted and unboosted people also contributes to the unreliability of testing data as a means to assess effectiveness.
-SARS-CoV-2 testing was not widely available early in the pandemic so evidence of prior infection is likely to be under-ascertained.
-We may not have all records of negative tests so counts of number of tests are also incomplete.
-Hospital admission records are only completed after discharge, so some very long hospital stays commencing within the follow-up period may not have been included.
-
-We excluded a number of groups, such as health care workers and care home residents, where testing, vaccination, and infection risk characteristics were unusual or had substantial within-group heterogeneity that could not be adequately accounted for.
-The generalisability of our results to these excluded groups is unclear.
-
-### Findings in context
-
-<!-- [ cov-boost trial] -->
-
-A phase III trial assessing BNT162b2 booster efficacy in two-dose BNT162b2 recipients without prior infection reported a relative efficacy of 95.6% (95% CI 89.3-98.6) against Delta infection @PfizerBioNTechAnnounce2021.
-Observational studies in the UK have provided evidence for increased protection from the booster vaccine against symptomatic COVID-19 infection in comparison to second dose recipients using a test negative design.
-Andrews et al.
-reported estimates of relative vaccine effectiveness against syptomatic disease in the 14 days after a BNT162b2 booster dose in those aged over 50 years of 87.4% (95%CI 84.9-89.4) where the primary course was ChAdOx1-S and 84.4% (95%CI 82.8-85.8) for BNT162b2 @andrews2021.
-A study in Scotland reported comparable estimates of relative vaccine effectiveness for S gene positive symptomatic infection (83% (95%CI 81-84) 6-49 years, 88% (95%CI 86-89)  over 50 years) in the 14 days after a BNT162b2 or mRNA-1273 booster dose, but lower estimates for S gene negative symptomatic infection (56% (95%CI 51-60) 6-49 years, 57% (95%CI 52-62) over 50 years) @sheikhSeverityOmicronVariant2021.
-
-Elsewhere, a study in Israeli health registry data @barda2021 found strong protection against admission (1 -- risk ratio (95%CI) = 93% (88-97)), severe disease (92% (82-97)), and COVID-19 death (81% (59-97)), with similar results in specific demographic and clinical subgroups @barda2021.
-Our results broadly agree with these findings and provide important additional insights into effectiveness in specific clinical subgroups and over longer follow-up.
-
-<!-- @1050236/technical-briefing-34-14-january-2022.pdf -->
-
-### Conclusion
-
-This study of over 8 million people in England found high protection for BNT162b2 boosting against documented infection, COVID-19 hospitalisation and death during a period in which the Delta variant was dominant.
-
-\newpage
-
-## Acknowledgements
-
-We are very grateful for all the support received from the TPP Technical Operations team throughout this work, and for generous assistance from the information governance and database teams at NHS England / NHSX.
-
-## Conflicts of Interest
-
-All authors have completed the ICMJE uniform disclosure form and declare the following: BG has received research funding from Health Data Research UK (HDRUK), the Laura and John Arnold Foundation, the Wellcome Trust, the NIHR Oxford Biomedical Research Centre, the NHS National Institute for Health Research School of Primary Care Research, the Mohn-Westlake Foundation, the Good Thinking Foundation, the Health Foundation, and the World Health Organisation; he also receives personal income from speaking and writing for lay audiences on the misuse of science.
-IJD holds shares in GlaxoSmithKline (GSK).
-
-## Funding
-
-This work was supported by the Medical Research Council MR/V015737/1 and the Longitudinal Health and wellbeing strand of the National Core Studies programme.
-The OpenSAFELY platform is funded by the Wellcome Trust.
-TPP provided technical expertise and infrastructure within their data centre pro bono in the context of a national emergency.
-BGs work on clinical informatics is supported by the NIHR Oxford Biomedical Research Centre and the NIHR Applied Research Collaboration Oxford and Thames Valley.
-BG's work on better use of data in healthcare more broadly is currently funded in part by: NIHR Oxford Biomedical Research Centre, NIHR Applied Research Collaboration Oxford and Thames Valley, the Mohn-Westlake Foundation, NHS England, and the Health Foundation; all DataLab staff are supported by BG's grants on this work.
-LS reports grants from Wellcome, MRC, NIHR, UKRI, British Council, GSK, British Heart Foundation, and Diabetes UK outside this work.
-KB holds a Wellcome Senior Research Fellowship (220283/Z/20/Z).
-HIM is funded by the NIHR Health Protection Research Unit in Immunisation, a partnership between Public Health England and London School of Hygiene & Tropical Medicine.
-AYSW holds a fellowship from the British Heart Foundation.
-EJW holds grants from MRC.
-RM holds a Sir Henry Wellcome Fellowship funded by the Wellcome Trust (201375/Z/16/Z).
-HF holds a UKRI fellowship.
-IJD holds grants from NIHR and GSK.
-
-Funders had no role in the study design, collection, analysis, and interpretation of data; in the writing of the report and in the decision to submit the article for publication.
-
-The views expressed are those of the authors and not necessarily those of the NIHR, NHS England, Public Health England or the Department of Health and Social Care.
-
-For the purpose of Open Access, the author has applied a CC BY public copyright licence to any Author Accepted Manuscript (AAM) version arising from this submission.
-
-## Information governance and ethical approval
-
-NHS England is the data controller for OpenSAFELY-TPP; TPP is the data processor; and study authors using OpenSAFELY have the approval of NHS England.
-This implementation of OpenSAFELY is hosted within the TPP environment which is accredited to the ISO 27001 information security standard and is NHS IG Toolkit compliant @betad @datasec ; Patient data has been pseudonymised for analysis and linkage using industry standard cryptographic hashing techniques; all pseudonymised datasets transmitted for linkage onto OpenSAFELY are encrypted; access to the platform is via a virtual private network (VPN) connection, restricted to a small group of researchers; the researchers hold contracts with NHS England and only access the platform to initiate database queries and statistical models; all database activity is logged; only aggregate statistical outputs leave the platform environment following best practice for anonymisation of results such as statistical disclosure control for low cell counts @isb1523 .
-The OpenSAFELY research platform adheres to the obligations of the UK General Data Protection Regulation (GDPR) and the Data Protection Act 2018.
-In March 2020, the Secretary of State for Health and Social Care used powers under the UK Health Service (Control of Patient Information) Regulations 2002 (COPI) to require organisations to process confidential patient information for the purposes of protecting public health, providing healthcare services to the public and monitoring and managing the COVID-19 outbreak and incidents of exposure; this sets aside the requirement for patient consent @coronavi2020 .
-Taken together, these provide the legal bases to link patient datasets on the OpenSAFELY platform.
-General practices, from which the primary care data are obtained, are required to share relevant health information to support the public health response to the pandemic, and have been informed of the OpenSAFELY analytics platform.
-
-This study was approved by the Health Research Authority (REC reference 20/LO/0651) and by the LSHTM Ethics Board (reference 21863).
-
-BG is guarantor.
-
-\newpage
-
 ## References

### 2022-03-14 manuscript/draft-supplement.Rmd
update manuscript + supplement

@@
 
-### Table S1: Inclusion criteria
+\newpage
 
-```{r flowchart, echo=FALSE, warning=FALSE, message=FALSE}
+#### Table S1: Characteristics of candidate participants
 
-matchflowchart %>%
-  transmute(
-    treatment,
-    stage,
-    criteria = str_wrap(criteria, 40),
-    n,
-    n_exclude,
-    pct_all
+Candidate matching characteristics as on the day of recruitment into the treatment or control group.
+
+```{r table1, echo=FALSE, warning=FALSE, message=FALSE}
+table1_body <- read_csv(fs::path(output_dir_os, "matchtable1.csv")) %>% filter(treatment=="pfizer")
+table1_by <- read_csv(fs::path(output_dir_os, "matchtable1by.csv")) %>% filter(treatment=="pfizer")
+
+table1 <- 
+  table1_body %>%
+  #filter(outcome %in% outcomes) %>%
+  group_by(variable) %>%
+  mutate(
+    label = if_else(row_number()==1, label, str_c("&nbsp;&nbsp;&nbsp;", label)),
   ) %>%
-  pivot_wider(
-    id_cols = c(stage, criteria),
-    names_from = treatment,
-    values_from = c(n, n_exclude, pct_all)
+  mutate(
+    across(
+      starts_with("stat_"),
+      ~if_else(is.na(.x) & row_number()==1, "", .x)
+    )
   ) %>%
-  gt() %>%
-  cols_label(
-    stage="",
-    criteria = "",
+  ungroup() %>%
+  mutate(
+    order = cumsum(lag(variable, 1, first(variable))!=variable)
+  ) %>%
+  select(
+    variable, var_label, label, starts_with("stat_"), order
+  ) %>%
+  rename(!!!set_names(table1_by$by_col, table1_by$by_chr))
 
-    n_pfizer = "N",
-    #n_moderna = "N",
-
-    n_exclude_pfizer = "N excluded",
-    #n_exclude_moderna = "N excluded",
-
-    pct_all_pfizer = "% remaining",
-    #pct_all_moderna = "% remaining",
-  ) %>%
-  tab_spanner(
-    label = "BNT162b2",
-    columns = ends_with("pfizer")
-  ) %>%
-  # tab_spanner(
-  #   label = "mRNA-1273",
-  #   columns = ends_with("moderna")
-  # ) %>%
-  fmt_percent(
-    columns = starts_with(c("pct_")),
-    decimals = 1
-  ) %>%
-  fmt_number(
-    columns = starts_with(c("n_")),
-    decimals= 0
-  ) %>%
-  fmt_missing(
-    everything(),
-    missing_text="-"
+table1 %>%
+  select(-`Boosted, ineligible`, -`Boosted, eligible, unmatched`) %>%
+  gt(groupname_col="treatment_descr") %>%
+  fmt_markdown(columns = "label") %>%
+  tab_style(
+    style = list(
+      cell_fill(color = "gray96")
+    ),
+    locations = cells_body(
+      rows = order%%2 == 0
+    )
   ) %>%
   tab_style(
-    style=cell_text(whitespace ="pre-wrap"),
-    #style=cell_text(style ="italic"),
-    locations = cells_body(columns = "criteria")
-  )
+    style = list(
+      cell_text(weight = "bold")
+    ),
+    locations = cells_column_labels(everything())
+  ) %>%
+  cols_label(
+     label=md("Characteristic")
+  ) %>%
+  cols_hide(c("variable", "var_label", "order"))
 ```
@@
 
-### Figure S1: Standardised Mean Differences between Boosted and Unboosted groups
+#### Figure S1: Booster uptake and matching success
+
+Cumulative booster vaccination over the duration of the study period, by matching eligibility and matching success.
+
+```{r, vaxdate, echo=FALSE, message=FALSE, warning=FALSE, out.width='90%', out.height='50%', results='asis', out.extra=''}
+#read_rds(fs::path(output_dir_os, "descriptive", "vaxdate", "plot_vaxdate_stack.rds"))
+
+xmin <- min(matchcoverage$vax3_date )
+xmax <- max(matchcoverage$vax3_date )+1
+data_coverage <- matchcoverage  %>%
+  mutate(
+    status = factor(status, levels=c("ineligible", "unmatched", "matched")),
+    status_descr = fct_relabel(status, str_to_title)
+  ) %>%
+  arrange(status) %>%
+  mutate(
+    status_descr = fct_inorder(status_descr)
+  ) %>%
+  filter(treatment=="pfizer") %>%
+  mutate(
+    # vax12_type = fct_case_when(
+    #   vax12_type == "pfizer-pfizer" ~ "BNT162b2",
+    #   vax12_type == "az-az" ~ "ChAdOx1-S",
+    #   vax12_type == "moderna-moderna" ~ "mRNA-1273",
+    #   TRUE ~ NA_character_
+    # ),
+    course = fct_case_when(
+      vax12_type=="pfizer-pfizer" & treatment=="pfizer" ~ "2× BNT162b2  / 1× BNT162b2",
+      vax12_type=="az-az" & treatment=="pfizer" ~ "2× ChAdOx1-S / 1× BNT162b2",
+      vax12_type=="pfizer-pfizer" & treatment=="moderna" ~ "2× BNT162b2 / 1× mRNA-1273",
+      vax12_type=="az-az" & treatment=="moderna" ~ "2× ChAdOx1-S / 1× mRNA-1273",
+      TRUE ~ NA_character_
+    )
+  )
+
+data_annotate <- 
+  data_coverage %>%
+  filter(vax12_type=="pfizer-pfizer") %>%
+  group_by(status_descr, course) %>%
+  summarise(
+    xpos = first(lubridate::floor_date((vax3_date), "1 month")),
+    ypos= max(cumuln)
+  ) %>%
+  ungroup() %>%
+  arrange(desc(status_descr)) %>%
+  mutate(
+    ypos=(cumsum(ypos) + lag(cumsum(ypos), 1, 0))/2
+  ) %>%
+  mutate(
+    course="2× ChAdOx1-S / 1× BNT162b2"
+  )
+
+ggplot(data_coverage) +
+  geom_col(
+    aes(
+      x=vax3_date+0.5,
+      y=cumuln,
+      group=status_descr,
+      fill=status_descr,
+      colour=NULL
+    ),
+    position=position_stack(),
+    alpha=0.5,
+    width=1
+  )+
+  geom_rect(xmin=xmin, xmax= xmax+1, ymin=0, ymax=6, fill="grey", colour="transparent")+
+  geom_text(data=data_annotate, mapping=aes(x = xpos, y = (ypos), label = status_descr, colour=status_descr), hjust=0, show.legend = FALSE)+
+  facet_wrap(vars(course))+
+  scale_x_date(
+    breaks = unique(lubridate::ceiling_date(matchcoverage$vax3_date, "1 month")),
+    limits = c(lubridate::floor_date((xmin), "1 month"), NA),
+    labels = scales::label_date("%d/%m"),
+    expand = expansion(add=1),
+  )+
+  scale_y_continuous(
+    labels = scales::label_number(accuracy = 1, big.mark=","),
+    expand = expansion(c(0, NA))
+  )+
+  scale_fill_brewer(type="qual", palette="Set2", direction = -1)+
+  scale_colour_brewer(type="qual", palette="Set2", direction = -1)+
+  labs(
+    x="Date",
+    y="Cumulative booster vaccines",
+    colour=NULL,
+    fill=NULL,
+    alpha=NULL
+  ) +
+  theme_minimal()+
+  theme(
+    axis.line.x.bottom = element_line(),
+    axis.text.x.top=element_text(hjust=0),
+    strip.text.y.right = element_text(angle = 0),
+    axis.ticks.x=element_line(),
+    legend.position = "none"
+  )+
+  NULL
+cat("  \n\n")
+```
+
+### Figure S2: Standardised Mean Differences between Boosted and Unboosted groups
+
+For variables that were matched exactly the SMD is zero, and so are not shown.
 
@@
   smd %>%
+  filter(
+    !(var_label %in% c("Primary vaccine course", "Age", "Region", "Prior documented SARS-CoV-2 infection", "Clinically extremely vulnerable"))
+  ) %>%
   mutate(
     level=fct_rev(fct_inorder(level))
+  ) %>%
+  droplevels() %>%
+  mutate(
+    variable = fct_inorder(variable),
+    variable_card = as.numeric(variable)%%2,
+    variable_levels = replace_na(as.character(variable_levels), ""),
+  ) %>%
+  arrange(variable) %>%
+  mutate(
+    level = fct_rev(fct_inorder(str_replace(paste(variable, variable_levels, sep=": "),  "\\:\\s$", ""))),
+    cardn = row_number()
   ) %>%
@@
 plot_smd
-
-```
-
-### Figure S2: Kaplan-Meier cumulative incidence curves
-
-```{r plot_km, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}
-
-plot_km <- km %>%
-  filter(
-    outcome %in% outcomes,
-    treatment=="pfizer"
-  ) %>%
-    ggplot()+
-    geom_step(aes(x=time, y=(1-surv)*1000, group=treated_descr, colour=treated_descr))+
-    geom_rect(aes(xmin=time, xmax=leadtime, ymin=(1-surv.ll)*1000, ymax=(1-surv.ul)*1000, group=treated_descr, fill=treated_descr), alpha=0.1)+
-    geom_hline(aes(yintercept=0), colour='black')+
-    facet_grid(rows=vars(outcome_descr_wrap), cols=vars(subgroup_level_descr), switch="y", scales="free_y")+
-    scale_x_continuous(
-      breaks = seq(0,7*max(postbaselinecuts),by=14),
-      expand = expansion(0)
-    )+
-    scale_y_continuous(
-      expand = expansion(0),
-      #limits = c(0L,NA)
-    )+
-    scale_colour_brewer(type="qual", palette="Set1", na.value="grey")+
-    scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey")+
-    labs(
-      x=NULL,
-      y="Cumulative incidence\nper 1000",
-      colour=NULL
-    )+
-    theme_minimal()+
-    theme(
-      legend.position=c(0.15,0.85),
-      legend.justification = c(1,0),
-      legend.direction = "vertical",
-      axis.text.x.top=element_text(hjust=0),
-
-      strip.placement = "outside",
-      strip.background = element_blank(),
-      strip.text.y.left = element_text(angle = 0),
-
-      panel.border = element_blank(),
-      axis.line.y = element_line(colour = "black"),
-      axis.line.x.top = element_line(colour = "black"),
-      panel.grid.minor.x = element_blank(),
-      panel.grid.minor.y = element_blank(),
-
-      panel.spacing = unit(0.8, "lines"),
-
-    )+
-    NULL
-
-plot_km
 

### 2022-03-14 analysis/design.R

@@ -15,7 +15,7 @@ fs::dir_create(here("lib", "design"))
 
 study_dates <- lst(
   index_date = "2021-09-16", # index date for dates as "time since index date" format
-  pfizerstart_date = "2021-09-16", #start of recruitment thursday 16 sepetember first pfizer booster jabs administered in england
+  pfizerstart_date = "2021-09-16", #start of recruitment thursday 16 september first pfizer booster jabs administered in england
   pfizerend_date = "2021-12-16", # end of recruitment (13 weeks later)
   #pfizerfollowend_date = "2021-12-31", # end of follow-up
 

### 2022-03-14 analysis/merge_seqtrialcox.R

@@ -461,7 +461,7 @@ write_csv(tab_summary_baseline_redacted$table_body, fs::path(output_dir, "merge_
 write_csv(tab_summary_baseline_redacted$df_by, fs::path(output_dir, "merge_table1by.csv"))
 gtsave(as_gt(tab_summary_baseline_redacted), fs::path(output_dir, "merge_table1.html"))
 
-# love plot ----
+# love / smd plot ----
 
 
 data_smd <- tab_summary_baseline$meta_data %>%

### 2022-03-14 manuscript/draft-manuscript.Rmd
do not keep md

@@
   html_document:
-    keep_md: yes
     self_contained: yes

### 2022-03-14 manuscript/draft-manuscript.Rmd
rename chunks

@@
 
-```{r table1, echo=FALSE, warning=FALSE, message=FALSE}
+```{r table-table1, echo=FALSE, warning=FALSE, message=FALSE}
 table1_body <- read_csv(fs::path(output_dir_os, "matchtable1.csv")) %>% filter(treatment=="pfizer")
@@
 
-```{r table2, echo=FALSE, warning=FALSE, message=FALSE}
+```{r table-incidence, echo=FALSE, warning=FALSE, message=FALSE}
 incidences %>%
@@
 
-```{r table_effects, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}
+```{r table-effects, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}
 
@@
 
-```{r flowchart, echo=FALSE, warning=FALSE, message=FALSE}
+```{r figure-flowchart, echo=FALSE, warning=FALSE, message=FALSE}
 
@@
 
-```{r cmlinc, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', out.width='90%', fig.height=7.5, results='asis'}
+```{r figure-cmlinc, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', out.width='90%', fig.height=7.5, results='asis'}
 
@@
 
-```{r plot_effects, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}
+```{r figure-effects, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}
 

### 2022-03-14 analysis/report_seqtrialcox.R
replace redacted counts with rounded counts

@@ -207,6 +207,7 @@ format_ratio = function(numer,denom, width=7){
   )
 }
 
+threshold <- 5
 
 incidence_rate_redacted <- local({
 
@@ -260,38 +261,36 @@ incidence_rate_redacted <- local({
       rrECI = paste0(rrE, " ", rrCI)
     )
 
-  redacted <-
+  rounded <-
     unredacted_wide %>%
     mutate(
-      n_0 = redactor2(n_0, 5, n_0),
-      n_1 = redactor2(n_0, 5, n_1),
+      n_0 = ceiling_any(n_0, threshold+1),
+      n_1 = ceiling_any(n_1, threshold+1),
 
-      rate_0 = redactor2(events_1, 5, rate_0),
-      rate_1 = redactor2(events_1, 5, rate_1),
+      rate_0 = ceiling_any(events_0, threshold+1),
+      rate_1 = ceiling_any(events_1, threshold+1),
 
-      rr = redactor2(pmin(events_1, events_0), 5, rr),
-      rrE = redactor2(pmin(events_1, events_0), 5, rrE),
-      rrCI = redactor2(pmin(events_1, events_0), 5, rrCI),
-      rrECI = redactor2(pmin(events_1, events_0), 5, rrECI),
+      rr =  rate_1 / rate_0,
+      rrE = scales::label_number(accuracy=0.01, trim=FALSE)(rr),
+      rrCI = rrCI_exact(events_1, yearsatrisk_1, events_0, yearsatrisk_0, 0.01),
+      rrECI = paste0(rrE, " ", rrCI),
 
-      events_0 = redactor2(events_0, 5),
-      events_1 = redactor2(events_1, 5),
+      events_0 = ceiling_any(events_0, threshold+1),
+      events_1 = ceiling_any(events_1, threshold+1),
 
       q_0 = format_ratio(events_0, yearsatrisk_0),
       q_1 = format_ratio(events_1, yearsatrisk_1),
     )
 
-  redacted
+  rounded
 
 })
 
-write_csv(incidence_rate_redacted, fs::path(output_dir, "report_incidence.csv"))
+write_csv(incidence_rate_rounded, fs::path(output_dir, "report_incidence.csv"))
 
 
 ## kaplan meier cumulative risk differences ----
 
-
-threshold <- 5
 
 dat_surv <-
   data_seqtrialcox %>%

### 2022-03-16 analysis/report_seqtrialcox.R
bug

@@ -209,7 +209,7 @@ format_ratio = function(numer,denom, width=7){
 
 threshold <- 5
 
-incidence_rate_redacted <- local({
+incidence_rate_rounded <- local({
 
   unredacted_treated <-
     data_seqtrialcox %>%

### 2022-03-16 analysis/report_seqtrialcox.R
decommission model-based marginal risk for now

@@ -478,102 +478,106 @@ ggsave(filename=fs::path(output_dir, "report_cifplot.png"), plot_cif, width=20, 
 
 ## marginal cumulative risk differences ----
 
-model3 <- read_rds(fs::path(output_dir, "model_obj3.rds"))
+# do not run for now--
+if(FALSE){
 
-data_tidy_surv <- broom::tidy(survfit(model3)) %>%
-  group_by(strata) %>%
-  summarise(
-    n.event=sum(n.event),
-    n.risk=first(n.risk),
-    n.censor = sum(n.censor),
-  )
+  model3 <- read_rds(fs::path(output_dir, "model_obj3.rds"))
 
-if(any(data_tidy_surv$n.event==0)){
-  message("some strata have zero events -- cannot calculate marginalised risk.", "\n")
-} else {
-
-
-  #
-  # formula1_pw
-  #
-  # model3 <- coxph(
-  #   formula = Surv(tstop, ind_outcome) ~ treated_period_1 + strata(region),# +
-  #     #strata(region) + strata(jcvi_group) + strata(vax12_type),
-  #   data = data_seqtrialcox,
-  #  # robust = TRUE,
-  #  # id = patient_id,
-  #   na.action = "na.fail"
-  # )
-
-
-  ## need to redo treatment-time indicators too, as overwriting `treated`` variable not enough
-
-  data_seqtrialcox1 <-
-    data_seqtrialcox %>%
-    select(-starts_with("treated_period_")) %>%
-    rename(treated_period=fup_period) %>%
-    fastDummies::dummy_cols(select_columns = c("treated_period")) %>%
-    mutate(treated=1L)
-
-  data_seqtrialcox0 <-
-    data_seqtrialcox %>%
-    mutate(
-      across(
-        starts_with("treated_period_"),
-        ~0L
-      ),
-      treated = 0L
+  data_tidy_surv <- broom::tidy(survfit(model3)) %>%
+    group_by(strata) %>%
+    summarise(
+      n.event=sum(n.event),
+      n.risk=first(n.risk),
+      n.censor = sum(n.censor),
     )
 
-  times <- seq_len(last(postbaselinecuts)+1)
-
-  surv1 <- survexp(~1, data=data_seqtrialcox1, ratetable = model3, method="ederer", times=times)# times=seq_len(end(postbaselinecuts)))
-  surv0 <-  survexp(~1, data=data_seqtrialcox0, ratetable = model3, method="ederer", times=times)# times=seq_len(end(postbaselinecuts)))
-
[…]

### 2022-03-16 manuscript/draft-manuscript.Rmd
add risk ratio and risk difference to table 2

@@
 incidences <- read_csv(fs::path(output_dir_os, "none", "incidence_all.csv")) %>% mutate(events=events_1+events_0)
-
 pfizerdurationweeks <- as.numeric(as.duration(study_dates$pfizerend_date - study_dates$pfizerstart_date +1) , "weeks")
@@
 ```{r table-incidence, echo=FALSE, warning=FALSE, message=FALSE}
+
+km70 <- km %>% 
+  filter(leadtime==max(postbaselinecuts)) %>%
+  mutate(
+    treated=(treated_descr=="Boosted")*1L
+  ) %>%
+  select(
+    treatment, outcome , subgroup, 
+    treated, surv
+  ) %>%
+  pivot_wider(
+    id_cols= c("treatment", "outcome", "subgroup"),
+    names_from=treated,
+    names_prefix="surv_",
+    values_from=surv
+  ) %>%
+  mutate(
+    risk_1 = 1 - surv_1,
+    risk_0 = 1 - surv_0,
+
+    kmrr = risk_1 / risk_0,
+    kmrd = risk_1 - risk_0,
+    VErr = 1-kmrr,
+    kmrrE = label_number(accuracy=0.01, scale=1)(kmrr),
+    kmrdE = label_number(accuracy=0.01, scale=1000)(kmrd)
+  )
+
 incidences %>%
+  left_join(km70) %>%
   filter(
     fup_period == "All",
-    outcome %in% c("postest", "covidadmitted", "coviddeath", "noncoviddeath"),
+    outcome %in% outcomes,
     treatment == "pfizer"
-  ) %>%
-  left_join(
-    matchsummary %>% select(treatment, n_eligible, prop_matched_eligible),
-    by=c("treatment")
-  ) %>%
-  mutate(
-    treatment_descr = paste0(
-      #treatment_descr, " - ",
-      label_number(1, big.mark=",")(n_eligible), " eligible, ",
-      label_number(1, big.mark=",")(n_1), " matched (",
-      label_percent(0.1)(prop_matched_eligible), ")."
-    )
   ) %>%
@@
     q_1,
-    rate_1,
+    risk_1,
     q_0, 
-    rate_0,
+    risk_0,
+    kmrrE,
+    kmrdE
     #rrECI
@@
   mutate(
-    rate_1=if_else(rate_1<0.001, "<0.001", label_number(0.001)(rate_1)),
-    rate_0=if_else(rate_0<0.001, "<0.001", label_number(0.001)(rate_0)),
+    risk_1=if_else(risk_1<0.001, "<0.001", label_number(0.001)(risk_1)),
+    risk_0=if_else(risk_0<0.001, "<0.001", label_number(0.001)(risk_0)),
   ) %>%
@@
 
-    rate_0 = "Incidence rate",
-    rate_1 = "Incidence rate",
+    #rate_0 = "Incidence rate",
+    #rate_1 = "Incidence rate",
+
+    risk_0 = "Cumulative risk",
+    risk_1 = "Cumulative risk",
+    kmrrE = "Risk ratio",
+    kmrdE = "Risk difference (per 1000 people)"
     #rrECI = "Incidence rate ratio (95% CI)"

### 2022-03-16 analysis/report_seqtrialcox.R
change where treated descr is made, and keep treated

@@ -299,10 +299,7 @@ dat_surv <-
     tte_outcome = max(tstop) - min(tstart),
     ind_outcome = as.integer(last(ind_outcome))
   ) %>%
-  mutate(
-    treated_descr = if_else(treated==1L, "Boosted", "Unboosted"),
-  ) %>%
-  group_by(treated_descr) %>%
+  group_by(treated) %>%
   nest() %>%
   mutate(
     n_events = map_int(data, ~sum(.x$ind_outcome, na.rm=TRUE)),
@@ -311,8 +308,11 @@ dat_surv <-
     }),
     surv_obj_tidy = map(surv_obj, ~tidy_surv(.x, addtimezero = TRUE)),
   ) %>%
-  select(treated_descr, n_events, surv_obj_tidy) %>%
-  unnest(surv_obj_tidy)
+  select(treated, n_events, surv_obj_tidy) %>%
+  unnest(surv_obj_tidy) %>%
+  mutate(
+    treated_descr = if_else(treated==1L, "Boosted", "Unboosted"),
+  )
 
 data_surv_rounded <-
   dat_surv %>%
@@ -328,7 +328,7 @@ data_surv_rounded <-
     n.censor = c(NA, diff(cml.censor)),
     n.risk = ceiling_any(max(n.risk, na.rm=TRUE), threshold+1) - (cml.event + cml.censor)
   ) %>%
-  select(treated_descr, time, leadtime, interval, surv, surv.ll, surv.ul, n.risk, n.event, n.censor)
+  select(treated, treated_descr, time, leadtime, interval, surv, surv.ll, surv.ul, n.risk, n.event, n.censor)
 
 
 write_csv(data_surv_rounded, fs::path(output_dir, "report_km.csv"))

### 2022-03-16 analysis/combine_model.R
add cif plot

@@ -465,3 +465,36 @@ cif <- model_metaparams %>%
 
 write_csv(cif, fs::path(output_dir, "cif.csv"))
 
+
+
+plot_cif <-
+  cif %>%
+  mutate(
+    treatment_subgroup_descr = paste0(subgroup_descr, " \n", treatment_descr)
+  ) %>%
+  ggplot(aes(group=treated_descr, colour=treated_descr, fill=treated_descr)) +
+  geom_step(aes(x=time, y=cmlinc))+
+  geom_rect(aes(xmin=time, xmax=leadtime, ymin=cmlinc.ll, ymax=cmlinc.ul), alpha=0.1, colour="transparent")+
+  facet_grid(rows=vars(outcome_descr), cols=vars(treatment_subgroup_descr), switch="y")+
+  scale_color_brewer(type="qual", palette="Set1", na.value="grey") +
+  scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey") +
+  scale_x_continuous(breaks = seq(0,600,14), limits=c(min(postbaselinecuts), max(postbaselinecuts)+1), expand = c(0, 0))+
+  scale_y_continuous(expand = expansion(mult=c(0,0.01)))+
+  coord_cartesian(xlim=c(0, NA))+
+  labs(
+    x="Days",
+    y="Cumulative incidence",
+    colour=NULL,
+    title=NULL
+  )+
+  theme_minimal()+
+  theme(
+    axis.line.x = element_line(colour = "black"),
+    panel.grid.minor.x = element_blank(),
+    legend.position=c(.05,.95),
+    legend.justification = c(0,1),
+  )+
+  NULL
+plot_cif
+
+ggsave(filename=fs::path(output_dir, "cifplot.png"), plot_cif, width=20, height=15, units="cm")

### 2022-03-16 analysis/data_process.R
add cev_cv variable

@@ -170,6 +170,12 @@ data_processed <- data_extract %>%
     cv = immunosuppressed | chronic_kidney_disease | chronic_resp_disease | diabetes | chronic_liver_disease |
       chronic_neuro_disease | chronic_heart_disease | asplenia | learndis | sev_mental,
 
+    cev_cv = fct_case_when(
+      cev ~ "Clinically extremely vulnerable",
+      cv ~ "Clinically at-risk",
+      TRUE ~ "Not clinically at-risk"
+    ) %>% fct_rev(),
+
     multimorb =
       (sev_obesity) +
       (chronic_heart_disease) +
@@ -188,6 +194,15 @@ data_processed <- data_extract %>%
     # original priority groups https://assets.publishing.service.gov.uk/government/uploads/system/uploads/attachment_data/file/1007737/Greenbook_chapter_14a_30July2021.pdf#page=15
     # new priority groups https://www.england.nhs.uk/coronavirus/wp-content/uploads/sites/52/2021/07/C1327-covid-19-vaccination-autumn-winter-phase-3-planning.pdf
     # group 10 split into 16-39 and 40-49 because of earlier roll-out in 40+ from 15 Nov https://www.gov.uk/government/news/jcvi-issues-advice-on-covid-19-booster-vaccines-for-those-aged-40-to-49-and-second-doses-for-16-to-17-year-olds
+
+    jcvi_ageband = cut(
+      age_august2021,
+      breaks=c(-Inf, 18, 40, 50, 55, 60, 65, 70, 75, 80, Inf),
+      labels=c("under 18", "18-39", "40-49", "50-54", "55-59", "60-64", "65-69", "70-74", "75-79", "80+"),
+      right=FALSE
+    ),
+
+
     jcvi_group = fct_case_when(
       care_home_combined | hscworker  ~ "1",
       age_august2021>=80 ~ "2",

### 2022-03-16 analysis/design.R

@@ -87,13 +87,14 @@ write_rds(postbaselinecuts, here("lib", "design", "postbaselinecuts.rds"))
 # what matching variables
 exact_variables <- c(
 
-  "jcvi_group",
+  "jcvi_ageband",
+  "cev_cv",
   "vax12_type",
   #"vax2_week",
   "region",
   #"sex",
-  "cev",
-  "cv",
+  #"cev_cv",
+
   #"multimorb",
   "prior_covid_infection",
   #"immunosuppressed",

### 2022-03-16 analysis/match_seqtrialcox.R

@@ -258,6 +258,7 @@ data_baseline <-
     sev_mental,
     cev,
     cv,
+    cev_cv,
     sev_obesity,
     chronic_heart_disease,
     chronic_kidney_disease,

### 2022-03-16 analysis/table1.R

@@ -54,8 +54,8 @@ var_labels <-
     rural_urban_group ~ "Rural/urban category",
     stp ~ "STP",
     jcvi_group ~ "JCVI group",
-    cev ~ "Clinically extremely vulnerable",
-
+    #cev ~ "Clinically extremely vulnerable",
+    cev_cv ~ "Risk group",
     sev_obesity ~ "Body Mass Index > 40 kg/m^2",
 
     chronic_heart_disease ~ "Chronic heart disease",
@@ -80,6 +80,14 @@ var_labels <-
 
 ## baseline variables
 tab_summary_baseline <- data_cohort %>%
+  mutate(
+    #TODO can remove once data_process has rerun
+    cev_cv = case_when(
+      cev ~ "Clinically extremely vulnerable",
+      cv ~ "Clinically at-risk",
+      TRUE ~ "Not Clinically at-risk"
+    ),
+  ) %>%
   select(
     all_of(names(var_labels)),
     -age, -stp, -jcvi_group

### 2022-03-16 create-project.R

@@ -411,6 +411,25 @@ actions_list <- splice(
   action_model("pfizer", "noncoviddeath", "cev-FALSE"),
   action_model("pfizer", "noncoviddeath", "cev-TRUE"),
 
+  comment("### Models by prior infection ('prior_covid_infection')"),
+
+  action_model("pfizer", "postest", "prior_covid_infection-FALSE"),
+  action_model("pfizer", "postest", "prior_covid_infection-TRUE"),
+  action_model("pfizer", "covidemergency", "prior_covid_infection-FALSE"),
+  action_model("pfizer", "covidemergency", "prior_covid_infection-TRUE"),
+  action_model("pfizer", "covidadmittedproxy1", "prior_covid_infection-FALSE"),
+  action_model("pfizer", "covidadmittedproxy1", "prior_covid_infection-TRUE"),
+  action_model("pfizer", "covidadmitted", "prior_covid_infection-FALSE"),
+  action_model("pfizer", "covidadmitted", "prior_covid_infection-TRUE"),
+  action_model("pfizer", "noncovidadmitted", "prior_covid_infection-FALSE"),
+  action_model("pfizer", "noncovidadmitted", "prior_covid_infection-TRUE"),
+  action_model("pfizer", "covidcc", "prior_covid_infection-FALSE"),
+  action_model("pfizer", "covidcc", "prior_covid_infection-TRUE"),
+  action_model("pfizer", "coviddeath", "prior_covid_infection-FALSE"),
+  action_model("pfizer", "coviddeath", "prior_covid_infection-TRUE"),
+  action_model("pfizer", "noncoviddeath", "prior_covid_infection-FALSE"),
+  action_model("pfizer", "noncoviddeath", "prior_covid_infection-TRUE"),
+
   comment("### Models by age ('age65plus')"),
 
   action_model("pfizer", "postest", "age65plus-FALSE"),
@@ -481,6 +500,25 @@ actions_list <- splice(
   action_model("moderna", "coviddeath", "cev-TRUE"),
   action_model("moderna", "noncoviddeath", "cev-FALSE"),
   action_model("moderna", "noncoviddeath", "cev-TRUE"),
+
+  comment("### Models by prior infection ('prior_covid_infection')"),
+
+  action_model("moderna", "postest", "prior_covid_infection-FALSE"),
+  action_model("moderna", "postest", "prior_covid_infection-TRUE"),
+  action_model("moderna", "covidemergency", "prior_covid_infection-FALSE"),
[…]

### 2022-03-16 analysis/merge_seqtrialcox.R
add cev_cv variable

@@ -405,7 +405,8 @@ var_labels <- list(
   imd_Q5 ~ "IMD",
   region ~ "Region",
   #rural_urban_group ~ "Rural/urban category",
-  cev ~ "Clinically extremely vulnerable",
+  #cev ~ "Clinically extremely vulnerable",
+  cev_cv ~ "JCVI clinical risk group",
 
   sev_obesity ~ "Body Mass Index > 40 kg/m^2",
 
@@ -437,6 +438,12 @@ tab_summary_baseline <-
   mutate(
     N = 1L,
     status_descr = fct_recode(status, !!!status_recode),
+    #TODO can remove once data_process has rerun
+    cev_cv = case_when(
+      cev ~ "Clinically extremely vulnerable",
+      cv ~ "Clinically at-risk",
+      TRUE ~ "Not Clinically at-risk"
+    ),
   ) %>%
   select(
     status_descr,

### 2022-03-16 analysis/manuscript_objects.R
add prior covid subgroup

@@ -59,7 +59,7 @@ fs::file_copy(here("output", "match", "combined", "flowchart.csv"), here("output
 
 ## models ----
 
-for(subgroup_variable in c("none", "vax12_type", "cev", "age65plus")){
+for(subgroup_variable in c("none", "vax12_type", "cev", "age65plus", "prior_covid_infection")){
   fs::dir_create(here("output", "manuscript-objects", subgroup_variable))
   #fs::file_copy(here("output", "models", "seqtrialcox", "combined", subgroup_variable, "incidence.csv"), here("output", "manuscript-objects", subgroup_variable, "incidence.csv"), overwrite = TRUE)
   fs::file_copy(here("output", "models", "seqtrialcox", "combined", subgroup_variable, "km.csv"), here("output", "manuscript-objects", subgroup_variable, "km.csv"), overwrite = TRUE)

### 2022-03-16 manuscript/draft-manuscript.Rmd

@@
 
+# incidence <- 
+#   bind_rows(
+#     read_csv(fs::path(output_dir_os, "none", "incidence.csv")),
+#     read_csv(fs::path(output_dir_os, "vax12_type", "incidence.csv")),
+#     read_csv(fs::path(output_dir_os, "cev", "incidence.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
+#     read_csv(fs::path(output_dir_os, "age65plus", "incidence.csv")) %>% mutate(subgroup_level=as.character(subgroup_level))
+#   ) %>%
+#   filter(
+#     outcome %in% outcomes
+#   ) %>%
+#   mutate(
+#     subgroup_descr = fct_inorder(subgroup_descr),
+#     subgroup_level = fct_inorder(subgroup_level),
+#     subgroup_level_descr = fct_inorder(fct_explicit_na(subgroup_level_descr, "All")),
+#     treatment_descr = fct_inorder(treatment_descr),
+#     outcome_descr = fct_inorder(outcome_descr),
+#     outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=15)),
+#     #model_name = fct_inorder(str_wrap(model_name, 15)),
+#   )
+
 km <- 
@@
   )
+
+km70 <- km %>% 
+  filter(leadtime==max(postbaselinecuts)) %>%
+  mutate(
+    treated=(treated_descr=="Boosted")*1L
+  ) %>%
+  select(
+    treatment, outcome , subgroup, 
+    treated, surv
+  ) %>%
+  pivot_wider(
+    id_cols= c("treatment", "outcome", "subgroup"),
+    names_from=treated,
+    names_prefix="surv_",
+    values_from=surv
+  ) %>%
+  mutate(
+    risk_1 = 1 - surv_1,
+    risk_0 = 1 - surv_0,
+
+    kmrr = risk_1 / risk_0,
+    kmrd = risk_1 - risk_0,
+    VErr = 1-kmrr,
+    kmrrE = label_number(accuracy=0.01, scale=1)(kmrr),
+    kmrdE = label_number(accuracy=0.01, scale=1000)(kmrd)
+  )
+
 # 
@@
     ve_display = paste0(label_percent(accuracy=0.1)(ve), " (", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
+    ve_display95 = paste0(label_percent(accuracy=0.1)(ve), " (95% CI", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
   )
@@
 
-The UK COVID-19 vaccination programme delivered its first "booster" doses in mid September 2021, initially in groups at high risk of severe disease and then across the adult population.
-The BNT162b2 mRNA Pfizer-BioNTech vaccine was used initially, with Moderna mRNA-1273 subsequently also used.
+The UK COVID-19 vaccination programme delivered its first "booster" doses in September 2021, initially in groups at high risk of severe disease then across the adult population.
+The BNT162b2 Pfizer-BioNTech vaccine was used initially, with Moderna mRNA-1273 subsequently also used.
 
@@
 
-With the approval of NHS England we used the OpenSAFELY-TPP database, covering 40% of GP practices in England and linked to national coronavirus surveillance, hospital episodes, and death registry data, to estimate the effectiveness of BNT162b2 in eligible adults between `r format(as.Date(study_dates$pfizerstart_date), "%e %B")` and `r format(as.Date(study_dates$pfizerend_date), "%e %B %Y")`.
-Follow up was for up to 10 weeks, during a period when the Delta variant of SARS-CoV-2 was dominant.
+We used the OpenSAFELY-TPP database, covering 40% of English primary care practices in England and linked to national coronavirus surveillance, hospital episodes, and death registry data, to estimate the effectiveness of BNT162b2 in eligible adults between `r format(as.Date(study_dates$pfizerstart_date), "%e %B")` and `r format(as.Date(study_dates$pfizerend_date), "%e %B %Y")`, a period when the Delta variant of SARS-CoV-2 was dominant.
+Follow up was for up to 10 weeks.
 Each booster recipient was matched with an unboosted control on factors relating to booster priority status and prior immunisation.
+Additional factors were adjusted for in a Cox model to estimate hazard ratios.
 Outcomes were positive SARS-CoV-2 test, COVID-19 hospital admission, COVID-19 death and non-COVID-9 death.
-Vaccine effectiveness, defined as 1 minus the adjusted hazard ratio, was estimated using Cox models.
+Booster effectiveness was defined as 1 minus the hazard ratio.
 
@@
 
-The national COVID-19 vaccination programme in England administered its first booster doses in September 2021 in adults who had already received their two-dose primary vaccination course @NHSEnglandNHS.
+In mid-September 2021, the national COVID-19 vaccination programme in England administered its first booster doses in adults who had already received their two-dose primary vaccination course @NHSEnglandNHS.
 Eligibility was initially restricted to those at highest risk of severe disease, then progressively extended.
@@
 
-We included adults aged 18 years and over who were eligible for booster vaccination between `r format(as.Date(study_dates$pfizerstart_date), "%e %B")` and `r format(as.Date(study_dates$pfizerend_date), "%e %B %Y")` inclusive (the "recruitment period").
+We included adults aged $\geq$ 18 years who were eligible for booster vaccination between `r format(as.Date(study_dates$pfizerstart_date), "%e %B")` and `r format(as.Date(study_dates$pfizerend_date), "%e %B %Y")` inclusive (the "recruitment period").
 On each day within this period, eligibility for the study was determined as follows: registered at a GP practice using TPP's SystmOne clinical information system; received a two-dose primary vaccination course of either BNT162b2 or ChAdOx1-S (mixed dosing and Moderna mRNA-1273 were not considered due to small numbers); not a health or social care worker, not resident in a care or nursing home; not medically housebound or receiving end-of-life care; no evidence of SARS-CoV-2 infection or COVID-19 disease within the previous 90 days; not undergoing an unplanned hospital admission; complete information on sex, ethnicity, deprivation, and NHS region.
 
-Additionally, if the rolling weekly average count of BNT162b2 booster doses within strata defined by region, JCVI group, and the week of second dose was below 50 then recruitment did not occur for that strata on that day.
-This helped to ensure that recruitment was restricted to those who were boosted in line with national prioritisation schedules and that the matched controls were both eligible and able to receive a booster dose at the time of potential recruitment.
+Additionally, if the rolling weekly average count of BNT162b2 booster doses within strata defined by region, JCVI group, and the week of second dose was below 50 then recruitment did not occur for that stratum on that day.
+This helped ensure that recruitment was restricted to those boosted in line with national prioritisation schedules and that the matched controls were both eligible and able to receive a booster dose at the time of potential recruitment.
+
+## Matching and treatment groups
+
+On each day of the recruitment period, each eligible person receiving a booster dose was recruited to the treatment group and matched, if possible, with an eligible unboosted control person.
+Pairs were matched on: primary course vaccine brand; date of second vaccine dose, within 7 days; NHS region (East of England, Midlands, London, North East and Yorkshire, North West, South East, South West); evidence of prior SARS-CoV-2 infection (positive SARS-CoV-2 test, "probable" infection documented in primary care, or COVID-19 hospital admission); clinical risk groups used for prioritisation (clinically extremely vulnerable, clinically at-risk, neither); age, within 3 years; and age groups used by JCVI for prioritisation.
+<!-- The matching and exclusion criteria jointly ensure that treatment groups are also matched within JCVI priority group. -->
+
+People selected as controls were not eligible to be included again as a control, but were eligible for subsequent inclusion in the booster group. Any unmatched boosted people were excluded.
 
 The supplementary materials provide further details of the matching process.
-
-## Matching and treatment groups
-
-On each day of the recruitment period, each eligible person receiving a booster dose was recruited to the treatment group and matched, if possible, with an unboosted control person who had received only two doses.
-Pairs were matched on the following characteristics: primary course vaccine brand; date of second vaccine dose, within 7 days; NHS region (East of England, Midlands, London, North East and Yorkshire, North West, South East, South West); evidence of prior SARS-CoV-2 infection (positive SARS-CoV-2 test, "probable" infection documented in primary care, or COVID-19 hospital admission); clinical risk groups used for prioritisation (clinically extremely vulnerable, clinically at-risk, neither); age, within 3 years; and age groups used by JCVI for prioritisation.
-<!-- The matching and exclusion criteria jointly ensure that treatment groups are also matched within JCVI priority group. -->
-
-People selected as controls were not eligible to be included again as a control, but were eligible for subsequent selection into the treatment group. Any unmatched boosted people were excluded.
-
 <!-- We excluded matched pairs from days in which fewer than 10% of eligible treated people were successfully. -->
@@
 SARS-CoV-2 infections were identified using SGSS testing records and based on swab date.
-Both polymerase chain reaction (PCR) and lateral flow tests are included, without differentiation between symptomatic and asymptomatic infection.
+Both polymerase chain reaction (PCR) and lateral flow test results are included, without differentiation between symptomatic and asymptomatic infection.
 COVID-19 hospital admissions were identified using HES in-patient hospital records with U07.1 or U07.2 reason for admission ICD-10 codes.
-COVID-19 deaths with the same COVID-19 ICD-10 codes mentioned anywhere on the death certificate (i.e., as an underlying or contributing cause of death) were included.
+Deaths were classified as from COVID-19 if deaths with the same ICD-10 codes were mentioned anywhere on the death certificate (i.e., as an underlying or contributing cause of death).
 
@@
 
-Each matched pair was followed from the day of recruitment (i.e., time zero) until the earliest of death, GP practice de-registration, booster receipt by the control, 10 weeks, or `r format(as.Date(study_dates$studyend_date), "%e %B %Y")`.
+Each person was followed from the day of recruitment (i.e., time zero) until the earliest of outcome, death, practice de-registration, booster receipt by the control in the matching pair, 10 weeks, or `r format(as.Date(study_dates$studyend_date), "%e %B %Y")`.
 
@@
 
-The following variables were included as covariates in Cox models (see below): sex (male or female); English Index of Multiple Deprivation (IMD, grouped by quintile); ethnicity (Black, Mixed, South Asian, White, Other, as per the UK census); morbidity count (diabetes, BMI \> 40kg/m^2^, chronic heart disease, chronic kidney disease, chronic liver disease, chronic respiratory disease or severe asthma, chronic neurological disease); learning disabilities; severe mental illness; the number of SARS-CoV-2 tests in the 6 months prior to the study start date; elective in-hospital episode at the time of recruitment.
-
 The supplementary materials provide further details of the codelists and data sources used for all variables in the study.
@@
 
-We estimated Kaplan-Meier cumulative incidence curves separately in the boosted and unboosted groups.
-Rates of competing events (due to all-cause or cause-specific death) were low and so were not explicitly accounted for.
-
-We used Cox models, stratified by recruitment day, NHS region, and brand of second vaccine dose, to estimate adjusted hazard ratios comparing boosted with unboosted people, overall and separately for days 1-28 and 29-70.
-A robust variance estimator was used to account for participant clustering.
+We estimated Kaplan-Meier cumulative incidence curves separately in the boosted and unboosted groups, and the risk difference and risk ratio at `r label_number(accuracy=1)(max(postbaselinecuts)/7)` weeks.
+
+We used Cox models, stratified by recruitment day, NHS region, and brand of second vaccine dose, to estimate hazard ratios (HRs) comparing boosted with unboosted people, overall and separately for days 1-28 and 29-70.
+As we were not able to match on all potential confounders, these models included the following additional covariates: sex (male or female); English Index of Multiple Deprivation (IMD, grouped by quintile); ethnicity (Black, Mixed, South Asian, White, Other, as per the UK census); prior morbidity count (diabetes, BMI \> 40kg/m^2^, chronic heart disease, chronic kidney disease, chronic liver disease, chronic respiratory disease or severe asthma, chronic neurological disease); learning disabilities; severe mental illness; the number of SARS-CoV-2 tests in the 6 months prior to the study start date; elective in-hospital episode at the time of recruitment.A robust variance estimator was used to account for inclusion of some people in both the boosted and unboosted groups.
+Booster vaccine effectiveness was calculated as one minus the hazard ratio, expressed as a percentage.
 
@@
 
-We also estimated booster effectiveness separately in the following subgroups: primary vaccine course; age (18-64 or 65 and over); clinically extremely vulnerable or not.
-
-Booster vaccine effectiveness was calculated as one minus the hazard ratio, expressed as a percentage.
+We also estimated booster effectiveness separately in the following subgroups: primary vaccine course; age (18-64 or $\geq$ 65 years); clinically extremely vulnerable or not.
 
@@
 All code is shared openly for review and re-sue under MIT open license at <https://github.com/opensafely/booster-effectiveness>.
-Codelists are available at <https://www.opencodelists.org/>.
+The supplementary materials provide further details of the codelists and data sources used for all variables in the study.
 Detailed pseudonymised patient data is potentially re-identifiable and therefore not shared.
@@
 
-`r label_number(1, big.mark=",")(filter(matchflowchart, crit=="c0", treatment=="pfizer") %>% pull(n))` adults registered at a TPP pracitce received a BNT162b2 booster vaccination during the recruitment period, with `r label_number(1, big.mark=",")(filter(matchflowchart, treatment=="pfizer", crit=="c7") %>% pull(n))` (`r label_percent(0.1)(filter(matchflowchart, treatment=="pfizer", crit=="c7") %>% pull(pct_all))`) eligible for inclusion into the treatment group, of whom `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` (`r label_percent(0.1)(filter(matchsummary, treatment=="pfizer") %>% pull(prop_matched_eligible))`) were matched with unboosted controls (Figure 1).
+`r label_number(1, big.mark=",")(filter(matchflowchart, crit=="c0", treatment=="pfizer") %>% pull(n))` adults registered at a TPP practice received a BNT162b2 booster vaccination during the recruitment period, with `r label_number(1, big.mark=",")(filter(matchflowchart, treatment=="pfizer", crit=="c7") %>% pull(n))` (`r label_percent(0.1)(filter(matchflowchart, treatment=="pfizer", crit=="c7") %>% pull(pct_all))`) eligible for inclusion in the treatment group, of whom `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` (`r label_percent(0.1)(filter(matchsummary, treatment=="pfizer") %>% pull(prop_matched_eligible))`) were matched with unboosted controls (Figure 1).
 
@@
 
-As expected, the matching factors has similar distributions in the boosted and unboosted groups at the start of follow up (Table 1), and the distribution of comorbidities was generally similar between the groups.
+As expected, the matching factors has similar distributions in the boosted and unboosted groups at the start of follow up (Table 1), and the proportion of people with prior comorbidities was generally similar between the groups.
 However, unboosted controls had higher levels of deprivation, higher rates of learning disabilities and severe mental illness, and lower SARS-CoV-2 testing frequency, though the standardised mean difference was consistently below 0.1 (Supplementary Figure S2).
@@
 
-There were `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="postest", fup_period=="All") %>% pull(events))` positive SARS-CoV-2 tests, `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="covidadmitted", fup_period=="All") %>% pull(events))` COVID-19 hospital admissions, and `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="coviddeath", fup_period=="All") %>% pull(events))` COVID-19 deaths and `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="noncoviddeath", fup_period=="All") %>% pull(events))` non-COVID-19 deaths across `r label_number(1, big.mark=",")((filter(matchsummary, treatment=="pfizer") %>% pull(fup_years)) *365.25/7)` person-weeks of follow-up (Table 2).
-
-The cumulative incidence of positive SARS-CoV-2 tests in boosted people appears to taper substantially around 7 days after boosting (Figure 2), something not observed in unboosted controls (Figure 2).
-The marginal risk difference at 70 days for boosted versus unboosted was `r riskdiff %>% filter(treatment=="pfizer", outcome=="postest", time==70, subgroup=="none") %>% pull(survdiff_inv_display)` per 1000 people.
-For more severe outcomes the cumulative incidence between treatment groups also diverged substantially.
-At 70 days, there were  `r (riskdiff %>% filter(treatment=="pfizer", outcome=="covidadmitted", time==70, subgroup=="none") %>% pull(survdiff_display))` fewer COVID-19 admissions, `r (riskdiff %>% filter(treatment=="pfizer", outcome=="coviddeath", time==70, subgroup=="none") %>% pull(survdiff_display))` fewer COVID-19 deaths, and  `r (riskdiff %>% filter(treatment=="pfizer", outcome=="covidadmitted", time==70, subgroup=="none") %>% pull(survdiff_display))` fewer non-COVID-19 deaths per 1000 people in the boosted group compared with the unboosted group.
-
-The overall estimated effectiveness (95% CI) comparing a BNT162b2 booster dose to two doses only was `r filter(overallhazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for non-COVID-19 death.
-Estimated effectiveness of BNT162b2 boosting in days 1-28 and days 29-70 was `r filter(meta2hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` and `r filter(meta2hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(meta2hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` and `r filter(meta2hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(meta2hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` and `r filter(meta2hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` for COVID-19 death, and `r filter(meta2hazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` and `r filter(meta2hazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` for non-COVID-19 death.
+There were `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="postest", fup_period=="All") %>% pull(events))` positive SARS-CoV-2 tests, 
+`r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="covidadmitted", fup_period=="All") %>% pull(events))` COVID-19 hospital admissions, 
+`r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="coviddeath", fup_period=="All") %>% pull(events))` COVID-19 deaths, 
+and `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="noncoviddeath", fup_period=="All") %>% pull(events))` non-COVID-19 deaths 
+across `r label_number(1, big.mark=",")((filter(matchsummary, treatment=="pfizer") %>% pull(fup_years)) *365.25/7)` person-weeks of follow-up. 
+The 10-week risk of positive SARS-CoV-2 test was `r km70 %>% filter(treatment=="pfizer", outcome=="postest", subgroup=="none") %>% pull(risk_1) %>% (label_number(0.1, scale=1000))` per 1000 in the boosted group
+and `r km70 %>% filter(treatment=="pfizer", outcome=="postest", subgroup=="none") %>% pull(risk_0) %>% (label_number(0.1, scale=1000))` in the unboosted group, 
+representing a risk difference of `r km70 %>% filter(treatment=="pfizer", outcome=="postest", subgroup=="none") %>% pull(kmrd) %>% (label_number(0.1, scale=1000))` per 1000. 
+For COVID-19 hospital admission, this was `r km70 %>% filter(treatment=="pfizer", outcome=="covidadmitted", subgroup=="none") %>% pull(risk_1) %>% (label_number(0.1, scale=1000))` 
+and `r km70 %>% filter(treatment=="pfizer", outcome=="covidadmitted", subgroup=="none") %>% pull(risk_0) %>% (label_number(0.1, scale=1000))` 
+respectively (`r km70 %>% filter(treatment=="pfizer", outcome=="covidadmitted", subgroup=="none") %>% pull(kmrd) %>% (label_number(0.1, scale=-1000))` fewer events). 
+For COVID-19 death it was `r km70 %>% filter(treatment=="pfizer", outcome=="coviddeath", subgroup=="none") %>% pull(risk_1) %>% (label_number(0.1, scale=1000))` 
+and `r km70 %>% filter(treatment=="pfizer", outcome=="coviddeath", subgroup=="none") %>% pull(risk_0) %>% (label_number(0.1, scale=1000))` 
+(`r km70 %>% filter(treatment=="pfizer", outcome=="coviddeath", subgroup=="none") %>% pull(kmrd) %>% (label_number(0.1, scale=-1000))` fewer events), 
+and for non-COVID-19 death it was `r km70 %>% filter(treatment=="pfizer", outcome=="noncoviddeath", subgroup=="none") %>% pull(risk_1) %>% (label_number(0.1, scale=1000))` 
+and `r km70 %>% filter(treatment=="pfizer", outcome=="noncoviddeath", subgroup=="none") %>% pull(risk_0) %>% (label_number(0.1, scale=1000))` 
+(`r km70 %>% filter(treatment=="pfizer", outcome=="noncoviddeath", subgroup=="none") %>% pull(kmrd) %>% (label_number(0.1, scale=-1000))` fewer events) respectively  (Table 2).
+
+The rate of increase in the cumulative incidence of positive SARS-CoV-2 tests in boosted people tapered substantially around 7 days after boosting, something not observed in unboosted controls (Figure 2).
+The cumulative incidence of COVID-19 hospital admission and COVID-19 death also diverged substantially between the boosted and unboosted groups.
+We also see substantial differences for non-COVID-19 deaths between the boosted and unboosted.
+
+The overall (days 1-70) estimated booster effectiveness (95% CI) comparing the boosted and unboosted gorups was `r filter(overallhazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3) %>% pull(ve_display95)` for positive SARS-CoV-2 test, `r filter(overallhazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for non-COVID-19 death.
+Estimated booster effectiveness against COVID-19 outcomes was generally lower during days 1-28 than days 29-70: `r filter(meta2hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` and `r filter(meta2hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` for positive SARS-CoV-2 test; `r filter(meta2hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` and `r filter(meta2hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` respectively for COVID-19 hospital admission; `r filter(meta2hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` and `r filter(meta2hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` respectively for COVID-19 death.
+Estimated booster effectiveness against non-COVID19 death was similar during days 1-28 and 29-70 (`r filter(meta2hazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` and `r filter(meta2hazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)`).
 
@@
 Estimated booster effectiveness was similar between subgroups defined by primary vaccine course.
-For those receiving 2× BNT162b2, estimated effectiveness was `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3) %>% pull(ve_display)` for non-COVID-19 death.
-For 2× ChAdOx1-S recipients this was `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup_level=="az-az", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup_level=="az-az", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup_level=="az-az", model==3) %>% pull(ve_display)`, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup_level=="az-az", model==3) %>% pull(ve_display)` respectively.
-
-For people aged 65 years and over, estimated effectiveness was `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for non-COVID-19 death.
-Estimated effectiveness was slightly lower in people aged under 65 years, at `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` respectively.
-
-For clinically extremely vulnerable people estimated effectiveness was `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for non-COVID-19 death.
-For people who were not clinically extremely vulnerable effectiveness was slightly higher, at `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="cev-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="cev-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="cev-FALSE", model==3) %>% pull(ve_display)`, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="cev-FALSE", model==3) %>% pull(ve_display)` respectively.
+For those who had received BNT162b2, estimated booster effectiveness was `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3) %>% pull(ve_display)` for non-COVID-19 death.
+Corresponding estimates of booster effectiveness for ChAdOx1-S recipients were `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup_level=="az-az", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup_level=="az-az", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup_level=="az-az", model==3) %>% pull(ve_display)`, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup_level=="az-az", model==3) %>% pull(ve_display)` respectively.
+
+Amongst people classified as clinically extremely vulnerable estimated booster effectiveness was `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for COVID-19 death.
+For people who not classified as clinically extremely vulnerable estimated booster effectiveness was slightly higher, at `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="cev-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="cev-FALSE", model==3) %>% pull(ve_display)`, and `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="cev-FALSE", model==3) %>% pull(ve_display)`, respectively.
+Estimated booster effectiveness for non-COVID-19 death was similar for those who were and were not classified as clinically extremely vulnerable: `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="cev-FALSE", model==3) %>% pull(ve_display)` respectively.
+
+Amongst people aged $\geq$ 65 years estimated booster effectiveness was `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for non-COVID-19 death.
+Corresponding effectiveness was slightly lower in people aged 18-64 years: `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` respectively.
 
@@
 ```{r table-incidence, echo=FALSE, warning=FALSE, message=FALSE}
-
-km70 <- km %>% 
-  filter(leadtime==max(postbaselinecuts)) %>%
-  mutate(
-    treated=(treated_descr=="Boosted")*1L
-  ) %>%
-  select(
-    treatment, outcome , subgroup, 
-    treated, surv
-  ) %>%
-  pivot_wider(
-    id_cols= c("treatment", "outcome", "subgroup"),
-    names_from=treated,
-    names_prefix="surv_",
-    values_from=surv
-  ) %>%
-  mutate(
-    risk_1 = 1 - surv_1,
-    risk_0 = 1 - surv_0,
-
-    kmrr = risk_1 / risk_0,
-    kmrd = risk_1 - risk_0,
-    VErr = 1-kmrr,
-    kmrrE = label_number(accuracy=0.01, scale=1)(kmrr),
-    kmrdE = label_number(accuracy=0.01, scale=1000)(kmrd)
-  )
 

### 2022-03-16 manuscript/draft-supplement.Rmd

@@
 
-```{r, jcvi, echo=FALSE, message=FALSE, warning=FALSE}
+```{r, table-jcvi, echo=FALSE, message=FALSE, warning=FALSE}
 
@@
 
-```{r, codelists, echo=FALSE, message=FALSE, warning=FALSE}
+```{r, table-codelists, echo=FALSE, message=FALSE, warning=FALSE}
 
@@
 
-```{r, vaxdate, echo=FALSE, message=FALSE, warning=FALSE, out.width='90%', out.height='50%', results='asis', out.extra=''}
+```{r, figure-vaxdate, echo=FALSE, message=FALSE, warning=FALSE, out.width='90%', out.height='50%', results='asis', out.extra=''}
 #read_rds(fs::path(output_dir_os, "descriptive", "vaxdate", "plot_vaxdate_stack.rds"))
@@
 
-```{r plot_smd, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}
+```{r figure-smd, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}
 
@@
 
-```{r plot_effects, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}
+```{r figure-effects-alladjustment, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}
 
@@
       treatment %in% c("pfizer")
+    ) %>%
+    mutate(
+      model_descr = fct_inorder(model_descr),
     ) %>%
@@
 
-      legend.position = "none"
+      legend.position = "bottom",
     ) +

### 2022-03-16 analysis/model_seqtrialcox.R
remove models 1 and 2

@@ -287,8 +287,8 @@ formula3_pw <- formula_vaxonly %>% update(formula_strata) %>% update(formula_dem
 
 model_descr = c(
   "Unadjusted" = "0",
-  "region- and trial-stratified" = "1",
-  "Demographic adjustment" = "2",
+  #"region- and trial-stratified" = "1",
+  #"Demographic adjustment" = "2",
   "Full adjustment" = "3"
 )
 
@@ -453,16 +453,16 @@ cox_model <- function(timesplit, number, formula_cox){
 }
 
 summary0 <- cox_model("", 0, formula0_pw)
-summary1 <- cox_model("", 1, formula1_pw)
-summary2 <- cox_model("", 2, formula2_pw)
+#summary1 <- cox_model("", 1, formula1_pw)
+#summary2 <- cox_model("", 2, formula2_pw)
 summary3 <- cox_model("", 3, formula3_pw)
 
 # combine results
 model_glance <-
   bind_rows(
     summary0$glance,
-    summary1$glance,
-    summary2$glance,
+    #summary1$glance,
+    #summary2$glance,
     summary3$glance,
   ) %>%
   mutate(
@@ -473,8 +473,8 @@ write_csv(model_glance, fs::path(output_dir, "model_glance.csv"))
 model_tidy <-
   bind_rows(
     summary0$tidy,
-    summary1$tidy,
-    summary2$tidy,
+    #summary1$tidy,
+    #summary2$tidy,
     summary3$tidy,
   ) %>%
   mutate(
@@ -496,16 +496,16 @@ formula3_overall <- formula_vaxonly_overall %>% update(formula_strata) %>% updat
 
 
 summary0overall <- cox_model("overall", 0, formula0_overall)
-summary1overall <- cox_model("overall", 1, formula1_overall)
-summary2overall <- cox_model("overall", 2, formula2_overall)
+#summary1overall <- cox_model("overall", 1, formula1_overall)
+#summary2overall <- cox_model("overall", 2, formula2_overall)
 summary3overall <- cox_model("overall", 3, formula3_overall)
 
 # combine results
 model_overallglance <-
   bind_rows(
     summary0overall$glance,
-    summary1overall$glance,
-    summary2overall$glance,
+    #summary1overall$glance,
+    #summary2overall$glance,
     summary3overall$glance,
   ) %>%
   mutate(
[…]

### 2022-03-16 analysis/report_seqtrialcox.R
add prior infection subgroup actions

@@ -22,7 +22,7 @@ if(length(args)==0){
   removeobs <- FALSE
   treatment <- "pfizer"
   outcome <- "postest"
-  subgroup <- "none"
+  subgroup <- "prior_covid_infection-TRUE"
 } else {
   removeobs <- TRUE
   treatment <- args[[1]]

### 2022-03-16 create-project.R

@@ -622,3 +622,4 @@ names(actions_list) %>% tibble(action=.) %>%
 } else {
   stop("Backend not recognised")
 }
+

### 2022-03-23 manuscript/draft-manuscript.Rmd
updates

@@
 ---
-title: "Effectiveness of BNT162b2 booster doses in England: a observational study in OpenSAFELY-TPP"
+title: "Effectiveness of BNT162b2 booster doses in England: an observational study in OpenSAFELY-TPP"
 output:
@@
     read_csv(fs::path(output_dir_os, "cev", "km.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
-    read_csv(fs::path(output_dir_os, "age65plus", "km.csv")) %>% mutate(subgroup_level=as.character(subgroup_level))
+    read_csv(fs::path(output_dir_os, "age65plus", "km.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
+    read_csv(fs::path(output_dir_os, "prior_covid_infection", "km.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
   ) %>%
@@
 
-km70 <- km %>% 
-  filter(leadtime==max(postbaselinecuts)) %>%
+kmdiff <- km %>% 
   mutate(
@@
     treatment, outcome , subgroup, 
-    treated, surv
+    treated, surv, time, leadtime
   ) %>%
   pivot_wider(
-    id_cols= c("treatment", "outcome", "subgroup"),
+    id_cols= c("treatment", "outcome", "subgroup", "time", "leadtime"),
     names_from=treated,
@@
 
+km70 <- kmdiff %>% filter(leadtime==max(postbaselinecuts))
 # 
@@
     read_csv(fs::path(output_dir_os, "age65plus", "effects.csv"))  %>% mutate(subgroup_level=as.character(subgroup_level)),
+    read_csv(fs::path(output_dir_os, "prior_covid_infection", "effects.csv"))  %>% mutate(subgroup_level=as.character(subgroup_level)),
   ) %>%
@@
     read_csv(fs::path(output_dir_os, "age65plus", "overalleffects.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
+    read_csv(fs::path(output_dir_os, "prior_covid_infection", "overalleffects.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
   ) %>%
@@
     ve_display = paste0(label_percent(accuracy=0.1)(ve), " (", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
-    ve_display95 = paste0(label_percent(accuracy=0.1)(ve), " (95% CI", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
+    ve_display95 = paste0(label_percent(accuracy=0.1)(ve), " (95%CI ", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
   )
@@
 We matched `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` BNT162b2 booster recipients to unboosted controls.
-During the first 28 days after the booster, estimated booster effectiveness (95% CI) comparing a BNT162b2 booster dose to two doses only was `r filter(overallhazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for non-COVID-19 death.
+During the first 28 days after the booster, estimated booster effectiveness comparing a BNT162b2 booster dose to two doses only was `r filter(overallhazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3) %>% pull(ve_display95)` for COVID-19 hospital admission, `r filter(overallhazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for non-COVID-19 death.
 
@@
 Data management and analyses were conducted in Python version 3.8.10 and R version 4.0.2.
-All code is shared openly for review and re-sue under MIT open license at <https://github.com/opensafely/booster-effectiveness>.
+All code is shared openly for review and re-use under MIT open license at <https://github.com/opensafely/booster-effectiveness>.
 The supplementary materials provide further details of the codelists and data sources used for all variables in the study.
@@
 and `r km70 %>% filter(treatment=="pfizer", outcome=="covidadmitted", subgroup=="none") %>% pull(risk_0) %>% (label_number(0.1, scale=1000))` 
-respectively (`r km70 %>% filter(treatment=="pfizer", outcome=="covidadmitted", subgroup=="none") %>% pull(kmrd) %>% (label_number(0.1, scale=-1000))` fewer events). 
+respectively (`r km70 %>% filter(treatment=="pfizer", outcome=="covidadmitted", subgroup=="none") %>% pull(kmrd) %>% (label_number(0.1, scale=-1000))` fewer events in the boosted group). 
 For COVID-19 death it was `r km70 %>% filter(treatment=="pfizer", outcome=="coviddeath", subgroup=="none") %>% pull(risk_1) %>% (label_number(0.1, scale=1000))` 
@@
 
-The overall (days 1-70) estimated booster effectiveness (95% CI) comparing the boosted and unboosted gorups was `r filter(overallhazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3) %>% pull(ve_display95)` for positive SARS-CoV-2 test, `r filter(overallhazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for non-COVID-19 death.
+The overall (days 1-70) estimated booster effectiveness comparing the boosted and unboosted groups was `r filter(overallhazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3) %>% pull(ve_display95)` for positive SARS-CoV-2 test, `r filter(overallhazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for non-COVID-19 death.
 Estimated booster effectiveness against COVID-19 outcomes was generally lower during days 1-28 than days 29-70: `r filter(meta2hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` and `r filter(meta2hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` for positive SARS-CoV-2 test; `r filter(meta2hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` and `r filter(meta2hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` respectively for COVID-19 hospital admission; `r filter(meta2hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` and `r filter(meta2hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` respectively for COVID-19 death.
@@
 Corresponding effectiveness was slightly lower in people aged 18-64 years: `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` respectively.
+
+Amongst people with evidence of prior SARS-CoV-2 infection, estimated booster effectiveness was `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="prior_covid_infection-TRUE", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="prior_covid_infection-TRUE", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="prior_covid_infection-TRUE", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="prior_covid_infection-TRUE", model==3) %>% pull(ve_display)` for non-COVID-19 death.
+Corresponding effectiveness was ...
+in people with no evidence of prior infection: `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="prior_covid_infection-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="prior_covid_infection-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="prior_covid_infection-FALSE", model==3) %>% pull(ve_display)`, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="prior_covid_infection-FALSE", model==3) %>% pull(ve_display)` respectively.
 
@@
 
-#### Table 1: Participant characteristics
-
-Participant characteristics as on the day of recruitment into the treatment or control group.
+#### Table 1: Characteristics (%) of `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` boosted people included in the study, and their matched controls, on the day of recruitment.
 
@@
 
-#### Table 2: Follow-up and outcomes
-
-Total follow-up and incidence rates for each outcome, by treatment group. The cumulative incidence of each outcome is provided in Figure 2.
+#### Table 2: Total events and person-years of follow-up, and incidence rates, for each outcome, in the boosted and unboosted groups.
 
@@
 incidences %>%
-  left_join(km70) %>%
+  left_join(kmdiff %>% filter(time = postbaselinecuts[-1] -1)) %>%
   filter(
@@
 
-#### Table 3: Estimated booster effectiveness
-
-Vaccine effectiveness, $(1-\text{hazard ratio})\times100\%$, for the main and subgroup analyses across all outcomes.
+#### Table 3: Estimated booster vaccine effectiveness
+
+Effectiveness $(100 \times (1-\text{hazard ratio}))$ for the main and subgroup analyses and for all outcomes.
 
@@
   plot_hazards <-
-    meta2hazards %>%
+    hazards %>%
     filter(

### 2022-03-23 manuscript/reports.Rmd
updates

@@
 matchcoverage <- read_csv(fs::path(output_dir_os, "matchcoverage.csv"))
-incidences <- read_csv(fs::path(output_dir_os, "none", "incidence_all.csv")) %>% mutate(events=events_1+events_0)
-
-studydurationweeks <- as.numeric(as.duration(study_dates$studyend_date - study_dates$studystart_date +1) , "weeks")
+incidences <- read_csv(fs::path(output_dir_os, "none", "incidence.csv")) %>% mutate(events=events_1+events_0)
+
 ```
@@
     read_csv(fs::path(output_dir_os, "cev", "km.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
-    read_csv(fs::path(output_dir_os, "age65plus", "km.csv")) %>% mutate(subgroup_level=as.character(subgroup_level))
+    read_csv(fs::path(output_dir_os, "age65plus", "km.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
+    read_csv(fs::path(output_dir_os, "prior_covid_infection", "km.csv")) %>% mutate(subgroup_level=as.character(subgroup_level))
   ) %>%
@@
 
-#### Table 1: Participant characteristics
+#### Table 1a: Participant characteristics for Pfizer boosting
 
@@
 
+#### Table 1b: Participant characteristics for Moderna boosting
+
+Participant characteristics as on the day of recruitment into the treatment or control group.
+
+```{r table1, echo=FALSE, warning=FALSE, message=FALSE}
+table1_body <- read_csv(fs::path(output_dir_os, "matchtable1.csv")) %>% filter(treatment=="moderna")
+table1_by <- read_csv(fs::path(output_dir_os, "matchtable1by.csv")) %>% filter(treatment=="moderna")
+
+table1 <- 
+  table1_body %>%
+  #filter(outcome %in% outcomes) %>%
+  group_by(variable) %>%
+  mutate(
+    label = if_else(row_number()==1, label, str_c("&nbsp;&nbsp;&nbsp;", label)),
+  ) %>%
+  mutate(
+    across(
+      starts_with("stat_"),
+      ~if_else(is.na(.x) & row_number()==1, "", .x)
+    )
+  ) %>%
+  ungroup() %>%
+  mutate(
+    order = cumsum(lag(variable, 1, first(variable))!=variable)
+  ) %>%
+  select(
+    variable, var_label, label, starts_with("stat_"), order
+  ) %>%
+  rename(!!!set_names(table1_by$by_col, table1_by$by_chr))
+
+table1 %>%
+  gt(groupname_col="treatment_descr") %>%
+  fmt_markdown(columns = "label") %>%
+  tab_style(
+    style = list(
+      cell_fill(color = "gray96")
+    ),
+    locations = cells_body(
+      rows = order%%2 == 0
+    )
+  ) %>%
+  tab_style(
+    style = list(
+      cell_text(weight = "bold")
+    ),
+    locations = cells_column_labels(everything())
+  ) %>%
+  cols_label(
+     label=md("Characteristic")
+  ) %>%
+  cols_hide(c("variable", "var_label", "order"))
+```
+
 \newpage
 
-#### Table 2: Follow-up and outcomes
-
-Total follow-up and incidence rates for each outcome, by treatment group. The cumulative incidence of each outcome is provided in Supplementary Figure S2.
+#### Table 2a: Follow-up and outcomes for Pfizer boosting
+
+Total follow-up and incidence rates for each outcome, by treatment group.
 
@@
 
-## Table 3a: Estimated booster effectiveness (5 period)
+#### Table 2b: Follow-up and outcomes for Moderna boosting
+
+Total follow-up and incidence rates for each outcome, by treatment group.
+
+```{r table2, echo=FALSE, warning=FALSE, message=FALSE}
+incidences %>%
+  filter(
+    fup_period == "All",
+    outcome %in% c("postest", "covidadmitted", "coviddeath", "noncoviddeath"),
+    treatment == "moderna"
+  ) %>%
+  left_join(
+    matchsummary %>% select(treatment, n_eligible, prop_matched_eligible),
+    by=c("treatment")
+  ) %>%
+  mutate(
+    treatment_descr = paste0(
+      #treatment_descr, " - ",
+      label_number(1, big.mark=",")(n_eligible), " eligible, ",
+      label_number(1, big.mark=",")(n_1), " matched (",
+      label_percent(0.1)(prop_matched_eligible), ")."
+    ),
+  ) %>%
+  select(
+    treatment_descr, 
+    outcome_descr,
+    q_1,
+    rate_1,
+    q_0, 
+    rate_0,
+    rrECI
+  ) %>%
+  mutate(
+    rate_1=if_else(rate_1<0.001, "<0.001", label_number(0.001)(rate_1)),
+    rate_0=if_else(rate_0<0.001, "<0.001", label_number(0.001)(rate_0)),
+  ) %>%
+  gt(
+   groupname_col = c("treatment_descr"),
+  ) %>%
+  cols_label(
+    #treatment_descr = "Treatment",
+    outcome_descr = "Outcome",
+
+    q_0 = "Events / person-years",
+    q_1   = "Events / person-years",
+
+    rate_0 = "Incidence rate",
+    rate_1 = "Incidence rate",
+    rrECI = "Incidence rate ratio (95% CI)"
+  ) %>%
+  tab_spanner(
+    label = "Boosted",
+    columns = ends_with("1")
+  ) %>%
+  tab_spanner(
+    label = "Unboosted",
+    columns = ends_with("0")
+  ) %>%
+  fmt_missing(
+    everything(),
+    missing_text="--"
+  ) %>%
+  cols_align(
+    align = "right",
+    columns = everything()
+  )
+
+```
+\newpage
+
+## Table 3a: Estimated Pfizer booster effectiveness (5 period)
 
@@
 
-hazards %>%
+bind_rows(
+  hazards,
+  overallhazards %>% mutate(term_left = 0, term_right=70)
+) %>%
   filter(model==3, treatment=="pfizer") %>%
   mutate(
-    period = paste(term_left+1, "-", term_right),
+    period = fct_inorder(paste(term_left+1, "-", term_right)),
     ve_display = str_remove(ve_display, fixed("%"))
   ) %>%
-  arrange(treatment_descr, outcome_descr, subgroup_level_descr, term_left) %>%
+  arrange(treatment_descr, outcome_descr, subgroup_level_descr, period) %>%
   group_by(outcome, treatment, subgroup_level_descr, model) %>%
   mutate(
-    subgroup_level_descr=if_else(row_number()==1, subgroup_level_descr, factor(NA, levels=levels(subgroup_level_descr)))
+    #subgroup_level_descr=if_else(row_number()==1, subgroup_level_descr, factor(NA, levels=levels(subgroup_level_descr))),
+    period = paste0(period, " days")
   ) %>%
   ungroup() %>%
-  select(outcome_descr, subgroup_level_descr, period, hr_display, ve_display) %>%
+  select(outcome_descr, subgroup_level_descr, period, ve_display) %>%
+  pivot_wider(
+    id_cols = c("outcome_descr", "subgroup_level_descr"),
+    names_from = "period",
+    #names_prefix ="period",
+    values_from = "ve_display"
+  ) %>%
   gt(
@@
     outcome_descr = "Outcome",
-
-    period = "Days since booster",
-    hr_display   = "HR (95% CI)",
-    ve_display   = "VE (95% CI)",
   ) %>%
@@
   ) %>%
+  tab_spanner(
+    label = "Booster vaccine effectiveness (95% CI)",
+    columns = ends_with("days")
+  ) %>%
   fmt_missing(
@@
 
-## Table 3b: Estimated booster effectiveness (2 period split at 28 days)
-
-```{r table_effects28, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}
-
-meta28hazards %>%
-  filter(model==3, treatment=="pfizer") %>%
-  mutate(
-    period = paste(term_left+1, "-", term_right),
+## Table 3a: Estimated Moderna booster effectiveness (5 period)
+
+Main and subgroup analyses
+```{r table_effects, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}
+
+bind_rows(
+  hazards,
+  overallhazards %>% mutate(term_left = 0, term_right=70)
+) %>%
+  filter(model==3, treatment=="moderna") %>%
+  mutate(
+    period = fct_inorder(paste(term_left+1, "-", term_right)),
     ve_display = str_remove(ve_display, fixed("%"))
   ) %>%
-  arrange(treatment_descr, outcome_descr, subgroup_level_descr, term_left) %>%
+  arrange(treatment_descr, outcome_descr, subgroup_level_descr, period) %>%
   group_by(outcome, treatment, subgroup_level_descr, model) %>%
   mutate(
-    subgroup_level_descr=if_else(row_number()==1, subgroup_level_descr, factor(NA, levels=levels(subgroup_level_descr)))
+    #subgroup_level_descr=if_else(row_number()==1, subgroup_level_descr, factor(NA, levels=levels(subgroup_level_descr))),
+    period = paste0(period, " days")
   ) %>%
   ungroup() %>%
-  select(outcome_descr, subgroup_level_descr, period, hr_display, ve_display) %>%
+  select(outcome_descr, subgroup_level_descr, period, ve_display) %>%
+  pivot_wider(
+    id_cols = c("outcome_descr", "subgroup_level_descr"),
+    names_from = "period",
+    #names_prefix ="period",
+    values_from = "ve_display"
+  ) %>%
   gt(
@@
 
-## Table 3c: Estimated booster effectiveness (2 period split at 42 days)
-
-```{r table_effects42, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}
-
-meta28hazards %>%
-  filter(model==3, treatment=="pfizer") %>%
-  mutate(
-    period = paste(term_left+1, "-", term_right),
-    ve_display = str_remove(ve_display, fixed("%"))
-  ) %>%
-  arrange(treatment_descr, outcome_descr, subgroup_level_descr, term_left) %>%
-  group_by(outcome, treatment, subgroup_level_descr, model) %>%
-  mutate(
-    subgroup_level_descr=if_else(row_number()==1, subgroup_level_descr, factor(NA, levels=levels(subgroup_level_descr)))
-  ) %>%
-  ungroup() %>%
-  select(outcome_descr, subgroup_level_descr, period, hr_display, ve_display) %>%
-  gt(
-    groupname_col = c("outcome_descr"),
-  ) %>%
-  cols_label(
-    #treatment_descr = "Treatment",
-    subgroup_level_descr="",
-    outcome_descr = "Outcome",
-
-    period = "Days since booster",
-    hr_display   = "HR (95% CI)",
-    ve_display   = "VE (95% CI)",
-  ) %>%
-  cols_align(
-    align = "right",
-    columns = everything()
-  ) %>%
-  fmt_missing(
-    c("subgroup_level_descr"),
-    missing_text=""
-  )
-
-```
-
-\newpage
-
 #### Figure 1: Booster vaccination uptake, treatment group eligibility and matching success
@@
   ) %>%
-  filter(treatment=="pfizer") %>%
+  #filter(treatment=="pfizer") %>%
   mutate(

### 2022-03-28 manuscript/draft-manuscript.Rmd
v3 updates

@@
 matchcoverage <- read_csv(fs::path(output_dir_os, "matchcoverage.csv"))
-incidences <- read_csv(fs::path(output_dir_os, "none", "incidence_all.csv")) %>% mutate(events=events_1+events_0)
+incidence_all <- read_csv(fs::path(output_dir_os, "none", "incidence_all.csv")) %>% mutate(events=events_1+events_0)
 pfizerdurationweeks <- as.numeric(as.duration(study_dates$pfizerend_date - study_dates$pfizerstart_date +1) , "weeks")
@@
     read_csv(fs::path(output_dir_os, "age65plus", "km.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
-    read_csv(fs::path(output_dir_os, "prior_covid_infection", "km.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
+    #read_csv(fs::path(output_dir_os, "prior_covid_infection", "km.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
   ) %>%
@@
     #model_name = fct_inorder(str_wrap(model_name, 15)),
-  )
-
-kmdiff <- km %>% 
-  mutate(
-    treated=(treated_descr=="Boosted")*1L
+    treated = (treated_descr=="Boosted")*1L
+  ) %>%
+  group_by(treatment, treated_descr, outcome, subgroup) %>%
+  mutate(
+    cml.risk = cumsum(n.risk),
+    cml.event = cumsum(replace_na(n.event, 0)),
+  )
+
+groupvariables <- c("treatment", "treatment_descr", "outcome", "outcome_descr", "outcome_descr_wrap", "subgroup", "subgroup_descr", "subgroup_level", "subgroup_level_descr")
+
+km_incidence <-
+  km %>%
+  mutate(
+    n.atrisk = n.risk,
+    cml.atrisk = cumsum(replace_na(n.atrisk, 0)),
+    cml.event = cumsum(replace_na(n.event, 0)),
+    cml.censor = cumsum(replace_na(n.censor, 0)),
+    rate = n.event/n.atrisk,
+    cml.rate = cml.event / cml.atrisk,
+    risk = 1 - surv,
   ) %>%
   select(
-    treatment, outcome , subgroup, 
-    treated, surv, time, leadtime
+    treated,
+    all_of(groupvariables),
+    time, interval,
+    surv, risk, n.atrisk, n.event, n.censor, rate, cml.atrisk, cml.event, cml.censor, cml.rate
+  )
+
+kmconstrast <- function(data, cuts=NULL){
+
+  if(is.null(cuts)){cuts <- unique(data$time)}
+
+  data %>%
+    filter(time!=0) %>%
+    mutate(
+      period_start = cut(time, cuts, right=TRUE, label=cuts[-length(cuts)]),
+      period_end = cut(time, cuts, right=TRUE, label=cuts[-1]),
+      period = cut(time, cuts, right=TRUE, label=paste0(cuts[-length(cuts)]+1, " - ", cuts[-1]))
+    ) %>%
+    group_by(across(all_of(groupvariables))) %>%
+    group_by(
+      treated, period_start, period_end, period,
+      .add = TRUE
+    ) %>%
+    summarise(
+      interval = last(time) - first(time) + 1,
+      time=last(time),
+      cml.atrisk = last(cml.atrisk),
+      cml.event = last(cml.event),
+      cml.censor = last(cml.censor),
+      cml.rate = last(cml.rate),
+      persontime = sum(n.atrisk),
+      n.atrisk = first(n.atrisk),
+      n.event = sum(n.event, na.rm=TRUE),
+      n.censor = sum(n.censor, na.rm=TRUE),
+      surv = last(surv),
+      risk = last(risk),
+      rate = n.event/persontime,
     ) %>%
     pivot_wider(
-    id_cols= c("treatment", "outcome", "subgroup", "time", "leadtime"),
+      id_cols= c(groupvariables, "period_start", "period_end", "period",  "interval", "time"),
       names_from=treated,
-    names_prefix="surv_",
-    values_from=surv
-  ) %>%
-  mutate(
-    risk_1 = 1 - surv_1,
-    risk_0 = 1 - surv_0,
-
+      names_glue="{.value}_{treated}",
+      values_from=c(surv, risk, n.atrisk, n.event, n.censor, rate, cml.atrisk, cml.event, cml.censor, cml.rate)
+    ) %>%
+    mutate(
       kmrr = risk_1 / risk_0,
       kmrd = risk_1 - risk_0,
-    VErr = 1-kmrr,
-    kmrrE = label_number(accuracy=0.01, scale=1)(kmrr),
-    kmrdE = label_number(accuracy=0.01, scale=1000)(kmrd)
-  )
-
-km70 <- kmdiff %>% filter(leadtime==max(postbaselinecuts))
+      irr = rate_1 / rate_0,
+      ird = rate_1 - rate_0,
+      cmlirr = cml.rate_1 / cml.rate_0,
+      cmlird = cml.rate_1 - cml.rate_0
+    ) %>%
+    mutate(
+      kmrd_display = label_number(accuracy=0.1, scale=1000, big.mark =",")(kmrd),
+      kmrd_inv_display = label_number(accuracy=0.1, scale=1000, big.mark =",")(-kmrd),
+    ) %>%
+    mutate(
+      cml.events = cml.event_0 + cml.events_1,
+      cml.atrisk = cml.atrisk_0 + cml.atrisk_1
+    )
+}
+
+kmtreated_daily <- kmconstrast(km_incidence)
+kmtreated_cuts <- kmconstrast(km_incidence, postbaselinecuts)
+kmtreated_overall <- kmconstrast(km_incidence, c(0,last(postbaselinecuts)))
+
+km70 <- kmtreated_daily %>% filter(time==max(postbaselinecuts))
+
 # 
-# cmlinc <- read_csv(fs::path(output_dir_os, "cmlinc.csv")) %>%
+# incidence <- 
+#   bind_rows(
+#     read_csv(fs::path(output_dir_os, "none", "incidence.csv")),
+#     read_csv(fs::path(output_dir_os, "vax12_type", "incidence.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
+#     #read_csv(fs::path(output_dir_os, "cev", "incidence.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
+#     #read_csv(fs::path(output_dir_os, "age65plus", "incidence.csv"))  %>% mutate(subgroup_level=as.character(subgroup_level)),
+#     #read_csv(fs::path(output_dir_os, "prior_covid_infection", "incidence.csv"))  %>% mutate(subgroup_level=as.character(subgroup_level)),
+#   ) %>%
+#   filter(
+#     outcome %in% outcomes
+#   ) %>%
 #   mutate(
+#     subgroup_descr = fct_inorder(subgroup_descr),
+#     subgroup_level = fct_inorder(subgroup_level),
+#     #subgroup_level_descr = fct_inorder(fct_explicit_na(subgroup_level_descr, "All")),
 #     treatment_descr = fct_inorder(treatment_descr),
 #     outcome_descr = fct_inorder(outcome_descr),
-#     model_name = fct_inorder(str_wrap(model_name, 15)),
-#     inc_CI_to = paste0(label_number(0.1, scale=1000)(1-survival.ul), " to ", label_number(0.1, scale=1000)(1-survival.ll))
-#   )
-
-# riskdiff <- read_csv(fs::path(output_dir_os, "riskdiff.csv")) %>%
-#   mutate(
-#     treatment_descr = fct_inorder(treatment_descr),
-#     outcome_descr = fct_inorder(outcome_descr),
-#     model_name = fct_inorder(str_wrap(model_name, 15)),
-#     diff_CI_to = paste0(label_number(0.01, scale=1000)(diff.ll), " to ", label_number(0.01, scale=1000)(diff.ul))
+#     outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=15)),
 #   )
@@
     read_csv(fs::path(output_dir_os, "none", "effects.csv")),
-    read_csv(fs::path(output_dir_os, "vax12_type", "effects.csv")),
+    read_csv(fs::path(output_dir_os, "vax12_type", "effects.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
     read_csv(fs::path(output_dir_os, "cev", "effects.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
@@
 
-#cmlinc3 <- cmlinc %>% filter(model==3)
 hazards3 <- hazards %>% filter(model==3)
 
-riskdiff <- km %>% 
-  group_by(treatment, treatment_descr, outcome, outcome_descr, outcome_descr_wrap, subgroup, subgroup_descr, subgroup_variable, subgroup_variable_descr, subgroup_level, subgroup_level_descr, time) %>%
-  summarise(
-    survdiff = diff(surv),
-  ) %>% 
-  mutate(
-    survdiff_display = label_number(accuracy=0.1, scale=1000, big.mark =",")(survdiff),
-    survdiff_inv_display = label_number(accuracy=0.1, scale=1000, big.mark =",")(-survdiff),
-  ) %>%
-  ungroup()
 ```
@@
 <!-- https://www.england.nhs.uk/2021/09/nhs-begins-covid-19-booster-vaccination-campaign/ -->
-
 <!-- 13 december 30-39 become eligible (https://www.england.nhs.uk/2021/12/nhs-to-roll-out-life-saving-booster-jab-to-people-aged-30-plus/) -->
-
 <!-- 15 december everyone eligible https://www.england.nhs.uk/2021/12/nhs-booster-bookings-open-to-every-eligible-adult/ -->
@@
 The supplementary materials provide further details of the matching process.
-<!-- We excluded matched pairs from days in which fewer than 10% of eligible treated people were successfully. -->
 
@@
 
-## Additional variables
-
 The supplementary materials provide further details of the codelists and data sources used for all variables in the study.
@@
 
-We used Cox models, stratified by recruitment day, NHS region, and brand of second vaccine dose, to estimate hazard ratios (HRs) comparing boosted with unboosted people, overall and separately for days 1-28 and 29-70.
+We used Cox models, stratified by recruitment day, NHS region, and brand of second vaccine dose, to estimate hazard ratios (HRs) comparing boosted with unboosted people, overall and separately for days 1-28 and 29-70, and also for days 1-7, 8-14, 15-28, 29-42, and 43-70.
 As we were not able to match on all potential confounders, these models included the following additional covariates: sex (male or female); English Index of Multiple Deprivation (IMD, grouped by quintile); ethnicity (Black, Mixed, South Asian, White, Other, as per the UK census); prior morbidity count (diabetes, BMI \> 40kg/m^2^, chronic heart disease, chronic kidney disease, chronic liver disease, chronic respiratory disease or severe asthma, chronic neurological disease); learning disabilities; severe mental illness; the number of SARS-CoV-2 tests in the 6 months prior to the study start date; elective in-hospital episode at the time of recruitment.A robust variance estimator was used to account for inclusion of some people in both the boosted and unboosted groups.
@@
 
-There were `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="postest", fup_period=="All") %>% pull(events))` positive SARS-CoV-2 tests, 
-`r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="covidadmitted", fup_period=="All") %>% pull(events))` COVID-19 hospital admissions, 
-`r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="coviddeath", fup_period=="All") %>% pull(events))` COVID-19 deaths, 
-and `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="noncoviddeath", fup_period=="All") %>% pull(events))` non-COVID-19 deaths 
-across `r label_number(1, big.mark=",")((filter(matchsummary, treatment=="pfizer") %>% pull(fup_years)) *365.25/7)` person-weeks of follow-up. 
+There were `r label_number(1, big.mark=",")(filter(km70, treatment=="pfizer", outcome=="postest") %>% pull(cml.events))` positive SARS-CoV-2 tests,
+`r label_number(1, big.mark=",")(filter(km70, treatment=="pfizer", outcome=="covidadmitted") %>% pull(cml.events))` COVID-19 hospital admissions, 
+`r label_number(1, big.mark=",")(filter(km70, treatment=="pfizer", outcome=="coviddeath") %>% pull(cml.event))` COVID-19 deaths, 
+and `r label_number(1, big.mark=",")(filter(km70, treatment=="pfizer", outcome=="noncoviddeath") %>% pull(cml.events))` non-COVID-19 deaths 
+across `r label_number(1, big.mark=",")((filter(km70, treatment=="pfizer", outcome="coviddeath") %>% pull(cml.atrisk)) /7)` person-weeks of follow-up. 
 The 10-week risk of positive SARS-CoV-2 test was `r km70 %>% filter(treatment=="pfizer", outcome=="postest", subgroup=="none") %>% pull(risk_1) %>% (label_number(0.1, scale=1000))` per 1000 in the boosted group
@@
 Amongst people classified as clinically extremely vulnerable estimated booster effectiveness was `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for COVID-19 death.
-For people who not classified as clinically extremely vulnerable estimated booster effectiveness was slightly higher, at `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="cev-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="cev-FALSE", model==3) %>% pull(ve_display)`, and `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="cev-FALSE", model==3) %>% pull(ve_display)`, respectively.
+For people who were not classified as clinically extremely vulnerable estimated booster effectiveness was slightly higher at `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="cev-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="cev-FALSE", model==3) %>% pull(ve_display)`, and `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="cev-FALSE", model==3) %>% pull(ve_display)`, respectively.
 Estimated booster effectiveness for non-COVID-19 death was similar for those who were and were not classified as clinically extremely vulnerable: `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="cev-FALSE", model==3) %>% pull(ve_display)` respectively.
@@
 
-incidences %>%
-  left_join(kmdiff %>% filter(time = postbaselinecuts[-1] -1)) %>%
+incidence_table <- 
+  incidence %>%
+  left_join(
+    kmdiff %>% filter(time == postbaselinecuts[-1] -1),
+    by = c("treatment", "outcome", "subgroup", )
+  ) %>%
   filter(
-    fup_period == "All",
+    #fup_period == "All",
     outcome %in% outcomes,
@@
     risk_0=if_else(risk_0<0.001, "<0.001", label_number(0.001)(risk_0)),
-  ) %>%
+  )
+
   gt(
+    incidence_table
    #groupname_col = c("treatment_descr"),

### 2022-03-28 manuscript/draft-manuscript.Rmd
Update draft-manuscript.Rmd

@@
 We matched `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` BNT162b2 booster recipients to unboosted controls.
-During the first 28 days after the booster, estimated booster effectiveness comparing a BNT162b2 booster dose to two doses only was `r filter(overallhazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3) %>% pull(ve_display95)` for COVID-19 hospital admission, `r filter(overallhazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for non-COVID-19 death.
+Estimated booster effectiveness comparing a BNT162b2 booster dose to two doses only was `r filter(overallhazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3) %>% pull(ve_display95)` for COVID-19 hospital admission, `r filter(overallhazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for non-COVID-19 death.
+Estimated effectiveness was similar in those receiving BNT162b2 or ChAdOx1-S two-dose primary vaccination course, but slightly lower in those classified as clinically extremely vulnerable (`r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for COVID-19 death) and in those aged 18-65 years (`r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` respectively).
+Period-specific effects were also estimated, showing some evidence of waning booster effectiveness against positive SARS-CoV-2 tests (`r filter(hazards, outcome=="postest", treatment=="pfizer", term_right=28, subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` at 15-28 days versus `r filter(hazards, outcome=="postest", treatment=="pfizer", term_right=70, subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` at 43-70 days).
 
@@
 Booster doses were initially available no earlier than six months after dose two, but this was reduced to three months on 8 December 2021, following concerns over the emergence of the Omicron variant @NHSEnglandNHSa1 @NHSEnglandNHSa2.
-Vaccine prioritisation schedules were guided by recommendations from the Joint Committee for Vaccine and Immunisation (JCVI) expert working group.
+Vaccine prioritisation schedules were guided by recommendations from the Joint Committee for Vaccine and Immunisation (JCVI) expert working group @covid19.
 
@@
 COVID-19 vaccination history is available in the GP record directly via the National Immunisation Management System (NIMS).
-Data include pseudonymized data such as coded diagnoses, medications and physiological parameters.
-No free text data are included.
 
@@
 
-We included adults aged $\geq$ 18 years who were eligible for booster vaccination between `r format(as.Date(study_dates$pfizerstart_date), "%e %B")` and `r format(as.Date(study_dates$pfizerend_date), "%e %B %Y")` inclusive (the "recruitment period").
+We included adults aged \ge 18 years who were eligible for booster vaccination between `r format(as.Date(study_dates$pfizerstart_date), "%e %B")` and `r format(as.Date(study_dates$pfizerend_date), "%e %B %Y")` inclusive (the "recruitment period").
 On each day within this period, eligibility for the study was determined as follows: registered at a GP practice using TPP's SystmOne clinical information system; received a two-dose primary vaccination course of either BNT162b2 or ChAdOx1-S (mixed dosing and Moderna mRNA-1273 were not considered due to small numbers); not a health or social care worker, not resident in a care or nursing home; not medically housebound or receiving end-of-life care; no evidence of SARS-CoV-2 infection or COVID-19 disease within the previous 90 days; not undergoing an unplanned hospital admission; complete information on sex, ethnicity, deprivation, and NHS region.
 
-Additionally, if the rolling weekly average count of BNT162b2 booster doses within strata defined by region, JCVI group, and the week of second dose was below 50 then recruitment did not occur for that stratum on that day.
-This helped ensure that recruitment was restricted to those boosted in line with national prioritisation schedules and that the matched controls were both eligible and able to receive a booster dose at the time of potential recruitment.
+Additionally, if the rolling weekly average count of BNT162b2 booster doses within strata defined by region, JCVI priority group, and the week of second dose was below 50 then recruitment did not occur for that stratum on that day.
+This helped to ensure that recruitment was restricted to those eligible and able to receive a booster dose in line with national prioritisation schedules at the time of potential recruitment.
 
@@
 People selected as controls were not eligible to be included again as a control, but were eligible for subsequent inclusion in the booster group. Any unmatched boosted people were excluded.
-
-The supplementary materials provide further details of the matching process.
 
@@
 
-The supplementary materials provide further details of the codelists and data sources used for all variables in the study.
-
 ## Statistical Analysis
 
-We estimated Kaplan-Meier cumulative incidence curves separately in the boosted and unboosted groups, and the risk difference and risk ratio at `r label_number(accuracy=1)(max(postbaselinecuts)/7)` weeks.
-
-We used Cox models, stratified by recruitment day, NHS region, and brand of second vaccine dose, to estimate hazard ratios (HRs) comparing boosted with unboosted people, overall and separately for days 1-28 and 29-70, and also for days 1-7, 8-14, 15-28, 29-42, and 43-70.
+We estimated Kaplan-Meier cumulative incidence curves separately in the boosted and unboosted groups, and the corresponding 10-week risk difference and risk ratio.
+
+We used Cox models, stratified by recruitment day, NHS region, and two-dose primary vaccine course, to estimate hazard ratios (HRs) comparing boosted with unboosted people, overall and separately for days 1-28 and 29-70, and also for days 1-7, 8-14, 15-28, 29-42, and 43-70.
 As we were not able to match on all potential confounders, these models included the following additional covariates: sex (male or female); English Index of Multiple Deprivation (IMD, grouped by quintile); ethnicity (Black, Mixed, South Asian, White, Other, as per the UK census); prior morbidity count (diabetes, BMI \> 40kg/m^2^, chronic heart disease, chronic kidney disease, chronic liver disease, chronic respiratory disease or severe asthma, chronic neurological disease); learning disabilities; severe mental illness; the number of SARS-CoV-2 tests in the 6 months prior to the study start date; elective in-hospital episode at the time of recruitment.A robust variance estimator was used to account for inclusion of some people in both the boosted and unboosted groups.
@@
 
-We also estimated booster effectiveness separately in the following subgroups: primary vaccine course; age (18-64 or $\geq$ 65 years); clinically extremely vulnerable or not.
-
-## Missing data
-
-After exclusions for missing values on demographic variables, there were no missing values in the remaining variables as they were each defined by the presence or absence of clinical codes or events in the patient record.
+We also estimated booster effectiveness separately in the following subgroups: primary vaccine course; age (18-64 or \ge 65 years); clinically extremely vulnerable or not; evidence of prior SARS-CoV-2 infection or not.
+
+<!-- ## Missing data -->
+
+<!-- After exclusions for missing values on demographic variables, there were no missing values in the remaining variables as they were each defined by the presence or absence of clinical codes or events in the patient record.
+-->
 
@@
 
-There were `r label_number(1, big.mark=",")(filter(km70, treatment=="pfizer", outcome=="postest") %>% pull(cml.events))` positive SARS-CoV-2 tests,
-`r label_number(1, big.mark=",")(filter(km70, treatment=="pfizer", outcome=="covidadmitted") %>% pull(cml.events))` COVID-19 hospital admissions, 
-`r label_number(1, big.mark=",")(filter(km70, treatment=="pfizer", outcome=="coviddeath") %>% pull(cml.event))` COVID-19 deaths, 
-and `r label_number(1, big.mark=",")(filter(km70, treatment=="pfizer", outcome=="noncoviddeath") %>% pull(cml.events))` non-COVID-19 deaths 
-across `r label_number(1, big.mark=",")((filter(km70, treatment=="pfizer", outcome="coviddeath") %>% pull(cml.atrisk)) /7)` person-weeks of follow-up. 
+There were `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="postest", fup_period=="All") %>% pull(events))` positive SARS-CoV-2 tests,
+`r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="covidadmitted", fup_period=="All") %>% pull(events))` COVID-19 hospital admissions, 
+`r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="coviddeath", fup_period=="All") %>% pull(events))` COVID-19 deaths, 
+and `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="noncovidadmitted", fup_period=="All") %>% pull(events))` non-COVID-19 deaths 
+across `r label_number(1, big.mark=",")((filter(incidences, treatment=="pfizer", outcome=="coviddeath", fup_period=="All") %>% pull(yearsatrisk)) * 365.25/7)` person-weeks of follow-up. 
 The 10-week risk of positive SARS-CoV-2 test was `r km70 %>% filter(treatment=="pfizer", outcome=="postest", subgroup=="none") %>% pull(risk_1) %>% (label_number(0.1, scale=1000))` per 1000 in the boosted group
@@
 
-The overall (days 1-70) estimated booster effectiveness comparing the boosted and unboosted groups was `r filter(overallhazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3) %>% pull(ve_display95)` for positive SARS-CoV-2 test, `r filter(overallhazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for non-COVID-19 death.
+The overall (days 1-70) estimated booster effectiveness comparing the boosted and unboosted groups was `r filter(overallhazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3) %>% pull(ve_display95)` for positive SARS-CoV-2 test, `r filter(overallhazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for non-COVID-19 death (Table 3).
 Estimated booster effectiveness against COVID-19 outcomes was generally lower during days 1-28 than days 29-70: `r filter(meta2hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` and `r filter(meta2hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` for positive SARS-CoV-2 test; `r filter(meta2hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` and `r filter(meta2hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` respectively for COVID-19 hospital admission; `r filter(meta2hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` and `r filter(meta2hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` respectively for COVID-19 death.
 Estimated booster effectiveness against non-COVID19 death was similar during days 1-28 and 29-70 (`r filter(meta2hazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` and `r filter(meta2hazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)`).
-
-Estimated effectiveness for narrower time periods are provided in supplementary materials (Table S3).
 
@@
 
-Amongst people aged $\geq$ 65 years estimated booster effectiveness was `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for non-COVID-19 death.
+Amongst people aged \ge 65 years estimated booster effectiveness was `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for non-COVID-19 death.
 Corresponding effectiveness was slightly lower in people aged 18-64 years: `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` respectively.
@@
 
-#### Table 3: Estimated booster vaccine effectiveness
-
-Effectiveness $(100 \times (1-\text{hazard ratio}))$ for the main and subgroup analyses and for all outcomes.
+#### Table 3:  Estimated booster vaccine effectiveness ($100 \times (1-\text{hazard ratio})$) for the main and subgroup analyses and for all outcomes.
 
@@
 
-#### Figure 2: Cumulative outcome incidence
-
-Kaplan-Meier estimates of cumulative incidence in matched boosted and unboosted treatment groups, stratified by primary course and without further adjustment for potential confounders.
+#### Figure 2: Kaplan-Meier estimates of cumulative incidence in matched boosted and unboosted treatment groups, stratified by primary course and without further adjustment for potential confounders.
 
@@
 
-#### Figure 3: Estimated booster effectiveness
-
-For each outcome based on the fully adjusted model, the hazard ratios for boosting with BNT162b2 versus not boosting is shown, stratified by primary course and time since boosting.
+#### Figure 3: Estimated booster vaccine effectiveness ($100 \times (1-\text{hazard ratio})$) for each outcome based on the fully adjusted model, stratified by primary course and time since boosting.
 Models with less extensive confounder adjustment are provided in supplementary materials (Figure S3).

### 2022-03-28 manuscript/draft-manuscript.Rmd
Update draft-manuscript.Rmd

@@
     geom_linerange(aes(ymin=hr.ll, ymax=hr.ul, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+
-    geom_point(data = overallhazards_plotdata, mapping = aes(y=hr, x=max(postbaselinecuts28+14)), colour="darkgrey") +
-    geom_linerange(data = overallhazards_plotdata, mapping = aes(ymin=hr.ll, ymax=hr.ul, x=max(postbaselinecuts28+14)), colour="darkgrey") +
-    geom_rect(data = overallhazards_plotdata, ymin=-Inf, ymax=Inf, xmin=max(postbaselinecuts28), xmax=max(postbaselinecuts28+28), fill="grey95", alpha=0.5, colour="transparent") +
+    geom_point(data = overallhazards_plotdata, mapping = aes(y=hr, x=max(postbaselinecuts+14)), colour="black") +
+    geom_linerange(data = overallhazards_plotdata, mapping = aes(ymin=hr.ll, ymax=hr.ul, x=max(postbaselinecuts+14)), colour="black") +
+    geom_rect(data = overallhazards_plotdata, ymin=-Inf, ymax=Inf, xmin=max(postbaselinecuts), xmax=max(postbaselinecuts+28), fill="grey95", alpha=0.5, colour="transparent") +
     geom_hline(aes(yintercept=1), colour='grey')+
@@
     )+
-    scale_x_continuous(breaks=c(postbaselinecuts28, max(postbaselinecuts28+14)), labels=c(postbaselinecuts28, "All"), limits=c(min(postbaselinecuts28), max(postbaselinecuts28+14)+14), expand = c(0, 0))+
+    scale_x_continuous(breaks=c(postbaselinecuts, max(postbaselinecuts+14)), labels=c(postbaselinecuts, "All"), limits=c(min(postbaselinecuts), max(postbaselinecuts+14)+14), expand = c(0, 0))+
     scale_colour_brewer(type="qual", palette="Set2", guide=guide_legend(ncol=1))+

### 2022-03-28 manuscript/draft-manuscript.Rmd
Update draft-manuscript.Rmd

@@
 `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="coviddeath", fup_period=="All") %>% pull(events))` COVID-19 deaths, 
-and `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="noncovidadmitted", fup_period=="All") %>% pull(events))` non-COVID-19 deaths 
+and `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="noncoviddeath", fup_period=="All") %>% pull(events))` non-COVID-19 deaths 
 across `r label_number(1, big.mark=",")((filter(incidences, treatment=="pfizer", outcome=="coviddeath", fup_period=="All") %>% pull(yearsatrisk)) * 365.25/7)` person-weeks of follow-up. 

### 2022-03-29 analysis/report_seqtrialcox.R
get incidence with km risk table too

@@ -209,9 +209,9 @@ format_ratio = function(numer,denom, width=7){
 
 threshold <- 5
 
-incidence_rate_rounded <- local({
+local({
 
-  unredacted_treated <-
+  incidence_treated <-
     data_seqtrialcox %>%
     group_by(treated, fup_period) %>%
     summarise(
@@ -223,7 +223,7 @@ incidence_rate_rounded <- local({
     ungroup()
 
 
-  unredacted_all <-
+  incidence_all <-
     data_seqtrialcox %>%
     group_by(treated) %>%
     summarise(
@@ -234,10 +234,10 @@ incidence_rate_rounded <- local({
     ) %>%
     ungroup()
 
-  unredacted <-
+  incidence <-
     bind_rows(
-      unredacted_treated,
-      unredacted_all
+      incidence_treated,
+      incidence_all
     ) %>%
     mutate(
       fup_period = fct_explicit_na(fup_period, na_level="All")
@@ -246,8 +246,8 @@ incidence_rate_rounded <- local({
       treated, fup_period
     )
 
-  unredacted_wide <-
-    unredacted %>%
+  incidence_rate_unrounded <<-
+    incidence %>%
     pivot_wider(
       id_cols =c(fup_period),
       names_from = treated,
@@ -261,8 +261,8 @@ incidence_rate_rounded <- local({
       rrECI = paste0(rrE, " ", rrCI)
     )
 
-  rounded <-
-    unredacted_wide %>%
+  incidence_rate_rounded <<-
+    incidence_rate_unrounded %>%
     mutate(
       n_0 = ceiling_any(n_0, threshold+1),
       n_1 = ceiling_any(n_1, threshold+1),
@@ -282,8 +282,6 @@ incidence_rate_rounded <- local({
       q_1 = format_ratio(events_1, yearsatrisk_1),
     )
 
-  rounded
-
 })
 
 write_csv(incidence_rate_rounded, fs::path(output_dir, "report_incidence.csv"))
@@ -292,7 +290,7 @@ write_csv(incidence_rate_rounded, fs::path(output_dir, "report_incidence.csv"))
 ## kaplan meier cumulative risk differences ----
 
 
-dat_surv <-
+data_surv <-
   data_seqtrialcox %>%
   group_by(patient_id, treated) %>%
   summarise(
@@ -315,7 +313,7 @@ dat_surv <-
   )
 
 data_surv_rounded <-
[…]

### 2022-03-29 manuscript/draft-manuscript.Rmd
add redacted outputs

@@
 #output_dir_os <- here("output", "manuscript-objects")
-output_dir_os <- here("released_output", "manuscript-objects")
+output_dir_os <- here("released_output", "release-objects")
 
@@
 
-cohortsummary <- read_csv(fs::path(output_dir_os, "cohortsummary.csv"))
 matchsummary <- read_csv(fs::path(output_dir_os, "matchsummary.csv"))
@@
 matchcoverage <- read_csv(fs::path(output_dir_os, "matchcoverage.csv"))
-incidence_all <- read_csv(fs::path(output_dir_os, "none", "incidence_all.csv")) %>% mutate(events=events_1+events_0)
+incidence <- read_csv(fs::path(output_dir_os, "none", "incidence.csv")) %>% mutate(events=events_1+events_0, yearsatrisk=yearsatrisk_1+yearsatrisk_0)
 pfizerdurationweeks <- as.numeric(as.duration(study_dates$pfizerend_date - study_dates$pfizerstart_date +1) , "weeks")
@@
     read_csv(fs::path(output_dir_os, "age65plus", "km.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
-    #read_csv(fs::path(output_dir_os, "prior_covid_infection", "km.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
+    read_csv(fs::path(output_dir_os, "prior_covid_infection", "km.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
   ) %>%
@@
   mutate(
+    treatment = fct_inorder(treatment),
+    treatment_descr = fct_inorder(treatment_descr),
+    outcome = fct_inorder(outcome),
+    outcome_descr = fct_inorder(outcome_descr),
+    outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=15)),
+    subgroup = fct_inorder(subgroup),
     subgroup_descr = fct_inorder(subgroup_descr),
+    subgroup_variable = fct_inorder(subgroup_variable),
+    subgroup_variable_descr = fct_inorder(subgroup_variable_descr),
     subgroup_level = fct_inorder(subgroup_level),
     subgroup_level_descr = fct_inorder(fct_explicit_na(subgroup_level_descr, "All")),
-    treatment_descr = fct_inorder(treatment_descr),
-    outcome_descr = fct_inorder(outcome_descr),
-    outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=15)),
     #model_name = fct_inorder(str_wrap(model_name, 15)),
@@
 
-groupvariables <- c("treatment", "treatment_descr", "outcome", "outcome_descr", "outcome_descr_wrap", "subgroup", "subgroup_descr", "subgroup_level", "subgroup_level_descr")
+groupvariables <- c("treatment", "treatment_descr", "outcome", "outcome_descr", "outcome_descr_wrap", "subgroup", "subgroup_descr", "subgroup_variable", "subgroup_variable_descr", "subgroup_level", "subgroup_level_descr")
 
@@
       interval = last(time) - first(time) + 1,
-      time=last(time),
+      time = last(time),
       cml.atrisk = last(cml.atrisk),
@@
     pivot_wider(
-      id_cols= c(groupvariables, "period_start", "period_end", "period",  "interval", "time"),
-      names_from=treated,
-      names_glue="{.value}_{treated}",
-      values_from=c(surv, risk, n.atrisk, n.event, n.censor, rate, cml.atrisk, cml.event, cml.censor, cml.rate)
+      id_cols = c(groupvariables, "period_start", "period_end", "period",  "interval", "time"),
+      names_from = treated,
+      names_glue = "{.value}_{treated}",
+      values_from = c(surv, risk, n.atrisk, n.event, n.censor, rate, cml.atrisk, cml.event, cml.censor, cml.rate)
     ) %>%
@@
     mutate(
-      cml.events = cml.event_0 + cml.events_1,
+      cml.events = cml.event_0 + cml.event_1,
       cml.atrisk = cml.atrisk_0 + cml.atrisk_1
@@
     ve_display = paste0(label_percent(accuracy=0.1)(ve), " (", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
-    ve_display95 = paste0(label_percent(accuracy=0.1)(ve), " (95% CI", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
+    ve_display95 = paste0(label_percent(accuracy=0.1)(ve), " (95% CI ", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
   )
@@
     ve_display = paste0(label_percent(accuracy=0.1)(ve), " (", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
-    ve_display95 = paste0(label_percent(accuracy=0.1)(ve), " (95%CI ", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
+    ve_display95 = paste0(label_percent(accuracy=0.1)(ve), " (95% CI ", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
   )
@@
 
-William J Hulme^1^, Elizabeth J Williamson^2^, Elsie Horne^3,4^, Amelia Green^1^, Linda Nab^1^, Ruth Keogh^2^, Edward PK Parker^2^, Venexia Walker^3,4^, Helen Curtis^1^, Milan Wiedemann^1^, Christine Cunningham^1^, Alex J Walker^1^, Louis Fisher^1^, Christopher T Rentsch^2^, Anna Schultze^2^, Krishnan Bhaskaran^2^, Helen I McDonald^2^, John Tazare^2^, Laurie Tomlinson^2^, Tom Palmer^3,4^, Brian MacKenna^1^, Richard Croker^1^, Caroline E Morton^1^, Colm Andrews^1^, Robin Parks^1^, Lisa Hopcroft^1^, Jon Massey^1^, Jessica Morley^1^, Amir Mehrkar^1^, Seb Bacon^1^, Dave Evans^1^, Peter Inglesby^1^, George Hickman^1^, Simon Davy^1^, Tom Ward^1^, Rosalind M Eggo^2^, Angel YS Wong^2^, Rohini Mathur^2^, Kevin Wing^2^, Harriet Forbes^2^, Daniel Grint^2^, Viyasaan Mahalingasivam^2^, Bang Zheng^2^, Ian J Douglas^2^, Stephen JW Evans^2^, Liam Smeeth^2^, Chris Bates^5^, Jonathan Cockburn^5^, John Parry^5^, Frank Hester^5^, Sam Harper^5^, Jonathan AC Sterne^3,4^, Miguel A Hernán^6,7^, Ben Goldacre^1^.
+William J Hulme^1^, Elizabeth J Williamson^2^, Elsie Horne^3,4^, Amelia Green^1^, Linda Nab^1^, Ruth Keogh^2^, Edward PK Parker^2^, Venexia Walker^3,4^, Helen Curtis^1^, Milan Wiedemann^1^, Christine Cunningham^1^, Alex J Walker^1^, Louis Fisher^1^, Christopher T Rentsch^2^, Anna Schultze^2^, Krishnan Bhaskaran^2^, Helen I McDonald^2^, John Tazare^2^, Laurie Tomlinson^2^, Tom Palmer^3,4^, Brian MacKenna^1^, Richard Croker^1^, Caroline E Morton^1^, Colm Andrews^1^, Robin Parks^1^, Lisa Hopcroft^1^, Jon Massey^1^, Jessica Morley^1^, Amir Mehrkar^1^, Seb Bacon^1^, Dave Evans^1^, Peter Inglesby^1^, George Hickman^1^, Simon Davy^1^, Tom Ward^1^, Rosalind M Eggo^2^, Angel YS Wong^2^, Rohini Mathur^2^, Kevin Wing^2^, Harriet Forbes^2^, Daniel Grint^2^, Viyasaan Mahalingasivam^2^, Bang Zheng^2^, Ian J Douglas^2^, Stephen JW Evans^2^, Liam Smeeth^2^, Chris Bates^5^, Jonathan Cockburn^5^, John Parry^5^, Frank Hester^5^, Sam Harper^5^, Jonathan AC Sterne^3,4^\*, Miguel A Hernán^6,7^\*, Ben Goldacre^1^\*.
 
@@
 
+\* equal contribution.
 \newpage
@@
 We matched `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` BNT162b2 booster recipients to unboosted controls.
-Estimated booster effectiveness comparing a BNT162b2 booster dose to two doses only was `r filter(overallhazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3) %>% pull(ve_display95)` for COVID-19 hospital admission, `r filter(overallhazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for non-COVID-19 death.
-Estimated effectiveness was similar in those receiving BNT162b2 or ChAdOx1-S two-dose primary vaccination course, but slightly lower in those classified as clinically extremely vulnerable (`r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for COVID-19 death) and in those aged 18-65 years (`r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` respectively).
-Period-specific effects were also estimated, showing some evidence of waning booster effectiveness against positive SARS-CoV-2 tests (`r filter(hazards, outcome=="postest", treatment=="pfizer", term_right=28, subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` at 15-28 days versus `r filter(hazards, outcome=="postest", treatment=="pfizer", term_right=70, subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` at 43-70 days).
+Estimated booster effectiveness comparing a BNT162b2 booster dose to two doses only was `r filter(overallhazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3) %>% pull(ve_display95)` for positive SARS-CoV-2 test, `r filter(overallhazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for non-COVID-19 death.
+Estimated effectiveness was similar in those receiving a BNT162b2 or ChAdOx1-S two-dose primary vaccination course, but slightly lower in those classified as clinically extremely vulnerable (`r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for COVID-19 death) and in those aged 18-65 years (`r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` respectively).
+Period-specific effects were also estimated, showing some evidence of waning booster effectiveness against positive SARS-CoV-2 tests (`r filter(hazards, outcome=="postest", treatment=="pfizer", term_right==28, subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` at 15-28 days versus `r filter(hazards, outcome=="postest", treatment=="pfizer", term_right==70, subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` at 43-70 days after boosting).
 
@@
 All data were linked, stored and analysed securely within the OpenSAFELY platform: <https://opensafely.org/>.
-Primary care records managed by the GP software provider TPP were linked, using NHS numbers, to A&E attendance and in-patient hospital spell records via NHS Digital's Hospital Episode Statistics (HES), national coronavirus testing records via the Second Generation Surveillance System (SGSS), and national death registry records from the Office for National Statistics (ONS).
+With the approval of NHS England, primary care records managed by the GP software provider TPP were linked, using NHS numbers, to A&E attendance and in-patient hospital spell records via NHS Digital's Hospital Episode Statistics (HES), national coronavirus testing records via the Second Generation Surveillance System (SGSS), and national death registry records from the Office for National Statistics (ONS).
 COVID-19 vaccination history is available in the GP record directly via the National Immunisation Management System (NIMS).
@@
 
-We included adults aged \ge 18 years who were eligible for booster vaccination between `r format(as.Date(study_dates$pfizerstart_date), "%e %B")` and `r format(as.Date(study_dates$pfizerend_date), "%e %B %Y")` inclusive (the "recruitment period").
+We included adults aged $\ge$ 18 years who were eligible for booster vaccination between `r format(as.Date(study_dates$pfizerstart_date), "%e %B")` and `r format(as.Date(study_dates$pfizerend_date), "%e %B %Y")` inclusive (the "recruitment period").
 On each day within this period, eligibility for the study was determined as follows: registered at a GP practice using TPP's SystmOne clinical information system; received a two-dose primary vaccination course of either BNT162b2 or ChAdOx1-S (mixed dosing and Moderna mRNA-1273 were not considered due to small numbers); not a health or social care worker, not resident in a care or nursing home; not medically housebound or receiving end-of-life care; no evidence of SARS-CoV-2 infection or COVID-19 disease within the previous 90 days; not undergoing an unplanned hospital admission; complete information on sex, ethnicity, deprivation, and NHS region.
@@
 
-We estimated Kaplan-Meier cumulative incidence curves separately in the boosted and unboosted groups, and the corresponding 10-week risk difference and risk ratio.
+We estimated Kaplan-Meier cumulative incidence curves separately in the boosted and unboosted groups, and compared groups using the 10-week risk difference and risk ratio.
 
@@
 
-We also estimated booster effectiveness separately in the following subgroups: primary vaccine course; age (18-64 or \ge 65 years); clinically extremely vulnerable or not; evidence of prior SARS-CoV-2 infection or not.
+We also estimated booster effectiveness separately in the following subgroups: primary vaccine course; age (18-64 or $\ge$ 65 years); clinically extremely vulnerable or not; evidence of prior SARS-CoV-2 infection or not.
 
@@
 
-There were `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="postest", fup_period=="All") %>% pull(events))` positive SARS-CoV-2 tests,
-`r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="covidadmitted", fup_period=="All") %>% pull(events))` COVID-19 hospital admissions, 
-`r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="coviddeath", fup_period=="All") %>% pull(events))` COVID-19 deaths, 
-and `r label_number(1, big.mark=",")(filter(incidences, treatment=="pfizer", outcome=="noncoviddeath", fup_period=="All") %>% pull(events))` non-COVID-19 deaths 
-across `r label_number(1, big.mark=",")((filter(incidences, treatment=="pfizer", outcome=="coviddeath", fup_period=="All") %>% pull(yearsatrisk)) * 365.25/7)` person-weeks of follow-up. 
+There were `r label_number(1, big.mark=",")(filter(incidence, treatment=="pfizer", outcome=="postest", fup_period=="All") %>% pull(events))` positive SARS-CoV-2 tests,
+`r label_number(1, big.mark=",")(filter(incidence, treatment=="pfizer", outcome=="covidadmitted", fup_period=="All") %>% pull(events))` COVID-19 hospital admissions, 
+`r label_number(1, big.mark=",")(filter(incidence, treatment=="pfizer", outcome=="coviddeath", fup_period=="All") %>% pull(events))` COVID-19 deaths, 
+and `r label_number(1, big.mark=",")(filter(incidence, treatment=="pfizer", outcome=="noncoviddeath", fup_period=="All") %>% pull(events))` non-COVID-19 deaths 
+across `r label_number(1, big.mark=",")((filter(incidence, treatment=="pfizer", outcome=="coviddeath", fup_period=="All") %>% pull(yearsatrisk)) * 365.25/7)` person-weeks of follow-up.
+
 The 10-week risk of positive SARS-CoV-2 test was `r km70 %>% filter(treatment=="pfizer", outcome=="postest", subgroup=="none") %>% pull(risk_1) %>% (label_number(0.1, scale=1000))` per 1000 in the boosted group
 and `r km70 %>% filter(treatment=="pfizer", outcome=="postest", subgroup=="none") %>% pull(risk_0) %>% (label_number(0.1, scale=1000))` in the unboosted group, 
-representing a risk difference of `r km70 %>% filter(treatment=="pfizer", outcome=="postest", subgroup=="none") %>% pull(kmrd) %>% (label_number(0.1, scale=1000))` per 1000. 
+representing a risk difference of `r km70 %>% filter(treatment=="pfizer", outcome=="postest", subgroup=="none") %>% pull(kmrd) %>% (label_number(0.1, scale=1000))` per 1000, .
 For COVID-19 hospital admission, this was `r km70 %>% filter(treatment=="pfizer", outcome=="covidadmitted", subgroup=="none") %>% pull(risk_1) %>% (label_number(0.1, scale=1000))` 
@@
 The rate of increase in the cumulative incidence of positive SARS-CoV-2 tests in boosted people tapered substantially around 7 days after boosting, something not observed in unboosted controls (Figure 2).
-The cumulative incidence of COVID-19 hospital admission and COVID-19 death also diverged substantially between the boosted and unboosted groups.
-We also see substantial differences for non-COVID-19 deaths between the boosted and unboosted.
+Similar patterns were not observed for more severe outcomes, with differences in the cumulative incidence between the groups apparent in the first few days.
 
 The overall (days 1-70) estimated booster effectiveness comparing the boosted and unboosted groups was `r filter(overallhazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3) %>% pull(ve_display95)` for positive SARS-CoV-2 test, `r filter(overallhazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for non-COVID-19 death (Table 3).
-Estimated booster effectiveness against COVID-19 outcomes was generally lower during days 1-28 than days 29-70: `r filter(meta2hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` and `r filter(meta2hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` for positive SARS-CoV-2 test; `r filter(meta2hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` and `r filter(meta2hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` respectively for COVID-19 hospital admission; `r filter(meta2hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` and `r filter(meta2hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` respectively for COVID-19 death.
-Estimated booster effectiveness against non-COVID19 death was similar during days 1-28 and 29-70 (`r filter(meta2hazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` and `r filter(meta2hazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)`).
+Estimated booster effectiveness against COVID-19 outcomes was generally lower during days 1-28 than days 29-70: `r filter(meta2hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` and `r filter(meta2hazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` for respectively positive SARS-CoV-2 test; `r filter(meta2hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` and `r filter(meta2hazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` respectively for COVID-19 hospital admission; `r filter(meta2hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` and `r filter(meta2hazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` respectively for COVID-19 death; and `r filter(meta2hazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3, term_right==28) %>% pull(ve_display)` and `r filter(meta2hazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3, term_right==70) %>% pull(ve_display)` for non-COVID-19 death.
 
@@
 
-Estimated booster effectiveness was similar between subgroups defined by primary vaccine course.
+Outcomes were broadly similar between subgroups defined by primary vaccine course.
 For those who had received BNT162b2, estimated booster effectiveness was `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3) %>% pull(ve_display)` for non-COVID-19 death.
@@
 
-Amongst people classified as clinically extremely vulnerable estimated booster effectiveness was `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for COVID-19 death.
-For people who were not classified as clinically extremely vulnerable estimated booster effectiveness was slightly higher at `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="cev-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="cev-FALSE", model==3) %>% pull(ve_display)`, and `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="cev-FALSE", model==3) %>% pull(ve_display)`, respectively.
-Estimated booster effectiveness for non-COVID-19 death was similar for those who were and were not classified as clinically extremely vulnerable: `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="cev-FALSE", model==3) %>% pull(ve_display)` respectively.
-
-Amongst people aged \ge 65 years estimated booster effectiveness was `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for non-COVID-19 death.
-Corresponding effectiveness was slightly lower in people aged 18-64 years: `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` respectively.
-
-Amongst people with evidence of prior SARS-CoV-2 infection, estimated booster effectiveness was `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="prior_covid_infection-TRUE", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="prior_covid_infection-TRUE", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="prior_covid_infection-TRUE", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="prior_covid_infection-TRUE", model==3) %>% pull(ve_display)` for non-COVID-19 death.
-Corresponding effectiveness was ...
-in people with no evidence of prior infection: `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="prior_covid_infection-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="prior_covid_infection-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="prior_covid_infection-FALSE", model==3) %>% pull(ve_display)`, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="prior_covid_infection-FALSE", model==3) %>% pull(ve_display)` respectively.
+Amongst people classified as clinically extremely vulnerable, estimated booster effectiveness was `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for non-COVID-19 death.
+For people who were not classified as clinically extremely vulnerable, estimated booster effectiveness was similar for positive SARS-CoV-2 tests (`r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="cev-FALSE", model==3) %>% pull(ve_display)` and non-COVID-19 death (`r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="cev-FALSE", model==3) %>% pull(ve_display)`), but higher for COVID-19 hospital admission and death (`r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="cev-FALSE", model==3) %>% pull(ve_display)` and `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="cev-FALSE", model==3) %>% pull(ve_display)` respectively).
+
+Amongst people aged $\ge$ 65 years, estimated booster effectiveness was `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for non-COVID-19 death.
+Corresponding effectiveness was similar in people aged 18-64 years for positive SARS-CoV-2 tests (`r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`), but slightly lower for more severe outcomes: `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` for non-COVID-19 death.
+
+Outcomes were similar for COVID-19-related events when comparing people with or without evidence of prior SARS-CoV-2 infection.
+Estimated booster effectiveness for those with evidence of prior infection was `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="prior_covid_infection-TRUE", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="prior_covid_infection-TRUE", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="prior_covid_infection-TRUE", model==3) %>% pull(ve_display)` for COVID-19 death.
+Corresponding estimates in people with no evidence of prior infection were `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="prior_covid_infection-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="prior_covid_infection-FALSE", model==3) %>% pull(ve_display)`, and `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="prior_covid_infection-FALSE", model==3) %>% pull(ve_display)`, respectively.
+For non-COVID-19, estimated booster effectiveness was slightly lower in those with prior evidence of SARS-CoV-2 infection (`r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="prior_covid_infection-TRUE", model==3) %>% pull(ve_display)`) than in those without (`r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="prior_covid_infection-FALSE", model==3) %>% pull(ve_display)`).
+
+Booster effectiveness was also estimated using narrower periods, with effectiveness against positive SARS-CoV-2 tests appearing to wane in later periods (`r filter(hazards, outcome=="postest", treatment=="pfizer", term_right==28, subgroup=="none", model==3) %>% pull(ve_display)` during days 15-28 after boosting versus `r filter(hazards, outcome=="postest", treatment=="pfizer", term_right==70, subgroup=="none", model==3) %>% pull(ve_display)` during days 43-70), a pattern that persisted in primary course subgroup analyses (Figure 3).
+For more severe outcomes, period-specific effectiveness estimates are less clear due to fewer events, though were broadly similar between different primary course brands.
 
@@
 
-incidence_table <- 
-  incidence %>%
-  left_join(
-    kmdiff %>% filter(time == postbaselinecuts[-1] -1),
-    by = c("treatment", "outcome", "subgroup", )
-  ) %>%
+format_ratio = function(numer,denom, width=7){
+  paste0(
+    replace_na(scales::comma_format(accuracy=1)(numer), "--"),
+    " / ",
+    str_pad(replace_na(scales::comma_format(accuracy=1)(denom),"--"), width=width, pad=" ")
+  )
+}
+
+km70_table <- 
+  km70 %>%
   filter(
@@
     outcome %in% outcomes,
-    treatment == "pfizer"
+    treatment == "pfizer",
+    subgroup == "none",
+  ) %>%
+  ungroup() %>%
+  mutate(
+    q_0 = format_ratio(cml.event_0, cml.atrisk_0/365.25),
+    q_1 = format_ratio(cml.event_1, cml.atrisk_1/365.25),
+    risk_1=label_number(accuracy=0.01, scale=1000)(risk_1),
+    risk_0=label_number(accuracy=0.01, scale=1000)(risk_0),
+    kmrr = label_number(accuracy=0.001)(kmrr),
+    kmrd = label_number(accuracy=0.01, scale=1000)(kmrd)
   ) %>%
@@
     outcome_descr,
-    #fup_period,
     q_1,
@@
     risk_0,
-    kmrrE,
-    kmrdE
+    kmrr,
+    kmrd
     #rrECI
-  ) %>%
-  mutate(
-    risk_1=if_else(risk_1<0.001, "<0.001", label_number(0.001)(risk_1)),
-    risk_0=if_else(risk_0<0.001, "<0.001", label_number(0.001)(risk_0)),
   )
@@
   gt(
-    incidence_table
+    km70_table
    #groupname_col = c("treatment_descr"),
@@
 
-    risk_0 = "Cumulative risk",
-    risk_1 = "Cumulative risk",
-    kmrrE = "Risk ratio",
-    kmrdE = "Risk difference (per 1000 people)"
-    #rrECI = "Incidence rate ratio (95% CI)"
+    risk_0 = "Cumulative risk, per 1000 people",
+    risk_1 = "Cumulative risk, per 1000 people",
+    kmrr = "Risk ratio",
+    kmrd = "Risk difference, per 1000 people"
   ) %>%
@@
   ) %>%
-  fmt_missing(
-    everything(),
-    missing_text="--"
+  tab_spanner(
+    label = "Boosted versus unboosted",
+    columns = starts_with("km")
   ) %>%
@@
     columns = everything()
+  )
+
+```
+
+\newpage
+
+#### Table 2 (v2): Total events and person-years of follow-up, for each outcome, in the boosted and unboosted groups.
+
+```{r table-incidence2, echo=FALSE, warning=FALSE, message=FALSE}
+
+format_ratio = function(numer,denom, width=7){
+  paste0(
+    replace_na(scales::comma_format(accuracy=1)(numer), "--"),
+    " / ",
+    str_pad(replace_na(scales::comma_format(accuracy=1)(denom),"--"), width=width, pad=" ")
+  )
+}
+
+km_table <- 
+  kmtreated_cuts %>%
+  filter(
+    outcome %in% outcomes,
+    treatment == "pfizer",
+    #subgroup == "none",
+  ) %>%
+  #group_by(across(all_of(groupvariables))) %>%
+  mutate(
+    q_0 = format_ratio(cml.event_0, cml.atrisk_0/365.25),
+    q_1 = format_ratio(cml.event_1, cml.atrisk_1/365.25),
+  ) %>%
+  select(
+    #treatment_descr, 
+    outcome_descr,
+    period,
+    subgroup_variable_descr,
+    subgroup_level_descr,
+    Boosted=q_1,
+    Unboosted=q_0, 
+  ) %>%
+  pivot_wider(
+    id_cols = c("subgroup_variable_descr", "subgroup_level_descr", "period"),
+    names_from = outcome_descr,
+    names_sep = "_",
+    values_from = c("Boosted", "Unboosted")
+  ) %>%
+  group_by(subgroup_variable_descr, subgroup_level_descr) %>%
+  mutate(
+    subgroup_level_descr=if_else(row_number()==1, subgroup_level_descr, as.factor("")),
+    subgroup_level_descr=if_else(subgroup_variable_descr!="Main analysis", subgroup_level_descr, as.factor(""))
+  ) %>%
+  group_by(subgroup_variable_descr) %>%
+  mutate(
+    subgroup_variable_descr=if_else(row_number()==1, subgroup_variable_descr, as.factor(""))
+  ) %>% 
+  ungroup()
+
+  gt(
+    km_table
+   #groupname_col = c("treatment_descr"),
+  ) %>%
+  cols_label(
+    subgroup_variable_descr = "Subgroup",
+    subgroup_level_descr = "",
+    period = "Days since booster"
+  ) %>%
+  tab_spanner_delim(
+    delim = "_",
+    columns = c(starts_with("Boosted"), starts_with("Unboosted"))
+  ) %>%
+  cols_align(
+    align = "right",
+    columns = everything()
+  ) %>%
+  fmt_missing(
+    everything(),
+    missing_text=""
   )
@@
     geom_step(aes(x=time, y=(1-surv)*1000, group=treated_descr, colour=treated_descr))+
-    geom_rect(aes(xmin=time, xmax=leadtime, ymin=(1-surv.ll)*1000, ymax=(1-surv.ul)*1000, group=treated_descr, fill=treated_descr), alpha=0.1)+
+    geom_rect(aes(xmin=time, xmax=time+1, ymin=(1-surv.ll)*1000, ymax=(1-surv.ul)*1000, group=treated_descr, fill=treated_descr), alpha=0.1)+
     geom_hline(aes(yintercept=0), colour='black')+

### 2022-03-29 manuscript/reports.Rmd

@@
 # where are the outputs (ie the inputs for this manuscript!) saved?
-#output_dir_os <- here("output", "manuscript-objects")
-output_dir_os <- here("released_output", "manuscript-objects")
+#output_dir_os <- here("output", "release-objects")
+output_dir_os <- here("released_output", "release-objects")
 
@@
 matchcoverage <- read_csv(fs::path(output_dir_os, "matchcoverage.csv"))
-incidences <- read_csv(fs::path(output_dir_os, "none", "incidence.csv")) %>% mutate(events=events_1+events_0)
+incidence <- read_csv(fs::path(output_dir_os, "none", "incidence.csv")) %>% mutate(events=events_1+events_0, yearsatrisk=yearsatrisk_1+yearsatrisk_0)
 
@@
     read_csv(fs::path(output_dir_os, "age65plus", "km.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
-    read_csv(fs::path(output_dir_os, "prior_covid_infection", "km.csv")) %>% mutate(subgroup_level=as.character(subgroup_level))
+    read_csv(fs::path(output_dir_os, "prior_covid_infection", "km.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
   ) %>%
@@
   mutate(
+    treatment = fct_inorder(treatment),
+    treatment_descr = fct_inorder(treatment_descr),
+    outcome = fct_inorder(outcome),
+    outcome_descr = fct_inorder(outcome_descr),
+    outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=15)),
+    subgroup = fct_inorder(subgroup),
     subgroup_descr = fct_inorder(subgroup_descr),
+    subgroup_variable = fct_inorder(subgroup_variable),
+    subgroup_variable_descr = fct_inorder(subgroup_variable_descr),
     subgroup_level = fct_inorder(subgroup_level),
     subgroup_level_descr = fct_inorder(fct_explicit_na(subgroup_level_descr, "All")),
-    treatment_descr = fct_inorder(treatment_descr),
-    outcome_descr = fct_inorder(outcome_descr),
-    outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=12)),
     #model_name = fct_inorder(str_wrap(model_name, 15)),
-  )
+    treated = (treated_descr=="Boosted")*1L
+  ) %>%
+  group_by(treatment, treated_descr, outcome, subgroup) %>%
+  mutate(
+    cml.risk = cumsum(n.risk),
+    cml.event = cumsum(replace_na(n.event, 0)),
+  )
+
+groupvariables <- c("treatment", "treatment_descr", "outcome", "outcome_descr", "outcome_descr_wrap", "subgroup", "subgroup_descr", "subgroup_variable", "subgroup_variable_descr", "subgroup_level", "subgroup_level_descr")
+
+km_incidence <-
+  km %>%
+  mutate(
+    n.atrisk = n.risk,
+    cml.atrisk = cumsum(replace_na(n.atrisk, 0)),
+    cml.event = cumsum(replace_na(n.event, 0)),
+    cml.censor = cumsum(replace_na(n.censor, 0)),
+    rate = n.event/n.atrisk,
+    cml.rate = cml.event / cml.atrisk,
+    risk = 1 - surv,
+  ) %>%
+  select(
+    treated,
+    all_of(groupvariables),
+    time, interval,
+    surv, risk, n.atrisk, n.event, n.censor, rate, cml.atrisk, cml.event, cml.censor, cml.rate
+  )
+
+kmconstrast <- function(data, cuts=NULL){
+
+  if(is.null(cuts)){cuts <- unique(data$time)}
+
+  data %>%
+    filter(time!=0) %>%
+    mutate(
+      period_start = cut(time, cuts, right=TRUE, label=cuts[-length(cuts)]),
+      period_end = cut(time, cuts, right=TRUE, label=cuts[-1]),
+      period = cut(time, cuts, right=TRUE, label=paste0(cuts[-length(cuts)]+1, " - ", cuts[-1]))
+    ) %>%
+    group_by(across(all_of(groupvariables))) %>%
+    group_by(
+      treated, period_start, period_end, period,
+      .add = TRUE
+    ) %>%
+    summarise(
+      interval = last(time) - first(time) + 1,
+      time = last(time),
+      cml.atrisk = last(cml.atrisk),
+      cml.event = last(cml.event),
+      cml.censor = last(cml.censor),
+      cml.rate = last(cml.rate),
+      persontime = sum(n.atrisk),
+      n.atrisk = first(n.atrisk),
+      n.event = sum(n.event, na.rm=TRUE),
+      n.censor = sum(n.censor, na.rm=TRUE),
+      surv = last(surv),
+      risk = last(risk),
+      rate = n.event/persontime,
+    ) %>%
+    pivot_wider(
+      id_cols = c(groupvariables, "period_start", "period_end", "period",  "interval", "time"),
+      names_from = treated,
+      names_glue = "{.value}_{treated}",
+      values_from = c(surv, risk, n.atrisk, n.event, n.censor, rate, cml.atrisk, cml.event, cml.censor, cml.rate)
+    ) %>%
+    mutate(
+      kmrr = risk_1 / risk_0,
+      kmrd = risk_1 - risk_0,
+      irr = rate_1 / rate_0,
+      ird = rate_1 - rate_0,
+      cmlirr = cml.rate_1 / cml.rate_0,
+      cmlird = cml.rate_1 - cml.rate_0
+    ) %>%
+    mutate(
+      kmrd_display = label_number(accuracy=0.1, scale=1000, big.mark =",")(kmrd),
+      kmrd_inv_display = label_number(accuracy=0.1, scale=1000, big.mark =",")(-kmrd),
+    ) %>%
+    mutate(
+      cml.events = cml.event_0 + cml.event_1,
+      cml.atrisk = cml.atrisk_0 + cml.atrisk_1
+    )
+}
+
+kmtreated_daily <- kmconstrast(km_incidence)
+kmtreated_cuts <- kmconstrast(km_incidence, postbaselinecuts)
+kmtreated_overall <- kmconstrast(km_incidence, c(0,last(postbaselinecuts)))
+
+km70 <- kmtreated_daily %>% filter(time==max(postbaselinecuts))
+
 # 
-# cmlinc <- read_csv(fs::path(output_dir_os, "cmlinc.csv")) %>%
+# incidence <- 
+#   bind_rows(
+#     read_csv(fs::path(output_dir_os, "none", "incidence.csv")),
+#     read_csv(fs::path(output_dir_os, "vax12_type", "incidence.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
+#     #read_csv(fs::path(output_dir_os, "cev", "incidence.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
+#     #read_csv(fs::path(output_dir_os, "age65plus", "incidence.csv"))  %>% mutate(subgroup_level=as.character(subgroup_level)),
+#     #read_csv(fs::path(output_dir_os, "prior_covid_infection", "incidence.csv"))  %>% mutate(subgroup_level=as.character(subgroup_level)),
+#   ) %>%
+#   filter(
+#     outcome %in% outcomes
+#   ) %>%
 #   mutate(
+#     subgroup_descr = fct_inorder(subgroup_descr),
+#     subgroup_level = fct_inorder(subgroup_level),
+#     #subgroup_level_descr = fct_inorder(fct_explicit_na(subgroup_level_descr, "All")),
 #     treatment_descr = fct_inorder(treatment_descr),
 #     outcome_descr = fct_inorder(outcome_descr),
-#     model_name = fct_inorder(str_wrap(model_name, 15)),
-#     inc_CI_to = paste0(label_number(0.1, scale=1000)(1-survival.ul), " to ", label_number(0.1, scale=1000)(1-survival.ll))
-#   )
-
-# riskdiff <- read_csv(fs::path(output_dir_os, "riskdiff.csv")) %>%
-#   mutate(
-#     treatment_descr = fct_inorder(treatment_descr),
-#     outcome_descr = fct_inorder(outcome_descr),
-#     model_name = fct_inorder(str_wrap(model_name, 15)),
-#     diff_CI_to = paste0(label_number(0.01, scale=1000)(diff.ll), " to ", label_number(0.01, scale=1000)(diff.ul))
+#     outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=15)),
 #   )
@@
     ve_display = paste0(label_percent(accuracy=0.1)(ve), " (", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
-
+    ve_display95 = paste0(label_percent(accuracy=0.1)(ve), " (95% CI ", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
   )
@@
     ve_display = paste0(label_percent(accuracy=0.1)(ve), " (", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
-  )
-
-meta28hazards <- 
-  # derive directly from hazards as meta28effects table needs to be re-run
+    ve_display95 = paste0(label_percent(accuracy=0.1)(ve), " (95% CI ", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
+  )
+
+meta2hazards <- 
+  # derive directly from hazards as meta2effects table needs to be re-run
   bind_rows(
@@
     ve_display = paste0(label_percent(accuracy=0.1)(ve), " (", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
-  )
-
-meta42hazards <- 
-  # derive directly from hazards as meta28effects table needs to be re-run
-  bind_rows(
-    read_csv(fs::path(output_dir_os, "none", "effects.csv")),
-    read_csv(fs::path(output_dir_os, "vax12_type", "effects.csv")),
-    read_csv(fs::path(output_dir_os, "cev", "effects.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
-    read_csv(fs::path(output_dir_os, "age65plus", "effects.csv"))  %>% mutate(subgroup_level=as.character(subgroup_level)),
-    read_csv(fs::path(output_dir_os, "prior_covid_infection", "effects.csv"))  %>% mutate(subgroup_level=as.character(subgroup_level)),
-  ) %>%
-  filter(
-    outcome %in% outcomes
-  ) %>%
-  mutate(
-    period = (term_right>42) + 1,
-    subgroup_descr = fct_inorder(subgroup_descr),
-    subgroup_level = fct_inorder(subgroup_level),
-    subgroup_level_descr = fct_inorder(fct_explicit_na(subgroup_level_descr, "All")),
-    treatment_descr = fct_inorder(treatment_descr),
-    outcome_descr = fct_inorder(outcome_descr),
-    outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=15)),
-  ) %>%
-  group_by(
-    treatment, treatment_descr, outcome, outcome_descr, outcome_descr_wrap, subgroup, subgroup_variable, subgroup_descr, subgroup_level, subgroup_level_descr, model, model_descr, period, 
-  ) %>%
-  summarise(
-    term_left = min(term_left),
-    term_right = max(term_right),
-    term_midpoint = (term_left+term_right)/2,
-    estimate = weighted.mean(estimate, robust.se^-2),
-    std.error = sqrt(1/sum(std.error^-2)),
-    robust.se = sqrt(1/sum(robust.se^-2)),
-    statistic = estimate/robust.se,
-    p.value = 2 * pmin(pnorm(statistic), pnorm(-statistic)),
-    conf.low = estimate + qnorm(0.025)*robust.se,
-    conf.high = estimate + qnorm(0.975)*robust.se,
-    hr = exp(estimate),
-    hr.ll = exp(conf.low),
-    hr.ul = exp(conf.high),
-    ve = 1-hr,
-    ve.ll = 1-hr.ul,
-    ve.ul = 1-hr.ll
-  ) %>%
-  ungroup() %>%
-  mutate(
-    hr_display = paste0(label_number(accuracy=0.01)(hr), " (", label_number(accuracy=0.01)(hr.ll), "-", label_number(accuracy=0.01)(hr.ul), ")"),
-    ve_display = paste0(label_percent(accuracy=0.1)(ve), " (", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
-  )
+    ve_display95 = paste0(label_percent(accuracy=0.1)(ve), " (95% CI ", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
+  )
+
+# meta2hazards <- 
+#   bind_rows(
+#     read_csv(fs::path(output_dir_os, "none", "meta2effects.csv")),
+#     read_csv(fs::path(output_dir_os, "vax12_type", "meta2effects.csv")),
+#     read_csv(fs::path(output_dir_os, "cev", "meta2effects.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
+#     read_csv(fs::path(output_dir_os, "age65plus", "meta2effects.csv")) %>% mutate(subgroup_level=as.character(subgroup_level)),
+#   ) %>%
+#   filter(
+#     outcome %in% outcomes
+#   ) %>%
+#   mutate(
+#     subgroup_descr = fct_inorder(subgroup_descr),
+#     subgroup_level = fct_inorder(subgroup_level),
+#     subgroup_level_descr = fct_inorder(fct_explicit_na(subgroup_level_descr, "All")),
+#     treatment_descr = fct_inorder(treatment_descr),
+#     outcome_descr = fct_inorder(outcome_descr),
+#     outcome_descr_wrap = fct_relabel(outcome_descr, ~str_wrap(., width=20)),
+#     #model_name = fct_inorder(str_wrap(model_name, 15)),
+#     hr_display = paste0(label_number(accuracy=0.01)(hr), " (", label_number(accuracy=0.01)(hr.ll), "-", label_number(accuracy=0.01)(hr.ul), ")"),
+#     ve_display = paste0(label_percent(accuracy=0.1)(ve), " (", label_percent(accuracy=0.1, suffix="")(ve.ll), "-", label_percent(accuracy=0.1, suffix="")(ve.ul), ")"),
+#   )
 
@@
 
-<!-- TODO -->
-
-<!-- * full description of booster eligiblity over time -->
-
-<!-- * figure 1 chart with SMD / %-point differences -->
-
-<!-- * check for moderna references and remove if necessary -->
-
-<!-- * check final recruitment day and include in results -->
-
-<!-- * check if rolling average restrictions are ever used (need to output dataset) -->
-
-<!-- * references -->
-
-#### Table 0: Inclusion criteria
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
+
+### Table 0: Inclusion criteria
 
@@
   pivot_wider(
-    id_cols = c(stage, criteria),
+    id_cols = c(stage, criteria, treatment),
     names_from = treatment,
@@
     n_pfizer = "N",
-    #n_moderna = "N",
+    n_moderna = "N",
 
     n_exclude_pfizer = "N excluded",
-    #n_exclude_moderna = "N excluded",
+    n_exclude_moderna = "N excluded",
 
     pct_all_pfizer = "% remaining",
-    #pct_all_moderna = "% remaining",
+    pct_all_moderna = "% remaining",
   ) %>%
@@
     columns = ends_with("pfizer")
+  ) %>%
+   tab_spanner(
+    label = "ChAdOx1-S",
+    columns = ends_with("moderna")
   ) %>%
@@
 
-#### Table 1a: Participant characteristics for Pfizer boosting
+\newpage 
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
+
+### Table 1a: Participant characteristics for BNT162b2 boosting
 
@@
 
-```{r table1, echo=FALSE, warning=FALSE, message=FALSE}
+```{r table1a, echo=FALSE, warning=FALSE, message=FALSE}
 table1_body <- read_csv(fs::path(output_dir_os, "matchtable1.csv")) %>% filter(treatment=="pfizer")
@@
 
-#### Table 1b: Participant characteristics for Moderna boosting
+\newpage 
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
+
+### Table 1b: Participant characteristics for mRA-1273 boosting
 
@@
 
-```{r table1, echo=FALSE, warning=FALSE, message=FALSE}
+```{r table1b, echo=FALSE, warning=FALSE, message=FALSE}
 table1_body <- read_csv(fs::path(output_dir_os, "matchtable1.csv")) %>% filter(treatment=="moderna")
@@
 \newpage
-
-#### Table 2a: Follow-up and outcomes for Pfizer boosting
+&nbsp; &nbsp; &nbsp;
+
+### Table 2a: Follow-up and outcomes for BNT162b2 boosting
 
@@
 
-```{r table2, echo=FALSE, warning=FALSE, message=FALSE}
-incidences %>%
+Counts in the underlying risk table has been rounded up to the nearest 6, so numbers may not add up exactly.
+
+```{r table2a, echo=FALSE, warning=FALSE, message=FALSE}
+
+format_ratio = function(numer,denom, width=7){
+  paste0(
+    replace_na(scales::comma_format(accuracy=1)(numer), "--"),
+    " / ",
+    str_pad(replace_na(scales::comma_format(accuracy=1)(denom),"--"), width=width, pad=" ")
+  )
+}
+
+km70_table <- 
+  kmtreated_overall %>%
   filter(
-    fup_period == "All",
-    outcome %in% c("postest", "covidadmitted", "coviddeath", "noncoviddeath"),
-    treatment == "pfizer"
-  ) %>%
-  left_join(
-    matchsummary %>% select(treatment, n_eligible, prop_matched_eligible),
-    by=c("treatment")
-  ) %>%
-  mutate(
-    treatment_descr = paste0(
-      #treatment_descr, " - ",
-      label_number(1, big.mark=",")(n_eligible), " eligible, ",
-      label_number(1, big.mark=",")(n_1), " matched (",
-      label_percent(0.1)(prop_matched_eligible), ")."
-    ),
-  ) %>%
+    #fup_period == "All",
+    outcome %in% outcomes,
+    treatment == "pfizer",
+    #subgroup == "none",
+  ) %>%
+  ungroup() %>%
+  mutate(
+    n.atrisk = label_number(accuracy=1, big.mark = ",")(n.atrisk_1),
+    q_0 = format_ratio(cml.event_0, cml.atrisk_0/365.25),
+    q_1 = format_ratio(cml.event_1, cml.atrisk_1/365.25),
+    risk_1=label_number(accuracy=0.01, scale=1000)(risk_1),
+    risk_0=label_number(accuracy=0.01, scale=1000)(risk_0),
+    kmrr = label_number(accuracy=0.001)(kmrr),
+    kmrd = label_number(accuracy=0.01, scale=1000)(kmrd)
+  ) %>%
+  group_by(treatment, outcome, subgroup_variable_descr, subgroup_level_descr) %>%
+  mutate(
+    subgroup_level_descr=if_else(row_number()==1, subgroup_level_descr, as.factor("")),
+    subgroup_level_descr=if_else(subgroup_variable_descr!="Main analysis", subgroup_level_descr, as.factor(""))
+  ) %>%
+  group_by(treatment, outcome, subgroup_variable_descr) %>%
+  mutate(
+    subgroup_variable_descr=if_else(row_number()==1, subgroup_variable_descr, as.factor(""))
+  ) %>% 
+  ungroup() %>%
   select(
-    treatment_descr, 
     outcome_descr,
+    subgroup_variable_descr,
+    subgroup_level_descr,
+    n.atrisk,
     q_1,
-    rate_1,
+    risk_1,
     q_0, 
-    rate_0,
-    rrECI
-  ) %>%
-  mutate(
-    rate_1=if_else(rate_1<0.001, "<0.001", label_number(0.001)(rate_1)),
-    rate_0=if_else(rate_0<0.001, "<0.001", label_number(0.001)(rate_0)),
-  ) %>%
+    risk_0,
+    kmrr,
+    kmrd
+  )
+
   gt(
-   groupname_col = c("treatment_descr"),
+    km70_table,
+   groupname_col = c("outcome_descr"),
   ) %>%
   cols_label(
-    #treatment_descr = "Treatment",
-    outcome_descr = "Outcome",
-
+    subgroup_variable_descr = "Subgroup",
+    subgroup_level_descr = "",
+    #fup_period = "Days since booster",
     q_0 = "Events / person-years",
     q_1   = "Events / person-years",
-
-    rate_0 = "Incidence rate",
-    rate_1 = "Incidence rate",
-    rrECI = "Incidence rate ratio (95% CI)"
+    n.atrisk = "N per group",
+    risk_0 = "Cumulative risk, per 1000 people",
+    risk_1 = "Cumulative risk, per 1000 people",
+    kmrr = "Risk ratio",
+    kmrd = "Risk difference, per 1000 people"
   ) %>%
@@
   ) %>%
+  tab_spanner(
+    label = "Boosted versus unboosted",
+    columns = starts_with("km")
+  ) %>%
   fmt_missing(
     everything(),
-    missing_text="--"
+    missing_text=""
   ) %>%
@@
 \newpage
-
-#### Table 2b: Follow-up and outcomes for Moderna boosting
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
+
+### Table 2b: Follow-up and outcomes for mRNA-1273 boosting
 
@@
 
-```{r table2, echo=FALSE, warning=FALSE, message=FALSE}
-incidences %>%
+Counts in the underlying risk table has been rounded up to the nearest 6, so numbers may not add up exactly.
+
+```{r table2b, echo=FALSE, warning=FALSE, message=FALSE}
+
+format_ratio = function(numer,denom, width=7){
+  paste0(
+    replace_na(scales::comma_format(accuracy=1)(numer), "--"),
+    " / ",
+    str_pad(replace_na(scales::comma_format(accuracy=1)(denom),"--"), width=width, pad=" ")
+  )
+}
+
+km70_table <- 
+  kmtreated_overall %>%
   filter(
-    fup_period == "All",
-    outcome %in% c("postest", "covidadmitted", "coviddeath", "noncoviddeath"),
-    treatment == "moderna"
-  ) %>%
-  left_join(
-    matchsummary %>% select(treatment, n_eligible, prop_matched_eligible),
-    by=c("treatment")
-  ) %>%
-  mutate(
-    treatment_descr = paste0(
-      #treatment_descr, " - ",
-      label_number(1, big.mark=",")(n_eligible), " eligible, ",
-      label_number(1, big.mark=",")(n_1), " matched (",
-      label_percent(0.1)(prop_matched_eligible), ")."
-    ),
-  ) %>%
+    #fup_period == "All",
+    outcome %in% outcomes,
+    treatment == "moderna",
+    #subgroup == "none",
+  ) %>%
+  ungroup() %>%
+  mutate(
+    n.atrisk = label_number(accuracy=1, big.mark = ",")(n.atrisk_1),
+    q_0 = format_ratio(cml.event_0, cml.atrisk_0/365.25),
+    q_1 = format_ratio(cml.event_1, cml.atrisk_1/365.25),
+    risk_1=label_number(accuracy=0.01, scale=1000)(risk_1),
+    risk_0=label_number(accuracy=0.01, scale=1000)(risk_0),
+    kmrr = label_number(accuracy=0.001)(kmrr),
+    kmrd = label_number(accuracy=0.01, scale=1000)(kmrd)
+  ) %>%
+  group_by(treatment, outcome, subgroup_variable_descr, subgroup_level_descr) %>%
+  mutate(
+    subgroup_level_descr=if_else(row_number()==1, subgroup_level_descr, as.factor("")),
+    subgroup_level_descr=if_else(subgroup_variable_descr!="Main analysis", subgroup_level_descr, as.factor(""))
+  ) %>%
+  group_by(treatment, outcome, subgroup_variable_descr) %>%
+  mutate(
+    subgroup_variable_descr=if_else(row_number()==1, subgroup_variable_descr, as.factor(""))
+  ) %>% 
+  ungroup() %>%
   select(
-    treatment_descr, 
+    #treatment_descr, 
     outcome_descr,
+    subgroup_variable_descr,
+    subgroup_level_descr,
+    n.atrisk,
     q_1,
-    rate_1,
+    risk_1,
     q_0, 
-    rate_0,
-    rrECI
-  ) %>%
-  mutate(
-    rate_1=if_else(rate_1<0.001, "<0.001", label_number(0.001)(rate_1)),
-    rate_0=if_else(rate_0<0.001, "<0.001", label_number(0.001)(rate_0)),
-  ) %>%
+    risk_0,
+    kmrr,
+    kmrd
+  )
+
   gt(
-   groupname_col = c("treatment_descr"),
+    km70_table,
+   groupname_col = c("outcome_descr"),
   ) %>%
   cols_label(
-    #treatment_descr = "Treatment",
-    outcome_descr = "Outcome",
-
+    subgroup_variable_descr = "Subgroup",
+    subgroup_level_descr = "",
+    #fup_period = "Days since booster",
     q_0 = "Events / person-years",
     q_1   = "Events / person-years",
-
-    rate_0 = "Incidence rate",
-    rate_1 = "Incidence rate",
-    rrECI = "Incidence rate ratio (95% CI)"
+    n.atrisk = "N per group",
+    risk_0 = "Cumulative risk, per 1000 people",
+    risk_1 = "Cumulative risk, per 1000 people",
+    kmrr = "Risk ratio",
+    kmrd = "Risk difference, per 1000 people"
   ) %>%
@@
 ```
+
 \newpage
-
-## Table 3a: Estimated Pfizer booster effectiveness (5 period)
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
+
+### Table 3a: Estimated Pfizer booster effectiveness (5 period)
 
 Main and subgroup analyses
-```{r table_effects, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}
+```{r table_effectsa, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}
 
@@
 \newpage
-
-## Table 3a: Estimated Moderna booster effectiveness (5 period)
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
+
+### Table 3b: Estimated Moderna booster effectiveness (5 period)
 
 Main and subgroup analyses
-```{r table_effects, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}
+```{r table_effectsb, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}
 
@@
 
-#### Figure 1: Booster vaccination uptake, treatment group eligibility and matching success
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
+
+### Figure 1a: Pfizer booster vaccination uptake, treatment group eligibility and matching success
 
@@
 
-```{r, vaxdate, echo=FALSE, message=FALSE, warning=FALSE, out.width='60%', out.height='50%', results='asis', out.extra=''}
+```{r, vaxdatea, echo=FALSE, message=FALSE, warning=FALSE, out.width='60%', out.height='50%', results='asis', out.extra=''}
 #read_rds(fs::path(output_dir_os, "descriptive", "vaxdate", "plot_vaxdate_stack.rds"))
@@
   ) %>%
-  #filter(treatment=="pfizer") %>%
+  filter(treatment=="pfizer") %>%
   mutate(
@@
   scale_x_date(
-    breaks = unique(lubridate::ceiling_date(matchcoverage$vax3_date, "1 month")),
+    breaks = unique(lubridate::ceiling_date(data_coverage$vax3_date, "1 month")),
     limits = c(lubridate::floor_date((xmin), "1 month"), NA),
@@
 
-#### Figure 2a: Estimated booster effectiveness (5 period)
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
+
+### Figure 1b: Moderna booster vaccination uptake, treatment group eligibility and matching success
+
+Cumulative booster vaccination over the duration of the study period, by matching eligibility and matching success. Separately by primary vaccination course.
+
+```{r, vaxdateb, echo=FALSE, message=FALSE, warning=FALSE, out.width='60%', out.height='50%', results='asis', out.extra=''}
+#read_rds(fs::path(output_dir_os, "descriptive", "vaxdate", "plot_vaxdate_stack.rds"))
+
+xmin <- min(matchcoverage$vax3_date )
+xmax <- max(matchcoverage$vax3_date )+1
+data_coverage <- matchcoverage  %>%
+  mutate(
+    status = factor(status, levels=c("ineligible", "unmatched", "matched")),
+    status_descr = fct_relabel(status, str_to_title)
+  ) %>%
+  arrange(status) %>%
+  mutate(
+    status_descr = fct_inorder(status_descr)
+  ) %>%
+  filter(treatment=="moderna") %>%
+  mutate(
+    # vax12_type = fct_case_when(
+    #   vax12_type == "pfizer-pfizer" ~ "BNT162b2",
+    #   vax12_type == "az-az" ~ "ChAdOx1-S",
+    #   vax12_type == "moderna-moderna" ~ "mRNA-1273",
+    #   TRUE ~ NA_character_
+    # ),
+    course = fct_case_when(
+      #vax12_type=="pfizer-pfizer" & treatment=="pfizer" ~ "2× BNT162b2  / 1× BNT162b2",
+      #vax12_type=="az-az" & treatment=="pfizer" ~ "2× ChAdOx1-S / 1× BNT162b2",
+      vax12_type=="pfizer-pfizer" & treatment=="moderna" ~ "2× BNT162b2 / 1× mRNA-1273",
+      vax12_type=="az-az" & treatment=="moderna" ~ "2× ChAdOx1-S / 1× mRNA-1273",
+      TRUE ~ NA_character_
+    )
+  )
+
+data_annotate <- 
+  data_coverage %>%
+  filter(vax12_type=="pfizer-pfizer") %>%
+  group_by(status_descr, course) %>%
+  summarise(
+    xpos = first(lubridate::floor_date((vax3_date), "1 month")),
+    ypos= max(cumuln)
+  ) %>%
+  ungroup() %>%
+  arrange(desc(status_descr)) %>%
+  mutate(
+    ypos=(cumsum(ypos) + lag(cumsum(ypos), 1, 0))/2
+  ) %>%
+  mutate(
+    course="2× ChAdOx1-S / 1× mRNA-1273"
+  )
+
+ggplot(data_coverage) +
+  geom_col(
+    aes(
+      x=vax3_date+0.5,
+      y=cumuln,
+      group=status_descr,
+      fill=status_descr,
+      colour=NULL
+    ),
+    position=position_stack(),
+    alpha=0.5,
+    width=1
+  )+
+  geom_rect(xmin=xmin, xmax= xmax+1, ymin=0, ymax=6, fill="grey", colour="transparent")+
+  geom_text(data=data_annotate, mapping=aes(x = xpos, y = (ypos), label = status_descr, colour=status_descr), hjust=0, show.legend = FALSE)+
+  facet_wrap(vars(course))+
+  scale_x_date(
+    breaks = unique(lubridate::ceiling_date(data_coverage$vax3_date, "1 month")),
+    limits = c(lubridate::floor_date((xmin), "1 month"), NA),
+    labels = scales::label_date("%d/%m"),
+    expand = expansion(add=1),
+  )+
+  scale_y_continuous(
+    labels = scales::label_number(accuracy = 1, big.mark=","),
+    expand = expansion(c(0, NA))
+  )+
+  scale_fill_brewer(type="qual", palette="Set2", direction = -1)+
+  scale_colour_brewer(type="qual", palette="Set2", direction = -1)+
+  labs(
+    x="Date",
+    y="Cumulative booster vaccines",
+    colour=NULL,
+    fill=NULL,
+    alpha=NULL
+  ) +
+  theme_minimal()+
+  theme(
+    axis.line.x.bottom = element_line(),
+    axis.text.x.top=element_text(hjust=0),
+    strip.text.y.right = element_text(angle = 0),
+    axis.ticks.x=element_line(),
+    legend.position = "none"
+  )+
+  NULL
+cat("  \n\n")
+```
+
+\newpage
+
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
+
+### Figure 2: Estimated booster effectiveness (5 period)
 
@@
 
-```{r plot_effects, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}
+```{r plot_effects, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=9.5, out.width='90%', results='asis'}
 
@@
       outcome %in% outcomes,
-      treatment %in% c("pfizer"),
-      subgroup_variable %in% c("none", "vax12_type")
+      #treatment %in% c("pfizer"),
+      subgroup_variable %in% c("none", "vax12_type", "cev", "age65plus", "prior_covid_infection")
+    ) %>%
+    mutate(
+      subgroup_level_descr=fct_relabel(subgroup_level_descr, ~str_wrap(., width=10))
     )
@@
       outcome %in% outcomes,
-      treatment %in% c("pfizer"),
-      subgroup_variable %in% c("none", "vax12_type")
+      #treatment %in% c("pfizer"),
+      subgroup_variable %in% c("none", "vax12_type", "cev", "age65plus", "prior_covid_infection")
+    ) %>%
+    ungroup() %>%
+    mutate(
+      subgroup_level_descr_wrap=fct_relabel(subgroup_level_descr, ~str_wrap(.x, width=10))
     ) %>%
     ggplot()+
-    geom_point(aes(y=hr, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+
-    geom_linerange(aes(ymin=hr.ll, ymax=hr.ul, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+
-    geom_point(data = overallhazards_plotdata, mapping = aes(y=hr, x=max(postbaselinecuts+14)), colour="darkgrey") +
-    geom_linerange(data = overallhazards_plotdata, mapping = aes(ymin=hr.ll, ymax=hr.ul, x=max(postbaselinecuts+14)), colour="darkgrey") +
-    geom_rect(data = overallhazards_plotdata, ymin=-Inf, ymax=Inf, xmin=max(postbaselinecuts+14)-7, xmax=max(postbaselinecuts+14)+7, fill="grey95", alpha=0.5, colour="transparent") +
+    #geom_rect(data = overallhazards_plotdata, ymin=-Inf, ymax=Inf, xmin=max(postbaselinecuts+14)-7, xmax=max(postbaselinecuts+14)+7, fill="grey95", alpha=0.5, colour="transparent") +
+    geom_point(aes(y=hr, x=term_midpoint, colour=treatment_descr), position = position_dodge(width = 1.8))+
+    geom_linerange(aes(ymin=hr.ll, ymax=hr.ul, x=term_midpoint, colour=treatment_descr), position = position_dodge(width = 1.8))+
+    #geom_point(data = overallhazards_plotdata, mapping = aes(y=hr, x=max(postbaselinecuts+14), colour=treatment_descr),  position = position_dodge(width = 1.8)) +
+    #geom_linerange(data = overallhazards_plotdata, mapping = aes(ymin=hr.ll, ymax=hr.ul, x=max(postbaselinecuts+14), colour=treatment_descr),  position = position_dodge(width = 1.8)) +
+
     geom_hline(aes(yintercept=1), colour='grey')+
-    facet_grid(rows=vars(outcome_descr_wrap), cols=vars(subgroup_level_descr), switch="y")+
+    facet_grid(rows=vars(subgroup_level_descr_wrap), cols=vars(outcome_descr_wrap), switch="y")+
     scale_y_log10(
       breaks=y_breaks,
-      limits = c(0.01, 1.5),
+      limits = c(0.05, 1.5),
       oob = scales::oob_keep,
@@
     )+
-    scale_x_continuous(breaks=c(postbaselinecuts, max(postbaselinecuts+14)), labels=c(postbaselinecuts, "Overall"), limits=c(min(postbaselinecuts), max(postbaselinecuts+14)+14), expand = c(0, 0))+
-    scale_colour_brewer(type="qual", palette="Set2", guide=guide_legend(ncol=1))+
+    scale_x_continuous(breaks=c(postbaselinecuts), labels=c(postbaselinecuts), limits=c(min(postbaselinecuts), max(postbaselinecuts)), expand = c(0, 0))+
+    scale_colour_brewer(type="qual", palette="Set2", guide=guide_legend(ncol=2))+
     labs(
@@
 
-      legend.position = "none"
+      legend.position = "bottom",
+      legend.direction = "horizontal"
      ) +
@@
 
-#### Figure 2b: Estimated booster effectiveness (2 period split at 28 days)
-
-For each outcome based on the fully adjusted model, the period-specific hazard ratios for boosting with BNT162b2 versus not boosting is shown, stratified by primary course.
-
-```{r plot_effects28, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}
-
-formatpercent100 <- function(x,accuracy){
-  formatx <- scales::label_percent(accuracy, suffix="")(x)
-
-  if_else(
-    formatx==scales::label_percent(accuracy, suffix="")(1),
-    paste0(">",scales::label_percent(1, suffix="")((100-accuracy)/100)),
-    formatx
-  )
-}
- ## hazard ratio
-
-  y_breaks <- c(0.01, 0.02, 0.05, 0.1, 0.2, 0.5, 1, 2)
-
-  overallhazards_plotdata <- 
-    overallhazards %>%
-    filter(
-      model==3, 
-      outcome %in% outcomes,
-      treatment %in% c("pfizer"),
-      subgroup_variable %in% c("none", "vax12_type")
-    )
-
-  postbaselinecuts28 <- c(0,28,70)
-
-  plot_hazards <-
-    meta28hazards %>%
-    filter(
-      model==3,
-      outcome %in% outcomes,
-      treatment %in% c("pfizer"),
-      subgroup_variable %in% c("none", "vax12_type")
-    ) %>%
-    ggplot()+
-    geom_point(aes(y=hr, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+
-    geom_linerange(aes(ymin=hr.ll, ymax=hr.ul, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+
-    geom_point(data = overallhazards_plotdata, mapping = aes(y=hr, x=max(postbaselinecuts28+14)), colour="darkgrey") +
-    geom_linerange(data = overallhazards_plotdata, mapping = aes(ymin=hr.ll, ymax=hr.ul, x=max(postbaselinecuts28+14)), colour="darkgrey") +
-    geom_rect(data = overallhazards_plotdata, ymin=-Inf, ymax=Inf, xmin=max(postbaselinecuts28+14)-7, xmax=max(postbaselinecuts28+14)+7, fill="grey95", alpha=0.5, colour="transparent") +
-    geom_hline(aes(yintercept=1), colour='grey')+
-    facet_grid(rows=vars(outcome_descr_wrap), cols=vars(subgroup_level_descr), switch="y")+
-    scale_y_log10(
-      breaks=y_breaks,
-      limits = c(0.01, 1.5),
-      oob = scales::oob_keep,
-      sec.axis = sec_axis(
-        ~(1-.),
-        name="Effectiveness (%)",
-        breaks = 1-(y_breaks),
-        labels = function(x){formatpercent100(x, 1)}
-      )
-    )+
-    scale_x_continuous(breaks=c(postbaselinecuts28, max(postbaselinecuts28+14)), labels=c(postbaselinecuts28, "Overall"), limits=c(min(postbaselinecuts28), max(postbaselinecuts28+14)+14), expand = c(0, 0))+
-    scale_colour_brewer(type="qual", palette="Set2", guide=guide_legend(ncol=1))+
-    labs(
-      y="Hazard ratio",
-      x="Days since booster dose",
-      colour=NULL
-     ) +
-    theme_minimal()+
-    theme(
-      panel.border = element_blank(),
-      axis.line.y = element_line(colour = "black"),
-      axis.line.y.right = element_blank(),
-
-      panel.grid.minor.x = element_blank(),
-      panel.grid.minor.y = element_blank(),
-      strip.background = element_blank(),
-      strip.placement = "outside",
-      strip.text.y.left = element_text(angle = 0),
-
-      panel.spacing = unit(0.8, "lines"),
-
-      plot.title = element_text(hjust = 0),
-      plot.title.position = "plot",
-      plot.caption.position = "plot",
-      plot.caption = element_text(hjust = 0, face= "italic"),
-
-      legend.position = "none"
-     ) +
-    NULL
-  plot_hazards
-
-```
-
-\newpage
-
-#### Figure 2c: Estimated booster effectiveness (2 period split at 42 days)
-
-For each outcome based on the fully adjusted model, the period-specific hazard ratios for boosting with BNT162b2 versus not boosting is shown, stratified by primary course.
-
-```{r plot_effects42, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}
-
-formatpercent100 <- function(x,accuracy){
-  formatx <- scales::label_percent(accuracy, suffix="")(x)
-
-  if_else(
-    formatx==scales::label_percent(accuracy, suffix="")(1),
-    paste0(">",scales::label_percent(1, suffix="")((100-accuracy)/100)),
-    formatx
-  )
-}
- ## hazard ratio
-
-  y_breaks <- c(0.01, 0.02, 0.05, 0.1, 0.2, 0.5, 1, 2)
-
-  overallhazards_plotdata <- 
-    overallhazards %>%
-    filter(
-      model==3, 
-      outcome %in% outcomes,
-      treatment %in% c("pfizer"),
-      subgroup_variable %in% c("none", "vax12_type")
-    )
-
-  postbaselinecuts42 <- c(0,42,70)
-
-  plot_hazards <-
-    meta42hazards %>%
-    filter(
-      model==3,
-      outcome %in% outcomes,
-      treatment %in% c("pfizer"),
-      subgroup_variable %in% c("none", "vax12_type")
-    ) %>%
-    ggplot()+
-    geom_point(aes(y=hr, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+
-    geom_linerange(aes(ymin=hr.ll, ymax=hr.ul, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+
-    geom_point(data = overallhazards_plotdata, mapping = aes(y=hr, x=max(postbaselinecuts42+14)), colour="darkgrey") +
-    geom_linerange(data = overallhazards_plotdata, mapping = aes(ymin=hr.ll, ymax=hr.ul, x=max(postbaselinecuts42+14)), colour="darkgrey") +
-    geom_rect(data = overallhazards_plotdata, ymin=-Inf, ymax=Inf, xmin=max(postbaselinecuts42+14)-7, xmax=max(postbaselinecuts42+14)+7, fill="grey95", alpha=0.5, colour="transparent") +
-    geom_hline(aes(yintercept=1), colour='grey')+
-    facet_grid(rows=vars(outcome_descr_wrap), cols=vars(subgroup_level_descr), switch="y")+
-    scale_y_log10(
-      breaks=y_breaks,
-      limits = c(0.01, 1.5),
-      oob = scales::oob_keep,
-      sec.axis = sec_axis(
-        ~(1-.),
-        name="Effectiveness (%)",
-        breaks = 1-(y_breaks),
-        labels = function(x){formatpercent100(x, 1)}
-      )
-    )+
-    scale_x_continuous(breaks=c(postbaselinecuts42, max(postbaselinecuts42+14)), labels=c(postbaselinecuts42, "Overall"), limits=c(min(postbaselinecuts42), max(postbaselinecuts42+14)+14), expand = c(0, 0))+
-    scale_colour_brewer(type="qual", palette="Set2", guide=guide_legend(ncol=1))+
-    labs(
-      y="Hazard ratio",
-      x="Days since booster dose",
-      colour=NULL
-     ) +
-    theme_minimal()+
-    theme(
-      panel.border = element_blank(),
-      axis.line.y = element_line(colour = "black"),
-      axis.line.y.right = element_blank(),
-
-      panel.grid.minor.x = element_blank(),
-      panel.grid.minor.y = element_blank(),
-      strip.background = element_blank(),
-      strip.placement = "outside",
-      strip.text.y.left = element_text(angle = 0),
-
-      panel.spacing = unit(0.8, "lines"),
-
-      plot.title = element_text(hjust = 0),
-      plot.title.position = "plot",
-      plot.caption.position = "plot",
-      plot.caption = element_text(hjust = 0, face= "italic"),
-
-      legend.position = "none"
-     ) +
-    NULL
-
-  plot_hazards
-
-```
-
-\newpage
-
-## Figure 3: Kaplan-Meier incidence curves
-
-```{r plot_km, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}
-
-plot_km <- km %>%
-  filter(
-    outcome %in% outcomes,
-    treatment=="pfizer",
-    subgroup_variable %in% c("none", "vax12_type")
-  ) %>%
-    ggplot()+
-    geom_step(aes(x=time, y=(1-surv)*1000, group=treated_descr, colour=treated_descr))+
-    geom_rect(aes(xmin=time, xmax=leadtime, ymin=(1-surv.ll)*1000, ymax=(1-surv.ul)*1000, group=treated_descr, fill=treated_descr), alpha=0.1)+
-    geom_hline(aes(yintercept=0), colour='black')+
-    facet_grid(rows=vars(outcome_descr_wrap), cols=vars(subgroup_level_descr), switch="y", scales="free_y")+
-    scale_x_continuous(
-      breaks = seq(0,7*max(postbaselinecuts),by=14),
-      expand = expansion(0)
-    )+
-    scale_y_continuous(
-      expand = expansion(0),
-      #limits = c(0L,NA)
-    )+
-    scale_colour_brewer(type="qual", palette="Set1", na.value="grey")+
-    scale_fill_brewer(type="qual", palette="Set1", guide="none", na.value="grey")+
-    labs(
-      x=NULL,
-      y="Cumulative incidence\nper 1000",
-      colour=NULL
-    )+
-    theme_minimal()+
-    theme(
-      legend.position=c(0.05,0.85),
-      legend.justification = c(0,0),
-      legend.direction = "vertical",
-      axis.text.x.top=element_text(hjust=0),
-
-      strip.placement = "outside",
-      strip.background = element_blank(),
-      strip.text.y.left = element_text(angle = 0),
-
-      panel.border = element_blank(),
-      axis.line.y = element_line(colour = "black"),
-      axis.line.x.top = element_line(colour = "black"),
-      panel.grid.minor.x = element_blank(),
-      panel.grid.minor.y = element_blank(),
-
-      panel.spacing = unit(0.8, "lines"),
-
-    )+
-    NULL
-
-plot_km
-
-```
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
+
+&nbsp; &nbsp; &nbsp;
+
+<!-- #### Figure 2b: Estimated booster effectiveness (2 period split at 28 days) -->
+
+<!-- For each outcome based on the fully adjusted model, the period-specific hazard ratios for boosting with BNT162b2 versus not boosting is shown, stratified by primary course. -->
+
+<!-- ```{r plot_effects28, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'} -->
+
+<!-- formatpercent100 <- function(x,accuracy){ -->
+<!--   formatx <- scales::label_percent(accuracy, suffix="")(x) -->
+
+<!--   if_else( -->
+<!--     formatx==scales::label_percent(accuracy, suffix="")(1), -->
+<!--     paste0(">",scales::label_percent(1, suffix="")((100-accuracy)/100)), -->
+<!--     formatx -->
+<!--   ) -->
+<!-- } -->
+<!--  ## hazard ratio -->
+
+<!--   y_breaks <- c(0.01, 0.02, 0.05, 0.1, 0.2, 0.5, 1, 2) -->
+
+<!--   overallhazards_plotdata <-  -->
+<!--     overallhazards %>% -->
+<!--     filter( -->
+<!--       model==3,  -->
+<!--       outcome %in% outcomes, -->
+<!--       treatment %in% c("pfizer"), -->
+<!--       subgroup_variable %in% c("none", "vax12_type") -->
+<!--     ) -->
+
+<!--   postbaselinecuts28 <- c(0,28,70) -->
+
+<!--   plot_hazards <- -->
+<!--     meta28hazards %>% -->
+<!--     filter( -->
+<!--       model==3, -->
+<!--       outcome %in% outcomes, -->
+<!--       treatment %in% c("pfizer"), -->
+<!--       subgroup_variable %in% c("none", "vax12_type") -->
+<!--     ) %>% -->
+<!--     ggplot()+ -->
+<!--     geom_point(aes(y=hr, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+ -->
+<!--     geom_linerange(aes(ymin=hr.ll, ymax=hr.ul, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+ -->
+<!--     geom_point(data = overallhazards_plotdata, mapping = aes(y=hr, x=max(postbaselinecuts28+14)), colour="darkgrey") + -->
+<!--     geom_linerange(data = overallhazards_plotdata, mapping = aes(ymin=hr.ll, ymax=hr.ul, x=max(postbaselinecuts28+14)), colour="darkgrey") + -->
+<!--     geom_rect(data = overallhazards_plotdata, ymin=-Inf, ymax=Inf, xmin=max(postbaselinecuts28+14)-7, xmax=max(postbaselinecuts28+14)+7, fill="grey95", alpha=0.5, colour="transparent") + -->
+<!--     geom_hline(aes(yintercept=1), colour='grey')+ -->
+<!--     facet_grid(rows=vars(outcome_descr_wrap), cols=vars(subgroup_level_descr), switch="y")+ -->
+<!--     scale_y_log10( -->
+<!--       breaks=y_breaks, -->
+<!--       limits = c(0.01, 1.5), -->
+<!--       oob = scales::oob_keep, -->
+<!--       sec.axis = sec_axis( -->
+<!--         ~(1-.), -->
+<!--         name="Effectiveness (%)", -->
+<!--         breaks = 1-(y_breaks), -->
+<!--         labels = function(x){formatpercent100(x, 1)} -->
+<!--       ) -->
+<!--     )+ -->
+<!--     scale_x_continuous(breaks=c(postbaselinecuts28, max(postbaselinecuts28+14)), labels=c(postbaselinecuts28, "Overall"), limits=c(min(postbaselinecuts28), max(postbaselinecuts28+14)+14), expand = c(0, 0))+ -->
+<!--     scale_colour_brewer(type="qual", palette="Set2", guide=guide_legend(ncol=1))+ -->
+<!--     labs( -->
+<!--       y="Hazard ratio", -->
+<!--       x="Days since booster dose", -->
+<!--       colour=NULL -->
+<!--      ) + -->
+<!--     theme_minimal()+ -->
+<!--     theme( -->
+<!--       panel.border = element_blank(), -->
+<!--       axis.line.y = element_line(colour = "black"), -->
+<!--       axis.line.y.right = element_blank(), -->
+
+<!--       panel.grid.minor.x = element_blank(), -->
+<!--       panel.grid.minor.y = element_blank(), -->
+<!--       strip.background = element_blank(), -->
+<!--       strip.placement = "outside", -->
+<!--       strip.text.y.left = element_text(angle = 0), -->
+
+<!--       panel.spacing = unit(0.8, "lines"), -->
+
+<!--       plot.title = element_text(hjust = 0), -->
+<!--       plot.title.position = "plot", -->
+<!--       plot.caption.position = "plot", -->
+<!--       plot.caption = element_text(hjust = 0, face= "italic"), -->
+
+<!--       legend.position = "none" -->
+<!--      ) + -->
+<!--     NULL -->
+<!--   plot_hazards -->
+
+<!-- ``` -->
+
+<!-- \newpage  -->
+
+<!-- #### Figure 2c: Estimated booster effectiveness (2 period split at 42 days) -->
+
+<!-- For each outcome based on the fully adjusted model, the period-specific hazard ratios for boosting with BNT162b2 versus not boosting is shown, stratified by primary course. -->
+
+<!-- ```{r plot_effects42, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'} -->
+
+<!-- formatpercent100 <- function(x,accuracy){ -->
+<!--   formatx <- scales::label_percent(accuracy, suffix="")(x) -->
+
+<!--   if_else( -->
+<!--     formatx==scales::label_percent(accuracy, suffix="")(1), -->
+<!--     paste0(">",scales::label_percent(1, suffix="")((100-accuracy)/100)), -->
+<!--     formatx -->
+<!--   ) -->
+<!-- } -->
+<!--  ## hazard ratio -->
+
+<!--   y_breaks <- c(0.01, 0.02, 0.05, 0.1, 0.2, 0.5, 1, 2) -->
+
+<!--   overallhazards_plotdata <-  -->
+<!--     overallhazards %>% -->
+<!--     filter( -->
+<!--       model==3,  -->
+<!--       outcome %in% outcomes, -->
+<!--       treatment %in% c("pfizer"), -->
+<!--       subgroup_variable %in% c("none", "vax12_type") -->
+<!--     ) -->
+
+<!--   postbaselinecuts42 <- c(0,42,70) -->
+
+<!--   plot_hazards <- -->
+<!--     meta42hazards %>% -->
+<!--     filter( -->
+<!--       model==3, -->
+<!--       outcome %in% outcomes, -->
+<!--       treatment %in% c("pfizer"), -->
+<!--       subgroup_variable %in% c("none", "vax12_type") -->
+<!--     ) %>% -->
+<!--     ggplot()+ -->
+<!--     geom_point(aes(y=hr, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+ -->
+<!--     geom_linerange(aes(ymin=hr.ll, ymax=hr.ul, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+ -->
+<!--     geom_point(data = overallhazards_plotdata, mapping = aes(y=hr, x=max(postbaselinecuts42+14)), colour="darkgrey") + -->
+<!--     geom_linerange(data = overallhazards_plotdata, mapping = aes(ymin=hr.ll, ymax=hr.ul, x=max(postbaselinecuts42+14)), colour="darkgrey") + -->
+<!--     geom_rect(data = overallhazards_plotdata, ymin=-Inf, ymax=Inf, xmin=max(postbaselinecuts42+14)-7, xmax=max(postbaselinecuts42+14)+7, fill="grey95", alpha=0.5, colour="transparent") + -->
+<!--     geom_hline(aes(yintercept=1), colour='grey')+ -->
+<!--     facet_grid(rows=vars(outcome_descr_wrap), cols=vars(subgroup_level_descr), switch="y")+ -->
+<!--     scale_y_log10( -->
+<!--       breaks=y_breaks, -->
+<!--       limits = c(0.01, 1.5), -->
+<!--       oob = scales::oob_keep, -->
+<!--       sec.axis = sec_axis( -->
+<!--         ~(1-.), -->
+<!--         name="Effectiveness (%)", -->
+<!--         breaks = 1-(y_breaks), -->
+<!--         labels = function(x){formatpercent100(x, 1)} -->
+<!--       ) -->
+<!--     )+ -->
+<!--     scale_x_continuous(breaks=c(postbaselinecuts42, max(postbaselinecuts42+14)), labels=c(postbaselinecuts42, "Overall"), limits=c(min(postbaselinecuts42), max(postbaselinecuts42+14)+14), expand = c(0, 0))+ -->
+<!--     scale_colour_brewer(type="qual", palette="Set2", guide=guide_legend(ncol=1))+ -->
+<!--     labs( -->
+<!--       y="Hazard ratio", -->
+<!--       x="Days since booster dose", -->
+<!--       colour=NULL -->
+<!--      ) + -->
+<!--     theme_minimal()+ -->
+<!--     theme( -->
+<!--       panel.border = element_blank(), -->
+<!--       axis.line.y = element_line(colour = "black"), -->
+<!--       axis.line.y.right = element_blank(), -->
+
+<!--       panel.grid.minor.x = element_blank(), -->
+<!--       panel.grid.minor.y = element_blank(), -->
+<!--       strip.background = element_blank(), -->
+<!--       strip.placement = "outside", -->
+<!--       strip.text.y.left = element_text(angle = 0), -->
+
+<!--       panel.spacing = unit(0.8, "lines"), -->
+
+<!--       plot.title = element_text(hjust = 0), -->
+<!--       plot.title.position = "plot", -->
+<!--       plot.caption.position = "plot", -->
+<!--       plot.caption = element_text(hjust = 0, face= "italic"), -->
+
+<!--       legend.position = "none" -->
+<!--      ) + -->
+<!--     NULL -->
+
+<!--   plot_hazards -->
+
+<!-- ``` -->
+
+<!-- \newpage -->
+
+<!-- ## Figure 3: Kaplan-Meier incidence curves -->
+
+<!-- ```{r plot_km, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'} -->
+
+<!-- plot_km <- km %>% -->
+<!--   filter( -->
+<!--     outcome %in% outcomes, -->
+<!--     #treatment=="pfizer", -->
+<!--     subgroup_variable %in% c("none", "vax12_type") -->
+<!--   ) %>% -->
+<!--     ggplot()+ -->
+<!--     geom_step(aes(x=time, y=(1-surv)*1000, group=paste0(treated_descr,treatment_descr), alpha=treated_descr, colour=treatment_descr))+ -->
+<!--     geom_rect(aes(xmin=time, xmax=time+1, ymin=(1-surv.ll)*1000, ymax=(1-surv.ul)*1000, group=treated_descr, fill=treatment_descr), alpha=0.1)+ -->
+<!--     geom_hline(aes(yintercept=0), colour='black')+ -->
+<!--     facet_grid(rows=vars(subgroup_level_descr), cols=vars(outcome_descr_wrap), switch="y", scales="free_y")+ -->
+<!--     scale_x_continuous( -->
+<!--       breaks = seq(0,7*max(postbaselinecuts),by=14), -->
+<!--       expand = expansion(0) -->
+<!--     )+ -->
+<!--     scale_y_continuous( -->
+<!--       expand = expansion(0), -->
+<!--       #limits = c(0L,NA) -->
+<!--     )+ -->
+<!--     scale_colour_brewer(type="qual", palette="Set2", na.value="grey")+ -->
+<!--     scale_fill_brewer(type="qual", palette="Set2", guide="none", na.value="grey")+ -->
+<!--     labs( -->
+<!--       x=NULL, -->
+<!--       y="Cumulative incidence\nper 1000", -->
+<!--       colour=NULL, -->
+<!--       alpha=NULL -->
+<!--     )+ -->
+<!--     theme_minimal()+ -->
+<!--     theme( -->
+<!--       legend.position="bottom", -->
+<!--       legend.justification = c(0,0), -->
+<!--       legend.direction = "horizontal", -->
+<!--       axis.text.x.top=element_text(hjust=0), -->
+
+<!--       strip.placement = "outside", -->
+<!--       strip.background = element_blank(), -->
+<!--       strip.text.y.left = element_text(angle = 0), -->
+
+<!--       panel.border = element_blank(), -->
+<!--       axis.line.y = element_line(colour = "black"), -->
+<!--       axis.line.x.top = element_line(colour = "black"), -->
+<!--       panel.grid.minor.x = element_blank(), -->
+<!--       panel.grid.minor.y = element_blank(), -->
+
+<!--       panel.spacing = unit(0.8, "lines"), -->
+
+<!--     )+ -->
+<!--     NULL -->
+
+<!-- plot_km -->
+
+<!-- ``` -->

### 2022-03-29 analysis/study_definition.py
fix emergencyhosp definition

@@ -571,7 +571,6 @@ study = StudyDefinition(
     name = "emergencyhosp",
     n = 4,
     index_date = "index_date",
-    with_these_diagnoses = codelists.covid_emergency,
     discharged_to = codelists.discharged_to_hospital
   ),
   

### 2022-05-04 manuscript/draft-manuscript.Rmd
manuscript updates

@@
 output:
-  html_document:
-    self_contained: yes
+  word_document: default
   pdf_document:
@@
     toc: no
-  word_document: default
+  html_document:
+    self_contained: yes
 bibliography: references.bib
@@
 kmtreated_cuts <- kmconstrast(km_incidence, postbaselinecuts)
+kmtreated_28 <- kmconstrast(km_incidence, c(0,28, last(postbaselinecuts)))
 kmtreated_overall <- kmconstrast(km_incidence, c(0,last(postbaselinecuts)))
@@
 
+test<- read_csv(fs::path(output_dir_os, "prior_covid_infection", "effects.csv"))  %>% mutate(subgroup_level=as.character(subgroup_level))
+
 meta2hazards <- 
@@
     treatment, treatment_descr, outcome, outcome_descr, outcome_descr_wrap, subgroup, subgroup_variable, subgroup_descr, subgroup_level, subgroup_level_descr, model, model_descr, period, 
+  ) %>%
+  mutate(
+    # set prior if no events
+    std.error = if_else(is.na(estimate), 100, std.error),
+    robust.se = if_else(is.na(estimate), 100, robust.se),
+    estimate = replace_na(estimate, 0),
   ) %>%
@@
 
-William J Hulme^1^, Elizabeth J Williamson^2^, Elsie Horne^3,4^, Amelia Green^1^, Linda Nab^1^, Ruth Keogh^2^, Edward PK Parker^2^, Venexia Walker^3,4^, Helen Curtis^1^, Milan Wiedemann^1^, Christine Cunningham^1^, Alex J Walker^1^, Louis Fisher^1^, Christopher T Rentsch^2^, Anna Schultze^2^, Krishnan Bhaskaran^2^, Helen I McDonald^2^, John Tazare^2^, Laurie Tomlinson^2^, Tom Palmer^3,4^, Brian MacKenna^1^, Richard Croker^1^, Caroline E Morton^1^, Colm Andrews^1^, Robin Parks^1^, Lisa Hopcroft^1^, Jon Massey^1^, Jessica Morley^1^, Amir Mehrkar^1^, Seb Bacon^1^, Dave Evans^1^, Peter Inglesby^1^, George Hickman^1^, Simon Davy^1^, Tom Ward^1^, Rosalind M Eggo^2^, Angel YS Wong^2^, Rohini Mathur^2^, Kevin Wing^2^, Harriet Forbes^2^, Daniel Grint^2^, Viyasaan Mahalingasivam^2^, Bang Zheng^2^, Ian J Douglas^2^, Stephen JW Evans^2^, Liam Smeeth^2^, Chris Bates^5^, Jonathan Cockburn^5^, John Parry^5^, Frank Hester^5^, Sam Harper^5^, Jonathan AC Sterne^3,4^\*, Miguel A Hernán^6,7^\*, Ben Goldacre^1^\*.
+William J Hulme^1^, Elizabeth J Williamson^2^, Elsie Horne^3,4^, Amelia Green^1^, Linda Nab^1^, Ruth Keogh^2^, Edward PK Parker^2^, Venexia Walker^3,4^, Helen Curtis^1^, Milan Wiedemann^1^, Christine Cunningham^1^, Alex J Walker^1^, Louis Fisher^1^, Christopher T Rentsch^2^, Anna Schultze^2^, Krishnan Bhaskaran^2^, Helen I McDonald^2^, John Tazare^2^, Laurie Tomlinson^2^, Tom Palmer^3,4^, Brian MacKenna^1^, Richard Croker^1^, Caroline E Morton^1^, Colm Andrews^1^, Robin Parks^1^, Lisa Hopcroft^1^, Jon Massey^1^, Jessica Morley^1^, Amir Mehrkar^1^, Seb Bacon^1^, Dave Evans^1^, Peter Inglesby^1^, George Hickman^1^, Simon Davy^1^, Tom Ward^1^, Rosalind M Eggo^2^, Angel YS Wong^2^, Rohini Mathur^2^, Kevin Wing^2^, Harriet Forbes^2^, Daniel Grint^2^, Viyasaan Mahalingasivam^2^, Bang Zheng^2^, Ian J Douglas^2^, Stephen JW Evans^2^, Liam Smeeth^2^, Chris Bates^5^, Jonathan AC Sterne^3,4^\*, Miguel A Hernán^6,7^\*, Ben Goldacre^1^\*.
 
@@
 
-\* equal contribution.
-\newpage
+\* equal contribution. \newpage
 
@@
 
-We used the OpenSAFELY-TPP database, covering 40% of English primary care practices in England and linked to national coronavirus surveillance, hospital episodes, and death registry data, to estimate the effectiveness of BNT162b2 in eligible adults between `r format(as.Date(study_dates$pfizerstart_date), "%e %B")` and `r format(as.Date(study_dates$pfizerend_date), "%e %B %Y")`, a period when the Delta variant of SARS-CoV-2 was dominant.
+We used the OpenSAFELY-TPP database, covering 40% of English primary care practices and linked to national coronavirus surveillance, hospital episodes, and death registry data, to estimate the effectiveness of BNT162b2 in eligible adults between `r format(as.Date(study_dates$pfizerstart_date), "%e %B")` and `r format(as.Date(study_dates$pfizerend_date), "%e %B %Y")` when the Delta variant of SARS-CoV-2 was dominant.
 Follow up was for up to 10 weeks.
 Each booster recipient was matched with an unboosted control on factors relating to booster priority status and prior immunisation.
-Additional factors were adjusted for in a Cox model to estimate hazard ratios.
+Additional factors were adjusted for in a Cox model to estimate hazard ratios (HRs).
 Outcomes were positive SARS-CoV-2 test, COVID-19 hospital admission, COVID-19 death and non-COVID-9 death.
-Booster effectiveness was defined as 1 minus the hazard ratio.
+Booster effectiveness was defined as 1-HR.
 
@@
 We matched `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` BNT162b2 booster recipients to unboosted controls.
-Estimated booster effectiveness comparing a BNT162b2 booster dose to two doses only was `r filter(overallhazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3) %>% pull(ve_display95)` for positive SARS-CoV-2 test, `r filter(overallhazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for non-COVID-19 death.
-Estimated effectiveness was similar in those receiving a BNT162b2 or ChAdOx1-S two-dose primary vaccination course, but slightly lower in those classified as clinically extremely vulnerable (`r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for COVID-19 death) and in those aged 18-65 years (`r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` respectively).
-Period-specific effects were also estimated, showing some evidence of waning booster effectiveness against positive SARS-CoV-2 tests (`r filter(hazards, outcome=="postest", treatment=="pfizer", term_right==28, subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` at 15-28 days versus `r filter(hazards, outcome=="postest", treatment=="pfizer", term_right==70, subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` at 43-70 days after boosting).
+Estimated effectiveness of a BNT162b2 booster dose compared with two doses only was `r filter(overallhazards, subgroup=="none", outcome=="postest", treatment=="pfizer", model==3) %>% pull(ve_display95)` for positive SARS-CoV-2 test, `r filter(overallhazards, subgroup=="none", outcome=="covidadmitted", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, subgroup=="none", outcome=="coviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, subgroup=="none", outcome=="noncoviddeath", treatment=="pfizer", model==3) %>% pull(ve_display)` for non-COVID-19 death.
+Estimated effectiveness was similar among those who had received a BNT162b2 or ChAdOx1-S two-dose primary vaccination course, but slightly lower in those classified as clinically extremely vulnerable (`r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for COVID-19 death) and in those aged 18-65 years (`r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` respectively).
+
+<!-- Period-specific effects were also estimated, showing some evidence of waning booster effectiveness against positive SARS-CoV-2 tests (`r filter(hazards, outcome=="postest", treatment=="pfizer", term_right==28, subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` at 15-28 days versus `r filter(hazards, outcome=="postest", treatment=="pfizer", term_right==70, subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` at 43-70 days after boosting).
+-->
 
@@
 
-We estimated high protection of BNT162b2 boosting against positive SARS-CoV-2 test, COVID-19 hospital admission, and COVID-19 death.
+BNT162b2 boosting was strongly protective against positive SARS-CoV-2 test, COVID-19 hospital admission, and COVID-19 death.
 
@@
 
-In mid-September 2021, the national COVID-19 vaccination programme in England administered its first booster doses in adults who had already received their two-dose primary vaccination course @NHSEnglandNHS.
+In mid-September 2021 the national COVID-19 vaccination programme in England administered its first booster doses in adults who had already received their two-dose primary vaccination course @NHSEnglandNHS.
 Eligibility was initially restricted to those at highest risk of severe disease, then progressively extended.
@@
 <!-- https://www.england.nhs.uk/2021/09/nhs-begins-covid-19-booster-vaccination-campaign/ -->
+
 <!-- 13 december 30-39 become eligible (https://www.england.nhs.uk/2021/12/nhs-to-roll-out-life-saving-booster-jab-to-people-aged-30-plus/) -->
+
 <!-- 15 december everyone eligible https://www.england.nhs.uk/2021/12/nhs-booster-bookings-open-to-every-eligible-adult/ -->
@@
 We used the OpenSAFELY-TPP linked primary care database covering around 40% of English residents.
-Recruitment and follow-up covers an era where the Delta variant is dominant.
+The Delta SARS-CoV-2 variant was dominant during recruitment and follow-up.
 
@@
 We included adults aged $\ge$ 18 years who were eligible for booster vaccination between `r format(as.Date(study_dates$pfizerstart_date), "%e %B")` and `r format(as.Date(study_dates$pfizerend_date), "%e %B %Y")` inclusive (the "recruitment period").
-On each day within this period, eligibility for the study was determined as follows: registered at a GP practice using TPP's SystmOne clinical information system; received a two-dose primary vaccination course of either BNT162b2 or ChAdOx1-S (mixed dosing and Moderna mRNA-1273 were not considered due to small numbers); not a health or social care worker, not resident in a care or nursing home; not medically housebound or receiving end-of-life care; no evidence of SARS-CoV-2 infection or COVID-19 disease within the previous 90 days; not undergoing an unplanned hospital admission; complete information on sex, ethnicity, deprivation, and NHS region.
-
-Additionally, if the rolling weekly average count of BNT162b2 booster doses within strata defined by region, JCVI priority group, and the week of second dose was below 50 then recruitment did not occur for that stratum on that day.
-This helped to ensure that recruitment was restricted to those eligible and able to receive a booster dose in line with national prioritisation schedules at the time of potential recruitment.
-
-## Matching and treatment groups
+On each day within this period, eligibility for the study was determined as follows: registered at a GP practice using TPP's SystmOne clinical information system; received a two-dose primary vaccination course of either BNT162b2 or ChAdOx1-S (mixed dosing and Moderna mRNA-1273 were not considered due to small numbers); not a health or social care worker; not resident in a care or nursing home; not medically housebound or receiving end-of-life care; no evidence of SARS-CoV-2 infection or COVID-19 disease within the previous 90 days; not undergoing an unplanned hospital admission; complete information on sex, ethnicity, deprivation, and NHS region (East of England, Midlands, London, North East and Yorkshire, North West, South East, South West).
+
+Additionally, on days when the rolling weekly average count of BNT162b2 booster doses within strata defined by region, JCVI priority group, and the week of second dose was below 50, recruitment did not occur within the stratum.
+This helped to ensure that recruitment was restricted to those eligible and able to receive a booster dose in line with national prioritisation schedules.
+
+## Matching, treatment groups, and follow up
 
 On each day of the recruitment period, each eligible person receiving a booster dose was recruited to the treatment group and matched, if possible, with an eligible unboosted control person.
-Pairs were matched on: primary course vaccine brand; date of second vaccine dose, within 7 days; NHS region (East of England, Midlands, London, North East and Yorkshire, North West, South East, South West); evidence of prior SARS-CoV-2 infection (positive SARS-CoV-2 test, "probable" infection documented in primary care, or COVID-19 hospital admission); clinical risk groups used for prioritisation (clinically extremely vulnerable, clinically at-risk, neither); age, within 3 years; and age groups used by JCVI for prioritisation.
-<!-- The matching and exclusion criteria jointly ensure that treatment groups are also matched within JCVI priority group. -->
-
-People selected as controls were not eligible to be included again as a control, but were eligible for subsequent inclusion in the booster group. Any unmatched boosted people were excluded.
+Pairs were matched on: primary course vaccine brand; date of second vaccine dose, within 7 days; NHS region; evidence of prior SARS-CoV-2 infection (positive SARS-CoV-2 test, "probable" infection documented in primary care, or COVID-19 hospital admission); clinical risk groups used for prioritisation (clinically extremely vulnerable, clinically at-risk, neither); age within 3 years; and age groups used by JCVI for prioritisation.
+<!-- The matching and exclusion criteria jointly ensure that treatment groups are also matched within JCVI priority group.
+--> People selected as controls were not eligible to be included again as a control, but were eligible for subsequent inclusion in the booster group.
+Any unmatched boosted people were excluded.
+
+Each person was followed from the day of recruitment (i.e., time zero) until the earliest of outcome, death, practice de-registration, booster receipt by the control in the matching pair, 10 weeks, or `r format(as.Date(study_dates$studyend_date), "%e %B %Y")`.
 
@@
 
-## Follow-up
-
-Each person was followed from the day of recruitment (i.e., time zero) until the earliest of outcome, death, practice de-registration, booster receipt by the control in the matching pair, 10 weeks, or `r format(as.Date(study_dates$studyend_date), "%e %B %Y")`.
-
 ## Statistical Analysis
 
-We estimated Kaplan-Meier cumulative incidence curves separately in the boosted and unboosted groups, and compared groups using the 10-week risk difference and risk ratio.
-
-We used Cox models, stratified by recruitment day, NHS region, and two-dose primary vaccine course, to estimate hazard ratios (HRs) comparing boosted with unboosted people, overall and separately for days 1-28 and 29-70, and also for days 1-7, 8-14, 15-28, 29-42, and 43-70.
-As we were not able to match on all potential confounders, these models included the following additional covariates: sex (male or female); English Index of Multiple Deprivation (IMD, grouped by quintile); ethnicity (Black, Mixed, South Asian, White, Other, as per the UK census); prior morbidity count (diabetes, BMI \> 40kg/m^2^, chronic heart disease, chronic kidney disease, chronic liver disease, chronic respiratory disease or severe asthma, chronic neurological disease); learning disabilities; severe mental illness; the number of SARS-CoV-2 tests in the 6 months prior to the study start date; elective in-hospital episode at the time of recruitment.A robust variance estimator was used to account for inclusion of some people in both the boosted and unboosted groups.
-Booster vaccine effectiveness was calculated as one minus the hazard ratio, expressed as a percentage.
+We estimated Kaplan-Meier cumulative incidence curves separately in the boosted and unboosted groups, and estimated the 10-week risk differences based on these.
+We used Cox models, stratified by recruitment day, NHS region, and two-dose primary vaccine course, to estimate hazard ratios (HRs) comparing boosted with unboosted people, overall and separately for days 1-28 and 29-70.
+We also estimated HRs within shorter time periods (days 1-7, 8-14, 15-28, 29-42, and 43-70).
+
+As we were not able to match on all potential confounders, these models included the following additional covariates: sex (male or female); English Index of Multiple Deprivation (IMD, grouped by quintile); ethnicity (Black, Mixed, South Asian, White, Other, as per the UK census); prior morbidity count (diabetes, BMI \> 40kg/m^2^, chronic heart disease, chronic kidney disease, chronic liver disease, chronic respiratory disease or severe asthma, chronic neurological disease); learning disabilities; severe mental illness; the number of SARS-CoV-2 tests in the 6 months prior to the study start date; elective in-hospital episode at baseline.
+There were no missing values in these covariates, because they were defined by the presence or absence of clinical codes or events.
+A robust variance estimator was used to account for inclusion of some people in both the boosted and unboosted groups.
+
+Booster vaccine effectiveness was calculated as 1-HR, expressed as a percentage.
+We also estimated booster effectiveness separately in the following subgroups: primary vaccine course; age (18-64 or $\ge$ 65 years); clinically extremely vulnerable or not; evidence of prior SARS-CoV-2 infection or not.
 
@@
 
-We also estimated booster effectiveness separately in the following subgroups: primary vaccine course; age (18-64 or $\ge$ 65 years); clinically extremely vulnerable or not; evidence of prior SARS-CoV-2 infection or not.
-
 <!-- ## Missing data -->
@@
 `r label_number(1, big.mark=",")(filter(matchflowchart, crit=="c0", treatment=="pfizer") %>% pull(n))` adults registered at a TPP practice received a BNT162b2 booster vaccination during the recruitment period, with `r label_number(1, big.mark=",")(filter(matchflowchart, treatment=="pfizer", crit=="c7") %>% pull(n))` (`r label_percent(0.1)(filter(matchflowchart, treatment=="pfizer", crit=="c7") %>% pull(pct_all))`) eligible for inclusion in the treatment group, of whom `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` (`r label_percent(0.1)(filter(matchsummary, treatment=="pfizer") %>% pull(prop_matched_eligible))`) were matched with unboosted controls (Figure 1).
+1,714,615 people were matched as controls then rematched in the booster group, resulting in 6,990,219 people included in the study in total.
 
@@
 
-As expected, the matching factors has similar distributions in the boosted and unboosted groups at the start of follow up (Table 1), and the proportion of people with prior comorbidities was generally similar between the groups.
+As expected, the matching factors were identically distributed in the boosted and unboosted groups at the start of follow up (Table 1), and the proportion of people with prior morbidities was generally similar between the groups.
 However, unboosted controls had higher levels of deprivation, higher rates of learning disabilities and severe mental illness, and lower SARS-CoV-2 testing frequency, though the standardised mean difference was consistently below 0.1 (Supplementary Figure S2).
@@
 
-There were `r label_number(1, big.mark=",")(filter(incidence, treatment=="pfizer", outcome=="postest", fup_period=="All") %>% pull(events))` positive SARS-CoV-2 tests,
-`r label_number(1, big.mark=",")(filter(incidence, treatment=="pfizer", outcome=="covidadmitted", fup_period=="All") %>% pull(events))` COVID-19 hospital admissions, 
-`r label_number(1, big.mark=",")(filter(incidence, treatment=="pfizer", outcome=="coviddeath", fup_period=="All") %>% pull(events))` COVID-19 deaths, 
-and `r label_number(1, big.mark=",")(filter(incidence, treatment=="pfizer", outcome=="noncoviddeath", fup_period=="All") %>% pull(events))` non-COVID-19 deaths 
-across `r label_number(1, big.mark=",")((filter(incidence, treatment=="pfizer", outcome=="coviddeath", fup_period=="All") %>% pull(yearsatrisk)) * 365.25/7)` person-weeks of follow-up.
-
-The 10-week risk of positive SARS-CoV-2 test was `r km70 %>% filter(treatment=="pfizer", outcome=="postest", subgroup=="none") %>% pull(risk_1) %>% (label_number(0.1, scale=1000))` per 1000 in the boosted group
-and `r km70 %>% filter(treatment=="pfizer", outcome=="postest", subgroup=="none") %>% pull(risk_0) %>% (label_number(0.1, scale=1000))` in the unboosted group, 
-representing a risk difference of `r km70 %>% filter(treatment=="pfizer", outcome=="postest", subgroup=="none") %>% pull(kmrd) %>% (label_number(0.1, scale=1000))` per 1000, .
-For COVID-19 hospital admission, this was `r km70 %>% filter(treatment=="pfizer", outcome=="covidadmitted", subgroup=="none") %>% pull(risk_1) %>% (label_number(0.1, scale=1000))` 
-and `r km70 %>% filter(treatment=="pfizer", outcome=="covidadmitted", subgroup=="none") %>% pull(risk_0) %>% (label_number(0.1, scale=1000))` 
-respectively (`r km70 %>% filter(treatment=="pfizer", outcome=="covidadmitted", subgroup=="none") %>% pull(kmrd) %>% (label_number(0.1, scale=-1000))` fewer events in the boosted group). 
-For COVID-19 death it was `r km70 %>% filter(treatment=="pfizer", outcome=="coviddeath", subgroup=="none") %>% pull(risk_1) %>% (label_number(0.1, scale=1000))` 
-and `r km70 %>% filter(treatment=="pfizer", outcome=="coviddeath", subgroup=="none") %>% pull(risk_0) %>% (label_number(0.1, scale=1000))` 
-(`r km70 %>% filter(treatment=="pfizer", outcome=="coviddeath", subgroup=="none") %>% pull(kmrd) %>% (label_number(0.1, scale=-1000))` fewer events), 
-and for non-COVID-19 death it was `r km70 %>% filter(treatment=="pfizer", outcome=="noncoviddeath", subgroup=="none") %>% pull(risk_1) %>% (label_number(0.1, scale=1000))` 
-and `r km70 %>% filter(treatment=="pfizer", outcome=="noncoviddeath", subgroup=="none") %>% pull(risk_0) %>% (label_number(0.1, scale=1000))` 
-(`r km70 %>% filter(treatment=="pfizer", outcome=="noncoviddeath", subgroup=="none") %>% pull(kmrd) %>% (label_number(0.1, scale=-1000))` fewer events) respectively  (Table 2).
-
+There were `r label_number(1, big.mark=",")(filter(incidence, treatment=="pfizer", outcome=="postest", fup_period=="All") %>% pull(events))` positive SARS-CoV-2 tests, `r label_number(1, big.mark=",")(filter(incidence, treatment=="pfizer", outcome=="covidadmitted", fup_period=="All") %>% pull(events))` COVID-19 hospital admissions, `r label_number(1, big.mark=",")(filter(incidence, treatment=="pfizer", outcome=="coviddeath", fup_period=="All") %>% pull(events))` COVID-19 deaths, and `r label_number(1, big.mark=",")(filter(incidence, treatment=="pfizer", outcome=="noncoviddeath", fup_period=="All") %>% pull(events))` non-COVID-19 deaths across `r label_number(1, big.mark=",")((filter(incidence, treatment=="pfizer", outcome=="coviddeath", fup_period=="All") %>% pull(yearsatrisk)) * 365.25/7)` person-weeks of follow-up.
 The rate of increase in the cumulative incidence of positive SARS-CoV-2 tests in boosted people tapered substantially around 7 days after boosting, something not observed in unboosted controls (Figure 2).
-Similar patterns were not observed for more severe outcomes, with differences in the cumulative incidence between the groups apparent in the first few days.
+Similar patterns were not observed for more severe outcomes, and there were differences in the cumulative incidence between the groups apparent in the first few days.
+
+The 10-week risk of positive SARS-CoV-2 test was `r km70 %>% filter(treatment=="pfizer", outcome=="postest", subgroup=="none") %>% pull(risk_1) %>% (label_number(0.1, scale=1000))` and `r km70 %>% filter(treatment=="pfizer", outcome=="postest", subgroup=="none") %>% pull(risk_0) %>% (label_number(0.1, scale=1000))` per 1000 in the boosted group and unboosted groups respectively (risk difference `r km70 %>% filter(treatment=="pfizer", outcome=="postest", subgroup=="none") %>% pull(kmrd) %>% (label_number(0.1, scale=1000))` per 1000).
+Corresponding 10-week risks per 1000 were `r km70 %>% filter(treatment=="pfizer", outcome=="covidadmitted", subgroup=="none") %>% pull(risk_1) %>% (label_number(0.1, scale=1000))` and `r km70 %>% filter(treatment=="pfizer", outcome=="covidadmitted", subgroup=="none") %>% pull(risk_0) %>% (label_number(0.1, scale=1000))` for COVID-19 hospital admission (risk difference `r km70 %>% filter(treatment=="pfizer", outcome=="covidadmitted", subgroup=="none") %>% pull(kmrd) %>% (label_number(0.1, scale=1000))`), `r km70 %>% filter(treatment=="pfizer", outcome=="coviddeath", subgroup=="none") %>% pull(risk_1) %>% (label_number(0.1, scale=1000))` and `r km70 %>% filter(treatment=="pfizer", outcome=="coviddeath", subgroup=="none") %>% pull(risk_0) %>% (label_number(0.1, scale=1000))` for COVID-19 death (`r km70 %>% filter(treatment=="pfizer", outcome=="coviddeath", subgroup=="none") %>% pull(kmrd) %>% (label_number(0.1, scale=1000))`), and `r km70 %>% filter(treatment=="pfizer", outcome=="noncoviddeath", subgroup=="none") %>% pull(risk_1) %>% (label_number(0.1, scale=1000))` and `r km70 %>% filter(treatment=="pfizer", outcome=="noncoviddeath", subgroup=="none") %>% pull(risk_0) %>% (label_number(0.1, scale=1000))` for non-COVID-19 death (`r km70 %>% filter(treatment=="pfizer", outcome=="noncoviddeath", subgroup=="none") %>% pull(kmrd) %>% (label_number(0.1, scale=1000))`).
 
@@
 For those who had received BNT162b2, estimated booster effectiveness was `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup_level=="pfizer-pfizer", model==3) %>% pull(ve_display)` for non-COVID-19 death.
-Corresponding estimates of booster effectiveness for ChAdOx1-S recipients were `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup_level=="az-az", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup_level=="az-az", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup_level=="az-az", model==3) %>% pull(ve_display)`, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup_level=="az-az", model==3) %>% pull(ve_display)` respectively.
-
-Amongst people classified as clinically extremely vulnerable, estimated booster effectiveness was `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for non-COVID-19 death.
-For people who were not classified as clinically extremely vulnerable, estimated booster effectiveness was similar for positive SARS-CoV-2 tests (`r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="cev-FALSE", model==3) %>% pull(ve_display)` and non-COVID-19 death (`r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="cev-FALSE", model==3) %>% pull(ve_display)`), but higher for COVID-19 hospital admission and death (`r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="cev-FALSE", model==3) %>% pull(ve_display)` and `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="cev-FALSE", model==3) %>% pull(ve_display)` respectively).
-
-Amongst people aged $\ge$ 65 years, estimated booster effectiveness was `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for non-COVID-19 death.
-Corresponding effectiveness was similar in people aged 18-64 years for positive SARS-CoV-2 tests (`r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)`), but slightly lower for more severe outcomes: `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` for non-COVID-19 death.
+Corresponding estimates for ChAdOx1-S recipients were `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup_level=="az-az", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup_level=="az-az", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup_level=="az-az", model==3) %>% pull(ve_display)`, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup_level=="az-az", model==3) %>% pull(ve_display)` respectively.
+
+Among people classified as clinically extremely vulnerable, estimated booster effectiveness was `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, and `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` for COVID-19 death.
+Among people not classified as clinically extremely vulnerable estimated booster effectiveness was slightly higher, at `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="cev-FALSE", model==3) %>% pull(ve_display)` , `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="cev-FALSE", model==3) %>% pull(ve_display)` and `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="cev-FALSE", model==3) %>% pull(ve_display)` respectively.
+Estimated booster effectiveness for non-COVID-19 death was similar for those who were and were not classified as clinically extremely vulnerable: `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="cev-TRUE", model==3) %>% pull(ve_display)` and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="cev-FALSE", model==3) %>% pull(ve_display)` respectively.
+
+Among people aged $\ge$ 65 years, estimated booster effectiveness was `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 test, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="age65plus-TRUE", model==3) %>% pull(ve_display)` for non-COVID-19 death.
+Corresponding effectiveness appeared lower in people aged 18-64 years: `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` for positive SARS-CoV-2 tests, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` for COVID-19 hospital admission, `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` for COVID-19 death, and `r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="age65plus-FALSE", model==3) %>% pull(ve_display)` for non-COVID-19 death.
 
@@
 Corresponding estimates in people with no evidence of prior infection were `r filter(overallhazards, outcome=="postest", treatment=="pfizer", subgroup=="prior_covid_infection-FALSE", model==3) %>% pull(ve_display)`, `r filter(overallhazards, outcome=="covidadmitted", treatment=="pfizer", subgroup=="prior_covid_infection-FALSE", model==3) %>% pull(ve_display)`, and `r filter(overallhazards, outcome=="coviddeath", treatment=="pfizer", subgroup=="prior_covid_infection-FALSE", model==3) %>% pull(ve_display)`, respectively.
-For non-COVID-19, estimated booster effectiveness was slightly lower in those with prior evidence of SARS-CoV-2 infection (`r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="prior_covid_infection-TRUE", model==3) %>% pull(ve_display)`) than in those without (`r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="prior_covid_infection-FALSE", model==3) %>% pull(ve_display)`).
-
-Booster effectiveness was also estimated using narrower periods, with effectiveness against positive SARS-CoV-2 tests appearing to wane in later periods (`r filter(hazards, outcome=="postest", treatment=="pfizer", term_right==28, subgroup=="none", model==3) %>% pull(ve_display)` during days 15-28 after boosting versus `r filter(hazards, outcome=="postest", treatment=="pfizer", term_right==70, subgroup=="none", model==3) %>% pull(ve_display)` during days 43-70), a pattern that persisted in primary course subgroup analyses (Figure 3).
-For more severe outcomes, period-specific effectiveness estimates are less clear due to fewer events, though were broadly similar between different primary course brands.
+For non-COVID-19 death, estimated booster effectiveness was slightly lower in those with prior evidence of SARS-CoV-2 infection (`r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="prior_covid_infection-TRUE", model==3) %>% pull(ve_display)`) than in those without (`r filter(overallhazards, outcome=="noncoviddeath", treatment=="pfizer", subgroup=="prior_covid_infection-FALSE", model==3) %>% pull(ve_display)`).
+
+When booster effectiveness was estimated within shorter time periods, effectiveness against positive SARS-CoV-2 tests appeared to wane towards the end of follow up (`r filter(hazards, outcome=="postest", treatment=="pfizer", term_right==28, subgroup=="none", model==3) %>% pull(ve_display95)` during days 15-28 versus `r filter(hazards, outcome=="postest", treatment=="pfizer", term_right==70, subgroup=="none", model==3) %>% pull(ve_display)` during days 43-70).
+This pattern was observed in each subgroup defined by primary course vaccine brand (Figure 3).
+For more severe outcomes, period-specific effectiveness estimates were imprecisely estimated but broadly similar between primary course vaccine brands.
 
@@
 
-This observational study investigated the effectiveness of a BNT162b2 booster dose in `r label_number(1, big.mark=",")(filter(matchsummary, treatment=="pfizer") %>% pull(n_matched))` adults, and their matched controls, in England in the Delta era.
-We found high protection against all studied outcomes.
-There were apparent waning protection against positive SARS-CoV-2 tests around 6 weeks after boosting, which is unlikely to be explained by increased transmissibility of the Omicron variant given Omicron was not yet dominant in the UK by the end of the follow-up window @COVID19OmicronDaily.
-Changing testing behaviours over time between boosted and unboosted people may influence these estimates, if for example persistently unboosted participants, who will contribute more person-time in later follow-up periods, were less likely to get tested.
-
-Protection was similar between 2× BNT162b2 primary course recipients and 2× ChAdOx1-S recipients, except for COVID-19 death in the first 28 days where protected appeared higher for those whose primary course was 2× BNT162b2 than for 2× ChAdOx1-S.
-For severe outcomes, protection was lower in the clinical extremely vulnerable sub-population than in those who were not vulnerable.
-
-We considered effectiveness in a population where Delta was the predominant circulating variant @COVID19OmicronDaily.
-Re-running the analysis against a backdrop of the now globally-dominant Omicron variant will be undertaken where valid concurrent comparisons between boosted and unboosted people can be made.
-Further, mRNA-1273 boosting began later than for BNT162b2 so we did not evaluate effectiveness of this vaccine due to insufficient follow-up.
-An analysis of mRNA-1273 boosting and comparative effectiveness between mRNA-1273 and BNT162b2 is planned.
+In this observational cohort study based on nearly 7 million adults in England during a period when the Delta variant was dominant @COVID19OmicronDaily, a BNT162b2 booster dose reduced rates of positive SARS-CoV-2 tests by approximately 50% over the first 10 weeks after booster vaccination.
+There were also substantial reductions in rates of COVID-19 hospitalisation, COVID-19 death, and non-COVID-19 death.
+Booster effectiveness against these severe outcomes appeared similar regardless of whether the primary vaccine course was with ChAdOx1-S or BNT162b2, and whether there was evidence of prior infection or not.
+However, protection appeared lower in people classified as clinically extremely vulnerable compared with those not so classified, and lower in those aged under 65 years.
+Protection against positive SARS-CoV-2 tests appeared to wane from around 6 weeks after booster vaccination.
 
@@
 
-We used routinely-collected health records with comprehensive coverage of primary care, hospital admissions, COVID-19 testing and vaccination, and death registrations.
-This provides substantial power to estimate effectiveness of booster doses against severe but rare outcomes such as hospital admission and death, including period-specific effects to investigate waning protection.
-
-We carefully matched booster recipients with unboosted controls using characteristics known at baseline to strengthen exchangeability between treatment groups and reduce dependence on modelling assumptions in the estimates of effectiveness.
-However, despite reasonable balance of important baseline characteristics and additional adjustment for a range of potential confounders, the possibility of unmeasured confounding remains.
-In particular, the apparent protective effect immediately after boosting is biologically implausible, and will be explained in part by the occurrence at baseline of COVID-19 symptoms in the unboosted controls that were not recorded in the health record, and so could not be identified and excluded.
-Such symptoms are much less likely in booster recipients as symptomatic people, or those with a recent positive test, were ineligible for boosting.
-
-The matching approach is robust but inefficient; a minority of boosted individuals are not retained after matching, and those who are retained are censored when their matched control is boosted, reducing the follow-up duration and therefore statistical power.
-Millions of eligible participants may be required to obtain precise estimates, particularly when event rates are low, and so extending the approach to smaller subgroups may not be feasible.
-For instance, due to small numbers, we did not study booster effectiveness in those who had received the mRNA-1273 vaccine as their primary course or those who had received a heterologous primary course, despite many thousands of such recipients in England.
-
-Positive SARS-CoV-2 test data likely underestimates the true incidence of infection.
-The mass availability of lateral flow tests in the UK during the study period, whose results are not routinely recorded in COVID-19 surveillance data, means many infections, including symptomatic infections, may be undocumented, despite encouragement to seek a confirmatory PCR test.
-Potential differences in testing behaviour between boosted and unboosted people also contributes to the unreliability of testing data as a means to assess effectiveness @glasziou2022.
-SARS-CoV-2 testing was not widely available early in the pandemic so evidence of prior infection is likely to be under-ascertained.
-We may not have all records of negative tests so counts of number of tests are also incomplete.
-Hospital admission records are only completed after discharge, so some very long hospital stays commencing within the follow-up period may not have been included.
-
+We used routinely-collected electronic health care records with comprehensive coverage of primary care, hospital admissions, COVID-19 testing and vaccination, and death registrations.
+This enabled us to control for a comprehensive set of potential confounding factors.
+<!-- provides substantial power to estimate effectiveness of booster doses against severe but rare outcomes such as hospital admission and death, including period-specific effects to investigate waning protection.
+--> We carefully matched booster recipients with unboosted controls whose characteristics, including prior SARS-CoV-2 infection and COVID-29 disease, were similar at time time of booster vaccination.
+However, unmeasured confounding remains possible, and is most likely because people with COVID-19 symptoms are unlikely to receive a booster vaccination due to ineligibility but such symptoms are not routinely recorded.
+<!-- In particular, the apparent protective effect immediately after boosting is biologically implausible, and will be explained in part by the occurrence at baseline of COVID-19 symptoms in the unboosted controls that were not recorded in the health record, and so could not be identified and excluded.
+Such symptoms are much less likely in booster recipients as symptomatic people and those with a recent positive test, were ineligible for boosting.
+-->
+
+More general healthy vaccinee bias was addressed by excluding hospitalised people and other very sick people likely to be near the end of life.
+This is evidenced here by the apparent premature effectiveness against hospitalisation and COVID-19 death during the first week after boosting.
+The implausibly strong protection of booster vaccination observed against non-COVID-19 death further suggests this bias was not completely removed, potentially leading to an over-estimation of effectiveness against COVID-19 related outcomes.
+We are unable to compare our non-COVID-19 death results with other recent studies of booster effectiveness as they were not investigated.
+
+<!--# An alternative, or complimentary, explanation for the spurious non-COVID-19 booster effect is the possible misattribution of COVID-19 deaths to other causes, which would lead to an under-estimation of effectiveness against COVID-19 death.
+However, English mortality data is generally considered to be reliable so misattribution of cause of death is unlikely to be a major factor.
+-->
+
+We addressed ongoing booster vaccination by censoring follow up for both the booster recipient and their unboosted control when the latter received booster vaccination.
 We excluded a number of groups, such as health care workers and care home residents, where testing use, vaccination uptake, and infection risk were unusual or had substantial within-group heterogeneity that could not adequately measured and controlled for.
 The generalisability of our results to these excluded groups is unclear.
+Due to small numbers, we did not study booster effectiveness in those who had received the mRNA-1273 vaccine as their primary course or those who had received a heterologous primary course.
+
+<!-- The matching approach is robust but inefficient; a minority of boosted individuals are not retained after matching, and those who are retained are censored when their matched control is boosted, reducing the follow-up duration and therefore statistical power.
+Millions of eligible participants may be required to obtain precise estimates, particularly when event rates are low, and so extending the approach to smaller subgroups may not be feasible.
+-->
+
+Positive SARS-CoV-2 test data likely underestimates the true incidence of infection.
+Despite the mass availability of lateral flow tests in the UK during the study period, many infections, including symptomatic infections, will not be recorded in national data, despite encouragement to seek a confirmatory PCR test.
+Potential differences in testing behaviour between boosted and unboosted people also contributes to the unreliability of testing data as a means to assess effectiveness @glasziou2022.
+SARS-CoV-2 testing was not widely available early in the pandemic so prior infection is also likely to be under-ascertained.
+Hospital admission records are only completed after discharge, so some very long hospital stays commencing within the follow-up period may not have been included.
 
@@
 
-A phase III trial assessing BNT162b2 booster efficacy in two-dose BNT162b2 recipients without prior infection reported a relative efficacy of 95.6% (95% CI 89.3-98.6) against Delta infection @PfizerBioNTechAnnounce2021.
-Observational studies in the UK have provided evidence for increased protection from the booster vaccine against symptomatic COVID-19 infection in comparison to second dose recipients using a test negative design.
+A phase III trial in two-dose BNT162b2 recipients without prior infection reported a relative boster vaccination efficacy over a median of 2.5 months of 95.6% (95% CI 89.3-98.6) against Delta infection @PfizerBioNTechAnnounce2021.
+Observational studies in the UK have used 'test-negative' designs to provide evidence of the effectiveness of booster vaccination against symptomatic COVID-19 infection, in comparison to second dose recipients.
 Andrews et al.
-reported estimates of relative vaccine effectiveness against symptomatic disease in the 14 days after a BNT162b2 booster dose in those aged over 50 years of 87.4% (95%CI 84.9-89.4) where the primary course was ChAdOx1-S and 84.4% (95%CI 82.8-85.8) for BNT162b2 @andrews2021.
-A study in Scotland reported comparable estimates of relative vaccine effectiveness for S gene positive symptomatic infection (83% (95%CI 81-84) 6-49 years, 88% (95%CI 86-89) over 50 years) in the 14 days after a BNT162b2 or mRNA-1273 booster dose, but lower estimates for S gene negative symptomatic infection (56% (95%CI 51-60) 6-49 years, 57% (95%CI 52-62) over 50 years) @sheikhSeverityOmicronVariant2021.
-Elsewhere, a study in Israeli health registry data found strong protection against admission (1 -- risk ratio (95%CI) = 93% (88-97)), severe disease (92% (82-97)), and COVID-19 death (81% (59-97)), with similar results in specific demographic and clinical subgroups @barda2021.
-
-Our results broadly agree with these findings and provide important additional insights into effectiveness in specific clinical subgroups, younger age groups, and more severe outcomes over a longer period of follow up.
-Encouragingly, we found effectiveness was broadly similar irrespective of whether BNT162b2 or ChAdOx1-S was given for the primary vaccination course, and the were only slight reductions in effectiveness in those who were clinically extremely vulnerable.
-
-<!-- @1050236/technical-briefing-34-14-january-2022.pdf -->
+reported estimates of relative vaccine effectiveness in the first 14 days after BNT162b2 booster vaccination of 87.4% (95%CI 84.9-89.4) where the primary course was ChAdOx1-S and 84.4% (95%CI 82.8-85.8) for BNT162b2, in those aged over 50 years @andrews2022.
+A study in Scotland reported comparable estimates of relative vaccine effectiveness for S gene positive (a surrogate the Delta variant in the latter half of 2021 in the United Kingdom) symptomatic infection (83% (95%CI 81-84) 6-49 years, 88% (95%CI 86-89) over 50 years) in the 14 days after a BNT162b2 or mRNA-1273 booster dose, but lower estimates for S gene negative symptomatic infection (56% (95%CI 51-60) 6-49 years, 57% (95%CI 52-62) over 50 years) @sheikhSeverityOmicronVariant2021.
+We assessed effectiveness of booster vaccination against test positivity, combining both asymptomatic and symptomatic testing from documented PCR and lateral flow tests, and found substantially lower protection for this outcome compared with symptomatic infection only as previously estimated using test-negative designs in UK data.
+This provides prima facie evidence that booster vaccination is effective at reducing disease severity but not solely by reducing susceptibility to infection itself.
+It may also suggest vulnerability of the test-negative case-control design to bias when booster vaccination reduces disease severity @lewnard2018.
+
+A study in Israeli health registry data using 1:1 matching found strong protection against admission (1--risk ratio (95%CI) = 93% (88-97)), severe disease (92% (82-97)), and COVID-19 death (81% (59-97)), with similar results in specific demographic and clinical subgroups @barda2021.
+Our study, which also used matching but additionally adjusted for potential confounders not included as matching factors, found slightly lower estimates of effectiveness against hospitalization but higher estimates for death.
+
+Our results also provide important additional insights into effectiveness in clinical subgroups, younger age groups, and more severe outcomes over a longer period of follow up.
+Encouragingly, we found effectiveness was broadly similar irrespective of whether BNT162b2 or ChAdOx1-S was given for the primary vaccination course, and there were only slight reductions in effectiveness in those who were clinically extremely vulnerable or aged under 65 years.
+The absence of large discrepancies in booster effectiveness in different subgroups may simplify the coordination of future booster campaigns in the UK, allowing boosting to be prioritised on vulnerability to infection and severe disease without consideration for differential vaccine escape.
+
+The analysis will be re-run against a backdrop of the now globally-dominant Omicron variant where valid concurrent comparisons between boosted and unboosted people can be made.
+An analysis of mRNA-1273 boosting and comparative effectiveness between mRNA-1273 and BNT162b2 is also underway.
 
@@
 
-This study of over 8 million people in England found high protection for BNT162b2 boosting against positive SARS-CoV-2 tests, COVID-19 hospital admission, and death during a period in which the Delta variant was dominant.
+This study of almost 7 million people in England found high protection for BNT162b2 boosting against positive SARS-CoV-2 tests, COVID-19 hospital admission, and death during a period in which the Delta variant was dominant.
+Some evidence of waning effectiveness against positive SARS-CoV-2 tests suggests a need for careful monitoring of vaccine protection over time.
 
@@
     replace_na(scales::comma_format(accuracy=1)(numer), "--"),
-    " / ",
+    "/",
     str_pad(replace_na(scales::comma_format(accuracy=1)(denom),"--"), width=width, pad=" ")
@@
 km_table <- 
-  kmtreated_cuts %>%
+  kmtreated_28 %>%
   filter(
@@
   mutate(
-    q_0 = format_ratio(cml.event_0, cml.atrisk_0/365.25),
-    q_1 = format_ratio(cml.event_1, cml.atrisk_1/365.25),
+    q_0 = stringr::str_replace(format_ratio(cml.event_0, cml.atrisk_0/365.25),fixed(" "), ""),
+    q_1 = stringr::str_replace(format_ratio(cml.event_1, cml.atrisk_1/365.25),fixed(" "), ""),
   ) %>%

### 2022-05-04 manuscript/reports.Rmd
manuscript updates

@@
       #"covidemergency",
+      "covidadmittedproxy1",
       "covidadmitted",
@@
 
-meta2hazards <- 
+meta28hazards <- 
   # derive directly from hazards as meta2effects table needs to be re-run
@@
   )
-
-# meta2hazards <- 
+# 
+# meta28hazards <-
 #   bind_rows(
@@
 #   )
-
-hazards3 <- hazards %>% filter(model==3)
+# 
+# hazards3 <- hazards %>% filter(model==3)
 
@@
 
-### Figure 2: Estimated booster effectiveness (5 period)
+### Figure 2a: Estimated booster effectiveness (5 period)
 
@@
 
-<!-- #### Figure 2b: Estimated booster effectiveness (2 period split at 28 days) -->
-
-<!-- For each outcome based on the fully adjusted model, the period-specific hazard ratios for boosting with BNT162b2 versus not boosting is shown, stratified by primary course. -->
-
-<!-- ```{r plot_effects28, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'} -->
-
-<!-- formatpercent100 <- function(x,accuracy){ -->
-<!--   formatx <- scales::label_percent(accuracy, suffix="")(x) -->
-
-<!--   if_else( -->
-<!--     formatx==scales::label_percent(accuracy, suffix="")(1), -->
-<!--     paste0(">",scales::label_percent(1, suffix="")((100-accuracy)/100)), -->
-<!--     formatx -->
-<!--   ) -->
-<!-- } -->
-<!--  ## hazard ratio -->
-
-<!--   y_breaks <- c(0.01, 0.02, 0.05, 0.1, 0.2, 0.5, 1, 2) -->
-
-<!--   overallhazards_plotdata <-  -->
-<!--     overallhazards %>% -->
-<!--     filter( -->
-<!--       model==3,  -->
-<!--       outcome %in% outcomes, -->
-<!--       treatment %in% c("pfizer"), -->
-<!--       subgroup_variable %in% c("none", "vax12_type") -->
-<!--     ) -->
-
-<!--   postbaselinecuts28 <- c(0,28,70) -->
-
-<!--   plot_hazards <- -->
-<!--     meta28hazards %>% -->
-<!--     filter( -->
-<!--       model==3, -->
-<!--       outcome %in% outcomes, -->
-<!--       treatment %in% c("pfizer"), -->
-<!--       subgroup_variable %in% c("none", "vax12_type") -->
-<!--     ) %>% -->
-<!--     ggplot()+ -->
-<!--     geom_point(aes(y=hr, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+ -->
-<!--     geom_linerange(aes(ymin=hr.ll, ymax=hr.ul, x=term_midpoint, colour=model_descr), position = position_dodge(width = 1.8))+ -->
-<!--     geom_point(data = overallhazards_plotdata, mapping = aes(y=hr, x=max(postbaselinecuts28+14)), colour="darkgrey") + -->
-<!--     geom_linerange(data = overallhazards_plotdata, mapping = aes(ymin=hr.ll, ymax=hr.ul, x=max(postbaselinecuts28+14)), colour="darkgrey") + -->
-<!--     geom_rect(data = overallhazards_plotdata, ymin=-Inf, ymax=Inf, xmin=max(postbaselinecuts28+14)-7, xmax=max(postbaselinecuts28+14)+7, fill="grey95", alpha=0.5, colour="transparent") + -->
-<!--     geom_hline(aes(yintercept=1), colour='grey')+ -->
-<!--     facet_grid(rows=vars(outcome_descr_wrap), cols=vars(subgroup_level_descr), switch="y")+ -->
-<!--     scale_y_log10( -->
-<!--       breaks=y_breaks, -->
-<!--       limits = c(0.01, 1.5), -->
-<!--       oob = scales::oob_keep, -->
-<!--       sec.axis = sec_axis( -->
-<!--         ~(1-.), -->
-<!--         name="Effectiveness (%)", -->
-<!--         breaks = 1-(y_breaks), -->
-<!--         labels = function(x){formatpercent100(x, 1)} -->
-<!--       ) -->
-<!--     )+ -->
-<!--     scale_x_continuous(breaks=c(postbaselinecuts28, max(postbaselinecuts28+14)), labels=c(postbaselinecuts28, "Overall"), limits=c(min(postbaselinecuts28), max(postbaselinecuts28+14)+14), expand = c(0, 0))+ -->
-<!--     scale_colour_brewer(type="qual", palette="Set2", guide=guide_legend(ncol=1))+ -->
-<!--     labs( -->
-<!--       y="Hazard ratio", -->
-<!--       x="Days since booster dose", -->
-<!--       colour=NULL -->
-<!--      ) + -->
-<!--     theme_minimal()+ -->
-<!--     theme( -->
-<!--       panel.border = element_blank(), -->
-<!--       axis.line.y = element_line(colour = "black"), -->
-<!--       axis.line.y.right = element_blank(), -->
-
-<!--       panel.grid.minor.x = element_blank(), -->
-<!--       panel.grid.minor.y = element_blank(), -->
-<!--       strip.background = element_blank(), -->
-<!--       strip.placement = "outside", -->
-<!--       strip.text.y.left = element_text(angle = 0), -->
-
-<!--       panel.spacing = unit(0.8, "lines"), -->
-
-<!--       plot.title = element_text(hjust = 0), -->
-<!--       plot.title.position = "plot", -->
-<!--       plot.caption.position = "plot", -->
-<!--       plot.caption = element_text(hjust = 0, face= "italic"), -->
-
-<!--       legend.position = "none" -->
-<!--      ) + -->
-<!--     NULL -->
-<!--   plot_hazards -->
-
-<!-- ``` -->
-
-<!-- \newpage  -->
+#### Figure 2b: Estimated booster effectiveness (2 period split at 28 days)
+
+For each outcome based on the fully adjusted model, the period-specific hazard ratios for boosting with BNT162b2 versus not boosting is shown, stratified by primary course.
+
+```{r plot_effects28, echo=FALSE, message=FALSE, warning=FALSE, out.extra='', fig.height=7.5, out.width='90%', results='asis'}
+
+formatpercent100 <- function(x,accuracy){
+  formatx <- scales::label_percent(accuracy, suffix="")(x)
+
+  if_else(
+    formatx==scales::label_percent(accuracy, suffix="")(1),
+    paste0(">",scales::label_percent(1, suffix="")((100-accuracy)/100)),
+    formatx
+  )
+}
+ ## hazard ratio
+
+  y_breaks <- c(0.01, 0.02, 0.05, 0.1, 0.2, 0.5, 1, 2)
+
+  overallhazards_plotdata <-
+    overallhazards %>%
+    filter(
+      model==3,
+      outcome %in% outcomes,
+      #treatment %in% c("pfizer"),
+      subgroup_variable %in% c("none", "vax12_type", "cev", "age65plus", "prior_covid_infection")
+    ) %>% mutate(
+      subgroup_level_descr_wrap=fct_relabel(subgroup_level_descr, ~str_wrap(.x, width=10))
+    )
+
+  postbaselinecuts28 <- c(0,28,70)
+
+  plot_hazards <-
+    meta28hazards %>%
+    filter(
+      model==3,
+      outcome %in% outcomes,
+      #treatment %in% c("pfizer"),
+      subgroup_variable %in% c("none", "vax12_type", "cev", "age65plus", "prior_covid_infection")
+    ) %>%
+    mutate(
+      subgroup_level_descr_wrap=fct_relabel(subgroup_level_descr, ~str_wrap(.x, width=10))
+    ) %>%
+    ggplot()+
+    geom_point(aes(y=hr, x=term_midpoint, colour=treatment_descr), position = position_dodge(width = 1.8))+
+    geom_linerange(aes(ymin=hr.ll, ymax=hr.ul, x=term_midpoint, colour=treatment_descr), position = position_dodge(width = 1.8))+
+    geom_point(data = overallhazards_plotdata, mapping = aes(y=hr, x=max(postbaselinecuts28+14), colour=treatment_descr), position = position_dodge(width = 1.8)) +
+    geom_linerange(data = overallhazards_plotdata, mapping = aes(ymin=hr.ll, ymax=hr.ul, x=max(postbaselinecuts28+14), colour=treatment_descr), position = position_dodge(width = 1.8)) +
+    geom_rect(data = overallhazards_plotdata, ymin=-Inf, ymax=Inf, xmin=max(postbaselinecuts28+14)-7, xmax=max(postbaselinecuts28+14)+7, fill="grey95", alpha=0.4, colour="transparent") +
+    geom_hline(aes(yintercept=1), colour='grey')+
+    facet_grid(rows=vars(subgroup_level_descr_wrap), cols=vars(outcome_descr_wrap), switch="y")+
+    scale_y_log10(
+      breaks=y_breaks,
+      limits = c(0.05, 1.5),
+      oob = scales::oob_keep,
+      sec.axis = sec_axis(
+        ~(1-.),
+        name="Effectiveness (%)",
+        breaks = 1-(y_breaks),
+        labels = function(x){formatpercent100(x, 1)}
+      )
+    )+
+    scale_x_continuous(breaks=c(postbaselinecuts28, max(postbaselinecuts28+14)), labels=c(postbaselinecuts28, ""), limits=c(min(postbaselinecuts28), max(postbaselinecuts28+14)+14), expand = c(0, 0))+
+    scale_colour_brewer(type="qual", palette="Set2", guide=guide_legend(ncol=1))+
+    labs(
+      y="Hazard ratio",
+      x="Days since booster dose",
+      colour=NULL
+     ) +
+    theme_minimal()+
+    theme(
+      panel.border = element_blank(),
+      axis.line.y = element_line(colour = "black"),
+      axis.line.y.right = element_blank(),
+
+      panel.grid.minor.x = element_blank(),
+      panel.grid.minor.y = element_blank(),
+      strip.background = element_blank(),
+      strip.placement = "outside",
+      strip.text.y.left = element_text(angle = 0),
+
+      panel.spacing = unit(0.8, "lines"),
+
+      plot.title = element_text(hjust = 0),
+      plot.title.position = "plot",
+      plot.caption.position = "plot",
+      plot.caption = element_text(hjust = 0, face= "italic"),
+
+      legend.position = "bottom"
+     ) +
+    NULL
+  plot_hazards
+
+```
+
+\newpage
 

### 2022-05-04 lib/functions/survival.R

@@ -154,59 +154,3 @@ tidy_surv <-
 
     return(output)
   }
-#
-#
-#
-# tidy_flexsurvspline <- function(
-#   flexsurvsplinefit,
-#   times=NULL,
-#   addtimezero=FALSE
-# ){
-#
-#   # tidy post-fit flexsurv dataset
-#
-#   if(is.null(times)){
-#     times <- unique(flexsurvsplinefit$data$Y[,"time"])
-#   }
-#
-#   if(addtimezero){
-#     mintime <- min(times)
-#     timezero <- min(0, mintime-1)
-#     times <- unique(c(timezero, times))
-#   }
-#
-#
-#   summaryfs_survival <- summary(flexsurvsplinefit, type="survival", t=times, tidy=TRUE)
-#   names(summaryfs_survival)[1:4] <- c("time", "smooth_surv", "smooth_surv.ll", "smooth_surv.ul")
-#
-#   summaryfs_hazard <- summary(flexsurvsplinefit, type="hazard", t=times, tidy=TRUE)
-#   names(summaryfs_hazard)[1:4] <- c("time", "smooth_haz", "smooth_haz.ll", "smooth_haz.ul")
-#
-#   summaryfs_cmlhazard <- summary(flexsurvsplinefit, type="cumhaz", t=times, tidy=TRUE)
-#   names(summaryfs_cmlhazard)[1:4] <- c("time", "smooth_cml.haz", "smooth_cml.haz.ll", "smooth_cml.haz.ul")
-#
-#   output <- bind_cols(summaryfs_survival, summaryfs_hazard[,c(2,3,4)], summaryfs_cmlhazard[,c(2,3,4)])
-#
-#   return(output)
-#
-# }
-#
-#
-# get_hr <- function(tidy_flexsurv_group, group){
-#
-#   hrdat <- tidy_flexsurv_group %>%
-#     mutate(
-#       .trt = tidy_flexsurv[[group]]
-#     ) %>%
-#     select(time, .trt, smooth_haz) %>%
-#     pivot_wider(
-#       id_cols=c(time),
-#       names_from = .trt,
-#       values_from=c(smooth_haz)
-#     )
-#
-#   hrdat
-# }
-#
-#
-#
